---
title: "Employment Opportunities Predictor, County-level"
author: "JP Walsh and Aaron R. Williams"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
params:
  year: "2024"
---

*Program for the creation of the Employment Opportunities Upward Mobility Predictor at the County-level*

ACS Code: Initially written by Tina Chelidze and updated by JP Walsh 2025-2026.

Primary data derived from the IPUMS API. Based on processes developed by Paul Johnson and Kevin Werner in SAS.

## Executive Summary

**What this script does:** Calculates county-level employment-to-population ratios for adults ages 25-54 using American Community Survey (ACS) microdata.

**Key metric:** `share_employed` - the proportion of prime working-age adults (25-54) who are employed, with confidence intervals and quality flags.

**Geographic aggregation:** PUMA-level data crosswalked to county boundaries using allocation factors.

**Output:** Longitudinal dataset covering all U.S. counties with annual employment opportunity metrics.

**Compute requirements:** Large microdata files require significant memory. Recommended for server environments.

---

### User Notes

*Parameters:* Update `params$year` in the YAML header to specify which ACS year to process.

*Extract dates:* If updating the IPUMS extract, remember to update `ipums_extract_date` and `ipums_repwt_extract_date` in the Housekeeping section.

*AWS storage:* Internal users can store intermediate files on AWS S3. Authentication steps are explained in this [blog](https://www.gormanalysis.com/blog/connecting-to-aws-s3-with-r/). Contact the project lead for access credentials.

*Performance:* ACS microdata files are large and processing is time-intensive. Strongly recommended to use a server with significant computing power (e.g., Urban Institute SAS server).


## Housekeeping

### Load Required Packages

Read in packages for use. If you do not have an IPUMS API key you will need to make one via the [IPUMS website](https://www.ipums.org/).

```{r}
library(tidyverse)   # Data manipulation and visualization
library(ipumsr)      # IPUMS API interface
library(reactable)   # Interactive tables
library(srvyr)       # Survey-weighted statistics
library(scales)      # Formatting functions
library(furrr)       # Parallel processing
library(tictoc)      # Timing functions

options(scipen = 999)

theme_set(theme_minimal())

# DO NOT PUSH YOUR API KEY. You only have to run this once and then comment it out as below.
#set_ipums_api_key("Your KEY", save = TRUE)


s3_dir <- "metric_name/data/acs"
my_bucket <- "mobility-from-poverty-test"

# Load custom functions
source(here::here("functions", "API", "extract_ipums_aws.R"))              # Retrieves ACS microdata via IPUMS API
source(here::here("functions", "API", "ipums_repwt_employment_aws.R"))    # Retrieves replicate weights
source(here::here("functions", "testing", "evaluate_final_data.R"))       # Validates final output
source(here::here("01_employment-opportunities", "R", "finalize_metric.R"))            # Cleans confidence intervals
source(here::here("01_employment-opportunities", "R", "calc_survey_mean.R"))           # Computes survey-weighted means
source(here::here("01_employment-opportunities", "R", "calc_survey_mean_subgroup.R"))  # Computes subgroup means

```

### Configuration Parameters

**IPUMS Extract Dates**

Date of IPUMS extract. Denote the latest date that the IPUMS extract was changed and pulled both for the overall data and the repweights. The recommended date notation format is "mm_dd_yy". 

**Important:** For internal reviewers using AWS, keep this date as it was last entered by the metric lead unless you intend to change the extract. Changing this date will trigger a new data pull from IPUMS.

```{r}
ipums_extract_date <- "01_20_26"        # Date of main ACS microdata extract
ipums_repwt_extract_date <- "01_20_26"  # Date of replicate weights extract
```

**Single Year Update Flag**

If this is to update a single year of new data, set `single_year_update` to "Yes". This will process only the year specified in `params$year` and skip the validation section. If processing multiple years or running full validation, set to "No".

```{r}
single_year_update = "No"  # Set to "Yes" for single-year updates
```

**Existing Years for Longitudinal Analysis**

Define all years with existing intermediate data files. These will be combined for validation and final longitudinal output.

```{r}
existing_years <- c("2014", "2015", "2016", "2017", "2018", "2019", "2021", "2022", "2023", "2024")
```

## Methodology Overview

### Population Definition

**Target population:** Adults ages 25-54 (prime working age)
- This age range captures the core labor force, excluding younger individuals who may be in school and older individuals approaching retirement
- Only household residents included (group quarters excluded)
- Person-level weights adjusted for PUMA-to-county allocation

### Employment Definition

Employment status based on IPUMS EMPSTAT variable:
- **Employed:** EMPSTAT == 1
- **Excludes:** Not in labor force, unemployed, not in universe (institutionalized populations)

### Geographic Aggregation

**Challenge:** ACS microdata provides geography at the PUMA level (100,000+ population), but we need county-level estimates.

**Solution:** PUMA-to-county crosswalk using allocation factors (`afact`):
- `afact` represents the proportion of a PUMA's population residing in each overlapping county
- Person weights multiplied by `afact` to distribute individuals across counties
- Some PUMAs map entirely to one county (afact = 1.0)
- Others span multiple counties (afact < 1.0 for each)

### Quality Thresholds

**Sample size requirement:** Minimum of 30 effective samples per county
- Effective sample = sum of `afact` values (proxy for survey record count)
- Counties below threshold are suppressed (NA values in final data)

**Quality flags (1-3 scale):**
- **1 (Best):** High sample size, clean geographic allocation
- **2 (Moderate):** Acceptable sample size, some crosswalk complexity
- **3 (Caution):** Suppressed due to low sample size OR complex allocation

### Expected Output Counts

- **Pre-2022:** 3,143 counties
- **2022+:** 3,144 counties (reflects census geography updates)

## Read ACS Data

Read in the ACS extracts using the `extract_ipums_aws()` function. Make sure to change the survey list to reflect what years you want to include in the data. The surveys being used in the code are defined in the code chunk below. For a list of available ACS surveys and their titles via the IPUMSR package you can run the following function: get_sample_info("acs").

```{r}

acs_surveys <- list(paste0("us", params$year, "a"))

```

### Person-Level Data

Run extract. This will trigger an import process through IPUMS. It is encouraged to add detail to your extract description so it is clear what the extract was used for.

Date of latest extract noted above. If a new IPUMS pull is required the extract_date argument below must be updated to track that new data was downloaded. The extract only needs to be updated if changes or additions are required from the IPUMS pull (such as new dates or variables). 

```{r}
acs <- extract_ipums_aws(
  extract_name = paste0("umf_data_", params$year, "_1year_acs"),
  extract_date = ipums_extract_date,
  extract_description = paste0("Microdata pull for Mobility Metric Predictors.
  American Community Survey, overall,", params$year, "(1-year)."),
  survey = acs_surveys
)

```

Look at the distribution of survey samples in the data. The number of unique samples in the data should match the number of surveys selected in the `extract_ipums()` function above.

```{r}
count(acs, sample)
```

**Filter to prime working age population**

Isolate the data to 25-54 year-olds. This is the standard prime working-age range used in employment analysis:
- Excludes younger adults who may be completing education
- Excludes older adults approaching retirement age
- Provides most stable measure of labor force participation

Note: The replicate weights are also limited to these ages during the data pull to match this population. 

```{r}
acs_age <- acs %>% 
  filter(age >= 25 & age <= 54) 

```

### Replicate Weights 

Read in person-level replicate weights for the 25 to 54 year old population. These will be used in creating standard errors for the employment opportunity calculation. 

Date of latest extract listed above. If a new IPUMS pull is required the extract_date argument below must be updated to track that new data was downloaded. The extract only needs to be updated if changes or additions are required from the IPUMS pull (such as new dates or variables). 

```{r}
repwts <- ipums_repwt_employment_aws(
  extract_name = paste0("employment_replicate_weights", params$year, "_1year"),
  extract_date = ipums_repwt_extract_date,
  extract_description = paste0("Person replicate weights for Employment Opportunities 
  Metric Predictors. American Community Survey,", params$year, "(1-year)."),
  survey = acs_surveys
)  %>% 
  select(-cbserial, -cbpernum)
```

Look at the distribution of ACS survey samples in the data. The number of different samples should align with the number of surveys selected in the extract ipums function.

```{r}
count(repwts, sample)
```

Remove the sample variable.

```{r}
repwts_person <- repwts %>% 
  select(-sample)

```

### Merge on repwts

Merge the replicate weights on to the ACS samples. The length calls let us know the count of records going in and coming out of the join. 
The number of unique persons in acs_age should be the same as acs_combined. Check that this is the case.
```{r}

acs_combined <- 
  left_join(
    acs_age,
    repwts_person,
    by = "unique_person_id"
  )

stopifnot(length(unique(pull(acs_combined, unique_person_id))) == length(unique(pull(acs_age, unique_person_id))))

rm(acs_age)
rm(repwts_person)

```

Check that the merge did not result in any missing values.
```{r}
stopifnot(all(map_dbl(acs_combined, ~sum(is.na(.x))) == 0))
```

## Clean Data

Missing data is reported in a variety of ways by IPUMS data. This step walks through the missing values in key variables and checks that we are dealing with them appropriately. 

* **Age:** No missing values reported in variable description. Run the following test to confirm all values are between 25 and 54.

```{r}

stopifnot(all(acs_combined$age %in% c(25:54)))
          
```

* **Vacancy:** reported as "0". Note there should be no vacancy results by default of the structure of census data read in (all should be "0"). Confirm this is true with the following test.

```{r}
stopifnot(all(acs_combined$vacancy == "0"))
```

* **Empstat:** N/As reported as "0".
There are many records with the value missing for `empstat`. 

```{r}
acs_combined %>% 
  count(empstat) %>% 
  ggplot(mapping = aes(x = factor(empstat), y = n)) +
  geom_col() +
  geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Empstat, All Records",
    x = "Employment Status",
    y = NULL
  )
```

Turn the missing/"not in universe" value from the empstat variable into NAs. Rename to the more meaningful "employment_status". 

```{r}
acs_clean <- acs_combined %>%
  mutate(
    employment_status = ifelse(
      empstat == 0, 
      NA_integer_,
      empstat)
  )

```

Look at distribution of `employment_status` after adjustment. There should be no more records with the value "0".

```{r}
acs_clean %>% 
  count(employment_status) %>% 
  ggplot(mapping = aes(x = factor(employment_status), y = n)) +
  geom_col() +
  geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Employment, All Records",
    x = "Employment Status",
    y = NULL
  )

```

Filter for group quarters (GQ), we want to keep only households including additional households under updated definitions.

```{r}
count(acs_clean, gq)

```

```{r}
acs_clean <- acs_clean %>%
  filter(gq %in% c("Households under 1970 definition", 
                   "Additional households under 1990 definition",
                   "Additional households under 2000 definition"))
```

Check that group quarters fall only into household categories.

```{r}
acs_clean %>% 
  count(gq) %>% 
  ggplot(mapping = aes(x = gq, y = n)) +
  geom_col() +
  geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_discrete(labels = label_wrap(10)) +
  labs(
    title = "GQ, All Records",
    x = "GQ",
    y = NULL
  )

```

```{r}
rm(acs_combined)

```

## Crosswalk

Read in the PUMA to county crosswalk file. This file is created by the program generate_puma_county_crosswalks.rmd in the geographic-crosswalks folder. 

```{r}
puma_county_crosswalk <- read_csv(
  here::here("geographic-crosswalks", "data", "crosswalk_puma_to_county.csv")
)
```

Create a version with just unique counties for creating a version of the data that matches the UMF county list.

```{r}
if(params$year < 2022){
  county_list <- puma_county_crosswalk %>%
    filter(crosswalk_period == "pre-2022") %>% 
    select(statefip, county) %>% 
    filter(statefip != 72) %>% 
    unique()

}else{
  county_list <- puma_county_crosswalk %>% 
  filter(crosswalk_period == "2022") %>% 
  select(statefip, county) %>% 
  filter(statefip != 72) %>% 
  unique()
}
```

## Merge Crosswalk 

**Handle PUMA geography changes**

Create a variable in the ACS data indicating whether the survey is pre or post-2022. This is necessary because PUMA boundaries were redefined in 2022 based on the 2020 Census, requiring different crosswalk files.

```{r}
acs_clean <- acs_clean %>% 
  mutate(crosswalk_period = ifelse(year < 2022, "pre-2022", "2022"))

```

Check that the assignment worked. All years prior to 2022 should be designated "pre-2022".

```{r}
count(acs_clean, crosswalk_period, year)

```

**Join PUMA-to-county crosswalk**

Join the cleaned ACS data onto the crosswalk. This is a many-to-many join because:
- One PUMA can overlap multiple counties (person assigned to multiple counties with different afact values)
- Multiple PUMAs can contribute to one county
- The `afact` variable will be used later to weight person records appropriately

```{r}
acs_crosswalked <- left_join(
  acs_clean, 
  puma_county_crosswalk, 
  by = c("crosswalk_period", "statefip","puma"),
  relationship = "many-to-many"  # Expected: PUMAs map to multiple counties
)

```

Run an anti_join to see what data does not combine. We keep all counties so there should be no observations resulting from the anti-join.

```{r}
anti_join(acs_clean, 
          puma_county_crosswalk, 
          by = c("crosswalk_period", "statefip","puma")
)
```

Drop any observations with NA for `afact` (i.e. there is no county of interest overlapping this PUMA).

Create a count of how many observations this applies to.
```{r}
count(acs_crosswalked, is.na(afact))

acs_crosswalked <- acs_crosswalked %>% 
  drop_na(afact)

```

Also filter out cases where `afact` is equal to zero. These cases represent PUMAs with no overlap to counties of interest and will not be counted in the metric calculation.

Create a count of how many observations this applies to.
```{r}
count(acs_crosswalked, afact == 0)

acs_crosswalked <- acs_crosswalked %>% 
  filter(afact > 0)

```

**Apply allocation factors to survey weights**

Adjust the person weight and all replicate weights to account for PUMA-to-county mapping. This is the key step that distributes individuals across counties when their PUMA spans multiple counties.

**Example:** If a PUMA overlaps 3 counties with afact values of 0.5, 0.3, and 0.2:
- Each person in that PUMA appears 3 times in the data (one row per county)
- Their weights are multiplied by afact so they contribute fractionally to each county
- Sum of weighted contributions across counties equals their original weight

Select an example of the weight variables before and after this step. The weighting should show in slightly adjusted weight variables after this process. 
```{r}
select(acs_crosswalked, perwt, repwtp1, repwtp80, afact)

acs_crosswalked <- acs_crosswalked %>%
  mutate(perwt = perwt * afact,  # Main person weight
         across(matches("repwtp[0-9]+"), ~.x * afact))  # All 80 replicate weights

select(acs_crosswalked, perwt, repwtp1, repwtp80, afact) 
```

## Prepare computational variables

Create a binary employed variable for calculating the share employed. IPUMS data codes "Employed" as equal to 1.

```{r}

acs_crosswalked <- acs_crosswalked %>% 
  mutate(employed = employment_status == 1) 

```

Check if `employment_status` is ever `NA`.

```{r}
stopifnot(sum(is.na(acs_crosswalked$employment_status)) == 0)
```

Confirm there are no vacant properties included in the data.

```{r}
stopifnot(sum(is.na(acs_crosswalked$vacancy)) == 0)
```

## Create employment metric: All

### Pre-drop suppressed cells

First, to save time when processing the metric, calculate the effective sample count for each county, defined as the sum of the `afact` variable. This is used to create a cutoff for data quality based on the actual number of survey records being used in the calculation. We will not include counties with less than 30 effective samples.

```{r}
#| label: all-effective-samples

acs_all <- acs_crosswalked %>%
  group_by(year, sample, crosswalk_period, statefip, county) %>%
  mutate(effective_sample = sum(afact)) %>% 
  ungroup()

# calculate the number of statistics after suppressing values
acs_all %>%
  group_by(year, sample, crosswalk_period, statefip, county) %>%
  summarize(
    effective_sample = max(effective_sample)
  ) %>%
  ungroup() %>%
  summarize(
    original_stats = n(),
    unsuppressed_stats = sum(effective_sample >= 30)
  )

```

Create the employment metric. 

Objective: get the percent of individuals between the ages 25 and 54 that are employed from the EMPSTAT variable.

Aggregation should be weighted by PERWT (this is a person level statistic).

Remove records from counties below the size quality cutoff. 

```{r}
#| label: subset-all

acs_all <- acs_all %>% 
  filter(effective_sample >= 30)

```

### Calculate metric

**Compute survey-weighted employment rates**

Calculate the county-level metrics using replicate weights for variance estimation. 

**Processing approach:** We use `group_split()` by state FIPS code to process each state individually, then combine results. This approach:
- Reduces memory usage (srvyr survey objects are very large)
- Prevents memory overflow on large datasets
- Allows parallel processing if needed in future
- Each state processes independently, then results row-bind together

**Statistical method:** The `calc_survey_mean()` function:
1. Converts data to srvyr survey object with replicate weights
2. Applies `survey_mean()` to calculate weighted employment rate
3. Uses replicate weights to estimate standard errors
4. Produces confidence intervals (lower bound, upper bound)
5. Accounts for complex survey design

```{r}
#| label: calc-all

results_all <- acs_all %>% 
  group_split(statefip) %>%  # Split by state for memory management
  map_dfr(~ calc_survey_mean(.data = .x, .geo_level = county))  # Process and combine

```

Clean up the confidence interval bounds and sort the data.

```{r}
#| label: finalize-all

results_all <- finalize_metric(results_all)

```

```{r}
rm(acs_all)

```

## Add Suppressed Rows to Data

To create the metric we suppressed counties that did not meet the threshold for sample size but we need to include these in the final data. Create an all version of the data that includes counties that were suppressed prior to the survey_mean calculation. 

Expand the county_list file so it includes every year in the results. Join the results onto this file and test that the number of records matches the total number of counties we anticipate.
```{r}
expanded_county_list <- county_list %>% 
  left_join(results_all, by = c("statefip", "county"), relationship = "many-to-many") %>%
  mutate(geoid = paste0(statefip, county)) %>% 
  expand(year, geoid) %>% 
  mutate(statefip = str_sub(geoid, 1, 2),
         county = str_sub(geoid, 3, 5)) 

results_all_expand <- expanded_county_list %>% 
  left_join(results_all, by = c("year", "statefip", "county")) %>% 
  filter(!is.na(year))

if(params$year < 2022){
stopifnot(nrow(results_all_expand) == 3143 * n_distinct(results_all_expand$year))
}else{
  stopifnot(nrow(results_all_expand) == 3144 * n_distinct(results_all_expand$year))
}
```

## Data Quality Flags

Add a flag for data quality, this is a numeric variable between 1 and 3 with 1 representing the best quality and 3 representing the worst. 

The `employed_quality `variable combines quality information on the quality of the crosswalk and the sample size (effective sample) to create a final quality flag. Note that any county with NA for the share_employed metric is being suppressed due to sample size and is given a quality flag of 3.  

```{r}
metrics_employment_all <- results_all_expand %>% 
  mutate(
    share_employed_quality = if_else(
      is.na(share_employed), 
      NA_real_,
      geographic_allocation_quality
    )
  )
```

### Write out intermediate data

```{r}
if(isTRUE(bucket_exists(bucket = my_bucket))){
    # write file to S3
    tmp <- tempfile()
    on.exit(unlink(tmp))
    write_csv(metrics_employment_all, file = tmp)
    
    # put object with an upload progress bar
    put_object(tmp, object = paste0(s3_dir, "/metrics_employment_all_county", "_", params$year, ".csv"), bucket = my_bucket, 
               show_progress = FALSE, multipart = FALSE)
} else {
  #write file to temporary intermediate folder
  write_csv(paste0(here::here("01_employment-opportunities", "data", "intermediate"), 
                  "/", "metrics_employment_all_county_", params$year, ".csv"))
  }
```

If updating a single year the program ends here. To see validation results re-run with selection that all years of intermediate data are available.

```{r}
if(single_year_update == "Yes"){
  knitr::knit_exit()
}
```

## Data Dictionary

### Key Variables in Final Output

| Variable | Type | Description | Range/Values |
|----------|------|-------------|-------------|
| `year` | Integer | ACS survey year | 2014-2024 |
| `state` | Character | State FIPS code (2-digit) | 01-56 |
| `county` | Character | County FIPS code (3-digit) | 001-999 |
| `share_employed` | Numeric | Employment-to-population ratio for ages 25-54 | 0.0-1.0 |
| `share_employed_lb` | Numeric | Lower bound of 90% confidence interval | 0.0-1.0 |
| `share_employed_ub` | Numeric | Upper bound of 90% confidence interval | 0.0-1.0 |
| `share_employed_quality` | Integer | Quality flag (1=best, 3=suppressed) | 1, 2, 3, NA |

### Intermediate Variables

| Variable | Description | Source |
|----------|-------------|--------|
| `effective_sample` | Sum of afact values (sample size proxy) | Calculated from crosswalk |
| `afact` | Allocation factor for PUMA-county overlap | PUMA-county crosswalk |
| `employment_status` | Employment status from EMPSTAT | IPUMS ACS |
| `employed` | Binary indicator (TRUE if EMPSTAT==1) | Calculated |
| `geographic_allocation_quality` | Crosswalk quality indicator | PUMA-county crosswalk |
| `perwt` | Person weight (adjusted by afact) | IPUMS ACS (modified) |
| `repwtp1-repwtp80` | Replicate weights (adjusted by afact) | IPUMS ACS (modified) |

## Validation

Validate results for all years of available output.

### Read intermediate files

Read in the intermediate files created for each ACS year available.  

Function reads the CSVs from intermediate folder or from AWS if the user has access.

```{r}
read_intermediate <- function(year, file_type){
  
  if(isTRUE(bucket_exists(bucket = my_bucket))) {
    
    data_year <- s3read_using(FUN=read_csv, 
                 bucket = my_bucket, 
                 object=paste0(s3_dir, "/metrics_employment_", file_type, "_", year, ".csv")) %>% 
      mutate(crosswalk_period = as.character(crosswalk_period))
  
  }
  else{
  
  read_csv(paste0(here::here("01_employment-opportunities", "data", "intermediate"), 
                  "/", "metrics_employment_", file_type, "_", year, ".csv"))
  }
  
}
```

```{r}

metrics_employment_all <- map_df(.x = existing_years, .f = ~read_intermediate(year = .x,
                                                                          file_type = "all_county"))
```

Tabulate count of results by year.

```{r}
metrics_employment_all %>% 
  group_by(year) %>% 
  count()
```

Show summary statistics for the metric across all counties in the 1-year data.
```{r}
metrics_employment_all %>% 
  select(share_employed:share_employed_quality) %>% 
  summary()
```

Calculate the coefficient of variation for each year for the `share_employed` variable.
The higher this number the greater variability in the data. Create a table descending by CV to see counties with the largest variability.
```{r}
metrics_employment_all %>% 
  group_by(statefip, county) %>% 
  summarise(sd_employ = round(sd(share_employed, na.rm = TRUE), 3),
         mean_employ = round(mean(share_employed, na.rm = TRUE), 3),
         cv = round(sd_employ/mean_employ, 3)) %>% 
   arrange(desc(cv)) %>% 
  reactable()
```

Look at distribution of `share_employed` for all counties. We generally see a normal distribution. 
```{r}
metrics_employment_all %>% 
  filter(year == params$year) %>% 
  select(share_employed) %>% 
  ggplot(aes(x = share_employed)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle(paste0("Distribution of share employed, by county,", params$year, "(1-year data)")) +
  ylab("Density")
```

Look at the counts of the quality flag. For the 1-year data at the county level there are still a fair amount of quality "3" observations that we would suggest users disregard or use with caution. 

```{r}
metrics_employment_all %>% 
  select(share_employed_quality) %>% 
  ggplot(aes(x = share_employed_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  ggtitle("Quality Flag (1-year data)") +
  ylab("Count")
```

Tabulate share of data suppressed by year.
```{r}
metrics_employment_all %>% 
  group_by(year) %>% 
  filter(is.na(share_employed)) %>% 
  count()
```

## Final Test and Export

Rename state.

```{r}
metrics_employment_all <- metrics_employment_all %>% 
  rename("state" = "statefip")

```

Order the variables how we want.

```{r}
metrics_employment_all <- metrics_employment_all %>% 
  select(year, state, county, share_employed, share_employed_lb,
         share_employed_ub, share_employed_quality)

```

Run file through evaluate final data function.

```{r}

evaluate_final_data(data = metrics_employment_all, exp_form_path = here::here("99a_final-evaluation", "evaluation_form_employment_overall_county.csv"), geography = "county", subgroups = FALSE, confidence_intervals = TRUE)

```

### Export as CSV.

Export as CSV.

```{r}
write_csv(metrics_employment_all, here::here("01_employment-opportunities", "data", "final", "metrics_employment_county_all_longitudinal.csv"))

```

## Expected Outputs

### Intermediate Files

Stored in `01_employment-opportunities/data/intermediate/` or AWS S3 (if configured):

- **File pattern:** `metrics_employment_all_county_YYYY.csv`
- **One file per year:** Created during single-year processing
- **Used for:** Building longitudinal dataset and validation

### Final Output

**File:** `01_employment-opportunities/data/final/metrics_employment_county_all_longitudinal.csv`

**Structure:**
- **Rows:** ~3,143-3,144 counties × number of years processed
- **Columns:** 7 (year, state, county, share_employed, share_employed_lb, share_employed_ub, share_employed_quality)
- **Coverage:** All U.S. counties (excluding Puerto Rico)
- **Time period:** 2014-2024 (excluding 2020 due to COVID-19 ACS disruption)

**Quality notes:**
- Counties with <30 effective samples have NA for share_employed
- Quality flags indicate reliability (1=best, 3=use with caution)
- Confidence intervals provided for all non-suppressed estimates

## Troubleshooting

### Common Issues and Solutions

**1. IPUMS API Authentication Fails**
- **Error:** "API key not found" or "Authentication failed"
- **Solution:** Run `set_ipums_api_key("YOUR_KEY", save = TRUE)` once, then restart R session
- **Get key:** Register at [IPUMS website](https://www.ipums.org/)

**2. AWS S3 Access Denied**
- **Error:** "Bucket does not exist" or "Access Denied"
- **Solution:** Set AWS credentials using `Sys.setenv(AWS_ACCESS_KEY_ID="...", AWS_SECRET_ACCESS_KEY="...")`
- **Alternative:** Script will fall back to local file storage if AWS unavailable
- **Documentation:** See [AWS S3 with R blog post](https://www.gormanalysis.com/blog/connecting-to-aws-s3-with-r/)

**3. Memory Errors During Processing**
- **Error:** "Cannot allocate vector of size..." or R session crashes
- **Solution:** Run on server with more RAM (minimum 16GB recommended, 32GB+ preferred)
- **Alternative:** Process one state at a time by modifying the `group_split()` approach

**4. Extract Date Confusion**
- **When to update:** Only change `ipums_extract_date` if you've modified variables or years in the IPUMS extract
- **AWS users:** Changing the date triggers a new download; coordinate with team
- **Format:** Use "mm_dd_yy" format (e.g., "01_20_26")

**5. Mismatched County Counts**
- **Error:** `stopifnot()` fails on expected 3,143 or 3,144 counties
- **Cause:** Crosswalk file may be outdated or mismatched to year
- **Solution:** Verify crosswalk file matches the census period (pre-2022 vs 2022+)
- **Check:** Ensure `crosswalk_period` variable correctly assigned

**6. Missing Replicate Weights**
- **Error:** Join results in NA values for repwtp variables
- **Cause:** Separate replicate weights extract may have failed
- **Solution:** Check `ipums_repwt_extract_date` and verify extract completed
- **Alternative:** Re-run replicate weights extract independently

**7. Long Processing Time**
- **Expected:** 1-3 hours per year depending on server
- **Speed up:** Ensure running on server, not local machine
- **Monitor:** Use `tictoc` package markers to identify slow sections
- **Note:** The `calc_survey_mean()` step is computationally intensive

### Data Quality Checks

If validation results look unexpected:

1. **Check employment rate distribution:** Should be roughly normal, centered around 0.75-0.85
2. **Review quality flags:** High proportion of quality=3 may indicate sample size issues
3. **Examine coefficient of variation:** CV > 0.10 suggests high variability
4. **Compare to prior years:** Large year-over-year changes warrant investigation
5. **Verify suppression counts:** Should be relatively consistent across years

## Processing Workflow Summary

Complete workflow for reference:

1. **Data Extraction** → Pull ACS microdata and replicate weights from IPUMS API
2. **Age Filtering** → Subset to ages 25-54 (prime working age)
3. **Data Cleaning** → Handle missing values, filter to households only
4. **Geographic Crosswalk** → Map PUMAs to counties using allocation factors
5. **Weight Adjustment** → Multiply person and replicate weights by afact
6. **Sample Quality Check** → Calculate effective samples, suppress <30
7. **Metric Calculation** → Compute survey-weighted employment ratios with SEs
8. **Quality Flags** → Assign quality indicators based on sample size and crosswalk
9. **Expand to Full County List** → Add suppressed counties with NA values
10. **Validation** → Review distributions, CVs, and year-over-year changes
11. **Export** → Save intermediate yearly files and final longitudinal dataset

## References and Documentation

### Primary Data Source
- **IPUMS USA:** Steven Ruggles, Sarah Flood, Matthew Sobek, et al. IPUMS USA: Version 13.0 [dataset]. Minneapolis, MN: IPUMS, 2023. https://doi.org/10.18128/D010.V13.0

### Methodology Sources
- **Original SAS implementation:** Paul Johnson and Kevin Werner (methodology foundation)
- **R translation:** Tina Chelidze (initial), JP Walsh (updates 2025-2026)

### Related Documentation
- **PUMA-County Crosswalk:** Generated by `geographic-crosswalks/generate_puma_county_crosswalks.Rmd`
- **Project documentation:** `mobility-from-poverty-documentation/` folder
- **Custom functions:** `functions/` folder with detailed comments

### External Resources
- **IPUMS ACS documentation:** https://usa.ipums.org/usa/acs.shtml
- **ACS replicate weights:** https://usa.ipums.org/usa/repwt.shtml
- **Srvyr package:** http://gdfe.co/srvyr/
- **Survey methodology:** Valliant, Dever, and Kreuter (2018). *Practical Tools for Designing and Weighting Survey Samples*

### Support
For questions about this metric, contact the Upward Mobility Framework team at the Urban Institute.
