---
title: "College Readiness Predictor, Place-level"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

*Program for the creation of the College Readiness Upward Mobility Predictor at the place-level*

ACS Code: Initially written by Tina Chelidze and updated by JP Walsh 2022-2023.

Primary data derived from the IPUMS API. This program produces the metric only for the 450 places included in the UMF data dashboard (other places are left out).This metric is the share of 19- and 20-year-olds in a community who have a high school degree.

Based on processes developed by Paul Johnson and Kevin Werner in SAS.

## Housekeeping

Read in packages for use. If you do not have an IPUMS API key you will need to make one via the [IPUMS website](https://www.ipums.org/).

```{r, results='hide'}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)
library(srvyr)
library(scales)

options(scipen = 999)

# DO NOT PUSH YOUR API KEY. You only have to run this once and then comment it out as below.
#set_ipums_api_key("Your KEY", save = TRUE)

source(here::here("functions", "API", "extract_ipums.R"))
source(here::here("functions", "API", "ipums_repwt_college.R"))
source(here::here("08_education", "R", "finalize_metric_college.R"))
source(here::here("08_education", "R", "calc_survey_mean_college.R"))
source(here::here("08_education", "R", "calc_survey_mean_subgroup_college.R"))
```

## Read Data

Read in the ACS extracts using the extract_ipums function. Make sure to change the survey list to reflect what years you want to include in the data. Currently we are including two years of 5-year data (2018c and 2021c). 5-year 2022 data will be added pending its release.

*Note this download is slow and will take some time to finish. The notes on "Checking extract status" are a feature of the IPUMSR package and are to be expected*
```{r}
acs <- extract_ipums(
  extract_name = "umf_data_18_22",
  extract_description = "Microdata pull for College Readiness Metric Predictors. American Community Survey, years 2018, 2019 and 2022 (5-year).",
  survey = list("us2018c", "us2021c"))

```

Look at the distribution of ACS survey samples in the data. The number of different samples should align with the number of surveys selected in the extract ipums function.

```{r}
count(acs, sample)
```

Isolate the data to 19 and 20 year-olds. This is the age-range included in the college readiness calculation.
```{r}
acs <- acs %>%
  filter(age == 19 | age == 20) 
```

### Replicate Weights 

Read in person-level replicate weights for the 19 and 20 year old population. These will be used in creating standard errors for the college ready calculation. 
```{r}
repwts_person <- ipums_repwt_college(
  extract_name = "college_replicate_weights",
  extract_description = "Person replicate weights for College Readiness Metric Predictors. American Community Survey, years 2018 and 2022 (5-year).",
  survey = list("us2018c", "us2021c")
)  %>% 
  select(-cbserial, -cbpernum)
```

Look at the distribution of ACS survey samples in the data. The number of different samples should align with the number of surveys selected in the extract ipums function.

```{r}
count(repwts_person, sample)
```

Remove the sample variable.

```{r}
repwts_person <- repwts_person %>% 
  select(-sample)
```

### Merge on repwts

```{r}
length(unique(pull(acs, unique_person_id)))
length(unique(pull(repwts_person, unique_person_id)))
acs_combined <- 
  left_join(
    acs,
    repwts_person,
    by = "unique_person_id"
  )
length(unique(pull(acs_combined, unique_person_id)))
rm(acs)
rm(repwts_person)
```

Check that the merge did not result in any missing values.
```{r}
stopifnot(all(map_dbl(acs_combined, ~sum(is.na(.x))) == 0))
```

## Clean Data

Missing data is reported in a variety of ways by IPUMS data. This step walks through the missing values in key variables and checks that we are dealing with them appropriately. 

Age: No missing values reported in variable description. Check the distribution to confirm there are no abnormalities: the distribution of age looks reasonable (no bunching or extreme outliers).

```{r}
acs_combined %>% 
  ggplot(aes(x = age)) +
  geom_density(color = "blue", fill = alpha("blue", 0.3)) +
  theme_minimal() +
  ggtitle("Age Variable Results") +
  ylab("Density")
```

EDUCD: N/As reported as "1" in all modern surveys. N/As were "000" in the 1980 and 1950 decennial census surveys (not used here).
Variable codes explained [here]([https://usa.ipums.org/usa-action/variables/EDUC#codes_section) & see "Detailed codes" at bottom).
Vacancy is reported as "0". 

There is a significant share of records reported as "1" for the education variable. 

```{r}
acs_combined %>% 
  mutate(educd = zap_labels(educd)) %>% 
  count(educd) %>% 
  ggplot(mapping = aes(x = factor(educd), y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Educd, All Records") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Educational Attainment",
       y = NULL)
```

Confirm the missing values from the value labels of the educd variable to ensure we are eliminating the correct values. "0", "1" and "999" are N/A or missing and should be turned into NA values that R recognizes. 

Note that there are no records with values of zero in the bar chart above which eliminates the need to deal with the "no schooling" option.
```{r}
head(ipums_val_labels(acs_combined$educd), 50) %>% 
  reactable()
```

Turn the missing/"not in universe" value from the educd variable into NAs and remove the labels for the rest of the analysis. Rename to the more meaningful "education_detailed". 
```{r}
acs_clean <- acs_combined %>%
  mutate(education_detailed = zap_labels(educd), 
         education_detailed = ifelse(education_detailed %in% c(1, 0, 999), NA_integer_,
                          education_detailed)
  )
```

Look at the distribution of education_detailed after adjustment for NAs. We can see that there are no longer records for "1" but we now see "NA".
```{r}
acs_clean %>% 
  count(education_detailed) %>% 
  ggplot(mapping = aes(x = factor(education_detailed), y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Education Detailed, All Records") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Educational Attainment",
       y = NULL)
```

Vacancy: Value is reported as "0" according to IPUMs. Note there are no vacancy results by default of the structure of census data read in (confirmed in the distribution of the variable).

```{r}
acs_clean %>% 
  ggplot(aes(x = vacancy)) +
  geom_density(color = "blue", fill = alpha("blue", 0.3)) +
  theme_minimal() +
  ggtitle("Vacancy Variable Results") +
  ylab("Density")
```

Examine the group quarters variable, we want to include both households and Group Quarters in this metric as both are relevant to the college readiness of a geography.

IPUMS includes the following categories:

-   0: Vacancy

-   1: Households under 1970 definition

-   2: Additional households under 1990 definition

-   3: Institutions

-   4: Other group quarters

-   5: Additional households under 2000 definition

-   6: Fragments


Check the distribution of GQ categories. Only households and group quarters should be included (no vacancy or fragments).
```{r}
acs_clean %>%  
  count(gq) %>% 
  ggplot(mapping = aes(x = gq, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("GQ, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "GQ",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))
```

### Create race variable

Evaluate the values in the RACE and HISPAN variables.

* IPUMS documented values for RACE: 
    + 1 White
    + 2 Black/African American/Negro
    + 3 American Indian or Alaska Native 
    + 4 Chinese 
    + 5 Japanese
    + 6 Other Asian or Pacific Islander 
    + 7 Other race 
    + 8 Two major races 
    + 9 Three or more major races

* IPUMS documented values for HISPAN: 
    + 0 Not Hispanic 
    + 1 Mexican 
    + 2 Puerto Rican 
    + 3 Cuban 
    + 4 Other 
    + 9 Not Reported 

Look at the distribution of values for the race and hispan variables.
```{r}
acs_clean %>% 
  count(race) %>% 
  ggplot(mapping = aes(x = race, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Race, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Race",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

acs_clean %>% 
  count(hispan) %>% 
  ggplot(mapping = aes(x = hispan, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Hispan, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Hispan",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))
```

Create the race/ethnicity variable. For race/ethnicity categories that are not coded as Hispanic, Hispan must be equal to "Not Hispanic". 

Rename the values from number category to race label: 
  +Race is Black/African American and Hispan is Not Hispanic = "Black, Non-Hispanic"
  +Hispan is not equal to Not Hispanic = "Hispanic"
  +Race is not Black or White and Hispan is Not Hispanic  = "Other Races and Ethnicities" 
  +Race is White and Hispan is Not Hispanic = "White, Non-Hispanic" 
```{r}
acs_clean <- acs_clean %>%
  mutate(subgroup_race = case_when((hispan == "Not Hispanic" & race == "White") ~ "White, Non-Hispanic",
                              (hispan == "Not Hispanic" & race == "Black/African American") ~ "Black, Non-Hispanic",
                              (hispan == "Not Hispanic" & !race %in% c("Black/African American", "White")) ~ "Other Races and Ethnicities",
                               hispan != "Not Hispanic" ~ "Hispanic"
  ))

```

Look at the race/ethnicity subgroup distribution.
```{r}

acs_clean %>% 
  count(subgroup_race) %>% 
  ggplot(mapping = aes(x = factor(subgroup_race), y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Race/ethnicity, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Race/ethnicity",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))
```

### Create gender variable

Evaluate the values in the gender variable.

* IPUMS documented values for SEX: 
    + 1 Male
    + 2 Female
    + 9 Missing/blank 

Look at the distribution of values for the sex variable. There are no missing values in the data and the records are reported in an appropriate format.
```{r}
acs_clean %>% 
  count(sex) %>% 
  ggplot(mapping = aes(x = sex, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("sex, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Sex",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))
```

Rename the variable to subgroup_gender. The variable terms remain the same but any values outside of Male and Female are mutated to NA character variables. Though there are no missing values in the current data let's leave this in the case of future ACS surveys including missing values for this variable. 
```{r}
acs_clean <- acs_clean %>% 
  mutate(subgroup_gender = case_when(sex == "Male" ~ "Male",
                                     sex == "Female" ~ "Female",
                                     TRUE ~ NA_character_))
```

### Create disability variable

The ACS includes 5 variables covering physical and mental disabilities (DIFFCARE, DIFFSENS, DIFFMOB, DIFFPHYS, DIFFREM). To make these variables more clear, first rename them.

```{r}
acs_clean <- acs_clean %>% 
  rename(self_care_difficulty = diffcare,
         vision_hearing_difficulty = diffsens,
         independent_living_difficulty = diffmob,
         ambulatory_difficulty = diffphys,
         cognitive_difficulty = diffrem)
```

Evaluate the values in these variables and create an overall subgroup for disability.
We will count an individual to be disabled if they answer yes or have any of these difficulties. 

* IPUMS documented values for self_care_difficulty: 
    + 0 N/A
    + 1 No
    + 2 Yes 

Plot the options for self_care_difficulty below.
```{r}

acs_clean %>% 
  count(self_care_difficulty) %>% 
  ggplot(mapping = aes(x = self_care_difficulty, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Self-care difficulty, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "self_care_difficulty",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

```

* IPUMS documented values for  vision_hearing_difficulty: 
    + 0 N/A
    + 1 No vision or hearing difficulty
    + 2 Has vision or hearing difficulty

Plot the options for  vision_hearing_difficulty below.
```{r}

acs_clean %>% 
  count( vision_hearing_difficulty) %>% 
  ggplot(mapping = aes(x =  vision_hearing_difficulty, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Vision or Hearing difficulty, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = " vision_hearing_difficulty",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

```

* IPUMS documented values for independent_living_difficulty: 
    + 0 N/A
    + 1 No independent living difficulty
    + 2 Has independent living difficulty

Plot the options for independent_living_difficulty below.
```{r}

acs_clean %>% 
  count(independent_living_difficulty) %>% 
  ggplot(mapping = aes(x = independent_living_difficulty, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Independent living difficulty, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "independent_living_difficulty",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

```

* IPUMS documented values for ambulatory_difficulty: 
    + 0 N/A
    + 1 No ambulatory difficulty
    + 2 Has ambulatory difficulty 
  
Plot the options for ambulatory_difficulty below.
```{r}

acs_clean %>% 
  count(ambulatory_difficulty) %>% 
  ggplot(mapping = aes(x = ambulatory_difficulty, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Ambulatory difficulty, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "ambulatory_difficulty",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

```
  
* IPUMS documented values for cognitive_difficulty: 
    + 0 N/A
    + 1 No cognitive difficulty
    + 2 Has cognitive difficulty 
  
Plot the options for cognitive_difficulty below.
```{r}
acs_clean %>% 
  count(cognitive_difficulty) %>% 
  ggplot(mapping = aes(x = cognitive_difficulty, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Cognitive difficulty, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "cognitive_difficulty",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

```

Create the subgroup for disability. 
```{r}
acs_clean <- acs_clean %>% 
  mutate(subgroup_disability = case_when(cognitive_difficulty == "Has cognitive difficulty" |
                                           ambulatory_difficulty == "Has ambulatory difficulty" |
                                           independent_living_difficulty == "Has independent living difficulty" |
                                            vision_hearing_difficulty == "Has vision or hearing difficulty" |
                                           self_care_difficulty == "Yes" ~ "with disabiliy",
                                         TRUE ~ "without disability"))
  
```

Check the variable breakout. 
```{r}
acs_clean %>% 
  count(subgroup_disability) %>% 
  ggplot(mapping = aes(x = subgroup_disability, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Subgroup Disability, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Disability",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))
```

## Crosswalk

Read in the PUMA to place crosswalk file. This file is created by the program generate_puma_place_crosswalks.rmd in the geographic-crosswalks folder. 
```{r}
puma_place_crosswalk <- read_csv(here::here("geographic-crosswalks", "data", "crosswalk_puma_to_place.csv"))
```

## Merge Crosswalk 

Create a variable in the ACS data indicating whether the survey year is pre or post-2022. This will be used to join on the correct crosswalk information.
```{r}
acs_clean <- acs_clean %>% 
  mutate(crosswalk_period = ifelse(year < 2022, "pre-2022", "2022"))

```

Check that the assignment worked. All years prior to 2022 should be designated "pre-2022".
```{r}
acs_clean %>% 
  group_by(crosswalk_period, year) %>% 
  count()
```

Join the cleaned ACS data onto the crosswalk.
```{r}

acs_crosswalked <- left_join(
  acs_clean, 
  puma_place_crosswalk, 
  by=c("crosswalk_period", "statefip","puma"), 
  relationship = "many-to-many"
)

```

Drop any observations with NA for afact (i.e. there is no place of interest overlapping this PUMA).
```{r}
acs_crosswalked <- acs_crosswalked %>% 
  drop_na(afact)
```

Also filter out cases where AFACT is equal to zero. These cases will not be counted in the metric calculation.
```{r}
acs_crosswalked <- acs_crosswalked %>% 
  filter(afact > 0)
```

Adjust the person weight to account for PUMA-to-place mapping (those where the PUMA is not entirely inside the place).

```{r}

select(acs_crosswalked, perwt, repwtp1, repwtp80, afact)
acs_crosswalked <- acs_crosswalked %>%
  mutate(perwt = perwt * afact) %>% 
  mutate(across(matches("repwtp[0-9]+"), ~.x * afact))
select(acs_crosswalked, perwt, repwtp1, repwtp80, afact) 

```

## Prepare comuptational variables

Create a binary college_ready variable indicating whether the individual has a HS degree or not (including GED as HS degree equivalent). IPUMS codes  HS graduate or GED (62) and above are all considered to meet this requirement.
```{r}
acs_crosswalked <- acs_crosswalked %>% 
  mutate(college_ready = ifelse(education_detailed >= 62, 1, 0)) 
```

Check how many records, if any, have NA reported for education_detailed after filtering (this should be zero).
```{r}
acs_crosswalked %>%
  filter(is.na(education_detailed)) %>% 
  group_by(pernum) %>% 
  count()
```

Check the distribution of the education_detailed variable one last time - there should be no NA values included. 
```{r}
acs_crosswalked %>% 
  count(education_detailed) %>% 
  ggplot(mapping = aes(x = factor(education_detailed), y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Education Detailed, All Records") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Educational Attainment",
       y = NULL)
```

Confirm there are no vacant properties included in the data.

```{r}
acs_crosswalked %>%
  filter(is.na(vacancy)) %>% 
  group_by(pernum) %>% 
  count()
```

## Create pre-K metric: All

Create the college ready metric. 

Objective: get the percent of individuals between 19 and 20 that have a HS degree.

Aggregation should be weighted by PERWT (this is a person level statistic).

### Pre-drop suppressed cells

First, to save time when processing the metric, calculate the effective sample count for each place, defined as the sum of the `afact` variable. This is used to create a cutoff for data quality based on the actual number of survey records being used in the calculation. We will not include counties with less than 30 effective samples.

```{r}
#| label: all-effective-samples
acs_all <- acs_crosswalked %>%
  group_by(year, sample, crosswalk_period, statefip, place) %>%
  mutate(effective_sample = sum(afact)) %>% 
  ungroup()
# calculate the number of statistics after suppressing values
acs_all %>%
  group_by(year, sample, crosswalk_period, statefip, place) %>%
  summarize(
    effective_sample = max(effective_sample)
  ) %>%
  ungroup() %>%
  summarize(
    original_stats = n(),
    unsuppressed_stats = sum(effective_sample >= 30)
  )
```

Remove records from counties below the size quality cutoff. 

```{r}
#| label: subset-all
acs_all <- acs_all %>% 
  filter(effective_sample >= 30)
```

### Calculate metric

Calculate the place-level metrics. 

```{r}
#| label: calc-all
results_all <- acs_all %>%
  group_split(statefip) %>% 
  map_dfr(~ calc_survey_mean_college(.data = .x, .geo_level = place))
```

Clean up the confidence interval bounds and sort the data.

```{r}
#| label: finalize-all
results_all <- finalize_metric_college(results_all)
```

```{r}
rm(acs_all)
```

## Add Suppressed Rows to Data

To create the metric we suppressed counties that did not meet the threshold for sample size but we need to include these in the final data. Create an all version of the data that includes counties that were suppressed prior to the survey_mean calculation. 
```{r}
results_all_expand <- acs_crosswalked %>% 
  mutate(geoid = paste0(statefip, place)) %>% 
  group_by(year, geoid) %>% 
  count() %>% 
  select(-n) %>% 
  mutate(statefip = str_sub(geoid, 1, 2),
         place = str_sub(geoid, 3, 7)) %>% 
  left_join(results_all, by = c("year", "statefip", "place")) %>% 
  ungroup()

stopifnot(nrow(results_all_expand) == 485 * 2)
```

## Data Quality Flags

Add a flag for data quality, this is a numeric variable between 1 and 3 with 1 representing the best quality and 3 representing the worst. 

The `share_hs_degree_quality` variable combines information on the quality of the crosswalk and the sample size (effective sample) to create a final quality flag. Note that any place with NA for the `share_hs_degree` metric is being suppressed due to sample size and is given a quality flag of 3.  

```{r}
metrics_education_all <- results_all_expand %>% 
  mutate(
    share_hs_degree_quality = if_else(
      is.na(share_hs_degree), 
      NA_real_,
      geographic_allocation_quality
    )
  )
```

## Create college metric: subgroups

## Race/Ethnicity

To save time when processing the metric, calculate the effective sample count for each race-ethnicity group in each place, defined as the sum of the `afact` variable. This is used to create a cutoff for data quality based on the actual number of survey records being used in the calculation. We will not include counties with less than 30 effective samples.

```{r}
#| label: race-ethnicity-effective-samples
acs_race_ethnicity <- acs_crosswalked %>%
  group_by(year, sample, crosswalk_period, statefip, place, subgroup_race) %>%
  mutate(effective_sample = sum(afact)) %>% 
  ungroup()
# calculate the number of statistics after suppressing values
acs_race_ethnicity %>%
  group_by(year, sample, crosswalk_period, statefip, place, subgroup_race) %>%
  summarize(
    effective_sample = max(effective_sample)
  ) %>%
  ungroup() %>%
  summarize(
    original_stats = n(),
    unsuppressed_stats = sum(effective_sample >= 30)
  )
```

Create the college readiness metric for the race/ethnicity subgroup. 

Remove records from counties below the size quality cutoff. 

```{r}
#| label: subset-race-ethnicity
acs_race_ethnicity <- acs_race_ethnicity %>% 
  filter(effective_sample >= 30)
```

### Calculate metric

```{r}
#| label: calc-race-ethnicity
results_race_ethnicity <- acs_race_ethnicity %>%
  group_split(statefip) %>% 
  map_dfr(~ calc_survey_mean_subgroup(.data = .x,
                                      .geo_level = place,
                                      .subgroup = subgroup_race))
```

```{r}
#| label: finalize-race-ethnicity
results_race_ethnicity <- finalize_metric(results_race_ethnicity)
```

```{r}
rm(acs_race_ethnicity)
```

### Add Suppressed Rows to Data

To create the metric we suppressed counties that did not meet the threshold for sample size but we need to include these in the final data. Create an all version of the data that includes counties that were suppressed prior to the survey_mean calculation. 

Create expanded version for race-ethnicity.
```{r}
results_race_ethnicity_expanded <- acs_crosswalked %>% 
  mutate(geoid = paste0(statefip, place)) %>% 
  group_by(year, geoid, subgroup_race) %>% 
  count() %>% 
  select(-n) %>% 
  ungroup() %>% 
  expand(year, geoid, subgroup_race) %>%  
  mutate(statefip = str_sub(geoid, 1, 2),
         place = str_sub(geoid, 3, 7)) %>% 
  left_join(results_race_ethnicity, by = c("year", "statefip", "place", "subgroup_race")) %>% 
  mutate(subgroup_type = "race-ethnicity") %>% 
  rename(subgroup = subgroup_race)
```

## Gender

To save time when processing the metric, calculate the effective sample count for each gender group in each place, defined as the sum of the `afact` variable. This is used to create a cutoff for data quality based on the actual number of survey records being used in the calculation. We will not include counties with less than 30 effective samples.

```{r}
#| label: gender-effective-samples
acs_gender <- acs_crosswalked %>%
  group_by(year, sample, crosswalk_period, statefip, place, subgroup_gender) %>%
  mutate(effective_sample = sum(afact)) %>% 
  ungroup()
# calculate the number of statistics after suppressing values
acs_gender %>%
  group_by(year, sample, crosswalk_period, statefip, place, subgroup_gender) %>%
  summarize(
    effective_sample = max(effective_sample)
  ) %>%
  ungroup() %>%
  summarize(
    original_stats = n(),
    unsuppressed_stats = sum(effective_sample >= 30)
  )
```

Create the college readiness metric for the gender subgroup. 

Remove records from counties below the size quality cutoff. 

```{r}
#| label: subset-gender
acs_gender <- acs_gender %>% 
  filter(effective_sample >= 30)
```

### Calculate metric

```{r}
#| label: calc-gender
results_gender <- acs_gender %>%
  group_split(statefip) %>% 
  map_dfr(~ calc_survey_mean_subgroup(.data = .x,
                                      .geo_level = place,
                                      .subgroup = subgroup_gender))
```

```{r}
#| label: finalize-gender
results_gender <- finalize_metric(results_gender)
```

```{r}
rm(acs_gender)
```

### Add Suppressed Rows to Data

To create the metric we suppressed counties that did not meet the threshold for sample size but we need to include these in the final data. Create an all version of the data that includes counties that were suppressed prior to the survey_mean calculation. 

Create expanded version for gender data.
```{r}
results_gender_expanded <- acs_crosswalked %>% 
  mutate(geoid = paste0(statefip, place)) %>% 
  group_by(year, geoid, subgroup_gender) %>% 
  count() %>% 
  select(-n) %>% 
  ungroup() %>% 
  expand(year, geoid, subgroup_gender) %>%  
  mutate(statefip = str_sub(geoid, 1, 2),
         place = str_sub(geoid, 3, 7)) %>% 
  left_join(results_results_gender, by = c("year", "statefip", "place", "subgroup_gender")) %>% 
  mutate(subgroup_type = "gender") %>% 
  rename(subgroup = subgroup_gender)
```

## Disability

To save time when processing the metric, calculate the effective sample count for each disability group in each place, defined as the sum of the `afact` variable. This is used to create a cutoff for data quality based on the actual number of survey records being used in the calculation. We will not include counties with less than 30 effective samples.

```{r}
#| label: disability-effective-samples
acs_disability <- acs_crosswalked %>%
  group_by(year, sample, crosswalk_period, statefip, place, subgroup_disability) %>%
  mutate(effective_sample = sum(afact)) %>% 
  ungroup()
# calculate the number of statistics after suppressing values
acs_disability %>%
  group_by(year, sample, crosswalk_period, statefip, place, subgroup_disability) %>%
  summarize(
    effective_sample = max(effective_sample)
  ) %>%
  ungroup() %>%
  summarize(
    original_stats = n(),
    unsuppressed_stats = sum(effective_sample >= 30)
  )
```

Create the college readiness metric for the disability subgroup. 

Remove records from counties below the size quality cutoff. 

```{r}
#| label: subset-disability
acs_disability <- acs_disability %>% 
  filter(effective_sample >= 30)
```

### Calculate metric

```{r}
#| label: calc-disability
results_disability <- acs_disability %>%
  group_split(statefip) %>% 
  map_dfr(~ calc_survey_mean_subgroup(.data = .x,
                                      .geo_level = place,
                                      .subgroup = subgroup_disability))
```

```{r}
#| label: finalize-disability
results_disability <- finalize_metric(results_disability)
```

```{r}
rm(acs_disability)
```

### Add Suppressed Rows to Data

To create the metric we suppressed counties that did not meet the threshold for sample size but we need to include these in the final data. Create an all version of the data that includes counties that were suppressed prior to the survey_mean calculation. 

Create expanded version for disability data.
```{r}
results_disability_expanded <- acs_crosswalked %>% 
  mutate(geoid = paste0(statefip, place)) %>% 
  group_by(year, geoid, subgroup_disability) %>% 
  count() %>% 
  select(-n) %>% 
  ungroup() %>% 
  expand(year, geoid, subgroup_disability) %>%  
  mutate(statefip = str_sub(geoid, 1, 2),
         place = str_sub(geoid, 3, 7)) %>% 
  left_join(results_disability, by = c("year", "statefip", "place", "subgroup_disability")) %>% 
  mutate(subgroup_type = "disability") %>% 
  rename(subgroup = subgroup_disability)
```

## Data Quality Flags

Add a flag for data quality, this is a numeric variable between 1 and 3 with 1 representing the best quality and 3 representing the worst. 

First combine the subgroup data sets.
```{r}
metrics_education_sub <-results_all_expand %>% 
  mutate(subgroup = "All",
         subgroup_type = "all") %>% 
  bind_rows(results_race_ethnicity_expanded, results_disability_expanded, results_gender_expanded)
stopifnot(nrow(metrics_education_sub) == 3143 * 2 * 5)
```

Now create the metric quality variable. The `share_hs_degree_quality` variable combines information on the quality of the crosswalk and the sample size (effective sample) to create a final quality flag. Note that any place with NA for the `share_hs_degree` metric is being suppressed due to sample size and is given a quality flag of 3.  

```{r}
metrics_education_sub <- metrics_education_sub %>% 
  mutate(
    share_hs_degree_quality = if_else(
      is.na(share_hs_degree), 
      NA_real_,
      geographic_allocation_quality
    )
  )
```

### Evaluate results

Summarize the predictors the all cateogry and categories by subgroups.
```{r}
metrics_education_all_clean %>% 
  select(share_hs_degree:share_hs_degree_quality) %>% 
  summary()

metrics_education_sub_clean %>% 
  select(share_hs_degree:share_hs_degree_quality) %>% 
  summary()
```

Look at distribution of hs_degree share by place.
```{r}
metrics_education_all_clean %>% 
  filter(year == 2021) %>% 
  select(share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("Distribution by all palces of share hs_degree, 2021 (5-year data)") +
  ylab("Density")
```

### Race distribution
Look at distribution of hs_degree share by place by race.
```{r}

metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "Black, Non-Hispanic") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, Black") +
  ylab("Density")

metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "Hispanic") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, Hispanic") +
  ylab("Density")

metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "Other Races and Ethnicities") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, Other Races and Ethnicities") +
  ylab("Density")

metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "White, Non-Hispanic") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, White, Non-Hispanic") +
  ylab("Density")

```

### Gender distribution
Look at distribution of hs_degree share by place by gender.
```{r}
metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "Male") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, males") +
  ylab("Density")

metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "Female") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, females") +
  ylab("Density")
```

### Disability distribution

Look at distribution of hs_degree share by place by disability status.
```{r}

metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "with disabiliy") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, with disabilities") +
  ylab("Density")

metrics_education_sub_clean %>% 
  filter(year == 2021, subgroup == "without disability") %>% 
  select(place, share_hs_degree) %>% 
  ggplot(aes(x = share_hs_degree)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, without disabilities") +
  ylab("Density")

```

Look at the counts of the quality flag.
```{r}
metrics_education_all_clean %>% 
  select(share_hs_degree_quality) %>% 
  ggplot(aes(x = share_hs_degree_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  ggtitle("Quality Flag (5-year data no subgroups)") +
  ylab("Count")

metrics_education_sub_clean %>% 
  select(share_hs_degree_quality) %>% 
  ggplot(aes(x = share_hs_degree_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  ggtitle("Quality Flag (5-year data with subgroups") +
  ylab("Count")
```

Tabulate share of data suppressed by year.
```{r}
metrics_education_all_clean %>% 
  group_by(year) %>% 
  filter(is.na(share_hs_degree)) %>% 
  count()

metrics_education_sub_clean %>% 
  group_by(year) %>% 
  filter(is.na(share_hs_degree)) %>% 
  count()
```

## Export 

Rename state.
```{r}
metrics_education_all_clean <- metrics_education_all_clean %>% 
  rename("state" = "statefip")

metrics_education_sub_clean <- metrics_education_sub_clean %>% 
  rename("state" = "statefip")
```

Order the variables how we want.
```{r}
metrics_education_all_clean <- metrics_education_all_clean %>% 
  select(year, state, place, share_hs_degree, 
         share_hs_degree_lb, share_hs_degree_ub, share_hs_degree_quality)

metrics_education_sub_clean <- metrics_education_sub_clean %>% 
  select(year, state, place, subgroup_type, subgroup, share_hs_degree,
         share_hs_degree_lb, share_hs_degree_ub, share_hs_degree_quality)
```

Export as CSV

```{r}
write_csv(metrics_education_all_clean, here::here("08_education", "data", "final", "metrics_college_place_all_longitudinal.csv"))
write_csv(metrics_education_sub_clean, here::here("08_education", "data", "final", "metrics_college_place_sub-groups_longitudinal.csv"))
```





