---
title: "Share of households with broadband access in the home"
author: "Manu Alcalá Kovalski"
date: now
format:
  html:
    self-contained: true
    toc: true
editor_options: 
  chunk_output_type: source
execute:
  warning: false
  message: false
editor:
  markdown:
    wrap: 80
    canonical: true
---

```{r}
options(scipen = 999)

library(tidyverse)
library(urbnthemes)
library(tidycensus)
library(censusapi)
library(glue)

set_urbn_defaults(style = "print")

```

This metric shows the ratio of households with the presence of a computer and a
type of broadband internet subscription in their household (we calculate the
overall ratio, as well as by racial subgroup - measured according to the race of
each head of the household).

## Process

1.  Housekeeping

2.  Pull demographics for Census Places and Census Counties

3.  Clean and reshape to move data into the vars we want

4.  Calculate the digital access metric

5.  Create a data quality flag

6.  Cleaning and export

## Housekeeping

```{r}
#| label: housekeeping
# Explore where to pull data from
acs5_vars <- listCensusMetadata(name="2021/acs/acs5", type = "variables")

acs_geos <- listCensusMetadata(name = "acs/acs5", vintage = 2021, type = "geography")

```

## Pull demographics for Census Places and Census Counties

```{r}
#| label: "Pull Census data"

# First, list & save variables of interest as a vector

popvars <-
    c("total_people" = "B02001_001",
    "white" = "B02001_002",
    "black" = "B02001_003", 
    "aian" = "B02001_004",
    "asian" = "B02001_005",
    "nhpi" = "B02001_006",
    "other" = "B02001_007",
    "two_or_more" = "B02001_008",
    "hispanic" = "B03001_003")

digitalvars <- c(
    "white_digital" = "B28009A_004",
    "black_digital" = "B28009B_004", 
    "aian_digital" = "B28009C_004",
    "asian_digital" = "B28009D_004",
    "nhpi_digital" = "B28009E_004",
    "other_digital" = "B28009F_004",
    "two_or_more_digital" = "B28009G_004",
    "hispanic_digital" = "B28009I_004"
  )
# Pull ACS data at the Census Place and Census County levels
# first, all the demographic populations
years = c(2018, 2019, 2021, 2022)

pull_census_data <- function(vars, geography, years){
  
  census_data <- map_dfr(years, ~get_acs(geography = {{geography}},
                      variables = vars,
                      year = .x) %>% 
                        mutate(year = .x) %>% 
                          pivot_wider(names_from = variable, values_from = c(estimate, moe))) %>% 
      rename_with(~str_remove(.x, "estimate_"), 
                  starts_with("estimate_")) 
      
  
  if(geography == "county"){
    census_data <- census_data %>% 
      filter(!str_detect(GEOID, "^72"))
  }
  
  # If population vars are inputted, create asian_other column
  if("B02001_001" %in% vars){
   census_data <- census_data %>% 
       mutate(
    asian_other = aian + asian + nhpi + other + two_or_more
  ) %>% 
    select(GEOID,  NAME, year,
                                              total_people, 
                                              asian_other, 
                                              black, 
                                              hispanic, 
                                              white)
  }
  
  return(census_data)
}

places_pop <- pull_census_data(popvars, "place", years) 
county_pop <- pull_census_data(popvars, "county", years) 

places_digital <- pull_census_data(digitalvars, "place", years)
county_digital <- pull_census_data(digitalvars, "county", years)

places_digital
```

We combine margins of errors for detailed groups using rules from slide 51
[here](https://www.census.gov/content/dam/Census/programs-surveys/acs/guidance/training-presentations/20180418_MOE_Webinar_Transcript.pdf).
We then calculate coefficients of variation:

1.  Convert the MOE into a standard error by dividing by 1.645 because Census
    uses 90% confidence interval

2.  Divide the standard error by the estimate

```{r}
#| label: calculate-coefficients-of-variation

calculate_coefficients_of_variation <- function(data){
    data %>% 
        mutate(
    asian_other_digital = aian_digital + asian_digital + nhpi_digital + other_digital + two_or_more_digital,
    # don't include hispanic when calculating total!
    total_people_digital = asian_other_digital + black_digital + white_digital,
    moe_asian_other_digital = sqrt(
      moe_aian_digital ^ 2 + 
        moe_asian_digital ^ 2 + 
        moe_nhpi_digital ^ 2 +
        moe_other_digital ^ 2 + 
        moe_two_or_more_digital ^ 2
    ),
    # don't include hispanic when calculating total!
    moe_total_people_digital = sqrt(
      moe_asian_other_digital ^ 2 + 
        moe_black_digital ^ 2 + 
        moe_white_digital ^ 2
    )
  ) %>%
  mutate(
    cv_total_people_digital = (moe_total_people_digital / 1.645) / total_people_digital, 
    cv_asian_other_digital = (moe_asian_other_digital / 1.645) / asian_other_digital, 
    cv_black_digital = (moe_black_digital / 1.645) / black_digital, 
    cv_hispanic_digital = (moe_hispanic_digital / 1.645) / hispanic_digital, 
    cv_white_digital = (moe_white_digital / 1.645) / white_digital
  ) %>% 
        select(GEOID, 
                                                      NAME, 
               year,
                                                      total_people_digital, 
                                                      asian_other_digital, 
                                                      black_digital, 
                                                      hispanic_digital, 
                                                      white_digital,
                                                      cv_total_people_digital, 
                                                      cv_asian_other_digital, 
                                                      cv_black_digital, 
                                                      cv_hispanic_digital, 
                                                      cv_white_digital) %>% 
        arrange(GEOID, NAME, year)
        
}
county_digital <- county_digital %>% calculate_coefficients_of_variation() 
 
places_digital <- places_digital %>% calculate_coefficients_of_variation()


```

## Calculate the digital access metric

```{r}
#| label: calculate-digital-access-metric

calculate_digital_access_metric <- function(pop_data, digital_data){
  pop_data %>% 
    left_join(digital_data, by=c("GEOID", "NAME", "year")) %>% 
    mutate(
      digital_access_total = total_people_digital / total_people,
      digital_access_asian_other = asian_other_digital / asian_other, 
      digital_access_black = black_digital / black, 
      digital_access_hispanic = hispanic_digital / hispanic,
      digital_access_white = white_digital / white
    ) %>% 
        arrange(GEOID, NAME, year) %>% 
        # replace instances with division by zero
          mutate(across(everything(), ~if_else(is.nan(.x) | is.infinite(.x), NA, .x)))
}

digital_access_county <- calculate_digital_access_metric(county_pop, county_digital)

digital_access_city <- calculate_digital_access_metric(places_pop, places_digital)

map_dbl(digital_access_county, ~sum(is.nan(.x)))
map_dbl(digital_access_city, ~sum(is.nan(.x)))
```

# Create data quality flags

We only look at the CVs for the numerator, which is less precise than the
denominator. We use 0.1, 0.2, and 0.5 as thresholds because the CV only accounts
for sampling error in the numerator. This is stricter than for some other
metrics.

```{r}
#| label: county-data-quality-flags
# Generate the quality var
set_quality <- function(cv, digital_access) {
  
  quality <- case_when(
    cv < 0.1 ~ 1,
    cv < 0.2 ~ 2,
    cv < 0.5 ~ 3,
    TRUE ~ NA_real_
  )
  
  quality <- if_else(is.na(digital_access), NA_real_, quality)
  
}

digital_access_county <- digital_access_county %>% 
  mutate(
    digital_access_total_quality = set_quality(cv_total_people_digital, total_people_digital),
    digital_access_asian_other_quality = set_quality(cv_asian_other_digital, asian_other_digital),
    digital_access_black_quality = set_quality(cv_black_digital, black_digital),
    digital_access_hispanic_quality = set_quality(cv_hispanic_digital, hispanic_digital),
    digital_access_white_quality = set_quality(cv_white_digital, white_digital)
  )

count(digital_access_county, digital_access_total_quality)
count(digital_access_county, digital_access_asian_other_quality)
count(digital_access_county, digital_access_black_quality)
count(digital_access_county, digital_access_hispanic_quality)
count(digital_access_county, digital_access_white_quality)


```

```{r}
#| label: city-data-quality-flags


digital_access_city <- digital_access_city %>% 
  mutate(
    digital_access_total_quality = set_quality(cv_total_people_digital, total_people_digital),
    digital_access_asian_other_quality = set_quality(cv_asian_other_digital, asian_other_digital),
    digital_access_black_quality = set_quality(cv_black_digital, black_digital),
    digital_access_hispanic_quality = set_quality(cv_hispanic_digital, hispanic_digital),
    digital_access_white_quality = set_quality(cv_white_digital, white_digital)
  )

count(digital_access_city, digital_access_total_quality)
count(digital_access_city, digital_access_asian_other_quality)
count(digital_access_city, digital_access_black_quality)
count(digital_access_city, digital_access_hispanic_quality)
count(digital_access_city, digital_access_white_quality)


```

# Prepare the data for saving & export final Metrics files

```{r}
#| label: read-pop-files
# Read final County or Places population file to isolate data needed for each geography
# 
# This function also cleans these files the following way:
# - add in the lost leading zeroes for the state/county FIPs & state/place FIPs
# - keep the most recent year of population data (not 2022, but 2020)
# - add a geoid column
#' 
#' @param path 
read_pop_file <- function(path){
    data <- read_csv(here::here(path))
    geo_var <- ifelse("county" %in% colnames(data), sym("county"), sym("place"))
    digits <- ifelse("county" %in% colnames(data), 3, 5)
    
    data %>%
        mutate(state = sprintf("%0.2d", as.numeric(state)),
               !!geo_var := sprintf(glue("%0.{digits}d"), as.numeric(!!geo_var)),
               GEOID = paste0(state, !!geo_var)) 

}
county_file <- read_pop_file("geographic-crosswalks/data/county-populations.csv")
places_file <- read_pop_file("geographic-crosswalks/data/place-populations.csv")
```

```{r}
#| label: merge-pop-files

# merge the data files into the population files (left join, since data files have more observations)
county_digital_access_by_race <- left_join(county_file, digital_access_county, by=c("GEOID", "year"))

county_digital_access_by_race %>% 
    count(year)

place_digital_access_by_race <- left_join(places_file, digital_access_city, by=c("GEOID", "year"))

place_digital_access_by_race %>% 
    count(year)
```

```{r}
#| label: reshape-to-export

# RESHAPE TO PREPARE EXPORT
# reshape each of these from wide to long (so that there are 5 obs per place -- all, white, black, asian, other)

reshape_for_export <- function(data){
    geo_var <- ifelse("county" %in% colnames(data), sym("county"), sym("place"))
    
    data %>% 
        select(year, state, !!geo_var, digital_access_total, digital_access_asian_other, 
           digital_access_black, digital_access_hispanic, digital_access_white,
           ends_with("quality")
    ) %>% 
    pivot_longer(-c(year, state, !!geo_var),
                 values_to = "digital_access",
                 names_prefix = "digital_access_") %>%
    mutate(name = if_else(!str_ends(name, "_quality"), paste0(name, "_share"), name)) %>% 
    separate(name, c("subgroup", "value_type"), sep = "_(?!.*_)") %>% 
    pivot_wider(names_from = value_type, values_from = digital_access) %>% 
    mutate(subgroup = case_match(subgroup,
                                 "total" ~ "All",
                                 "asian_other" ~ "Other Races and Ethnicities", 
                                 .default = str_to_title(subgroup))) %>% 
    mutate(subgroup_type = case_when((subgroup %in% c("All")) ~ "all",
                                     (subgroup %in% c("All",
                                                      "Other Races and Ethnicities",
                                                      "Black",
                                                      "White",
                                                      "Hispanic")) ~ "race-ethnicity"), .before = "subgroup") %>% 
    arrange(year, state, !!geo_var) %>% 
    rename(share_digital_access = share,
           share_digital_access_quality = quality)
}
county_digital_access <-
    county_digital_access_by_race %>% 
    reshape_for_export()

count(county_digital_access, year)
# 3144 obs (counties) * 5 race groups = 15,720 obs -- accurate

place_digital_access <-
    place_digital_access_by_race %>% 
    reshape_for_export()

count(place_digital_access, year)
# 486 obs (places) * 5 race groups = 2,430 obs -- accurate
```

We suppress showing metric values for observations with missing CVs

```{r}
#| label: suppress-bad-cvs

suppress_bad_cvs <- function(data){
    data %>% 
        mutate(
            share_digital_access = if_else(is.na(share_digital_access_quality), NA_real_, share_digital_access)
        )
}
county_digital_access <- suppress_bad_cvs(county_digital_access) 
place_digital_access <- suppress_bad_cvs(place_digital_access) 
```

```{r}
#| label: export-csvs
# Export each of the files as CSVs

export_metric <- function(data, type = c("recent", "subgroup")){
    browser()
    geo <- ifelse("county" %in% colnames(data), "county", "city")
    
    if(type == "recent"){
        data <- 
            data %>% 
            filter(subgroup == "All", year == max(year)) %>% 
            select(-c(subgroup_type, subgroup)) 
        
        write_csv(data,
                  here::here("08_education", glue("digital_access_{geo}_{unique(data$year)}.csv")))
    } else if(type == "subgroup"){
        write_csv(data,
                  here::here("08_education", glue("digital_access_{geo}_subgroup_all.csv")))
    }
    else{
        stop("type must be 'recent' or 'subgroup'")
    }
}


export_metric(county_digital_access, "recent")
export_metric(county_digital_access, "subgroup")
export_metric(place_digital_access, "recent")
export_metric(place_digital_access, "subgroup")
```
