---
title: "Share of households with broadband access in the home"
author: "Manu Alcalá Kovalski"
date: now
format:
  html:
    self-contained: true
    toc: true
editor_options: 
  chunk_output_type: source
execute:
  warning: false
  message: false
editor:
  markdown:
    wrap: 80
    canonical: true
---

```{r}
options(scipen = 999)

library(tidyverse)
library(urbnthemes)
library(tidycensus)
library(censusapi)
library(glue)
library(janitor)
library(tidylog)
library(skimr)

set_urbn_defaults(style = "print")
```

This metric shows the ratio of households with the presence of a computer and a
type of broadband internet subscription in their household (we calculate the
overall ratio, as well as by racial subgroup - measured according to the race of
each head of the household).

## Process

1.  Housekeeping

2.  Pull demographics for Census Places and Census Counties

3.  Clean and reshape to move data into the vars we want

4.  Calculate the digital access metric

5.  Create a data quality flag

6.  Cleaning and export

## Housekeeping

```{r}
#| label: housekeeping
# Explore where to pull data from
acs5_vars <- load_variables(2021, "acs5", cache = TRUE)
acs1_vars <- load_variables(2021, "acs1", cache = TRUE)
acs1_subject_vars <- load_variables(2021, "acs1/subject", cache = TRUE)
```

## Pull demographics for Census Places and Census Counties

```{r}
#| label: census-pull-function
pull_census_data <- function(vars, geography, survey, years) {
  census_data <- map_dfr(years, ~ get_acs(
    geography = {{ geography }},
    variables = vars,
    survey = survey,
    year = .x,
    moe_level = 95
  ) %>%
    mutate(year = .x) %>%
    pivot_wider(names_from = variable, values_from = c(estimate, moe))) %>%
    rename_with(
      ~ str_remove(.x, "estimate_"),
      starts_with("estimate_")
    ) |>
    clean_names()


  if (geography == "county") {
    census_data <- census_data %>%
      filter(!str_detect(geoid, "^72"))
  }


  return(census_data)
}
```

### Communities overall

-   Unit of analysis: Households,
-   Survey: ACS 5
-   Census table: B28003
-   Years: 2017, 2018, 2019, 2021, 2022

We are using the 5-yr ACS for the communities overall (and subgroups), from 2017
to 2022. This will be based on table B28003.

```{r}
#| label: pull-overall-communities
# "Population" variables refer to households/people in households in a
# given category.

# "Digital" variables (those with a "digital" prefix) refer to households/people in households in a given category that have a computer with a broadband internet subscription.

years <- c(2017, 2018, 2019, 2021, 2022)

overall_vars <-
  c( # Population variables
    "total_households" = "B28003_001",
    # Digital variables
    "total_households_digital" = "B28003_004"
  )

# Apply our function to multiple datasets using map2 and assigns the results
# back to the specified object names in the global environment, reducing repetitive code such as:
# county_overall <- pull_census_data(geography = "county",
#                                    vars = overall_vars,
#                                    years = acs1_years,
#                                    survey = "acs1")
# place_overall <- pull_census_data(geography = "place",
#                                    vars = overall_vars,
#                                    years = acs1_years,
#                                    survey = "acs1")
map(
  .x = c("county", "place"),
  .f = ~ assign(
    x = paste0(.x, "_overall"),
    value = pull_census_data(
      geography = .x,
      vars = overall_vars,
      years = years,
      survey = "acs5"
    ),
    envir = .GlobalEnv
  )
)

# 4,969 rows
skim(county_overall)
# 3,763 rows
skim(place_overall)
```

### Race/ethnicity subgroup

-   Unit of analysis: Population in households,
-   Survey: ACS 5
-   Census table(s): B28009, S2802 (subject table)
-   Years: 2018, 2021

```{r}
acs5_years <- c(2018, 2021)

raceeth_vars <-
  c( # Population variables
    "total_people_in_households" = "S2802_C01_001",
    "white" = "B28009A_001",
    "black" = "B28009B_001",
    "aian" = "B28009C_001",
    "asian" = "B28009D_001",
    "nhpi" = "B28009E_001",
    "other" = "B28009F_001",
    "two_or_more" = "B28009G_001",
    "hispanic" = "B28009I_001",
    # Digital variables
    "total_people_in_households_digital" = "S2802_C02_001",
    "white_digital" = "S2802_C02_005",
    "black_digital" = "S2802_C02_006",
    "aian_digital" = "S2802_C02_007",
    "asian_digital" = "S2802_C02_008",
    "nhpi_digital" = "S2802_C02_009",
    "other_digital" = "S2802_C02_010",
    "two_or_more_digital" = "S2802_C02_011",
    "hispanic_digital" = "S2802_C02_012"
  )

map(
  .x = c("county", "place"),
  .f = ~ assign(
    x = paste0(.x, "_raceeth"),
    value = pull_census_data(
      geography = .x,
      vars = raceeth_vars,
      years = acs5_years,
      survey = "acs5"
    ),
    envir = .GlobalEnv
  )
)

skim(county_raceeth)
skim(place_raceeth)
```

We aggregate the race/ethnicity categories into Other races and ethnicities,
Black, Hispanic, white, and All (total population in households).

We combine margins of errors for detailed groups using rules from slide 51
[here](https://www.census.gov/content/dam/Census/programs-surveys/acs/guidance/training-presentations/20180418_MOE_Webinar_Transcript.pdf).

```{r}
#| label: aggregate-raceeth
aggregate_raceeth <- function(data) {
  data |>
    mutate(
      other = asian + aian + nhpi + other + two_or_more
    ) |>
    mutate(
      other_digital = asian_digital + aian_digital + nhpi_digital + other_digital + two_or_more_digital
    ) |>
    mutate(
      moe_other_digital = sqrt(
        moe_aian_digital^2 +
          moe_asian_digital^2 +
          moe_nhpi_digital^2 +
          moe_other_digital^2 +
          moe_two_or_more_digital^2
      )
    ) |>
    select(
      geoid, name, year,
      matches("total_people|other$|other_digital$|black|hispanic|white$|white_digital")
    )
}

lst(county_raceeth, place_raceeth) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = aggregate_raceeth(.y),
      envir = .GlobalEnv
    )
  )
```

### Income subgroup

-   Unit of analysis: Households,
-   Survey: ACS 5,
-   Years: 2018, 2021,
-   Table: B28004

```{r}
#| label: pull-income-data

income_vars <-
  c( # Population variables
    "total_households" = "B28004_001",
    "less_than_10k" = "B28004_002",
    "10k_to_20k" = "B28004_006",
    "20k_to_35k" = "B28004_010",
    "35k_to_50k" = "B28004_014",
    "50k_to_75k" = "B28004_018",
    "75k_or_more" = "B28004_022",
    "less_than_10k_digital" = "B28004_004",
    "10k_to_20k_digital" = "B28004_008",
    "20k_to_35k_digital" = "B28004_012",
    "35k_to_50k_digital" = "B28004_016",
    "50k_to_75k_digital" = "B28004_020",
    "75k_or_more_digital" = "B28004_024"
  )


aggregate_income_subgroup <- function(data) {
  data |>
    mutate(
      income_less_than_50k = less_than_10k + x10k_to_20k + x20k_to_35k + x35k_to_50k,
      moe_income_less_than_50k = sqrt(
        moe_less_than_10k^2 +
          moe_10k_to_20k^2 +
          moe_20k_to_35k^2 +
          moe_35k_to_50k^2
      ),
      income_more_than_50k = x50k_to_75k + x75k_or_more,
      moe_income_more_than_50k = sqrt(moe_50k_to_75k^2 + moe_75k_or_more^2),
      income_less_than_50k_digital = less_than_10k_digital + x10k_to_20k_digital + x20k_to_35k_digital + x35k_to_50k_digital,
      moe_income_less_than_50k_digital = sqrt(
        moe_less_than_10k_digital^2 +
          moe_10k_to_20k_digital^2 +
          moe_20k_to_35k_digital^2 +
          moe_35k_to_50k_digital^2
      ),
      income_more_than_50k_digital =
        x50k_to_75k_digital + x75k_or_more_digital,
      moe_income_more_than_50k_digital = sqrt(
        moe_50k_to_75k_digital^2 +
          moe_75k_or_more_digital^2
      ),
      total_households_digital =
        income_less_than_50k_digital + income_more_than_50k_digital,
      moe_total_households_digital = sqrt(
        moe_less_than_10k_digital^2 +
          moe_10k_to_20k_digital^2 +
          moe_20k_to_35k_digital^2 +
          moe_35k_to_50k_digital^2 +
          moe_50k_to_75k_digital^2 +
          moe_75k_or_more_digital^2
      )
    ) |>
    select(year, geoid, name, total_households, moe_total_households_digital, total_households_digital, income_less_than_50k, moe_income_less_than_50k, income_more_than_50k, moe_income_more_than_50k, income_less_than_50k_digital, moe_income_less_than_50k_digital, income_more_than_50k_digital, moe_income_more_than_50k_digital)
}

map(
  .x = c("county", "place"),
  .f = ~ assign(
    x = paste0(.x, "_income"),
    value = pull_census_data(
      geography = .x,
      vars = income_vars,
      years = acs5_years,
      survey = "acs5"
    ) |>
      aggregate_income_subgroup(),
    envir = .GlobalEnv
  )
)

skim(county_income)
skim(place_income)
```

## Confidence Intervals

We then calculate confidence intervals at the 95% level by adding/subtracting
the MOE estimate from our digital access variables:

```{r}
#| label: calculate-confidence-intervals

calculate_confidence_intervals <- function(data) {
  data |>
    mutate(across(
      .cols = any_of(c(
        "total_households_digital",
        "total_people_in_households_digital",
        "other_digital",
        "black_digital",
        "hispanic_digital",
        "white_digital",
        "income_less_than_50k_digital",
        "income_more_than_50k_digital"
      )),
      .fns = list(
        lb = ~ .x - get(paste0("moe_", cur_column())),
        ub = ~ .x + get(paste0("moe_", cur_column()))
      ),
      .names = "{.col}_{.fn}"
    ))
}


lst(
  county_overall,
  place_overall,
  county_raceeth,
  place_raceeth,
  county_income,
  place_income
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = calculate_confidence_intervals(.y),
      envir = .GlobalEnv
    )
  )
```

## Calculate the digital access metric

```{r}
#| label: calculate-digital-access-metric

calculate_digital_access_metric <- function(data) {
  data |>
    mutate(across(
      .cols = any_of(c(
        "total_households",
        "total_people_in_households",
        "other",
        "black",
        "hispanic",
        "white",
        "income_less_than_50k",
        "income_more_than_50k"
      )),
      .fns = list(
        share_digital_access = ~ get(paste0(cur_column(), "_digital")) / .x,
        share_digital_access_lb = ~ get(paste0(cur_column(), "_digital_lb")) / .x,
        share_digital_access_ub = ~ get(paste0(cur_column(), "_digital_ub")) / .x
      ),
      .names = "{.fn}_{.col}"
    )) |>
    arrange(geoid, name, year) %>%
    # replace instances with division by zero
    mutate(across(everything(), ~ if_else(is.nan(.x) | is.infinite(.x), NA, .x)))
}


lst(
  county_overall,
  place_overall,
  county_raceeth,
  place_raceeth,
  county_income,
  place_income
) %>%
  map2(
    .x = paste0(names(.), "_digital_access"),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = calculate_digital_access_metric(.y),
      envir = .GlobalEnv
    )
  )


map_dbl(county_overall_digital_access, ~ sum(is.nan(.x)))
map_dbl(place_overall_digital_access, ~ sum(is.nan(.x)))
map_dbl(county_raceeth_digital_access, ~ sum(is.nan(.x)))
map_dbl(place_raceeth_digital_access, ~ sum(is.nan(.x)))
map_dbl(county_income_digital_access, ~ sum(is.nan(.x)))
map_dbl(place_income_digital_access, ~ sum(is.nan(.x)))
```

Inspect summary statistics and missigness for counties:

```{r}
map(
  list(
    county_overall_digital_access,
    county_raceeth_digital_access,
    county_income_digital_access
  ),
  skim
)
```

and for places:

```{r}
map(
  list(
    place_overall_digital_access,
    place_raceeth_digital_access,
    place_income_digital_access
  ),
  skim
)
```

# Create data quality flags

We can use the width of the MOE to determine quality. If the value minus the
lower bound is \<0.1, then the quality is strong (\_quality=1); if it is =\>.1
and \<.2, the quality is marginal (\_quality=2), and if it is \>=.2, then the
quality is weak (\_quality=3). Otherwise, if any of the census table values are

suppressed, those should have the metric value show as "NA" and the data quality
should also be "NA."

```{r}
#| label: data-quality-flags

add_quality_flag <- function(data) {
  data |>
    mutate(across(
      .cols = any_of(c(
        "share_digital_access_lb_total_households",
        "share_digital_access_lb_total_people_in_households",
        "share_digital_access_lb_income_less_than_50k",
        "share_digital_access_lb_income_more_than_50k",
        "share_digital_access_lb_other",
        "share_digital_access_lb_black",
        "share_digital_access_lb_hispanic",
        "share_digital_access_lb_white"
      )),
      .fns = ~ case_when(
        get(str_remove(cur_column(), "_lb")) - .x < 0.1 ~ 1,
        get(str_remove(cur_column(), "_lb")) - .x < 0.2 ~ 2,
        get(str_remove(cur_column(), "_lb")) - .x >= 0.2 ~ 3,
        .default = NA_real_
      ),
      .names = "{.col}_quality"
    )) |>
    rename_with(~ str_replace(., "_lb_", "_"), ends_with("_quality"))
}

lst(
  county_overall_digital_access,
  place_overall_digital_access,
  county_raceeth_digital_access,
  place_raceeth_digital_access,
  county_income_digital_access,
  place_income_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = add_quality_flag(.y),
      envir = .GlobalEnv
    )
  )

plot_quality_flags <- function(data) {
  data |>
    select(starts_with("share_digital_access") & ends_with("quality")) |>
    pivot_longer(everything()) |>
    mutate(name = str_remove(name, "share_digital_access_") |>
      str_remove("_quality") |>
      str_replace_all("_", " ") |>
      str_to_title()) |>
    count(name, value) |>
    ggplot() +
    geom_col(aes(
      x = value,
      y = n
    )) +
    facet_wrap(. ~ name) +
    labs(
      x = "Quality flag value",
      y = "Count",
      title = "Share Digital Access Quality Flag"
    )
}

count(
  county_overall_digital_access,
  share_digital_access_total_households_quality
)

count(
  place_overall_digital_access,
  share_digital_access_total_households_quality
)

plot_quality_flags(county_raceeth_digital_access) +
  labs(subtitle = "By Race/Ethnicity")

plot_quality_flags(place_raceeth_digital_access) +
  labs(subtitle = "By Race/Ethnicity")

plot_quality_flags(county_income_digital_access) +
  labs(subtitle = "By Income")

plot_quality_flags(place_income_digital_access) +
  labs(subtitle = "By Income")
```

We also cap the lower bound at 0 and upper bound at 1

```{r}
cap_ci <- function(data) {
  data |>
    rowwise() |>
    mutate(
      across(
        contains("share_digital_access_lb"),
        ~ max(.x, 0, na.rm = TRUE)
      ),
      across(
        contains("share_digital_access_ub"),
        ~ min(.x, 1, na.rm = TRUE)
      )
    ) |>
    ungroup()
}

lst(
  county_overall_digital_access,
  place_overall_digital_access,
  county_raceeth_digital_access,
  place_raceeth_digital_access,
  county_income_digital_access,
  place_income_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = cap_ci(.y),
      envir = .GlobalEnv
    )
  )

skim(county_overall_digital_access)
skim(county_raceeth_digital_access)
skim(county_income_digital_access)

skim(place_overall_digital_access)
skim(place_raceeth_digital_access)
skim(place_income_digital_access)
```

# Prepare the data for saving & export final Metrics files

```{r}
#| label: read-pop-files
# Read final County or Places population file to isolate data needed for each geography
#
# This function also cleans these files the following way:
# - add in the lost leading zeroes for the state/county FIPs & state/place FIPs
# - keep the most recent year of population data (not 2022, but 2020)
# - add a geoid column
#'
#' @param path
read_pop_file <- function(path) {
  data <- read_csv(here::here(path))
  geo_var <- ifelse("county" %in% colnames(data), sym("county"), sym("place"))
  digits <- ifelse("county" %in% colnames(data), 3, 5)

  data %>%
    mutate(
      state = sprintf("%0.2d", as.numeric(state)),
      !!geo_var := sprintf(glue("%0.{digits}d"), as.numeric(!!geo_var)),
      geoid = paste0(state, !!geo_var)
    ) %>%
    filter(year >= 2017) |>
    clean_names()
}
county_file <- read_pop_file("geographic-crosswalks/data/county-populations.csv")
places_file <- read_pop_file("geographic-crosswalks/data/place-populations.csv")
```

```{r}
#| label: merge-pop-files

# merge the data files into the population files (left join, since data files have more observations)

lst(
  county_overall_digital_access,
  county_raceeth_digital_access,
  county_income_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = left_join(county_file, .y, by = c("geoid", "year")),
      envir = .GlobalEnv
    )
  )

count(county_overall_digital_access, year)
count(county_raceeth_digital_access, year)
count(county_income_digital_access, year)

lst(
  place_overall_digital_access,
  place_raceeth_digital_access,
  place_income_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = left_join(places_file, .y, by = c("geoid", "year")),
      envir = .GlobalEnv
    )
  )

count(place_overall_digital_access, year)
count(place_raceeth_digital_access, year)
count(place_income_digital_access, year)
```

```{r}
#| label: reshape-to-export

# RESHAPE TO PREPARE EXPORT
# reshape each of these from wide to long (so that there are 5 obs per place -- all, white, black, asian, other)

reshape_for_export <- function(data, subgroup = "race-ethnicity") {
  geo_var <- ifelse("county" %in% colnames(data), sym("county"), sym("place"))



  data <-
    data %>%
    select(
      year, state, !!geo_var, state_name, paste0(geo_var, "_name"),
      starts_with("share_digital_access")
    )
  if (subgroup == "overall") {
    data <-
      data |>
      rename_with(~ str_remove(., "_total_households"), starts_with("share_digital_access"))
  } else if (subgroup == "race-ethnicity") {
    data <-
      data |>
      pivot_longer(starts_with("share_digital_access")) |>
      mutate(subgroup = str_remove(name, "share_digital_access_") |>
        str_remove("lb_") |>
        str_remove("ub_") |>
        str_remove("_quality") |>
        str_replace_all("_", " ") |>
        str_to_title()) |>
      mutate(
        name = case_when(
          str_detect(name, "quality") ~ "share_digital_access_quality",
          str_detect(name, "lb_") ~ "share_digital_access_lb",
          str_detect(name, "ub_") ~ "share_digital_access_ub",
          .default = "share_digital_access"
        ),
        subgroup = case_match(subgroup,
          "Total People In Households" ~ "All",
          "Other" ~ "Other Races and Ethnicities",
          .default = subgroup
        ),
        subgroup_type = case_match(subgroup,
          "All" ~ "all",
          .default = "race-ethnicity"
        )
      ) |>
      pivot_wider(
        names_from = name,
        values_from = value
      )
  } else if (subgroup == "income") {
    data <-
      data |>
      pivot_longer(starts_with("share_digital_access")) |>
      mutate(subgroup = str_remove(name, "share_digital_access_") |>
        str_remove("lb_") |>
        str_remove("ub_") |>
        str_remove("_quality") |>
        str_replace_all("_", " ") |>
        str_to_title()) |>
      mutate(
        name = case_when(
          str_detect(name, "quality") ~ "share_digital_access_quality",
          str_detect(name, "lb_") ~ "share_digital_access_lb",
          str_detect(name, "ub_") ~ "share_digital_access_ub",
          .default = "share_digital_access"
        )
      ) |>
      pivot_wider(
        names_from = name,
        values_from = value
      ) |>
      mutate(
        subgroup_type = "income",
        subgroup = str_replace(subgroup, "Total Households", "All")
      )
  }

  data |>
    arrange(year, state, !!geo_var, state_name, paste0(geo_var, "_name"))
}


lst(
  county_overall_digital_access,
  place_overall_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = reshape_for_export(.y, subgroup = "overall"),
      envir = .GlobalEnv
    )
  )

lst(
  county_raceeth_digital_access,
  place_raceeth_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = reshape_for_export(.y, subgroup = "race-ethnicity"),
      envir = .GlobalEnv
    )
  )

# 3144 obs (counties) * 5 race groups = 15,720 obs -- accurate
count(county_raceeth_digital_access, year)

# 486 obs (places) * 5 race groups = 2,430 obs -- accurate
count(place_raceeth_digital_access, year)

lst(
  county_income_digital_access,
  place_income_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = reshape_for_export(.y, subgroup = "income"),
      envir = .GlobalEnv
    )
  )
```

We suppress showing metric values for observations with missing quality flags.
However, since there aren't any missing quality flag values when the metric
value is not missing, the code chunk below doesn't actually remove any
observations currently.

```{r}
#| label: suppress-bad-cvs

# Use this to check that there are no missing quality flags
# when the metric value is not missing
count_bad_cvs <- function(data) {
  data |>
    filter(is.na(share_digital_access_quality) & !is.na(share_digital_access)) |>
    nrow()
}

county_overall_digital_access |> count_bad_cvs()

place_overall_digital_access |> count_bad_cvs()

county_raceeth_digital_access |> count_bad_cvs()

place_raceeth_digital_access |> count_bad_cvs()

county_income_digital_access |> count_bad_cvs()

place_income_digital_access |> count_bad_cvs()

suppress_bad_cvs <- function(data) {
  data %>%
    mutate(across(
      c(share_digital_access, share_digital_access_lb, share_digital_access_ub),
      ~ if_else(is.na(share_digital_access_quality), NA_real_, .x)
    ))
}


lst(
  county_overall_digital_access,
  place_overall_digital_access,
  county_raceeth_digital_access,
  place_raceeth_digital_access,
  county_income_digital_access,
  place_income_digital_access
) %>%
  map2(
    .x = names(.),
    .y = .,
    .f = ~ assign(
      x = .x,
      value = suppress_bad_cvs(.y),
      envir = .GlobalEnv
    )
  )

county_subgroup_digital_access <-
  bind_rows(
    county_raceeth_digital_access,
    county_income_digital_access
  )

place_subgroup_digital_access <-
  bind_rows(
    place_raceeth_digital_access,
    place_income_digital_access
  )
```

```{r}
#| label: skim-before-export

county_overall_digital_access |> skim()

place_overall_digital_access |> skim()

county_raceeth_digital_access |> skim()

place_raceeth_digital_access |> skim()

county_income_digital_access |> skim()

place_income_digital_access |> skim()
```

```{r}
#| label: export-csvs
# Export each of the files as CSVs

export_metric <- function(data, type = c("subgroup", "multi-year")) {
  geo <- ifelse("county" %in% colnames(data), "county", "city")
  geo_var <- ifelse("county" %in% colnames(data), "county", "place")

  data <-
    data |>
    # Transform metric from ratio to percent with 3 decimal places
    # mutate(across(
    #   .cols = starts_with("share_digital_access") & !ends_with("quality"),
    #   .fns = ~ scales::label_percent(accuracy = 0.001)(.x)
    # )) |>
    filter(!(year %in% c(2016, 2020))) |>
    select(-c(state_name, paste0({{ geo_var }}, "_name")))

  if (type == "multi-year") {
    write_csv(
      data,
      here::here("08_education", "final", glue("digital_access_{geo}_all.csv"))
    )
  } else if (type == "subgroup") {
    data <-
      data |>
      filter(year %in% c(2018, 2021)) |>
      relocate(subgroup_type, .before = "subgroup") |>
      mutate(subgroup = case_match(subgroup,
        "Income Less Than 50k" ~ "Less than $50,000",
        "Income More Than 50k" ~ "$50,000 or More",
        .default = subgroup
      ))
    write_csv(
      data,
      here::here("08_education", "final", glue("digital_access_{geo}_subgroup_all.csv"))
    )
  } else {
    stop("type must be 'multi-year' or 'subgroup'")
  }
}

export_metric(
  county_overall_digital_access,
  "multi-year"
)
export_metric(county_subgroup_digital_access, "subgroup")

export_metric(
  place_overall_digital_access,
  "multi-year"
)
export_metric(place_subgroup_digital_access, "subgroup")
```

Check that the outputted files match the README conventions.

```{r}
#| label: check-county-files
county_all <- read_csv(here::here("08_education/final/digital_access_county_all.csv"))
county_subgroup <- read_csv(here::here("08_education/final/digital_access_county_subgroup_all.csv"))

head(county_all)
head(county_subgroup)
```

```{r}
#| label: check-city-files
place_all <- read_csv(here::here("08_education/final/digital_access_city_all.csv"))
place_subgroup <- read_csv(here::here("08_education/final/digital_access_city_subgroup_all.csv"))

head(county_all)
head(county_subgroup)
```
