---
title: "Access to pre-kindergarten Predictor, County-level"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

*Program for the creation of the Access to Pre-Kindergarten Upward Mobility Predictor at the County-level*

ACS Code: Initially written by Tina Chelidze and updated by JP Walsh 2022-2023.

Primary data derived from the IPUMS API.

Based on processes developed by Paul Johnson and Kevin Werner in SAS.

-   [Housekeeping](#housekeeping)
-   [Read Data](#read-data)
-   [Clean Data](#clean-data)
-   [Merge Crosswalk](#merge-crosswalk)
-   [Setup Survey Design](#setup-survey-design)
-   [Create Preschool Access Metric](#create-preschool-access-metric)
-   [Evaluate Results](#evaluate-results)
-   [Create standard errors](#create-standard-errors)
-   [Export](#export)


## Housekeeping

Read in packages for use. If you do not have an IPUMS API key you will need to make one via the [IPUMS website](https://www.ipums.org/).

```{r, results='hide'}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)
library(srvyr)
library(scales)

options(scipen = 999)


#set_ipums_api_key("Your KEY", save = TRUE)
map(list.files(here::here("functions", "API")), 
    ~source(here::here("functions", "API", .x)))
```

## Read Data

Read in the ACS extracts using the extract_ipums function. Make sure to change the survey list to reflect what years you want to include in the data. Currently we are including three years of 1-year data (2018a, 2021a and 2022a) and two years of 5-year data (2018c and 2021c). 5-year 2022 data will be added pending its release.
```{r}
acs <- extract_ipums(
  extract_name = "umf_data_18_22",
  extract_description = "Microdata pull for Access to Pre-K Metric Predictors. American Community Survey, years 2018, 2019 and 2022 (1 and 5-year).",
  survey = list("us2018a", "us2021a", "us2022a", "us2018c", "us2021c")
) 
```

Look at the distribution of ACS survey samples in the data. The number of different samples should align with the number of surveys selected in the extract IPUMS function.
```{r}
acs %>% 
  count(sample) %>% 
  ggplot(mapping = aes(x = sample, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Samples, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Samples",
       y = NULL)
```

Read in the PUMA to county crosswalk file. This file is created by the program generate_puma_county_crosswalks.rmd in the geographic-crosswalks folder. 
```{r}
puma_county_crosswalk <- read_csv(here::here("geographic-crosswalks", "data", "crosswalk_puma_to_county.csv"))
```

Crate a version of the crosswalk with just the county-level values of data quality variables for each crosswalk period.
```{r}
county_puma <- puma_county_crosswalk %>%
  group_by(crosswalk_period, statefip, county) %>% 
  summarise(geographic_allocation_quality = mean(geographic_allocation_quality)) %>% 
  ungroup()
```

## Clean Data

Missing data is reported in a variety of ways by IPUMS data - listed below:

Age: No missing values reported in variable description.
GRADEATT: N/As reported as "0" in all modern surveys. In this analysis we assume that a zero reported means that the individual is not in school.
Variable codes explained [here](https://usa.ipums.org/usa-action/variables/GRADEATT#codes_section).
Vacancy is reported as "0". 

There is a significant share of records reported as "0" for the grade level attending variable. The distribution of age looks reasonable (no bunching or extreme outliers). Note there are no vacancy results by default of the structure of census data read in.

```{r}
acs %>% 
  mutate(gradeatt = zap_labels(gradeatt)) %>% 
  count(gradeatt) %>% 
  ggplot(mapping = aes(x = factor(gradeatt), y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("gradeatt, All Records") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Educational Attainment",
       y = NULL)

acs %>% 
  ggplot(aes(x = age)) +
  geom_density(color = "blue", fill = alpha("blue", 0.3)) +
  theme_minimal() +
  ggtitle("Age Variable Results") +
  ylab("Density")

acs %>% 
  ggplot(aes(x = vacancy)) +
  geom_density(color = "blue", fill = alpha("blue", 0.3)) +
  theme_minimal() +
  ggtitle("Vacancy Variable Results") +
  ylab("Density")
```

Isolate the data to 3 and 4 year-olds. This is the age-range included in the pre-K calculation.
```{r}
acs_clean <- acs %>% 
  filter(age == 3 | age == 4) 

rm(acs)
```


Confirm the values from the value labels of the GRADEATT variable to ensure we are interpreting them correctly. 1 is Nursery school/preschool and 0 is N/A - interpreting as not in school.
```{r}
head(ipums_val_labels(acs_clean$gradeatt), 50) %>% 
  reactable()
```

Examine the group quarters variable, we want to include both households and Group Quarters in this metric as both are relevant to the preschool access predictor in a geography.

IPUMS includes the following categories:
0: Vacancy
1: Households under 1970 definition
2: Additional households under 1990 definition
3: Institutions
4: Other group quarters
5: Additional households under 2000 definition
6: Fragments

Check the distribution of GQ categories. Only households and group quarters should be included (no vacancy or fragments).
```{r}
acs_clean %>%  
  count(gq) %>% 
  ggplot(mapping = aes(x = gq, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("GQ, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "GQ",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))
```

### Create race variable

Evaluate the values in the RACE and HISPAN variables.

* IPUMS documented values for RACE: 
  + 1 White
  + 2 Black/African American/Negro
  + 3 American Indian or Alaska Native 
  + 4 Chinese 
  + 5 Japanese
  + 6 Other Asian or Pacific Islander 
  + 7 Other race 
  + 8 Two major races 
  + 9 Three or more major races

* IPUMS documented values for HISPAN: 
  + 0 Not Hispanic 
  + 1 Mexican 
  + 2 Puerto Rican 
  + 3 Cuban 
  + 4 Other 
  + 9 Not Reported 

Look at the distribution of values for the race and hispan variables.
```{r}

acs_clean %>% 
  count(race) %>% 
  ggplot(mapping = aes(x = race, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Race, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Race",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

acs_clean %>% 
  count(hispan) %>% 
  ggplot(mapping = aes(x = hispan, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Hispan, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Hispan",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))

```

Create the race/ethnicity variable. For race/ethnicity categories that are not coded as Hispanic, Hispan must be equal to "Not Hispanic". 

Rename the values from number category to race label: 
  +Race is Black/African American and Hispan is Not Hispanic = "Black, Non-Hispanic"
  +Hispan is not equal to Not Hispanic = "Hispanic"
  +Race is not Black or White and Hispan is Not Hispanic  = "Other Races and Ethnicities" 
  +Race is White and Hispan is Not Hispanic = "White, Non-Hispanic" 
```{r}

acs_clean <- acs_clean %>%
  mutate(subgroup = case_when((hispan == "Not Hispanic" & race == "White") ~ "White, Non-Hispanic",
                              (hispan == "Not Hispanic" & race == "Black/African American") ~ "Black, Non-Hispanic",
                              (hispan == "Not Hispanic" & !race %in% c("Black/African American", "White")) ~ "Other Races and Ethnicities",
                               hispan != "Not Hispanic" ~ "Hispanic"
  ))

```

Look at the subgroup distribution.
```{r}

acs_clean %>% 
  count(subgroup) %>% 
  ggplot(mapping = aes(x = factor(subgroup), y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Subgroup, All Records") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Subgroup",
       y = NULL)+
  scale_x_discrete(labels = label_wrap(10))
```

## Merge Crosswalk 

Create a variable in the ACS data indicating whether the survey is pre or post-2022. This will be used to join on the correct crosswalk information.
```{r}
acs_clean <- acs_clean %>% 
  mutate(crosswalk_period = ifelse(year < 2022, "pre-2022", "2022"))

```

Check that the assignment worked. All years prior to 2022 should be designated "pre-2022".
```{r}
acs_clean %>% 
  group_by(crosswalk_period, year) %>% 
  count()
```

Join the cleaned ACS data onto the crosswalk.
```{r}

acs_clean <- left_join(acs_clean, puma_county_crosswalk, by=c("crosswalk_period", "statefip","puma"))

```

Drop any observations with NA for afact (i.e. there is no county of interest overlapping this PUMA).
```{r}
acs_clean <- acs_clean %>% 
  drop_na(afact)

```

Also filter out cases where AFACT is equal to zero. These cases will not be counted in the metric calculation.
```{r}
acs_clean <- acs_clean %>% 
  filter(afact > 0)
```

Adjust the person weight to account for PUMA-to-county mapping (those where the PUMA is not entirely inside the county).
```{r}

acs_clean <- acs_clean %>%
  mutate(perwt = perwt*afact) 

```

Create a binary variable to distinguish between 1 and 5 year ACS data sets. 5-year samples include the term "5-year".
```{r}
acs_clean <- acs_clean %>%
  mutate(sample_structure = ifelse(str_detect(sample, "5-year"), "5year", "1year")) 
```

### Create additional variables

Create a binary preschool variable indicating whether the child is in preschool or not. IPUMS codes  Nursery school/preschool (1) and N/A (0) (re: grade level attending).

```{r}
acs_clean <- acs_clean %>% 
  filter(gradeatt < 2) %>% 
  mutate(preschool = ifelse(gradeatt == 1, 1, 0)) 
```

Check how many records, if any, have NA reported for gradeatt after filtering (this should be zero).
```{r}
acs_clean %>%
  filter(is.na(gradeatt)) %>% 
  group_by(pernum) %>% 
  count()
```

Check the distribution of the gradeatt variable one last time - there should be no NA values included. 
```{r}
acs_clean %>% 
  count(gradeatt) %>% 
  ggplot(mapping = aes(x = factor(gradeatt), y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("gradeatt, All Records") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Educational Attainment",
       y = NULL)
```

Confirm there are no vacant properties included in the data.

```{r}
acs_clean %>%
  filter(is.na(vacancy)) %>% 
  group_by(pernum) %>% 
  count()
```

## Read replicate weights 

Read in person-level replicate weights for the 3 and 4 year old population. These will be used in creating standard errors for the preschool access calculation. 
```{r}
repwts_person <- ipums_repwt_pre_k(
  extract_name = "preschool_replicate_weights",
  extract_description = "Person replicate weights for Access to Preschool Metric Predictors. American Community Survey, years 2018, 2019 and 2022 (1 and 5-year).",
  survey = list("us2018a", "us2021a", "us2022a", "us2018c", "us2021c")
)  %>% 
  select(-cbserial, -cbpernum)
```

Look at the distribution of ACS survey samples in the data. The number of different samples should align with the number of surveys selected in the extract ipums function.
```{r}
repwts_person %>% 
  count(sample) %>% 
  ggplot(mapping = aes(x = sample, y = n)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Samples, All Records in Repwt File") +
 geom_text(mapping = aes(label = n), vjust = -1) +    
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Samples",
       y = NULL)
```

Remove the sample variable.
```{r}
repwts_person <- repwts_person %>% 
  select(-sample)
```

## Setup Survey Design 

Set up the survey design using instructions from IPUMS. To speed up the srvyr calculation we combine the data by grouping the college ready variable and summing the perwt variable - this limits the number of rows that have to go through the survey_mean function. 
```{r}
acs_survey_all <- acs_clean %>%
  filter(sample_structure == "1year") %>% 
  left_join(repwts_person, by = c("unique_person_id")) %>% 
  as_survey_design(weights = perwt, 
                   repweights = matches("repwtp[0-9]+"),
                   type = "JK1",
                   scale = 4/ 80,
                   rscales = rep(1, 80 ), mse = TRUE) %>% 
  ungroup()

acs_survey_subgroups <- acs_clean %>%
  filter(sample_structure == "5year") %>% 
  left_join(repwts_person, by = c("unique_person_id")) %>% 
  as_survey_design(weights = perwt, 
                   repweights = matches("repwtp[0-9]+"),
                   type = "JK1",
                   scale = 4/ 80,
                   rscales = rep(1, 80 ), mse = TRUE) %>% 
  ungroup()

```

## Create Preschool Access metric

Create the preschool access metric. 

Objective: calculate the percent of individuals between 3 and 4 that are in pre-K.

Aggregation should be weighted by PERWT (this is a person level statistic).

Check national pre-K access ratio. Compute the ratio (share_in_preschool) for 2018, 2021 and 2022 nationally and show results.
```{r}
acs_survey_all %>%
  group_by(year) %>%
  summarise(
    share_in_preschool = survey_mean(preschool, vartype = "ci"),
    count = unweighted(n())
  ) %>% 
  ungroup() %>% 
  reactable()
```

### All

Create share with preschool access for all 3 and 4 year-olds at the county-level (1-year data) using the survey_mean function.
```{r}
metrics_education_all <- acs_survey_all %>%
  group_by(year, crosswalk_period, statefip, county) %>% 
  summarise(
    share_in_preschool = survey_mean(preschool, vartype = "ci")
  ) %>% 
  ungroup()
```

Cap upper-bound confidence intervals at 100% (above would be an illogical option). If it is greater than 1, make the share equal to exactly 1.
```{r}
metrics_education_all <- metrics_education_all %>% 
  mutate(share_in_preschool_upp =  ifelse(share_in_preschool_upp > 1, 1, share_in_preschool_upp))
```

Sort by year, state and county.
```{r}
metrics_education_all <- metrics_education_all %>%
  arrange(year, statefip, county)
```

*Create an indicator for data quality and suppression*

Calculate the effective sample count for each county, defined as the sum of the afact variable. This is used to create a cutoff for data quality based on the actual number of survey records being used in the calculation.
```{r}
effective_sample <- acs_clean %>%
  filter(sample_structure == "1year") %>% 
  group_by(year, crosswalk_period, statefip, county) %>%
  summarise(effective_sample = sum(afact)) %>% 
  ungroup()
```

Merge the data quality flags from county_puma and effective sample onto the preschool access metric to create the final data quality metric. This metric is based on both size of county and the number of samples for that county.
```{r}
metrics_education_all <- left_join(metrics_education_all, county_puma, by=c("crosswalk_period", "statefip","county")) 

metrics_education_all <- left_join(metrics_education_all, effective_sample, by=c("year", "crosswalk_period", "statefip","county"))
```

Tabulate effective sample counts for each county, defined as the sum of the afact for all records in that county. Look at the distribution of counties by potential cutoffs (30, 100, 300). This is to measure data quality for each county based on the number of records reported. In the 1-year data for all there are no cases of a county with less than 100 observations for 2018, 2021 and 2022.
```{r}
reactable_data <- metrics_education_all %>%
  mutate(
    effective_sample_cutoffs = case_when(
      effective_sample < 30 ~ "Less than 30",
      effective_sample >= 30 &
        effective_sample <= 100 ~ "Over 30 but 100 or less",
      effective_sample > 100 &
        effective_sample <= 300 ~ "Over 100 but 300 or less",
      TRUE ~ "Greater than 300"
    ),
    effective_sample_cutoffs = factor(
      effective_sample_cutoffs,
      levels = c(
        "Less than 30",
        "Over 30 but 100 or less",
        "Over 100 but 300 or less",
        "Greater than 300"
      )
    )
  ) 

# Get suppression factor for each year of data
suppression_factor <- reactable_data %>%
  group_by(year) %>%
  summarize(yearly_total = n()) %>% 
  select(year, yearly_total)

# Create reactable for further examination
reactable_data %>%
  left_join(suppression_factor, by="year") %>% 
  group_by(year,effective_sample_cutoffs) %>%
  summarize(n = n(), suppression_share = n()/unique(yearly_total)) %>% # yearly value has one unique value per year
  reactable()

```

Generate the quality variable for each county.
```{r}
metrics_education_all <- metrics_education_all %>% 
  mutate(effective_sample_flag = ifelse(effective_sample < 100, 1, 0),
         share_in_preschool_quality = case_when(effective_sample_flag==0 & geographic_allocation_quality==1 ~ 1,
                                     effective_sample_flag==0 & geographic_allocation_quality==2 ~ 2,
                                     effective_sample_flag==0 & geographic_allocation_quality==3 ~ 3,
                                     effective_sample_flag==1 ~ 3)
  )
```

We do not want to show values that are calculated from less than 30 samples in the data. These points will be suppressed and shown as missing or NA. Suppress the results for values generated from less than 30 samples.
```{r}
metrics_education_all <- metrics_education_all %>% 
  mutate_at(c("share_in_preschool", "share_in_preschool_low", "share_in_preschool_upp"), ~if_else(effective_sample < 30, NA_integer_, .)) 
```

Keep only relevant variables.
```{r}
metrics_education_all <- metrics_education_all %>%
  select(year,
         statefip,
         county,
         share_in_preschool,
         share_in_preschool_lb = share_in_preschool_low,
         share_in_preschool_ub = share_in_preschool_upp,
         share_in_preschool_quality)
```

### Subgroup

Create share with preschool access variable by race/ethnicity at the county-level (5-year data).
```{r}
metrics_education_subgroup <- acs_survey_subgroups %>%
  group_by(year, crosswalk_period, statefip, county, subgroup) %>%
 summarise(
    share_in_preschool = survey_mean(preschool, vartype = "ci")
  ) %>% 
  ungroup() %>% 
  mutate(subgroup_type = "race-ethnicity")
```

Create the share with preschool access measure for all values in the 5-year data (non-subgroup). This value will be merged onto the subgroup data. 
```{r}
metrics_education_subgroup_all <- acs_survey_subgroups %>%
  group_by(year, crosswalk_period, statefip, county) %>%
  summarise(
   share_in_preschool = survey_mean(preschool, vartype = "ci")
  ) %>% 
  ungroup() %>% 
  mutate(subgroup_type = "all",
         subgroup = "All")
```

Append the "All" version of the 5-year data with the subgroup results.
```{r}
metrics_education_sub <- bind_rows(metrics_education_subgroup, metrics_education_subgroup_all)
```

Cap upper-bound confidence intervals at 100% (above would be an illogical option). If it is greater than 1, make the share equal to exactly 1.
```{r}
metrics_education_sub <- metrics_education_sub %>% 
  mutate(share_in_preschool_upp =  ifelse(share_in_preschool_upp > 1, 1, share_in_preschool_upp))
```

Sort by year, state and county.
```{r}
metrics_education_sub <- metrics_education_sub %>%
  arrange(year, statefip, county)
```

*Create an indicator for data quality and suppression*

Calculate the effective sample count for each subgroup in all counties as well as for all samples in every county in the 5-year data.
```{r}
effective_sample_sub_all <- acs_clean %>%
  filter(sample_structure == "5year") %>% 
  group_by(year, crosswalk_period, statefip, county) %>%
  summarise(effective_sample = sum(afact)) %>% 
  mutate(subgroup_type = "all",
         subgroup = "All")

effective_sample_sub <- acs_clean %>%
  filter(sample_structure == "5year") %>% 
  group_by(year, crosswalk_period, statefip, county, subgroup) %>%
  summarise(effective_sample = sum(afact)) %>% 
  mutate(subgroup_type = "race-ethnicity") %>% 
  bind_rows(effective_sample_sub_all) %>% 
  ungroup()
```

Merge the data quality flags and create the final data quality metric based on both size and geographic allocation.
```{r}
metrics_education_sub <- left_join(metrics_education_sub, county_puma, by=c("crosswalk_period", "statefip","county"))

metrics_education_sub <- left_join(metrics_education_sub, effective_sample_sub, by=c("year", "crosswalk_period", "statefip","county", "subgroup_type", "subgroup"))
```

Tabulate effective sample counts for each county by sugroup, defined as the sum of the afact for all records in that county. Look at the distribution of counties by potential cutoffs (30, 100, 300). This is to measure data quality for each county based on the number of records reported. 
```{r}
reactable_data <- metrics_education_sub %>%
  mutate(
    effective_sample_cutoffs = case_when(
      effective_sample < 30 ~ "Less than 30",
      effective_sample >= 30 &
        effective_sample <= 100 ~ "Over 30 but 100 or less",
      effective_sample > 100 &
        effective_sample <= 300 ~ "Over 100 but 300 or less",
      TRUE ~ "Greater than 300"
    ),
    effective_sample_cutoffs = factor(
      effective_sample_cutoffs,
      levels = c(
        "Less than 30",
        "Over 30 but 100 or less",
        "Over 100 but 300 or less",
        "Greater than 300"
      )
    )
  )

# Create reactable for further examination using suppression factor defined earlier
reactable_data %>%
  left_join(suppression_factor, by="year") %>% 
  group_by(year,subgroup,effective_sample_cutoffs) %>%
  summarize(n = n(), suppression_share = n()/unique(yearly_total)) %>% # yearly value has one unique value per year
  reactable()
```

Generate the quality variable for each county.
```{r}
metrics_education_sub <- metrics_education_sub %>% 
  mutate(effective_sample_flag = ifelse(effective_sample < 100, 1, 0),
         share_in_preschool_quality = case_when(effective_sample_flag==0 & geographic_allocation_quality==1 ~ 1,
                                     effective_sample_flag==0 & geographic_allocation_quality==2 ~ 2,
                                     effective_sample_flag==0 & geographic_allocation_quality==3 ~ 3,
                                     effective_sample_flag==1 ~ 3)
  )
```

We do not want to show values that are calculated from less than 30 samples in the data. These points will be suppressed and shown as missing or NA. Suppress the results for values generated from less than 30 samples.
```{r}
metrics_education_sub <- metrics_education_sub %>% 
  mutate_at(c("share_in_preschool", "share_in_preschool_low", "share_in_preschool_upp"), ~if_else(effective_sample < 30, NA_integer_, .)) 
```

Keep only relevant variables.
```{r}
metrics_education_sub <- metrics_education_sub %>%
  select(year,
         statefip,
         county,
         subgroup_type, 
         subgroup,
         share_in_preschool,
         share_in_preschool_lb = share_in_preschool_low,
         share_in_preschool_ub = share_in_preschool_upp,
         share_in_preschool_quality)
```

### Evaluate results

Summarize the predictors for the 1-year and 5-year data.
```{r}
metrics_education_all %>% 
  select(share_in_preschool:share_in_preschool_quality) %>% 
  summary()

metrics_education_sub %>% 
  select(share_in_preschool:share_in_preschool_quality) %>% 
  summary()
```

Look at distribution of hs_degree share by county.
```{r}

metrics_education_all %>% 
  filter(year == 2022) %>% 
  select(share_in_preschool) %>% 
  ggplot(aes(x = share_in_preschool)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("Distribution of share hs_degree, by county 2022 (1-year data)") +
  ylab("Density")

metrics_education_sub %>% 
  filter(year == 2021, subgroup == "All") %>% 
  select(share_in_preschool) %>% 
  ggplot(aes(x = share_in_preschool)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("Distribution of share hs_degree, by county 2022 (5-year data)") +
  ylab("Density")
  
```

Look at distribution of hs_degree share by county by race.
```{r}

metrics_education_sub %>% 
  filter(year == 2021, subgroup == "Black, Non-Hispanic") %>% 
  select(county, share_in_preschool) %>% 
  ggplot(aes(x = share_in_preschool)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, Black") +
  ylab("Density")

metrics_education_sub %>% 
  filter(year == 2021, subgroup == "Hispanic") %>% 
  select(county, share_in_preschool) %>% 
  ggplot(aes(x = share_in_preschool)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, Hispanic") +
  ylab("Density")

metrics_education_sub %>% 
  filter(year == 2021, subgroup == "Other Races and Ethnicities") %>% 
  select(county, share_in_preschool) %>% 
  ggplot(aes(x = share_in_preschool)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, Other Races and Ethnicities") +
  ylab("Density")

metrics_education_sub %>% 
  filter(year == 2021, subgroup == "White, Non-Hispanic") %>% 
  select(county, share_in_preschool) %>% 
  ggplot(aes(x = share_in_preschool)) +
  geom_density(alpha = 0.15) +
  theme_minimal() +
  ggtitle("hs_degree Share, White, Non-Hispanic") +
  ylab("Density")

```

Look at the counts of the quality flag.
```{r}
metrics_education_all %>% 
  select(share_in_preschool_quality) %>% 
  ggplot(aes(x = share_in_preschool_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  ggtitle("Quality Flag (1-year data)") +
  ylab("Count")

metrics_education_sub %>% 
  select(share_in_preschool_quality) %>% 
  ggplot(aes(x = share_in_preschool_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  ggtitle("Quality Flag (5-year data") +
  ylab("Count")
```

Tabulate share of data suppressed by year.
```{r}
metrics_education_all %>% 
  group_by(year) %>% 
  filter(is.na(share_in_preschool)) %>% 
  count()

metrics_education_sub %>% 
  group_by(year) %>% 
  filter(is.na(share_in_preschool)) %>% 
  count()
```

## Export 

Rename state.
```{r}
metrics_education_all <- metrics_education_all %>% 
  rename("state" = "statefip")

metrics_education_sub <- metrics_education_sub %>% 
  rename("state" = "statefip")
```

Order the variables how we want.
```{r}
metrics_education_all <- metrics_education_all %>% 
  select(year, state, county, share_in_preschool, 
         share_in_preschool_lb, share_in_preschool_ub, share_in_preschool_quality)

metrics_education_sub <- metrics_education_sub %>% 
  select(year, state, county, subgroup_type, subgroup, share_in_preschool,
         share_in_preschool_lb, share_in_preschool_ub, share_in_preschool_quality)
```

Export as CSV

```{r}
write_csv(metrics_education_all, here::here("08_education", "data", "final", "metrics_preschool_county_all_longitudinal.csv"))
write_csv(metrics_education_sub, here::here("08_education", "data", "final", "metrics_preschool_county_race-ethnicity_longitudinal.csv"))
```

