---
title: "Agencies - County"
author: "Vincent Pancini"
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: false
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    editor_options:
      chunk_output_type: console
---

<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup}
options(scipen = 999)

library(tidyverse)
library(tidycensus)
library(skimr)
library(tidylog)
library(urbnthemes)

set_urbn_defaults(style = "print")

```

This file is a prerequisite file to the following files:
* `reported-crime-county-all.Rmd` which creates the place-level metric for reported property crimes per 100,000 people and reported violent crimes per 100,000 people
* `juvenile-arrests-county-all.Rmd` which creates the place-level metric for juvenile arrests per 100,000 juveniles

You must run this file before either of the above files to create either metric at the county-level.

This file pulls in a universe of law enforcement agencies that have reported criminal activity and arrests in 2022. It then joins on county FIPS codes to those agencies. For agencies that serve more than one county, this program assigns a weight to that agency to determine the percentage of activity that should be associated with each county served by that agency. Next, it pulls in population and demographic information from the American Community Survey (ACS). Lastly, it creates several files that are used in the creation of each metric.

The rest of this file is organized as follows:
1. Background and NIBRS Batch Header File download
2. Load and clean NIBRS Batch Header File
3. Join county FIPS codes onto NIBRS agencies that are missing county using Law Enforcement Agency Identifiers Crosswalk
4. Construct County-Level Demographics Using ACS
5. Join Batch Header File and ACS County Demographics by County
6. Create County-Agency Weights
7. Write out data

All data used in this file and to create the two safety metrics are available on Box [here]().

## 1. Background and NIBRS Batch Header File download
This program and the safety metrics rely on National Incident-based Reporting System (NIBRS) data. NIBRS is an annual data collection that compiles information on criminal incidents and arrests reported by participating law enforcement agencies. It is a part of the Uniform Crime Reporting Program (UCR) which is administered by the Federal Bureau of Investigation (FBI).

NIBRS data as constructed and formatted by the FBI are stored in a single file. These data are organized by various segment levels (record types). Working with the NIBRS data in its single-file format is difficult. To facilitate use of NIBRS data, ICPSR created extract files. These files are available for download from the National Archive of Criminal Justice Data (NACJD), the criminal justice archive within ICPSR, and can be accessed here [here](https://www.icpsr.umich.edu/web/NACJD/series/128).

The segment of NIBRS data this file is concerned with is called the Batch Header File (also called the Batch Header Segment). Prior to the 2013 data, the Batch Header information was released as three segments. Due to the NIBRS data rapidly growing in size, the FBI released a single file instead. The Batch Header File identifies individual police agencies by originating agency identifier (ORI). An individual police agency (ORI) will appear in the batch header segment once. It also includes variables for each county served by each agency, which is what we are particularly interested in. The data used for both safety metrics (arrests and crimes) are reported at the agency-level. The purpose of using the Batch Header File is to be able to link the agencies to counties for those metrics.

To learn more about NIBRS data, consult Jacob Kaplan's book titled [National Incident-Based Reporting System (NIBRS) Data: A Practitioner's Guide](https://nibrsbook.com/). For these metrics, the most important thing to know about these data is that most police agencies do not report NIBRS data prior to 2021. In 2019, only 8,500 police agencies (covering about 45% of the US population) reported NIBRS data. This represented less than half of the approximately 18,000 police agencies in the United States. Instead, agencies primarily reported [UCR data](https://ucrbook.com/). However, the FBI has moved entirely to NIBRS data starting in 2021 and no longer collects UCR data, which is why we use NIBRS data instead ([source](https://nibrsbook.com/overview-of-the-data.html)). We do not want to use NIBRS data for some years and UCR data for others. This means that **there will be a lot of missing values prior to 2021**.

    Note that we only use the 2022 Batch Header File. Using the Batch Header File from each year might be more accurate, but for the most part it won't matter, especially because so few agencies reported through NIBRS prior to 2021. Additionally, downloading and reading in the Batch Header File for each year takes a lot of time and computing power, and must be done manually for each year which introduces room for error in replication. Future updates may consider using the Batch Header Files for each year, which can be accessed through the following links:
    * [2014](https://www.icpsr.umich.edu/web/ICPSR/studies/36421)
    * [2015](https://www.icpsr.umich.edu/web/ICPSR/studies/36851)
    * [2016](https://www.icpsr.umich.edu/web/ICPSR/studies/37066)
    * [2017](https://www.icpsr.umich.edu/web/ICPSR/studies/37650)
    * [2018](https://www.icpsr.umich.edu/web/ICPSR/studies/37649)
    * [2019](https://www.icpsr.umich.edu/web/ICPSR/studies/38565) 
    * [2020](https://www.icpsr.umich.edu/web/ICPSR/studies/38566) 
    * [2021](https://www.icpsr.umich.edu/web/ICPSR/studies/38807)
    * [2022](https://www.icpsr.umich.edu/web/ICPSR/studies/38925)
    
Accessing these data requires an ICPSR account. Manually downloading these data is a five-step process, but you will only need to do each of these steps once for the creation of this metric:
    1. Create an ICPSR account [here](https://login.icpsr.umich.edu/realms/icpsr/login-actions/registration?client_id=icpsr-web-prod&tab_id=BqotoE6wnqw).
    2. Navigate to the National Archive of Criminal Justice Data (NACJD) landing page for the year you are accessing (e.g., 2022 is [here](https://www.icpsr.umich.edu/web/NACJD/studies/38925)).
    3. Select `Download`
        + Select `Delimited`
        + Select `Agree` on the `Terms of Use`
    4. Sign into your ICPSR account
    5. This will download a zip file titled `ICPSR_38925-V2`. Unzip the file, navigate to `ICPSR_38925\DS0001\38925-0001-Data.tsv`, and move it to the appropriate directory/folder. This is the `Batch Header File`.



## 2. Load and clean NIBRS Batch Header File
This section reads in and cleans the reporting agency data contained in the Batch Header File.

### 2.1 Get universe of agencies from the 2022 NIBRS Batch Header File
Load Batch Header File
```{r, message=FALSE}
batch_header_file <- read_tsv(here::here("07_safety", "data", "38925-0001-Data.tsv"))

```

Limit file to necessary variables and rename. We use the codebook (`ICPSR_38925\DS0001\38925-0001-Codebook-ICPSR`) to determine what each variable represents.
```{r}
batch_header_file <- batch_header_file %>%
  select(
    ori = BH003, # ori is unique agency identifier used across multiple files
    agency_type = BH012,
    state_num_nibrs = BH002,
    state_abb = BH008,
    city = BH007,
    core_city = BH013,
    fips_county_1 = BH054,
    fips_county_2 = BH055,
    fips_county_3 = BH056,
    fips_county_4 = BH057,
    # fips_county_5 = BH058, # Exists in the data but all values are missing according to codebook
    # msa_1 = BH021,
    # msa_2 = BH025,
    # msa_3 = BH029,
    # msa_4 = BH033,
    # msa_5 = BH037,
    # pop_1 = BH019,
    # pop_2 = BH023,
    # pop_3 = BH027,
    # pop_4 = BH031,
    # pop_5 = BH035
  ) %>%
  mutate(
    # year = 2022,
    # pop_total = pop_1 + pop_2 + pop_3 + pop_4 + pop_5,
    city = str_to_title(city),
    # Edit NIBRS state and county variables to have two digits and three digits respectively
    state_num_nibrs = str_pad(state_num_nibrs, 2, pad = "0"),
    county = str_pad(fips_county_1, 3, pad = "0"
    )
  ) %>%
  # Drop US territories
  filter(!state_abb %in% c("GM", "PR", "VI"))

```

If an agency is only linked to one county, the other county variables have a value of `-6` for `Not Applicable`. We replace these values with `NA`.
```{r}
batch_header_file <- batch_header_file %>%
  mutate(across(c(fips_county_2, fips_county_3, fips_county_4), 
                ~ ifelse(. == "-6", NA, .)))

```

## 2.2 Convert NIBRS state codes to state FIPS codes
The state identifiers included in NIBRS are state abbreviation and a state code. NIBRS state codes are in alphabetical order and do not follow FIPS codes. We manually create a crosswalk between state FIPS codes and NIBRS codes by looking at the codebook (`ICPSR_38566\DS0001\38566-0001-Codebook`) and then join state FIPS codes onto the Batch Header File.
```{r, message=FALSE}
# Create "crosswalk" for NIBRS state codes to state FIPS codes.
nibrs_states <- tribble(
  ~state, ~state_abb, ~state_num_nibrs, ~state_name,
  "01",	"AL", "01", "Alabama",
  "02",	"AK", "50", "Alaska",
  "04",	"AZ", "02", "Arizona", 
  "05",	"AR", "03", "Arkansas",	
  "06",	"CA", "04", "California",	
  "08",	"CO", "05", "Colorado",
  "09",	"CT", "06", "Connecticut",
  "10",	"DE", "07", "Delaware",
  "11",	"DC", "08", "District of Columbia",
  "12",	"FL", "09", "Florida",
  "13",	"GA", "10", "Georgia",	
  "15",	"HI", "51", "Hawaii",
  "16",	"ID", "11", "Idaho",
  "17",	"IL", "12", "Illinois",
  "18",	"IN", "13", "Indiana",
  "19",	"IA", "14", "Iowa",	
  "20",	"KS", "15", "Kansas",	
  "21",	"KY", "16", "Kentucky",	
  "22",	"LA", "17", "Louisiana",
  "23",	"ME", "18", "Maine",
  "24",	"MD", "19", "Maryland",
  "25",	"MA", "20", "Massachusetts",
  "26",	"MI", "21", "Michigan",	
  "27",	"MN", "22", "Minnesota",	
  "28",	"MS", "23", "Mississippi",	
  "29",	"MO", "24", "Missouri",	
  "30",	"MT", "25", "Montana",	
  "31",	"NE", "26", "Nebraska",	
  "32",	"NV", "27", "Nevada",	
  "33",	"NH", "28", "New Hampshire",	
  "34",	"NJ", "29", "New Jersey",	
  "35",	"NM", "30", "New Mexico",	
  "36",	"NY", "31", "New York",	
  "37",	"NC", "32", "North Carolina",
  "38",	"ND", "33", "North Dakota",	
  "39",	"OH", "34", "Ohio",	
  "40",	"OK", "35", "Oklahoma",	
  "41",	"OR", "36", "Oregon",	
  "42",	"PA", "37", "Pennsylvania",	
  "44",	"RI", "38", "Rhode Island",	
  "45",	"SC", "39", "South Carolina",	
  "46",	"SD", "40", "South Dakota",
  "47",	"TN", "41", "Tennessee",
  "48",	"TX", "42", "Texas",	
  "49",	"UT", "43", "Utah",	
  "50",	"VT", "44", "Vermont",	
  "51",	"VA", "45", "Virginia",	
  "53",	"WA", "46", "Washington",	
  "54",	"WV", "47", "West Virginia",	
  "55",	"WI", "48", "Wisconsin",	
  "56",	"WY", "49", "Wyoming"
)

# The NIBRS data abbreviates Nebraska as "NB" instead of its correct abbreviation "NE"
batch_header_file <- batch_header_file %>%
  mutate(state_abb = if_else(state_abb=="NB", "NE", state_abb))

# Join state FIPS codes and full state name onto the NIBRS data
batch_header_file <- batch_header_file %>% 
  left_join(
    y = nibrs_states, by = c("state_abb", "state_num_nibrs")
  )

```


There are 51 unique combinations of `state`/`state_name`/`state_abb`/`state_num_nibrs`, so the manual crosswalk does not seem to have any errors.
```{r, message=FALSE}
# Check that crosswalk worked without errors
batch_header_file %>%
  group_by(state, state_name, state_abb, state_num_nibrs) %>%
  summarize()

```

### 2.3 Exclude state police and other state agencies
Check distribution of agency types. The agency types are as follows:

| Value | Label                     |
|:-----:|:--------------------------|
|   0   | Covered by another agency |
|   1   | City                      |
|   2   | County                    |
|   3   | University or college     |
|   4   | State Police              |
|   5   | Special Agency            |
|   6   | Other state agencies      |
|   7   | Tribal agencies           |
|   8   | Federal agencies          |

```{r}
batch_header_file %>%
  count(agency_type) %>%
  mutate(percent = n/nrow(batch_header_file))

```

State agencies cover multiple counties, so we remove these agencies for simplicity
```{r}
batch_header_file <- batch_header_file %>%
  filter(!agency_type %in% c(4, 6))

```

260 observations have a missing county FIPS value
```{r}
# Subset data to examine further
bhf_missing_county <- batch_header_file %>%
  filter(is.na(fips_county_1==0))

```


## 3. Join county FIPS codes onto NIBRS agencies that are missing county using Law Enforcement Agency Identifiers Crosswalk

The goal of this section is to match a county FIPS code onto every agency in the 2022 NIBRS Batch Header File. The Batch Header File already includes a county FIPS code for each county associated with a place, but we just showed that 260 agencies still do not have a county associated with them. We attempt to get county information for these agencies using the Law Enforcement Agency Identifiers Crosswalk (LEAIC).

### 3.1 Background on LEAIC and manual download
To address the agencies that are missing county information, we use the 2012 Law Enforcement Agency Identifiers Crosswalk (LEAIC)

The LEAIC facilitates linking reported crime data with socio-economic data. Note that 2012 is the most recent year for which this crosswalk is available. The LEAIC file is available for download from the National Archive of Criminal Justice Data (NACJD) available through ICPSR at the following link:
  * [LEAIC](https://www.icpsr.umich.edu/web/NACJD/studies/35158).
  
Accessing these data requires an ICPSR account. Manually downloading these data is a five-step process, but you will only need to do each of these steps once for the creation of this metric:
    1. Create an ICPSR account [here](https://login.icpsr.umich.edu/realms/icpsr/login-actions/registration?client_id=icpsr-web-prod&tab_id=BqotoE6wnqw).
    2. Navigate to the National Archive of Criminal Justice Data (NACJD) landing page for the LEAIC [here](https://www.icpsr.umich.edu/web/NACJD/studies/35158)).
    3. Select `Download`
        + Select `Delimited`
        + Select `Agree` on the `Terms of Use`
    4. Sign into your ICPSR account
    5. This will download a zip file titled `ICPSR_35158-V2`. Unzip the file, navigate to `ICPSR_35158\DS0001\35158-0001-Data.tsv`, and move it to the appropriate directory/folder.

### 3.2 Load and clean LEAIC
After downloading the LEAIC, read it in. This file has a county FIPS linked to every reporting agency.
```{r}
# Read in crosswalk
leaic <- read_tsv(here::here("07_safety", "data", "35158-0001-Data.tsv"))

```

232 of the 260 agencies with missing county information from the Batch Header File are in the LEAIC.
```{r}
# Check how many agencies with missing county are in LEAIC
sum(bhf_missing_county$ori %in% leaic$ORI9)

```

Limit file to necessary variables and rename. We use the codebook (`ICPSR_35158\DS0001\35158-0001-Codebook`) to determine what each variable represents.
```{r}
leaic <- leaic %>%
  select(
    ori = ORI9,
    county = FIPS_COUNTY
  ) %>%
  # Remove agencies with invalid ORI
  filter(ori != "-1") # Not in UCR/NCIC 

```

### 3.3 Join LEAIC onto NIBRS data with missing counties by agency identifier
Join the LEAIC to the 2022 NIBRS Batch Header File by agency.
28 agencies still have no county after using the county information from LEAIC. All of these observations have agency type codes of `5` or `7` indicating they are tribal or special agencies.
```{r}
bhf_missing_county <- left_join(
  x = bhf_missing_county %>%
    select(-county),
  y = leaic,
  by = c("ori")
)

bhf_missing_county %>%
  filter(is.na(county)) %>%
  count(agency_type)

```

### 3.4 Use LEAIC to create final agency-county data
Replace county information in NIBRS data for observations that initially had a missing county value of with the updated county information from LEAIC.
```{r}
batch_header_file <- batch_header_file %>%
  # Drop the 260 NIBRS agencies with a missing county value
  filter(!is.na(fips_county_1)) %>%
  # Append those 260 agencies back onto main data with the county info gleaned from LEAIC
  bind_rows(bhf_missing_county)

batch_header_file <- batch_header_file %>%
  # Drop the 28 NIBRS agencies that had no county info in LEAIC
  filter(!is.na(county)) %>%
  # Assign county fips from LEAIC to the NIBRS county fips variable
  mutate(fips_county_1 = county)

# Remove obsolete files
rm(leaic, bhf_missing_county)

```


## 4. Construct county-level population and demographics using ACS
Recall that this file serves as a precursor file to the files that create both of our metrics, which are:

* Reported violent crimes per 100,000 people and reported property crimes per 100,000 people
* Juvenile arrests per 100,000 juveniles

Additionally, for the juvenile arrests metric, we are interested in subgroups by age, race, and sex. To calculate these rates (both total and by subgroup), we need total population and demographic information for our denominators. We pull total population and demographic data from the 5-year ACS files for each year 2014-2022.

This section relies on the `tidycensus` package, which requires a Census API key. You can acquire a key [here](https://api.census.gov/data/key_signup.html) and learn more about installing your key [here](https://walker-data.com/tidycensus/reference/census_api_key.html).
```{r}
# set Census API once
# tidycensus::census_api_key("db0bd75bd829570bc973bd4b4fd66ed6f50b8371", install=TRUE, overwrite = TRUE)

```

### 4.1 Check ACS variable names and identify those we need
Load ACS variables for our first and last years. Manually explore each file and spot check several observations. The naming conventions of ACS variables do not seem to change during our time period.
    Note that if they did we would need to split up the code that reads in the years below.
```{r}
# Check ACS variables for 2014 and 2022 to see if they change over time
variables2014 <- tidycensus::load_variables(2014, "acs5")
variables2022 <- tidycensus::load_variables(2022, "acs5")

```

### 4.2 Create vectors for years and variables of interest
    Note that the `two or more race` variable is not appearing and may not be available at the county level.
    todo(): depending on the subgroups desired for the safety metrics, the sex/age group/race variables pulled here may need updated.
```{r}
# Create a list of all our years
years <- lst(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022)

# Create a vector for our ACS variables of interest
my_vars <- c(
  total_people = "B01003_001",
  age_m_1014 = "B01001_005", 
  age_m_1517 = "B01001_006",
  age_f_1014 = "B01001_029", 
  age_f_1517 = "B01001_030",
  age_m_1014_white = "B01001A_005", 
  age_m_1014_black = "B01001B_005",
  age_m_1014_aian = "B01001C_005", 
  age_m_1014_asin = "B01001D_005",
  age_m_1014_nhpi = "B01001E_005", 
  age_m_1014_othr = "B01001F_005",
  age_m_1014_twom = "B01001G_005",
  age_m_1014_white_nh = "B01001H_005",
  age_m_1014_hispanic = "B01001I_005",
  age_m_1517_white = "B01001A_006", 
  age_m_1517_black = "B01001B_006",
  age_m_1517_aian = "B01001C_006", 
  age_m_1517_asin = "B01001D_006",
  age_m_1517_nhpi = "B01001E_006", 
  age_m_1517_othr = "B01001F_006",
  age_m_1517_twom = "B01001G_006",
  age_m_1517_white_nh = "B01001H_006", 
  age_m_1517_hispanic = "B01001I_006",
  age_f_1014_white = "B01001A_020", 
  age_f_1014_black = "B01001B_020",
  age_f_1014_aian = "B01001C_020", 
  age_f_1014_asin = "B01001D_020",
  age_f_1014_nhpi = "B01001E_020", 
  age_f_1014_othr = "B01001F_020",
  age_f_1014_twom = "B01001G_020",
  age_f_1014_white_nh = "B01001H_020",
  age_f_1014_hispanic = "B01001I_020",
  age_f_1517_white = "B01001A_021", 
  age_f_1517_black = "B01001B_021",
  age_f_1517_aian = "B01001C_021", 
  age_f_1517_asin = "B01001D_021",
  age_f_1517_nhpi = "B01001E_021",
  age_f_1517_othr = "B01001F_021",
  age_f_1517_twom = "B01001G_021",
  age_f_1517_white_nh = "B01001H_021", 
  age_f_1517_hispanic = "B01001I_021"
)

```

### 4.3 Pull 2014-2022 5-year ACS data at the county level
Pull 5-year ACS data at the county level for each year
```{r}
# Pull data
county_demo_full <- map_dfr(
  years,
  ~ tidycensus::get_acs(
    geography = "county",
    variables = my_vars,
    year = .x,
    survey = "acs5",
    output = "wide",
    geometry = FALSE
  ),
  # Create a year variable for each year
  .id = "year"
)

# Remove Puerto Rico from ACS data
county_demo <- county_demo_full %>%
  mutate(state = substr(GEOID, 1, 2)) %>%
  filter(state != "72")

# Remove obsolete files
rm(variables2014, variables2022, years, my_vars)

```

### 4.4 Clean ACS data
Clean variable names and drop margins of error
```{r}
county_demo <- county_demo %>%
  rename_with(~ sub("E$", "", .x), everything()) %>%
  select(-c(ends_with("M")))

```

#### 4.4a Construct subgroups for age, race/ethnicity, and sex
    todo(): For the rate of juvenile justice arrests metric, we want additional subgroups of sex (male, female) and age groups (age 10-14, age 15-17) for all years.
```{r}
# Sum sex and age groups by race
county_demo <- county_demo %>%
  mutate(
    age_1017 = age_m_1014 + age_m_1517 + age_f_1014 + age_f_1517,
    age_1017_white = age_m_1014_white + age_m_1517_white + age_f_1014_white + age_f_1517_white,
    age_1017_black = age_m_1014_black + age_m_1517_black + age_f_1014_black + age_f_1517_black,
    age_1017_aian = age_m_1014_aian + age_m_1517_aian + age_f_1014_aian + age_f_1517_aian,
    age_1017_asin = age_m_1014_asin + age_m_1517_asin + age_f_1014_asin + age_f_1517_asin,
    age_1017_nhpi = age_m_1014_nhpi + age_m_1517_nhpi + age_f_1014_nhpi + age_f_1517_nhpi,
    age_1017_othr = age_m_1014_othr + age_m_1517_othr + age_f_1014_othr + age_f_1517_othr,
    # age_1017_twom = age_m_1014_twom + age_m_1517_twom + age_f_1014_twom + age_f_1517_twom,
    age_1017_hispanic = age_m_1014_hispanic + age_m_1517_hispanic + age_f_1014_hispanic + age_f_1517_hispanic,
    age_1017_white_nh = age_m_1014_white_nh + age_m_1517_white_nh + age_f_1014_white_nh + age_f_1517_white_nh,
  )

# Edit race variables  
county_demo <- county_demo %>%  
  mutate(
    # Manually create "two or more" race variable
    age_1017_twom = age_1017 - age_1017_white - age_1017_black - 
      age_1017_aian - age_1017_asin - age_1017_nhpi - age_1017_othr,
    # Combine all asian races, other races, and two or more races
    age_1017_asian_other = age_1017_aian + age_1017_asin + age_1017_nhpi +
      age_1017_othr + age_1017_twom
  )
  
```

Limit ACS data to relevant variables
```{r}
# Limit to relevant variables
county_demo <- county_demo %>%
  select(year, GEOID, total_people, starts_with("age_1017"))

```



## 5. Join Batch Header File onto ACS county demographics by county

### 5.1 Reshape BHF to be long by agency and county
The NIBRS batch header file is currently long by `ori` (agency), such that there is one observation for each agency. These agencies have variables that contain county FIPS codes for each of the counties that agency serves. 

686 agencies cover 2 counties and 66 agencies cover 3 counties
```{r}
sum(!is.na(batch_header_file$fips_county_2))
sum(!is.na(batch_header_file$fips_county_3))

```

We reshape the data to be long by `ori` and `county` such that each observation represents each unique `ori` and `county` pair.
```{r}
# Reshape NIBRS data to be long by agency AND county
batch_header_file <- batch_header_file %>%
  mutate(across(starts_with("fips"), ~str_pad(., 3, pad = "0"))) %>%
  select(c(ori, state, core_city, agency_type, starts_with("fips"))) %>%
  pivot_longer(-c(ori, state, core_city, agency_type), values_to = "county")

# Remove observations that are not relevant to more that one county
batch_header_file <- batch_header_file %>%
  filter(!is.na(county))

```

### 5.2 Join the NIBRS data (Batch Header File) onto the ACS county population and demographics
Join agency/county pairs onto the ACS county/year-level population and demographics
```{r}
# Create GEOID
batch_header_file <- batch_header_file %>%
  mutate(GEOID = str_c(state, county))

# Join data
joined_data <- left_join(
  x = county_demo %>% 
    select(year, GEOID, total_people),
  y = batch_header_file,
  by = "GEOID"
)

```


There are no observations that are only in the Batch Header File but not the ACS data. However, 61 observations (county/year pairs) are only in the ACS data and not in the Batch Header File. 
    todo(): 43 of these observations are from Alaska but I am not sure how to crosswalk/harmonize them.
    todo(): 9 of these observations are from Hawaii but I am not sure how to crosswalk/harmonize them.
    Note that 9 of these observations are from Connecticut. Connecticut county changes are reflected in the 2022 ACS data products [source](https://www.census.gov/programs-surveys/acs/technical-documentation/table-and-geography-changes/2022/geography-changes.html). However, we don't have a way to crosswalk these observations for now.
    Note that these may be counties that did not have any reporting agencies and are therefore excluded from the Batch Header File?
```{r}
# Check whether there are any unmatched observations when joining the data
test <- left_join(
  x = county_demo, y = batch_header_file, by = "GEOID"
)

# Check which observations do not match 
test <- anti_join(
  x = county_demo, y = batch_header_file, by = "GEOID"
)

# Check which states the unmatched observations come from
test %>%
  mutate(state = substr(GEOID, 1, 2)) %>%
  group_by(state) %>%
  count()

# Remove the test version of this file
rm(test)

```


## 6. Create county-agency eeights
For agencies that cover multiple counties, assign weight based on county population (e.g., if Agency A has jurisdiction in both County 1 which has 10,000 people and County 2 which has 5,000 people, the weights will be 0.667 and 0.333, respectively).
```{r}
# Make agency weights based on population of counties they cover
joined_data <- joined_data %>%
  filter(!is.na(total_people)) %>%  
  group_by(year, ori) %>%
  mutate(weight = total_people / sum(total_people),
         weight = if_else(is.na(ori), NA, weight)) %>%
  ungroup()

```


Count number of agencies in each county.
```{r}
county_agency <- joined_data %>%
  group_by(GEOID, year) %>%
  summarize(n_agencies = n_distinct(ori),
            n_core_city = sum(core_city==1),
            core_city = max(core_city),
            n_agen_city = sum(agency_type==1),
            n_agen_cnty = sum(agency_type==2),
            n_agen_univ = sum(agency_type==3),
            n_agen_spcl = sum(agency_type==5),
            n_agen_trbl = sum(agency_type==7)) %>%
  ungroup()

```


## 7. Write out data

Write out county demographics only
```{r}
write_csv(county_demo, file = "modified data/all_county_demo.csv")

```

Write out county demographics plus agency info
```{r}
# Join these data
county_demo_agency <- left_join(
  x = county_demo,
  y = county_agency, 
  by = c("year", "GEOID")
)

# Write out
write_csv(county_demo_agency, file = "modified data/all_county_demo_agency.csv")

```

Write out agency-county level file with weights
```{r}
write_csv(joined_data, file = "modified data/all_agency_county.csv")

```

