---
title: "Combine Disability County Files"
author: "Aaron R. Williams, JP Walsh & Ridhi Purohit"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
execute:
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---
## Update 

This file was updated by Ridhi Purohit in March 2025 for metric release version 2025.

## Setup

```{r quarto-setup, include = FALSE}
options(knitr.kable.NA = "")

```

```{r load-packages}
library(tidyverse)
library(tidylog)
library(here)
library(assertr)
library(gt)

source(here::here("10_construct-database", "R", "combine_functions.R"))
```

## Construct Database

### Disability files

Create population file with a variable for disability. 

```{r}

expand_grid(
  read_csv(here("geographic-crosswalks", "data", "county-populations.csv")),
           subgroup = c("All", "With Disability", "Without Disability")
) %>% 
   mutate(subgroup_type = ifelse(subgroup != "All", "disability", "all")) %>% 
  select(-population) %>% 
  write_csv(here("data", "temp", "county-populations_disability.csv"))

```

Read file paths to all county files with disability subgroup information. 

```{r}

filepaths <- c(
  
# Race-ethnicity populations
here("data", "temp", "county-populations_disability.csv"),

# 08 education
here("08_education/data/final/metrics_college_county_disability_longitudinal.csv"),

# 09 employment
here("09_employment/data/final/metrics_employment_county_disability_longitudinal.csv")

)

```

This code loads each file and then combines them using `left_join()`. The first file is a population file, so there should be 3,143 counties per year. 

```{r}

db_disability <- filepaths |>
  map(
    .f = ~safe_read_csv(.x)
  ) |>
  reduce(left_join, by = c("year", "state", "county", "subgroup", "subgroup_type"))

```

### Clean file

Some subgroup information is not available for entire years. To simplify the user experience, remove these years entirely instead of giving a bunch of NAs.

```{r}

years_data <- db_disability %>% 
  group_by(year) %>% 
  filter_at(vars(-year, -state, -state_name, -county, -county_name, -subgroup_type, -subgroup), any_vars(!is.na(.))) %>% 
  group_by(year) %>% 
  count() %>% 
  select(year) %>% 
  pull()

```

Filter for only years that have data available.

```{r}
db_disability  <- db_disability |>
  filter(year %in% years_data)
```

The quality flags are currently inputed as numbers 1 through 3. Update the data so these are labeled as text: 1="Strong", 2="Marginal", 3="Weak".

```{r}
db_disability  <- db_disability  |>
  mutate(across(ends_with("quality"), ~case_when(. == 1 ~ "Strong",
                                                . == 2 ~ "Marginal",
                                                . == 3 ~ "Weak")))
```

## Quality and Completeness

### Dimensions

At most there should be 9,432 in a year (3X3144) and at the least there should be 9,426 in a year (3X3142).

```{r}

db_disability |>
  count(year) |>
  assert(
    within_bounds(9426, 9432),
    n
  )
```

### Data Quality Flags

This section summarizes data quality flags for variables and years.

```{r}
db_disability |>
  filter(subgroup != "All") |>
  select(ends_with("_quality")) |>
  pivot_longer(everything(), names_to = "variable", values_to = "quality") |>
  count(variable, quality) |>
  filter(!is.na(quality)) |>
  print(n = 100)

db_disability  |>
  filter(subgroup != "All") |>
  select(year, ends_with("_quality")) |>
  pivot_longer(-year, names_to = "variable", values_to = "quality") |>
  count(year, quality) |>
  print(n = 100)

```

### Full Database

Evaluate the quality variables. 

```{r}
db_disability |>
  select(ends_with("_quality")) |>
  assert(in_set("Strong", "Marginal", "Weak"), everything())

```


### 08 Education

```{r}
db_disability |>
  assert(
    within_bounds(0, 1),
    share_hs_degree,
    share_hs_degree_lb,
    share_hs_degree_ub
  ) |>
  quiet()

test_bounds(
  data = db_disability, 
  estimate = share_hs_degree, 
  lb = share_hs_degree_lb, 
  ub = share_hs_degree_ub
)
```

### 09 Employment

```{r}
db_disability |>
  assert(
    within_bounds(0, 1),
    share_employed,
    share_employed_lb,
    share_employed_ub
  ) |>
  quiet()

test_bounds(
  data = db_disability, 
  estimate = share_employed, 
  lb = share_employed_lb, 
  ub = share_employed_ub
)

```

## Write the File

```{r}

write_csv(db_disability, here("data", "18_mobility-metrics_county_disability_longitudinal.csv"))
```

## Delete temp files

Delete the temp files created.

```{r}

tempfiles <- c(
  
# Race-ethnicity populations
here("data", "temp", "county-populations_disability.csv")

)

file.remove(tempfiles)
```
