---
title: "Combine Income Place Files"
author: "Aaron R. Williams & JP Walsh"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

## Setup

```{r quarto-setup, include = FALSE}
options(knitr.kable.NA = "")

```

```{r load-packages}
library(tidyverse)
library(tidylog)
library(here)
library(assertr)
library(gt)

```

This function is used to ensure that the lower bound of a 95% confidence interval is always below the estimate and the upper bound of the a 95% confidence interval is always above the estimate.

```{r}
#' Test the bounds of a confidence interval relative to the estimate
#'
#' @param data The data frame of interest
#' @param estimate The unquoted name of the estimate variable
#' @param lb The unquoted name of the lower bound variable
#' @param ub The unquoted name of the upper bound variable
#'
test_bounds <- function(data, estimate, lb, ub) {

  subset <- bind_rows(
    data |>
      filter({{ ub }} < {{ lb }}),
    data |>
      filter({{ estimate }} > {{ ub }}),
    data |>
      filter({{ estimate  }} < {{ lb }}),
  )

  stopifnot(nrow(subset) == 0)

}


#' Helper function to silence output from testing code
#'
#' @param data A data frame
#'
quiet <- function(data) {

  quiet <- data

}

```

## Construct Database

### Temporary Fixes

This section implements temporary fixes and saves the resulting data in a temp folder. The saved data are then included in the output data.

```{r}
read_csv(here("06_neighborhoods/environment/data/final/environment_place_poverty_longitudinal.csv")) |>
  mutate(subgroup_type = if_else(subgroup == "All", "all", "income")) |>
  write_csv(here("data", "temp", "place_envir_longitudinal_econ-well-being.csv"))

read_csv(here("08_education", "data", "built", "SEDA_all_subgroups_city_2014-2018.csv")) |>
    filter(subgroup_type == "income") |>
    write_csv(here("data", "temp", "SEDA_income_subgroup_city.csv"))
```


### Income and economic disadvantage

Create population file with a variable covering all income subgroup values.

```{r}

expand_grid(
  read_csv(here("geographic-crosswalks", "data", "place-populations.csv")),
           subgroup = c("All", "Economically Disadvantaged", "Not Economically Disadvantaged",
                        "Less than $50,000", "$50,000 or More", "High-Poverty", "Not High-Poverty")
) %>%
   mutate(subgroup_type = ifelse(subgroup != "All", "income", "all")) %>%
  select(-population) %>%
  write_csv(here("data", "temp", "place-populations_econ_dis.csv"))

```

Read file paths to all place files with income subgroup information.

```{r}
filepaths <- c(
  # Race-ethnicity populations
  here("data", "temp", "place-populations_econ_dis.csv"),

  # 06 neighborhoods
  here("data", "temp", "place_envir_longitudinal_econ-well-being.csv"),

  # 08 education
  here("data", "temp", "SEDA_income_subgroup_city.csv"),
  here("08_education/data/final/digital_access_place_income_longitudinal.csv")
)

```

This code loads each file and then combines them using `left_join()`. The first file is a population file, so there should be 486 counties per year.

```{r}

db_income <- filepaths |>
  map(
    .f = ~read_csv(.x)
  ) |>
  reduce(left_join, by = c("year", "state", "place", "subgroup", "subgroup_type"))

```

We used old learning rate data, so we rename the variables.

```{r}

db_income  <- db_income  |>
  rename(
    rate_learning = learning_rate,
    rate_learning_lb = learning_rate_lb,
    rate_learning_ub = learning_rate_ub,
    rate_learning_quality = learning_rate_quality
  )

```

### Clean file

Some subgroup information is not available for entire years. To simplify the user experience, remove these years entirely instead of giving a bunch of NAs.

```{r}

years_data <- db_income %>%
  group_by(year) %>%
  filter_at(vars(-year, -state, -state_name, -place, -place_name, -subgroup_type, -subgroup), any_vars(!is.na(.))) %>%
  group_by(year) %>%
  count() %>%
  select(year) %>%
  pull()

```

Filter for only years that have data available.

```{r}
db_income  <- db_income  |>
  filter(year %in% years_data)
```

The quality flags are currently inputed as numbers 1 through 3. Update the data so these are labeled as text: 1="Strong", 2="Marginal", 3="Weak".

```{r}
db_income  <- db_income  |>
  mutate(across(ends_with("quality"), ~case_when(. == 1 ~ "Strong",
                                                . == 2 ~ "Marginal",
                                                . == 3 ~ "Weak")))
```



## Quality and Completeness

### Dimensions

At most there should be 3,402 in a year (7X486) and at the least there should be 3,395 in a year (7X485).

```{r}

db_income |>
  count(year) |>
  assert(
    within_bounds(3395, 3402),
    n
  )

```

### Data Quality Flags

This section summarizes data quality flags for variables and years.

```{r}
db_income |>
  filter(subgroup != "All") |>
  select(ends_with("_quality")) |>
  pivot_longer(everything(), names_to = "variable", values_to = "quality") |>
  count(variable, quality) |>
  filter(!is.na(quality)) |>
  print(n = 100)

db_income  |>
  filter(subgroup != "All") |>
  select(year, ends_with("_quality")) |>
  pivot_longer(-year, names_to = "variable", values_to = "quality") |>
  count(year, quality) |>
  print(n = 100)

```

### Full Database

Evaluate the quality variables.

```{r}
db_income |>
  select(ends_with("_quality")) |>
  assert(in_set("Strong", "Marginal", "Weak"), everything())

```

### 06 Neighborhoods

```{r}
db_income |>
  assert(
    within_bounds(0, 100),
    index_air_hazard
  ) |>
  quiet()

```

### 08 Education

```{r}

db_income |>
  assert(
    within_bounds(-10, 10),
    rate_learning,
    rate_learning_lb,
    rate_learning_ub
  ) |>
  quiet()

test_bounds(
  data = db_income,
  estimate = rate_learning,
  lb = rate_learning_lb,
  ub = rate_learning_ub
)

db_income |>
  assert(
    within_bounds(0, 1),
    share_digital_access
  ) |>
  quiet()

```


## Write the File

```{r}

write_csv(db_income, here("data", "64_mobility-metrics_place_income_longitudinal.csv"))

```
