---
title: "Combine Tenure Place Files"
author: "Aaron R. Williams & JP Walsh"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
execute:
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

## Setup

```{r quarto-setup, include = FALSE}
options(knitr.kable.NA = "")

```

```{r load-packages}
library(tidyverse)
library(tidylog)
library(here)
library(assertr)
library(gt)

```

This function is used to ensure that the lower bound of a 95% confidence interval is always below the estimate and the upper bound of the a 95% confidence interval is always above the estimate. 

```{r}
#' Test the bounds of a confidence interval relative to the estimate
#'
#' @param data The data frame of interest
#' @param estimate The unquoted name of the estimate variable
#' @param lb The unquoted name of the lower bound variable
#' @param ub The unquoted name of the upper bound variable
#'
test_bounds <- function(data, estimate, lb, ub) {
  
  subset <- bind_rows(
    data |>
      filter({{ ub }} < {{ lb }}),
    data |>
      filter({{ estimate }} > {{ ub }}),
    data |>
      filter({{ estimate  }} < {{ lb }}),
  )
  
  stopifnot(nrow(subset) == 0)
  
}


#' Helper function to silence output from testing code
#'
#' @param data A data frame
#'
quiet <- function(data) {
  
  quiet <- data
  
}

```

## Construct Database

### Tenure files

Create population file with a variable for Tenure. 

```{r}

read_csv(here("geographic-crosswalks", "data", "place-populations.csv")) %>% 
  mutate(all = "All", 
         tenure_own = "Male", 
         tenure_rent = "Female") %>% 
  pivot_longer(cols = c(all:tenure_rent),
               names_to = "subgroup_type",
               values_to = "subgroup") %>% 
  mutate(subgroup_type = ifelse(subgroup_type != "all", "tenure", "all")) %>% 
  write_csv(here("data", "temp", "place-populations_tenure.csv"))

```

Read file paths to all place files with tenure subgroup information. 

```{r}

filepaths <- c(
  
# Race-ethnicity populations
here("data", "temp", "place-populations_Tenure.csv"),


)

```

This code loads each file and then combines them using `left_join()`. The first file is a population file, so there should be 3,143 counties per year. 

```{r}

db_Tenure <- filepaths |>
  map(
    .f = ~read_csv(.x) |> 
      select(-any_of(c("state_name", "place_name")))
  ) |>
  reduce(left_join, by = c("year", "state", "place", "subgroup", "subgroup_type"))

```

## Quality and Completeness

### Dimensions

At most there should be 9,432 in a year (3X3144) and at the least there should be 9,426 in a year (3X3142).

```{r}

db_Tenure |> 
  count(year) |>
  assert(
    within_bounds(9426, 9432), 
    n
  )
```

### Data Quality Flags

This section summarizes data quality flags for variables and years.

```{r}
db_Tenure |>
  select(ends_with("_quality")) |>
  pivot_longer(everything(), names_to = "variable", values_to = "quality") |>
  count(variable, quality) |>
  filter(!is.na(quality)) |>
  print(n = 100)

db_Tenure  |>
  select(year, ends_with("_quality")) |>
  pivot_longer(-year, names_to = "variable", values_to = "quality") |>
  count(year, quality) |>
  print(n = 100)

```

### Full Database

Evaluate the quality variables. 

```{r}
db_Tenure |>
  select(ends_with("_quality")) |>
  assert(in_set(1, 2, 3), everything())

```



### 07 Safety

```{r}
db_Tenure |>
  assert(
    within_bounds(0, Inf),
    rate_juv_arrest
  ) |>
  quiet()

```

### 08 Education

```{r}
db_Tenure |>
  assert(
    within_bounds(0, 1),
    share_hs_degree,
    share_hs_degree_lb,
    share_hs_degree_ub
  ) |>
  quiet()

test_bounds(
  data = db_Tenure, 
  estimate = share_hs_degree, 
  lb = share_hs_degree_lb, 
  ub = share_hs_degree_ub
)
```

### 09 Employment

```{r}
db_Tenure |>
  assert(
    within_bounds(0, 1),
    share_employed,
    share_employed_lb,
    share_employed_ub
  ) |>
  quiet()

test_bounds(
  data = db_Tenure, 
  estimate = share_employed, 
  lb = share_employed_lb, 
  ub = share_employed_ub
)

```

## Write the File

```{r}

write_csv(db_Tenure, here("data", "16_mobility-metrics_place_Tenure_longitudinal.csv"))
```
