---
title: "County Recent Data"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

source("R/summarize_metric.R")

mm_recent <- read_csv("../data/00_mobility-metrics_recent.csv")

```

## State

`state`
`state_name`

## County

`county`
`county_name`

## Population

`population`

## Predictor: Housing Affordablility

### Metric: Ratio of affordable and available housing units to households with low-, very low–, and extremely low income levels

**Variables:** `share_affordable_30_ami`, `share_affordable_50_ami`, `share_affordable_80_ami`, `share_affordable_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_affordable_30_ami, 
  share_affordable_quality
)

summarize_metric(
  .data = mm_recent, 
  var = share_affordable_50_ami, 
  share_affordable_quality
)

summarize_metric(
  .data = mm_recent, 
  var = share_affordable_80_ami, 
  share_affordable_quality
)

```

---

## Predictor: Housing Stability

### Metric: Number and share of public school children who are ever homeless during the school year

**Variables:** `count_homeless`, `share_homeless`, `count_homeless_lb`, `count_homeless_ub`, `homeless_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = count_homeless, 
  quality_var = homeless_quality,
  decimals = 0
)

summarize_metric(
  .data = mm_recent, 
  var = share_homeless, 
  quality_var = homeless_quality
)

```

---

## Predictor: Economic Inclusion

### Metric: Share of residents experiencing poverty living in high-poverty neighborhoods

**Variables:** `share_poverty_exposure`, `share_poverty_exposure_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_poverty_exposure, 
  quality_var = share_poverty_exposure_quality
)

```

---

## Predictor: Racial Diversity

### Metric: Neighborhood exposure index, or the share of a person’s neighbors who are people of other races and ethnicities

**Variables:** `share_black_nh_exposure`, `share_black_nh_exposure_quality`, `share_hispanic_exposure`, `share_hispanic_exposure_quality`, `share_other_nh_exposure`, `share_other_nh_exposure_quality`, `share_white_nh_exposure`, `share_white_nh_exposure_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_black_nh_exposure, 
  quality_var = share_black_nh_exposure_quality
)

summarize_metric(
  .data = mm_recent, 
  var = share_hispanic_exposure, 
  quality_var = share_hispanic_exposure_quality
)

summarize_metric(
  .data = mm_recent, 
  var = share_other_nh_exposure, 
  quality_var = share_other_nh_exposure_quality
)

summarize_metric(
  .data = mm_recent, 
  var = share_white_nh_exposure, 
  quality_var = share_white_nh_exposure_quality
)

```

---

## Predictor: Social Capital

### Metric: Number of membership associations per 10,000 people

**Variables:** `count_membership_associations_per_10k`, `count_membership_associations_per_10k_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = count_membership_associations_per_10k, 
  quality_var = count_membership_associations_per_10k_quality
)

```

### Metric: Ratio of Facebook friends with higher socioeconomic status to Facebook friends with lower socioeconomic status ("economic connectedness")

**Variables:** `ratio_high_low_ses_fb_friends`, `ratio_high_low_ses_fb_friends_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = ratio_high_low_ses_fb_friends, 
  quality_var = ratio_high_low_ses_fb_friends_quality
)

```

---

## Predictor: Transportation Access

### Metric: Transit trips index

**Variables:** `transportation_cost`, `transportation_cost_quality`, `index_transit_trips`, `index_transit_trips_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = index_transit_trips, 
  quality_var = index_transit_trips_quality
)

summarize_metric(
  .data = mm_recent, 
  var = transportation_cost, 
  quality_var = transportation_cost_quality
)

```

---

## Predictor: Access to Preschool

### Metric: Share of 3- to 4-year-olds enrolled in nursery school or preschool

**Variables:** `share_in_preschool`, `share_in_preschool_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_in_preschool, 
  quality_var = share_in_preschool_quality
)

```

---

## Predictor: Effective Public Education

### Metric: Average per grade change in English Language Arts achievement between third and eighth grades

**Variables:** `rate_learning`, `rate_learning_lb`, `rate_learning_ub`, `rate_learning_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = rate_learning, 
  quality_var = rate_learning_quality
)

```

---

## Predictor: School Economic Diversity

### Metric: Share of students attending high-poverty schools, by student race or ethnicity

**Variables:** `meps20_black`, `meps20_black_quality`, `meps20_hispanic`, `meps20_hispanic_quality`, `meps20_white`, `meps20_white_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = meps20_black, 
  quality_var = meps20_black_quality
)

summarize_metric(
  .data = mm_recent, 
  var = meps20_hispanic, 
  quality_var = meps20_hispanic_quality
)

summarize_metric(
  .data = mm_recent, 
  var = meps20_white, 
  quality_var = meps20_white_quality
)

```

---

## Predictor: Preparation for College

### Metric: Share of 19- and 20-year-olds with a high school degree

**Variables:** `share_hs_degree`, `share_hs_degree_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_hs_degree, 
  quality_var = share_hs_degree_quality
)

```

---

## Predictor: Digital Access

### Metric: Share of households with broadband access in the home

**Variables:** `share_digital_access`, `share_digital_access_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_digital_access, 
  quality_var = share_digital_access_quality
)

```

---

## Predictor: Employment Opportunities

### Metric: Employment-to-population ratio for adults ages 25 to 54

**Variables:** `share_employed`, `share_employed_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_employed, 
  quality_var = share_employed_quality
)

```

---

## Predictor: Jobs Paying Living Wages

### Metric: Ratio of pay on the average job to the cost of living

**Variables:** `ratio_average_to_living_wage`, `wage_ratio_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = ratio_average_to_living_wage, 
  quality_var = ratio_average_to_living_wage_quality
)

```

---

## Predictor: Opportunities for Income

### Metric: Household income at the 20th, 50th, and 80th percentiles

**Variables:** `pctl_income_20`, `pctl_income_50`, `pctl_income_80`, `pctl_income_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = pctl_income_20, 
  quality_var = pctl_income_quality
)

summarize_metric(
  .data = mm_recent, 
  var = pctl_income_50, 
  quality_var = pctl_income_quality
)

summarize_metric(
  .data = mm_recent, 
  var = pctl_income_80, 
  quality_var = pctl_income_quality
)

```

---

## Predictor: Financial Security

### Metric: Share with debt in collections

**Variables:** `share_debt_col`, `share_debt_col_lb`, `share_debt_col_ub`, `share_debt_col_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_debt_col, 
  quality_var = share_debt_col_quality
)

```

---

## Predictor: Wealth Building Opportunities

### Metric: Ratio of the share of a community’s housing wealth held by a racial or ethnic group to the share of households of the same group

**Variables:** `ratio_black_nh_house_value_households`, `ratio_hispanic_house_value_households`, `ratio_other_nh_house_value_households`, 
`ratio_white_nh_house_value_households`, `ratio_black_nh_house_value_households_quality`, `ratio_hispanic_house_value_households_quality`, `ratio_other_nh_house_value_households_quality`, `ratio_white_nh_house_value_households_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = ratio_black_nh_house_value_households, 
  quality_var = ratio_black_nh_house_value_households_quality
)

summarize_metric(
  .data = mm_recent, 
  var = ratio_hispanic_house_value_households, 
  quality_var = ratio_hispanic_house_value_households_quality
)

summarize_metric(
  .data = mm_recent, 
  var = ratio_other_nh_house_value_households, 
  quality_var = ratio_other_nh_house_value_households_quality
)

summarize_metric(
  .data = mm_recent, 
  var = ratio_white_nh_house_value_households, 
  quality_var = ratio_white_nh_house_value_households_quality
)

```

---

## Predictor: Access to Health Services

### Metric: Ratio of population per primary care physician

**Variables:** `ratio_population_pc_physician`, `ratio_population_pc_physician_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = ratio_population_pc_physician, 
  quality_var = ratio_population_pc_physician_quality
)

```

---

## Predictor: Neonatal Health

### Metric: Share of low-weight births

**Variables:** `rate_low_birth_weight`, `rate_low_birth_weight_lb`, `rate_low_birth_weight_ub`, `rate_low_birth_weight_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = rate_low_birth_weight, 
  quality_var = rate_low_birth_weight_quality
)

```

---

## Predictor: Environmental Quality

### Metric: Air quality index

**Variables:** `index_air_quality`, `index_air_quality_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = index_air_quality, 
  quality_var = index_air_quality_quality
)

```

---

## Predictor: Safety from Trauma

### Metric: Number of deaths caused by injury per 100,000 people

**Variables:** `rate_injury_death`, `rate_injury_death_lb`, `rate_injury_death_ub`, `rate_injury_death_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = rate_injury_death, 
  quality_var = rate_injury_death_quality
)

```

---

## Predictor: Political Participation

### Metric: Share of the voting-eligible population who turn out to vote

**Variables:** `share_election_turnout`, `share_election_turnoutt_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = share_election_turnout, 
  quality_var = share_election_turnout_quality
)

```

---

## Predictor: Descriptive Representation

### Metric: Ratio of the share of local elected officials of a racial or ethnic group to the share of residents of the same racial or ethnic group

**Variables:** `share_desc_rep_asian_other`, `share_desc_rep_asian_other_quality`, `share_desc_rep_black_nonhispanic`, `share_desc_rep_hispanic`, `share_desc_rep_white_nonhispanic`

**Disaggregation:** (link)

---

## Predictor: Safety from Crime

### Metric: Rates of reported violent crime and property crime

**Variables:** `rate_violent_crime`, `rate_property_crime`, `rate_crime_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = rate_violent_crime, 
  quality_var = rate_crime_quality
)

summarize_metric(
  .data = mm_recent, 
  var = rate_property_crime, 
  quality_var = rate_crime_quality
)

```

---

## Predictor: Just Policing

### Metric: Rate of juvenile justice arrests

**Variables:** `rate_juv_arrest`, `rate_juv_arrest_quality`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_recent, 
  var = rate_juv_arrest, 
  quality_var = rate_juv_arrest_quality
)

```
