---
output: html_document
editor_options: 
  chunk_output_type: console
---
# 00_mobility-metrics_longitudinal.csv

```{r}
library(tidyverse)

mm_longitudinal <- read_csv("../data/00_mobility-metrics_longitudinal.csv")

mm <- read_csv("../data/01_mobility-metrics.csv")

mm_race_ethnicity  <- read_csv("../data/02_mobility-metrics_race-ethnicity.csv")

mm_race  <- read_csv("../data/03_mobility-metrics_race.csv")

mm_race_share  <- read_csv("../data/04_mobility-metrics_race-share.csv")


```


```{r}
glimpse(mm_longitudinal)

```

# Other Variables

`year`
`state`
`county`
`state_name`
`county_name`
`population`

# Predictor: Housing Affordablility

Metric Name
Metric Description
Metric type
Data quality
Missingness
Years
Diaggregation (link)

`share_affordable_30_ami`
`share_affordable_50_ami`
`share_affordable_80_ami`
`housing_quality`

# Predictor: Housing Stability

`homeless_count`
`homeless_share`
`homeless_count_lb`
`homeless_count_ub`
`homeless_quality`

# Predictor: Economic Inclusion

`poverty_exposure`
`poverty_exposure_quality`

# Predictor: Racial Diversity

`black_nh_exposure`
`black_nh_exposure_quality`
`hispanic_exposure`
`hispanic_exposure_quality`
`other_nh_exposure`
`other_nh_exposure_quality`
`white_nh_exposure`
`white_nh_exposure_quality`

# Predictor: Social Capital

# Predictor: Transportation Access

`transit_cost`
`transit_cost_quality`
`transit_trips`
`transit_trips_quality`

# Predictor: Access to Preschool

`share_in_preschool`
`share_in_preschool_lb`
`share_in_preschool_ub`
`preschool_quality`

# Predictor: Effective Public Education

`learning_rate`
`learning_rate_lb`
`learning_rate_ub`
`learning_rate_quality`

# Predictor: School Economic Diversity

???

# Predictor: Preparation for College

`share_hs_degree`
`share_hs_degree_lb`
`share_hs_degree_ub`
`hs_degree_quality`

# Predictor: Digital Access

???

# Predictor: Employment Opportunities

`share_employed`
`share_employed_lb`
`share_employed_ub`
`employed_quality`

# Predictor: Jobs Paying Living Wages

`average_to_living_wage_ratio`
`wage_ratio_quality`

# Predictor: Opportunities for Income

`pctl_20`
`pctl_50`
`pctl_80`
`pctl_quality`

# Predictor: Financial Security

`share_debt_coll`
`share_debt_coll_lb`
`share_debt_coll_ub`
`share_debt_coll_quality`

# Predictor: Wealth Building Opportunities

???

# Predictor: Access to Health Services

???

# Predictor: Neonatal Health

`lbw`

# Predictor: Environmental Quality

`environmental`
`environmental_quality`

# Predictor: Safety from Trauma

???

# Predictor: Political Participation

`election_turnout`
`election_turnout_quality`

# Predictor: Descriptive Representation

`asian_other`
`asian_other_quality`
`black_nonhispanic`
`black_nonhispanic_quality`
`hispanic`
`hispanic_quality`
`white_nonhispanic`
`white_nonhispanic_quality`

# Predictor: Safety from Crime

Was `violent_crime_rate`, will be something new
Was `property_crime_rate`, will be something new
`crime_rate_quality`

# Predictor: Just Policing

`juvenile_arrest_rate`
`juvenile_arrest_rate_quality`
