---
title: "County Race Data"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

source("R/get_years.R")
source("R/summarize_metric.R")

mm_race <- read_csv("../data/02_mobility-metrics_race_longitudinal.csv")

```

## Year

`year`

## State

`state`
`state_name`

## County

`county`
`county_name`

## Population

`population`

## Predictor: Digital Access

### Metric: Share of households with broadband access in the home

**Variables:** `share_digital_access`, `share_digital_access_quality`

**Years:** `r get_years(mm_race, var = share_digital_access_quality)`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_race, 
  var = share_digital_access, 
  quality_var = share_digital_access_quality
)

```

## Predictor: Just Policing

### Metric: Rate of juvenile justice arrests

**Variables:** `rate_juv_arrest`, `rate_juv_arrest_quality`

**Years:** `r get_years(mm_race, var = rate_juv_arrest_quality)`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_race, 
  var = rate_juv_arrest, 
  quality_var = rate_juv_arrest_quality
)

```
