---
title: "County Race Share Data"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

source("R/get_years.R")
source("R/summarize_metric.R")

mm_race_share <- read_csv("../data/03_mobility-metrics_race-share_longitudinal.csv")

```

## Predictor: Transportation Access

### Metric: Transit trips index

**Variables:** `transportation_cost`, `transportation_cost_quality`, `index_transit_trips`, `index_transit_trips_quality`

**Years:** `r get_years(mm_race_share, var = index_transit_trips_quality)`

```{r}
#| results: asis
summarize_metric(
  .data = mm_race_share, 
  var = index_transit_trips, 
  quality_var = index_transit_trips_quality
)

summarize_metric(
  .data = mm_race_share, 
  var = transportation_cost, 
  quality_var = transportation_cost_quality
)

```

---

## Predictor: Financial Security

### Metric: Share with debt in collections

**Variables:** `share_debt_col`, `share_debt_col_lb`, `share_debt_col_ub`, `share_debt_col_quality`

**Years:** `r get_years(mm_race_share, var = share_debt_col_quality)`

```{r}
#| results: asis
summarize_metric(
  .data = mm_race_share, 
  var = share_debt_col, 
  quality_var = share_debt_col_quality
)

```

---

## Predictor: Environmental Quality

### Metric: Air quality index

**Variables:** `index_air_quality`, `index_air_quality_quality`

**Years:** `r get_years(mm_race_share, var = index_air_quality_quality)`

```{r}
#| results: asis
summarize_metric(
  .data = mm_race_share, 
  var = index_air_quality, 
  quality_var = index_air_quality_quality
)

```

---


