---
title: "City Race/Ethnicity Data"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

source("R/get_years.R")
source("R/summarize_metric.R")

mm_city_race_ethnicity  <- read_csv("../data/06_mobility-metrics_place_race-ethnicity_longitudinal.csv")

```

## Year

`year`

## State

`state`
`state_name`

## County

`county`
`county_name`

## Population

`population`

---

## Predictor: Access to Preschool

### Metric: Share of 3- to 4-year-olds enrolled in nursery school or preschool

**Variables:** `share_in_preschool`, `share_in_preschool_quality`

**Years:** `r get_years(mm_city_race_ethnicity, var = share_in_preschool_quality)`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_city_race_ethnicity, 
  var = share_in_preschool, 
  quality_var = share_in_preschool_quality
)

```

---

## Predictor: Effective Public Education

### Metric: Average per grade change in English Language Arts achievement between third and eighth grades

**Variables:** `rate_learning`, `rate_learning_lb`, `rate_learning_ub`, `rate_learning_quality`

**Years:** `r get_years(mm_city_race_ethnicity, var = rate_learning_quality)`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_city_race_ethnicity, 
  var = rate_learning, 
  quality_var = rate_learning_quality
)

```

---

## Predictor: Preparation for College

### Metric: Share of 19- and 20-year-olds with a high school degree

**Variables:** `share_hs_degree`, `share_hs_degree_quality`

**Years:** `r get_years(mm_city_race_ethnicity, var = share_hs_degree_quality)`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_city_race_ethnicity, 
  var = share_hs_degree, 
  quality_var = share_hs_degree_quality
)

```

---

## Predictor: Employment Opportunities

### Metric: Employment-to-population ratio for adults ages 25 to 54

**Variables:** `share_employed`, `share_employed_quality`

**Years:** `r get_years(mm_city_race_ethnicity, var = share_employed_quality)`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_city_race_ethnicity, 
  var = share_employed, 
  quality_var = share_employed_quality
)

```

---

## Predictor: Opportunities for Income

### Metric: Household income at the 20th, 50th, and 80th percentiles

**Variables:** `pctl_income_20`, `pctl_income_50`, `pctl_income_80`, `pctl_income_quality`

**Years:** `r get_years(mm_city_race_ethnicity, var = pctl_income_quality)`

**Disaggregation:** (link)

```{r}
#| results: asis
summarize_metric(
  .data = mm_city_race_ethnicity, 
  var = pctl_income_20, 
  quality_var = pctl_income_quality
)

summarize_metric(
  .data = mm_city_race_ethnicity, 
  var = pctl_income_50, 
  quality_var = pctl_income_quality
)

summarize_metric(
  .data = mm_city_race_ethnicity, 
  var = pctl_income_80, 
  quality_var = pctl_income_quality
)

```
