---
title: "Final Data Evaluation Function Example"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

## Final Data Test Function

The data team for the Mobility Metrics project has created a function that tests baseline requirements for the structure and content of data produced for each metric. This function relies in part on information provided in a metric-specific final data expectations form, which should be filled out collaboratively between the data team and metric lead prior to having the metric approved by their technical reviewer. 

Metric leads should include this function near the end of their programs, preferably just before the code writes out the final data as a CSv. 

The function lives in the "functions/testing" file in the repository and is a R script title "evaluate_final_data.R". 

The template for the final data expectation evaluation form is also in the "functions/testing" as a CSV document titled "final_data_evaluation_form.csv".

## Example

This function must be used inside the program that creates your metric. It should be ran on the final version of the data that your program will read out as a CSV and which will subsequently be read into the overall metric combined files.

The first step is to source the function at the top of your program using the following code. This will stage the function in your environment.

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```

In order to run the function you must fill out all relevant arguments to match the specifics of your code and final data file being tested. In this example, we evaluate the 2022 subgroup data for the housing availability metric.

As a first step, this example reads in the actual final data for the housing availability metric with subgroups at the county level for 2022. Note that this data would exist as a data frame within the program you are working on so you would not need to read it in.

```{r}
final_data <- read_csv(here("02_housing/data/available_2022_subgroups_county.csv"), show_col_types = FALSE)
```

The function's first argument asks for the path to the final data evaluation form. Note that we have filled out an example final data evaluation form for this example which you can find in the "functions/testing" folder. Generally, users should save their final data evaluation forms in the final data folder related to that metric.

Second, the function requires the name of the object containing the final metric data (in this example, `final_data`).

Third, the function asks whether there are subgroups in this data.

Finally, the function asks whether there are confidence intervals in this data.

```{r}
evaluate_final_data(exp_form_path = "functions/testing/example_final_data_evaluation_form_housing.csv",
                data = final_data,  geography = "county",
                subgroups = TRUE, confidence_intervals = FALSE)
```

It passed! Yay!
