---
title: "Validate Geographies Function Example"
author: "Manuel Alcala Kovalski"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute:
  warning: false
  error: true
editor_options:
  chunk_output_type: console
---

## Overview

`validate_geographies()` checks whether all geographic identifiers (GEOIDs)
in a crosswalk file are present in a metric's final data. The crosswalk is
the source of truth for which counties or places should exist. Running this
check at the end of a metric creation script catches silent geography drops
before they make it into the combined database.

The function lives in `functions/testing/check_crosswalk_geoids_present.R`.

```{r}
#| label: setup
library(tidyverse)
source(here::here("functions", "testing", "check_crosswalk_geoids_present.R"))
```

## Example 1: County-level check

This checks the housing affordability metric against the long county
crosswalk. The crosswalk has a `year` column, so the function checks each
year independently.

```{r}
#| label: county-check
result_county <- validate_geographies(
  crosswalk_path = here::here(
    "geographic-crosswalks", "data", "long-xwalk_county.csv"
  ),
  metric_path = here::here(
    "11_housing-affordability", "data", "final",
    "housing_county_2014_2015_2016_2017_2018_2019_2021_2022_2023.csv"
  )
)

result_county |> select(year, all_present, n_missing)
```

## Example 2: Place-level check

Same metric at the place level, using the long place crosswalk.

```{r}
#| label: place-check
result_place <- validate_geographies(
  crosswalk_path = here::here(
    "geographic-crosswalks", "data", "long-xwalk_place.csv"
  ),
  metric_path = here::here(
    "11_housing-affordability", "data", "final",
    "housing_place_2014_2015_2016_2017_2018_2019_2021_2022_2023.csv"
  )
)

result_place |> select(year, all_present, n_missing)
```

## Example 3: Crosswalk with `crosswalk_period`

Some crosswalks use `crosswalk_period` ("pre-2022" / "2022") instead of a
`year` column. The PUMA-to-county crosswalk is the main example. The function
maps each metric year to its corresponding period and checks GEOIDs
accordingly.

```{r}
#| label: puma-crosswalk-check
result_puma <- validate_geographies(
  crosswalk_path = here::here(
    "geographic-crosswalks", "data", "crosswalk_puma_to_county.csv"
  ),
  metric_path = here::here(
    "11_housing-affordability", "data", "final",
    "housing_county_2014_2015_2016_2017_2018_2019_2021_2022_2023.csv"
  )
)

result_puma |> select(year, all_present, n_missing)
```

## Example 4: Detecting per-year gaps

This is the key scenario the function is designed to catch. If a county is
present in most years but missing from one, the per-year check flags it. We
simulate this by dropping 5 Alabama counties from 2018 only.

```{r}
#| label: per-year-gap
metric <- read_csv(
  here::here(
    "11_housing-affordability", "data", "final",
    "housing_county_2014_2015_2016_2017_2018_2019_2021_2022_2023.csv"
  ),
  show_col_types = FALSE
)

# Drop 5 counties from 2018 only
dropped_counties <- metric |>
  filter(year == 2018, state == "01") |>
  distinct(state, county) |>
  slice_head(n = 5)

dropped_counties

metric_with_gap <- metric |>
  anti_join(
    dropped_counties |> mutate(year = 2018),
    by = c("state", "county", "year")
  )

# Write to a temp file and validate
tmp <- tempfile(fileext = ".csv")
write_csv(metric_with_gap, tmp)

result_gap <- validate_geographies(
  crosswalk_path = here::here(
    "geographic-crosswalks", "data", "long-xwalk_county.csv"
  ),
  metric_path = tmp
)

result_gap |> select(year, all_present, n_missing)
```

The function correctly identifies that 2018 is missing 5 GEOIDs while all
other years pass.

## Example 5: Filtering to specific years

The `years` parameter restricts the check to specific years. This is useful
when you only want to validate the years you just updated.

```{r}
#| label: year-filter
result_filtered <- validate_geographies(
  crosswalk_path = here::here(
    "geographic-crosswalks", "data", "long-xwalk_county.csv"
  ),
  metric_path = here::here(
    "11_housing-affordability", "data", "final",
    "housing_county_2014_2015_2016_2017_2018_2019_2021_2022_2023.csv"
  ),
  years = c(2022, 2023)
)

result_filtered |> select(year, all_present, n_missing)
```

## Example 6: Error handling

The function provides informative errors when inputs are wrong. Here we pass
a misspelled column name to show the fuzzy matching suggestion.

```{r}
#| label: misspelled-column
validate_geographies(
  crosswalk_path = here::here(
    "geographic-crosswalks", "data", "long-xwalk_county.csv"
  ),
  metric_path = here::here(
    "11_housing-affordability", "data", "final",
    "housing_county_2014_2015_2016_2017_2018_2019_2021_2022_2023.csv"
  ),
  metric_geo_cols = c("stte", "county")
)
```

And here we request a year that does not exist in the data.

```{r}
#| label: nonexistent-year
validate_geographies(
  crosswalk_path = here::here(
    "geographic-crosswalks", "data", "long-xwalk_county.csv"
  ),
  metric_path = here::here(
    "11_housing-affordability", "data", "final",
    "housing_county_2014_2015_2016_2017_2018_2019_2021_2022_2023.csv"
  ),
  years = 2050
)
```
