---
title: ""
author: ""
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: false
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: !expr here::here("04_health", "www", "web_report.css")
    editor_options:
      chunk_output_type: console
---

<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

# Exposure to Trauma
This metric is a county-level estimate of exposure to trauma, defined as the number of deaths due to injury per 100,000 population. We pull this data from the 2022 County Health Rankings, which uses 2016-2020 data from the National Center for Health Statistics - Mortality Files.

See the [County Health Rankings](https://www.countyhealthrankings.org/explore-health-rankings/measures-data-sources/county-health-rankings-model/health-factors/social-and-economic-factors/community-safety/injury-deaths) for more information.


**Process:**

1. Download Excel files for county-level data from each state

2. Add data quality flags

3. Save the data

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup}
options(scipen = 999)

library(urbnthemes)
library(tidyverse)
library(readxl)
library(writexl)

library(rvest)
library(glue)
library(fs)

set_urbn_defaults(style = "print")

```

## 1. Prepare for file download
```{r}
base_url <- "https://www.countyhealthrankings.org"
url <- paste0(base_url, "/explore-health-rankings/rankings-data-documentation")
download_urls <-
  read_html(url) %>% 
  html_elements('span.field-content') %>%  
  html_element('a') %>% 
  html_attr('href') %>%
  paste0(base_url, .)

out_state <-
  read_html(url) %>% 
  html_elements('span.field-content') %>% 
  html_text2() %>% 
  str_replace_all(' ', '-')

get_state_data_url <- function(url){
  read_html(url) %>% 
    html_elements('span.download-link') %>% 
    html_element('a') %>% 
    html_attr('href') %>% 
    str_subset('.xls') %>% 
    str_subset('2022') %>% 
    paste0(base_url, .)
}
state_data_url <- purrr::map_chr(.x = download_urls, .f = get_state_data_url)
```


## 2. Download files

Download all 51 Excel files from each state's URL
```{r}
# create a data folder for intermediate data files
intermediate_files <- here::here("04_health",
                                 "exposure-to-trauma",
                                 "data/")

if (!dir.exists(intermediate_files)) {
  
  dir.create(intermediate_files)
  
}

chr_data <- here::here("04_health",
                       "exposure-to-trauma",
                       "data",
                       out_state
)

download_data <- function(state_data_url, state_name){
  if (!file.exists(here::here(glue("04_health/exposure-to-trauma/data/{state_name}.xlsx")))) {
    
    download.file(state_data_url, 
                  destfile = here::here(glue("04_health/exposure-to-trauma/data/{state_name}.xlsx")))
    
  }
}

walk2(state_data_url, out_state, download_data)

```


## 3. Read in, clean, and append each Excel sheet
```{r}
injury_deaths <- map_df(out_state,
                        ~{readxl::read_excel(path = here::here(
                          paste0("04_health/exposure-to-trauma/data/", .x, ".xlsx")),
                          sheet = "Ranked Measure Data",
                          skip = 1
                        ) %>%
                            select(fips = 'FIPS',
                                   # number_injury_deaths = '# Injury Deaths',
                                   injury_death_rate = 'Injury Death Rate') %>%
                            filter(!str_detect(fips, '000$')) %>%
                            mutate(state = str_sub(fips, 1, 2),
                                   county = str_sub(fips, 3, 5),
                                   injury_death_rate = round(injury_death_rate, 2)) %>%
                            select(-fips)
                        }
)

```


## 4. Add data quality flags
```{r}


```


## 5. Save the data
```{r}
all_counties <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv")) %>%
  tidylog::filter(year == 2020) %>%
  select(-year)

all_counties <- left_join(all_counties, injury_deaths, by = c("state", "county"))

test <- anti_join(all_counties, injury_deaths, by = c("state", "county"))

all_counties <- all_counties %>%
  mutate(year = 2020) %>%
  select(year, state, county, injury_death_rate)

all_counties %>%
  write_csv(here::here("04_health",
                       "exposure-to-trauma",
                       "exposure-to-trauma.csv"))
```

