---
title: "Geo-crosswalk Update Quality Check"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

## Overview and House keeping

During the process of the tranche3 update to the Upward Mobility Metrics the data team made a small change to the geographic crosswalk process for metrics using 5-year ACS IPUMS data for 2014 calculations. This impacts the following metrics:

- 01_employment-opportunities
- 03_opportunities-for-income
- 06_access_to_preschool
- 07_preperation-for-college

The update includes pre-2012 PUMA crosswalks to have a more accurate match for the multi-years 2009-2011 that are included in the 5-year 2014 data. This correction does not apply to post-2022 data because IPUMS updated PUMAs to be unviersal in the 5-year data for this period.

Read in packages.

```{r}
library(tidyverse)
library(reactable)
library(scales)
library(furrr)
library(tictoc)
library(here)

options(scipen = 999)

source(here("functions", "construction", "compare_data.R"))
source(here("functions", "construction", "compare_metric_differences.R"))

theme_set(theme_minimal())
```


## Read data 

First read in the pre-update historical files from the remote data catalog, the updated version and combine into a list. Note the data catalog link will break when the files are updated next, this analysis is not intended to be updated.

```{r}
#| label: load-data

# Increase timeout for large file downloads
options(timeout = 300)

read_data <- function(file, date, local_path, root_url) {

  message("Reading: ", file)

  data <- 
    list(
      `local` = read_csv(here(local_path, str_remove(file, "_0")), show_col_types = FALSE),
      `catalog` = read_csv(paste0(root_url, date, file), show_col_types = FALSE)
    )
  
  return(data)
  
}

root_url <- "https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/"
local_path <- here("data")

data_urls <- tribble(
  ~date, ~file,
   "2026/06/03/", "61_mobility-metrics_place_longitudinal.csv",
  "2026/06/03/", "62_mobility-metrics_place_race-ethnicity_longitudinal.csv",
  "2026/06/03/", "66_mobility-metrics_place_gender_longitudinal.csv",
  "2026/06/03/", "68_mobility-metrics_place_disability_longitudinal.csv"
)

data <- map2(
  .x = data_urls$file |>
    purrr::set_names(),
  .y = data_urls$date,
  .f = ~read_data(
    file = .x, 
    date = .y,
    local_path = local_path, 
    root_url = root_url
  )
)

```


## Data quality

```{r}
#| label: compare-data
#| results: asis

walk2(
  .x = data,
  .y = names(data),
  .f = ~compare_data(.x, .y)
)

```

## Largest share_employed differences

```{r}
#| label: compare-share-employed-differences
#| results: asis

walk2(
  .x = data,
  .y = names(data),
  .f = ~compare_metric_differences(
    data_list = .x,
    header = .y,
    metric = "share_employed",
    top_n = 30
  )
)

```

