---
title: "Generate Puma to County Crosswalk"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

*Program for the creation of the PUMA to County crosswalk for UMF predictors*

## Executive Summary

This document generates PUMA-to-County crosswalk files for the UMF mobility metrics project. **NEW**: This version includes 2000 PUMA boundaries to support historical analysis back to the 2000 Census period.

**Three crosswalk periods:**
- **2000 PUMA → 2014 County** (for 2000-2011 ACS data)
- **2012 PUMA → 2014 County** (for 2012-2021 ACS data) 
- **2020 PUMA → 2020 County** (for 2022+ ACS data)

This enables consistent geographic analysis across 25+ years of Census data.

Program: Initially written by Tina Chelidze and updated by JP Walsh 2023-2024.

Latest update to bring in Connecticut Planning Regions for the 2025 metric update.

Primary data derived from the University of Missouri's [geographic correspondence engine](https://mcdc.missouri.edu/applications/geocorr.html).

Based on processes developed by Paul Johnson and Kevin Werner in SAS.

-   [Housekeeping](#housekeeping)
-   [Read data](#read-data)
-   [Clean crosswalks](#clean-crosswalks)
-   [Create quality variable](#create-quality-variable)
-   [Write files](#wite-files)


## Housekeeping

Read in packages for use. 

```{r}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)

options(scipen = 999)

cw_files <- c("puma_county_2020", "puma_county_2010", "puma_county_2k")
```

## Read data

Read the PUMA to County crosswalks from the raw data file in the geographic-crosswalks folder. The project is currently using three crosswalk files that are utilized based on the year of the Census data. For 2022 ACS and onward we use a crosswalk going from 2020 PUMA to 2020 County. For pre-2022 ACS data, we use a crosswalk from 2012 PUMA to 2014 County for 2012 to 2021 ACS and we use a crosswalk from 2K PUMA to 2014 County for pre-2012 ACS.


## Data Sources Detail

The GeoCorr website allows for a variety of selections in creating your crosswalk file. The selections used to create the files read here are detailed below:

### GeoCorr Configuration Parameters
| Parameter | 2000 PUMAs | 2012 PUMAs | 2020 PUMAs |
|-----------|------------|------------|------------|
| Source Geography | 2000 PUMA | 2012 PUMA | 2020 PUMA |
| Target Geography | 2014 County | 2014 County | 2020 County |
| Weight Variable | 2010 Census Pop | 2010 Census Pop | 2020 Census Pop |
| Coverage | All US States | All US States | All US States |

**Rationale**: Using 2010 population weights for both 2000 and 2012 PUMAs ensures consistency for pre-2020 Census period analysis.

[Geocorr2018](https://mcdc.missouri.edu/applications/geocorr2018.html)
*States:* All
*Source geography:* 2012 PUMA & 2k PUMA
*Target geography:* 2014 County
*Weighting variable:* Population (2010 census)
*Output Options:* Format CSV & Generate second allocation factor 

[Geocorr2022](https://mcdc.missouri.edu/applications/geocorr2022.html)
PUMA to County
*States:* All
*Source geography:* 2020 PUMA
*Target geography:* 2020 County (includes CT planning regions)
*Weighting variable:* Population (2020 census)
*Output Options:* Format CSV & Generate second allocation factor 


Read in the crosswalk files from the raw folder inside the geographic-crosswalk data folder.
```{r}
puma_county_2020 <- read_csv(here::here("geographic-crosswalks", "data", "raw", "geocorr2022_puma_county_2020.csv"))
puma_county_2010 <- read_csv(here::here("geographic-crosswalks", "data", "raw", "geocorr2018_puma_county_2010.csv")) %>% 
  rename(county = county14)
puma_county_2k <- read_csv(here::here("geographic-crosswalks", "data", "raw", "geocorr2018_puma_county_2k.csv")) %>% 
  rename(county = county14)
```

Check that all states are present in the raw data and count the number of unique counties in each state.
```{r}

check_unique_counties <- function(df_name) {
  df <- get(df_name)
  df %>% 
    select(state, county) %>% 
    unique() %>% 
    mutate(cw_period = df_name)
}

# Validation function
validate_crosswalk <- function(df, period_name) {
  cat("Validation for", period_name, "period:\n")
  cat("- Rows:", nrow(df), "\n")
  cat("- Unique PUMAs:", n_distinct(df$puma), "\n") 
  cat("- Unique Counties:", n_distinct(paste(df$statefip, df$county)), "\n")
  cat("- Missing afact values:", sum(is.na(df$afact)), "\n")
  cat("- Zero afact values:", sum(df$afact == 0, na.rm = TRUE), "\n\n")
}

coverage_summary <- map_dfr(cw_files, check_unique_counties) %>% 
  group_by(state, cw_period) %>% 
  summarise(unique_counties = n(), .groups = "drop") %>%
  pivot_wider(names_from = "cw_period", values_from = unique_counties) %>% 
  select(state, puma_county_2k, puma_county_2010, puma_county_2020)

coverage_summary %>% reactable()

# Coverage visualization
coverage_summary %>%
  pivot_longer(cols = -state, names_to = "cw_period", values_to = "unique_counties") %>%
  ggplot(aes(x = reorder(state, unique_counties), y = unique_counties, fill = cw_period)) +
  geom_col(position = "dodge", alpha = 0.8) +
  coord_flip() +
  scale_fill_viridis_d(name = "PUMA Period") +
  labs(
    title = "County Coverage by State Across PUMA Periods",
    x = "State", 
    y = "Number of Counties"
  ) +
  theme_minimal()
```

## Clean crosswalks

Filter out the row containing metadata and select only needed variables.

```{r}

clean_cw <- function(.x,  .cw_period) {
  .x %>% 
  rename_with(tolower) %>%
  filter(!str_detect(state, "State")) %>%
  select(state, starts_with("puma"), -ends_with("name"), county, starts_with("pop"), afact, afact2) %>% 
  mutate(crosswalk_period = .cw_period)
}

puma_county_2020 <- clean_cw(puma_county_2020, "2022")
puma_county_2010 <- clean_cw(puma_county_2010, "2012")
puma_county_2k <- clean_cw(puma_county_2k, "2k")
```

Rename GeoCorr variables for working purposes. Add leading zeros to state fips, PUMAs and counties to make them uniform in length.

```{r}
create_geo_indicators <- function(.x) {
  .x %>% 
    rename(
    puma = starts_with("puma"),
    statefip = state
    ) %>%
  mutate(
    statefip = sprintf("%0.2d", as.numeric(statefip)),
    puma = sprintf("%0.5d", as.numeric(puma)),
    county = sprintf("%0.5d", as.numeric(county)),
    across(contains("afact"), ~ as.double(.)),
    across(contains("pop"), ~ as.double(.))
  )
}

puma_county_2020 <- create_geo_indicators(puma_county_2020)
puma_county_2010 <- create_geo_indicators(puma_county_2010)
puma_county_2k <- create_geo_indicators(puma_county_2k)
```

Break out the county FIP code from the county variable which includes the state FIPs code as the first two digits.

```{r}
puma_county_2020 <- puma_county_2020 %>% 
  mutate(county = str_sub(county, start = 3L, end = 5L))

puma_county_2010 <- puma_county_2010 %>% 
  mutate(county = str_sub(county, start = 3L, end = 5L))

puma_county_2k <- puma_county_2k %>% 
  mutate(county = str_sub(county, start = 3L, end = 5L))
```


## Apply validation to each dataset

```{r}
validate_crosswalk(puma_county_2020, "2022")
validate_crosswalk(puma_county_2010, "2012")
validate_crosswalk(puma_county_2k, "2k")
```

Make changes to county definitions so they are consistent between the 2010 and 2020 crosswalks. All changes during the 2010 period are documented on this census [website page] (https://www.census.gov/programs-surveys/geography/technical-documentation/county-changes.2010.html#list-tab-957819518). Only need to address changes in 2014 and beyond.

*New Counties:*
- Chugach Census Area, Alaska (02-063): Created from part of former Valdez-Cordova Census Area (02-261) effective January 02, 2019.
- Copper River Census Area, Alaska (02-066): Created from part of former Valdez-Cordova Census Area (02-261) effective January 02, 2019.

For this change, use the ratio of afact from 2012 PUMA to 2020 counties to break the county into two.

*Code changes:*
- Kusilvak Census Area, Alaska (02-158): Changed name and code from Wade Hampton Census Area (02-270) effective July 1, 2015.
- Oglala Lakota County, South Dakota (46-102): Changed name and code from Shannon County (46-113) effective May 1, 2015.

```{r}
alaska_split <- tibble(
  statefip = c("02", "02"),
  puma = c("00300", "00300"),
  county = c("063", "066"),
  pop10 = c(9636*(0.0435/0.0595), 9636*(0.016/0.0595)),
  afact = c(0.0586*(0.0435/0.0595), 0.0586*(0.016/0.0595)),
  afact2 = c(1, 1)
)

puma_county_2010 <- puma_county_2010 %>% 
  mutate(county = ifelse(statefip == "02" & county == "270", "158", county),
         county = ifelse(statefip == "46" & county == "113", "102", county),
         flag = ifelse(statefip == "02" & county == "261", "1", "0")) %>% 
  filter(flag != "1") %>% 
  bind_rows(alaska_split %>% mutate(crosswalk_period = "2012")) %>% 
  select(-flag)

puma_county_2k <- puma_county_2k %>% 
  mutate(county = ifelse(statefip == "02" & county == "270", "158", county),
         county = ifelse(statefip == "46" & county == "113", "102", county),
         flag = ifelse(statefip == "02" & county == "261", "1", "0")) %>% 
  filter(flag != "1") %>% 
  bind_rows(alaska_split %>% mutate(crosswalk_period = "2k")) %>% 
  select(-flag)
```

## Create quality variable

Add the necessary components to the crosswalk for producing a data quality flag.

Create flags in the PUMA-county crosswalk for counties where a high percentage of the data comes from outside of the county. Per agreed upon guidance, 75% or more from the county is good, below 35% is bad, in between is marginal. This is calculated by taking the product of percentage of PUMA in the county and percentage of county in the PUMA for each county-PUMA pairing, and summing across the county.

First remove ZERO afact values and create new variable "products". This is the sum of the share of the county captured in the PUMA (afact2) and the share of the PUMA captured in the county (afact).

```{r}
puma_county20 <- puma_county_2020 %>%
  filter(afact > 0) %>%
  mutate(products = afact * afact2)

puma_county10 <- puma_county_2010 %>%
  filter(afact > 0) %>%
  mutate(products = afact * afact2)

puma_county2k <- puma_county_2k %>%
  filter(afact > 0) %>%
  mutate(products = afact * afact2)
```

Calculate the total county population by summing the GeoCorr pop10 and pop20 variables. These pop variables count the population in the PUMA that overlap with the county - note that GeoCorr presents these numbers so that it already accounts for afact (see What is a Correlation List? on the [GeoCorr website](https://mcdc.missouri.edu/applications/docs/geocorr-help.html#weightVar)).

```{r}
puma_county20 <- puma_county20 %>%
  group_by(statefip, county) %>%
  mutate(sum_products = sum(products),
         county_pop = sum(pop20)) %>% 
  ungroup() 

puma_county10 <- puma_county10 %>%
  group_by(statefip, county) %>%
  mutate(sum_products = sum(products),
         county_pop = sum(pop10)) %>% 
  ungroup() 

puma_county2k <- puma_county2k %>%
  group_by(statefip, county) %>%
  mutate(sum_products = sum(products),
         county_pop = sum(pop10)) %>% 
  ungroup() 
```

The average sum of products in the 2020 crosswalk is 54% and 53% in the 2010 crosswalk (more PUMAs have good quality than less than good).

```{r}

puma_county2k %>% 
  select(sum_products) %>% 
  summary()

puma_county10 %>% 
  select(sum_products) %>% 
  summary()

puma_county20 %>% 
  select(sum_products) %>% 
  summary()
```

Rank sum_product based on the guidance described above (1 to 3, with 1 being best) and create a new variable representing the quality of geographic allocation.

```{r}

geographic_allocation <- function(.x) {
  .x %>% 
    mutate(
    geographic_allocation_quality =
      case_when(
        sum_products >= 0.75 ~ 1,
        sum_products >= 0.35 ~ 2,
        sum_products < 0.35 ~ 3
      )
    )
}

puma_county20 <- geographic_allocation(puma_county20)
puma_county10 <- geographic_allocation(puma_county10)
puma_county2k <- geographic_allocation(puma_county2k)
```

We want to join the crosswalks together so they can be merged onto multi-year census files. 
Make the variable names universal.

```{r}

puma_county20 <- puma_county20 %>% 
 rename(population = pop20)

puma_county10 <- puma_county10 %>% 
  rename(population = pop10) 

puma_county2k <- puma_county2k %>% 
  rename(population = pop10) 

```

## Comparative Data Quality Analysis

Visualize key variables across all three crosswalk periods for comparison.

```{r}
# Create combined quality comparison
quality_comparison <- bind_rows(
  puma_county20 %>% select(crosswalk_period, geographic_allocation_quality) %>% mutate(type = "2020"),
  puma_county10 %>% select(crosswalk_period, geographic_allocation_quality) %>% mutate(type = "2010"),
  puma_county2k %>% select(crosswalk_period, geographic_allocation_quality) %>% mutate(type = "2000")
)

quality_comparison %>%
  ggplot(aes(x = factor(geographic_allocation_quality), fill = type)) +
  geom_bar(position = "dodge", alpha = 0.8) +
  scale_fill_viridis_d(name = "PUMA Period") +
  labs(
    title = "Geographic Allocation Quality Comparison Across PUMA Periods",
    subtitle = "1=High Quality (75%+), 2=Medium Quality (35-75%), 3=Low Quality (<35%)",
    x = "Quality Rating", 
    y = "Count of Counties"
  ) +
  theme_minimal()

# Combined population distribution
pop_combined <- bind_rows(
  puma_county20 %>% select(population, crosswalk_period) %>% mutate(period = "2020"),
  puma_county10 %>% select(population, crosswalk_period) %>% mutate(period = "2010"),
  puma_county2k %>% select(population, crosswalk_period) %>% mutate(period = "2000")
)

pop_combined %>%
  ggplot(aes(x = population, fill = period)) +
  geom_density(alpha = 0.6) +
  scale_x_log10(labels = scales::comma_format()) +
  scale_fill_viridis_d(name = "PUMA Period") +
  labs(
    title = "Population Distribution Across PUMA Periods",
    subtitle = "Log scale - weighted population in PUMA-County overlaps",
    x = "Population (log scale)",
    y = "Density"
  ) +
  theme_minimal()

# AFACT distribution comparison
afact_combined <- bind_rows(
  puma_county20 %>% select(afact, crosswalk_period) %>% mutate(period = "2020"),
  puma_county10 %>% select(afact, crosswalk_period) %>% mutate(period = "2010"),
  puma_county2k %>% select(afact, crosswalk_period) %>% mutate(period = "2000")
)

afact_combined %>%
  ggplot(aes(x = afact, fill = period)) +
  geom_histogram(alpha = 0.7, position = "identity", binwidth = 0.02) +
  scale_fill_viridis_d(name = "PUMA Period") +
  labs(
    title = "AFACT Distribution Comparison Across PUMA Periods",
    x = "AFACT (Allocation Factor)",
    y = "Count"
  ) +
  theme_minimal()
```

## Finalize 

Finally, bind the crosswalks together and sort by cross-walk period, state, and county. Clean up intermediate objects for memory management.

```{r}
# Clean up intermediate objects
rm(puma_county_2020, puma_county_2010, puma_county_2k)

# Combine all crosswalks with optimized memory usage
puma_county <- puma_county20 %>% 
  bind_rows(puma_county10, puma_county2k) %>% 
  arrange(crosswalk_period, statefip, county) %>%
  select(crosswalk_period, statefip, county, puma, afact, afact2, 
         population, products, sum_products, geographic_allocation_quality)
```


## Write files
Save as "puma_county.csv" in the intermediate file.

```{r}
write_csv(puma_county, here::here("geographic-crosswalks", "data", "crosswalk_puma_to_county.csv"))
```
