---
title: ""
author: ""
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: false
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: !expr here::here("04_health", "www", "web_report.css")
    editor_options:
      chunk_output_type: console
---

<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

# Access to Health Services
This metric is a county-level estimate of access to health services, defined as the ratio of population to primary care physicians. We pull this data from the 2022 County Health Rankings, which uses 2019 data from the Area Health Resource File.

See the [County Health Rankings](https://www.countyhealthrankings.org/explore-health-rankings/measures-data-sources/county-health-rankings-model/health-factors/clinical-care/access-to-care/primary-care-physicians) for more information.

**TODO:**

1. Add data quality flags - no confidence interval available, just z-score


**Process:**

1. Download Excel files for county-level data from each state

2. Add data quality flags

3. Save the data

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup}
options(scipen = 999)

library(urbnthemes)
library(tidyverse)
library(readxl)
library(writexl)

library(rvest)
library(glue)
library(fs)

set_urbn_defaults(style = "print")

```

## 1. Prepare for file download
```{r}
base_url <- "https://www.countyhealthrankings.org"
url <- paste0(base_url, "/explore-health-rankings/rankings-data-documentation")
download_urls <-
  read_html(url) %>% 
  html_elements('span.field-content') %>%  
  html_element('a') %>% 
  html_attr('href') %>%
  paste0(base_url, .)

out_state <-
  read_html(url) %>% 
  html_elements('span.field-content') %>% 
  html_text2() %>% 
  str_replace_all(' ', '-')

get_state_data_url <- function(url){
  read_html(url) %>% 
    html_elements('span.download-link') %>% 
    html_element('a') %>% 
    html_attr('href') %>% 
    str_subset('.xls') %>% 
    str_subset('2022') %>% 
    paste0(base_url, .)
}
state_data_url <- purrr::map_chr(.x = download_urls, .f = get_state_data_url)
```


## 2. Download files

Download all 51 Excel files from each state's URL
```{r}
# create a data folder for intermediate data files
intermediate_files <- here::here("04_health",
                                 "access-health-services",
                                 "data/")

if (!dir.exists(intermediate_files)) {
  
  dir.create(intermediate_files)
  
}

chr_data <- here::here("04_health",
                       "access-health-services",
                       "data",
                       out_state
)

download_data <- function(state_data_url, state_name){
  if (!file.exists(here::here(glue("04_health/access-health-services/data/{state_name}.xlsx")))) {
    
    download.file(state_data_url, 
                  destfile = here::here(glue("04_health/access-health-services/data/{state_name}.xlsx")))
    
  }
}

walk2(state_data_url, out_state, download_data)

```


## 3. Read in, clean, and append each Excel sheet
```{r}
physicians <- map_dfr(out_state,
                     ~{readxl::read_excel(path = here::here(
                       paste0("04_health/access-health-services/data/", .x, ".xlsx")),
                       sheet = "Ranked Measure Data",
                       skip = 1
                     ) %>%
                         select(fips = 'FIPS',
                                # number_pc_phys = '# Primary Care Physicians',
                                # pc_phys_rate = 'Primary Care Physicians Rate',
                                pc_phys_ratio = 'Primary Care Physicians Ratio'
                                ) %>%
                         filter(!str_detect(fips, '000$')) %>%
                         mutate(state = str_sub(fips, 1, 2),
                                county = str_sub(fips, 3, 5)
                                # pop_per_1_pcp = gsub(':1', '', pc_phys_ratio),
                                # pop_per_1_pcp = gsub(':0', '', 0)
                                )
                     }
)

physicians[c('antecedent', 'consequent')] <- str_split_fixed(physicians$pc_phys_ratio, ':', 2)

physicians <- physicians %>%
  mutate(pc_phys_ratio = case_when(consequent == "1" ~ pc_phys_ratio,
                                   consequent == "0" ~ "0",
                                   is.na(pc_phys_ratio) ~ "0"),
         year = 2020
         ) %>%
  select(year, state, county, pc_phys_ratio)

physicians <- physicians %>%
  arrange("year", "state", "county")


```


## 4. Add data quality flags
```{r}


```


## 5. Save the data
```{r}
# Is this join necessary?

# all_counties <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv")) %>%
#   tidylog::filter(year == 2020) %>%
#   select(-year)
# 
# all_counties <- left_join(all_counties, physicians, by = c("state", "county"))
# 
# test <- anti_join(all_counties, physicians, by = c("state", "county"))
# 
# all_counties <- all_counties %>%
#   mutate(year = 2020) %>%
#   select(year, state, county, pop_per_1_pcp)

physicians %>%
  write_csv(here::here("04_health",
                       "access-health-services",
                       "access-health-services-2020.csv"))
```

