---
title: "Access to Health Services"
author: "Vincent Pancini"
editor_options: 
  chunk_output_type: console
---

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup}
# library(urbnthemes)
library(tidyverse)
library(haven)
library(rvest)
library(glue)
library(fs)
library(tidycensus)

# set_urbn_defaults(style = "print")

options(scipen = 999)

```

For the 2022 Mobility Metrics, we used data from [County Health Rankings](https://www.countyhealthrankings.org/explore-health-rankings/county-health-rankings-model/health-factors/clinical-care/access-to-care/primary-care-physicians?year=2022). 

For the 2023 Mobility Metrics, we want to use the underlying source data used by CHR, which is the [Area Health Resource File, 2021-2022 County Level Data](https://data.hrsa.gov/data/download?data=AHRF#AHRF). The Area Health Resource File is a collection of data from more than 50 sources, including the American Medical Association, which maintains the Physician Masterfile. The Physician Masterfile contains information on nearly all the Doctors of Medicine and Doctors of Osteopathic Medicine in the nation.

The right side of the ratio is the number of primary care physicians in a county. Primary care physicians include practicing non-federal physicians (M.D.s and D.O.s) under age 75 specializing in general practice medicine, family medicine, internal medicine, and pediatrics. It does not include nurse practitioners, physician assistants, or other practitioners available for primary care services.


## 1. Download data
The data for this metric come from the [Area Health Resource File 2021-2022 County-Level Data](https://data.hrsa.gov/data/download). 
```{r}
# TODO: figure out how to download the data directly from the website in R
ahrf2022 <- haven::read_sas(here::here("04_health", "access-health-services", "ahrf2022.sas7bdat"))

ahrf2022 <- ahrf2022 %>%
  select(
    f00002, # state/county fips
    f00008, # state name
    f12424, # state abbreviation
    f00010, # county name
    f00011, # state fips code
    f00012, # county fips code
    f1198421, # population estimate, 2021
    f0453020, # census population, 2020
    f1198420, # population estimate, 2020
    f1467520, # phys, primary care, patient care - non-fed, excludes hospital residents & 75+
    
    f1390620, # pop total male, 2020, whole numbers, estimates, census county char file 
    f0482020, # Total MDs, male, non-fed, includes inactive
    f1470020, # Total DOs, male, non-fed, includes inactive
    f0490020, # MDs, inactive, male (non-fed)
    f1470220, # DOs, inactive, male (non-fed)
    
    f1390720, # pop total male, 2020, whole numbers, estimates, census county char file
    f0482120, # Total MDs, female, non-fed, includes inactive
    f1470120, # Total DOs, female, non-fed, includes inactive
    f0490120, # MDs, inactive, female (non-fed)
    f1470320, # DOs, inactive, female (non-fed)
  ) %>%
  filter(
    !f00011 %in% c('66', '72', '78')
  )

ahrf2022 <- ahrf2022 %>%
  mutate(
    male_pcp = (f0482020 + f1470020 - f0490020 - f1470220),
    fmle_pcp = (f0482120 + f1470120 - f0490120 - f1470320)
  ) %>%
  select(
    county_fips = f00002,
    state_fips = f00011,
    tot_pop_2020 = f0453020,
    tot_pcp = f1467520,
    male_pop_2020 = f1390620,
    fmle_pop_2020 = f1390720,
    male_pcp,
    fmle_pcp
  )


# Even after filtering out Guam, Puerto Rico, and the Virgin Islands, there seem to be 7 extra counties. I suspect these are weird Alaska counties
count(ahrf2022, state_fips) %>% view()

```


```{r}
# TODO: remove SAS labels from variables
ahrf2022 <- ahrf2022 %>%
  mutate(
    pc_phys_ratio = case_when(
      pc_physicians > 0 ~ round(population/pc_physicians, 1),
      pc_physicians == 0 ~ 0),
    low_pop_flag = case_when(
      population < 2000 ~ "1",
      population >= 2000 ~ "0"
    )
  )

```


I am checking our numbers against the County Health Rankings [data](https://www.countyhealthrankings.org/explore-health-rankings/rankings-data-documentation). These data define primary care physicians as practicing non-federal physicians (M.D.s and D.O.s) under age 75 specializing in general practice medicine, family medicine, internal medicine, and pediatrics [source](https://www.countyhealthrankings.org/explore-health-rankings/county-health-rankings-model/health-factors/clinical-care/access-to-care/primary-care-physicians?year=2023).
```{r}
chr2023_counties <- readxl::read_excel(
  path = here::here("04_health", "access-health-services", "2023 County Health Rankings Data - v2.xlsx"),
  range = "Ranked Measure Data!A2:C3195"
) %>%
  mutate(
    ID = row_number()
  )

chr2023_data <- readxl::read_excel(
  path = here::here("04_health", "access-health-services", "2023 County Health Rankings Data - v2.xlsx"),
  range = "Ranked Measure Data!DK2:DN3195"
) %>%
  mutate(
    ID = row_number()
  )

chr2023 <- left_join(
  x = chr2023_counties, y = chr2023_data, by = "ID"
) %>%
  filter(!is.na(County)) %>%
  select(
    county_fips = "FIPS",
    pc_physicians_chr = "# Primary Care Physicians",
    pc_physicians_rate_chr = "Primary Care Physicians Rate",
    pc_physicians_ratio_chr = "Primary Care Physicians Ratio",
    pc_physicians_quartile_chr = "Quartile"
  )

rm(chr2023_counties, chr2023_data)  

chr2023 <- chr2023 %>%
  separate(col = pc_physicians_ratio_chr,
           into = c('antecedent', 'consequent'),
           remove = FALSE) %>%
  mutate(pc_physicians_ratio_chr = case_when(consequent == "1" ~ pc_physicians_ratio_chr,
                                             consequent == "0" ~ "0",
                                             is.na(pc_physicians_ratio_chr) ~ "0"),
         antecedent = as.numeric(antecedent)
  )

joined_data <- joined_data %>%
  full_join(
    x = ahrf2022, y = chr2023, by = "county_fips"
  ) %>%
  mutate(diff = pc_phys_ratio-antecedent)

```



## 1. Prepare for file download
```{r}
base_url <- "https://www.countyhealthrankings.org"

url <- paste0(base_url, "/explore-health-rankings/rankings-data-documentation")

download_urls <-
  read_html(url) %>% 
  html_elements('span.field-content') %>%  
  html_element('a') %>% 
  html_attr('href') %>%
  paste0(base_url, .)

out_state <-
  read_html(url) %>% 
  html_elements('span.field-content') %>% 
  html_text2() %>% 
  str_replace_all(' ', '-')

get_state_data_url <- function(url){
  read_html(url) %>% 
    html_elements('span.download-link') %>% 
    html_element('a') %>% 
    html_attr('href') %>% 
    str_subset('.xls') %>% 
    str_subset('2022') %>% 
    paste0(base_url, .)
}

state_data_url <- purrr::map_chr(.x = download_urls, .f = get_state_data_url)

```


## 2. Download files

Download all 51 Excel files from each state's URL
```{r}
# # create a data folder for intermediate data files
# intermediate_files <- here::here("04_health",
#                                  "access-health-services",
#                                  "data/")
# 
# if (!dir.exists(intermediate_files)) {
#   
#   dir.create(intermediate_files)
#   
# }
# 
# chr_data <- here::here("04_health",
#                        "access-health-services",
#                        "data",
#                        out_state
# )
# 
# download_data <- function(state_data_url, state_name){
#   if (!file.exists(here::here(glue("04_health/access-health-services/data/{state_name}.xlsx")))) {
#     
#     download.file(state_data_url, 
#                   destfile = here::here(glue("04_health/access-health-services/data/{state_name}.xlsx")))
#     
#   }
# }
# 
# walk2(state_data_url, out_state, download_data)

```


## 3. Read in, clean, and append each Excel sheet
```{r, message = FALSE}
physicians <- map_dfr(out_state,
                     ~{readxl::read_excel(path = here::here(
                       paste0('04_health/access-health-services/data/', .x, '.xlsx')),
                       sheet = 'Ranked Measure Data',
                       skip = 1
                     ) %>%
                         select(fips = 'FIPS',
                                pc_phys_ratio = 'Primary Care Physicians Ratio',
                                number_pc_phys = '# Primary Care Physicians',
                                pc_phys_rate = 'Primary Care Physicians Rate'
                                ) %>%
                         filter(!str_detect(fips, '000$')) %>%
                         mutate(state = str_sub(fips, 1, 2),
                                county = str_sub(fips, 3, 5)
                                )
                     }
)

```

```{r}
physicians <- physicians %>%
  separate(col = pc_phys_ratio,
           into = c('antecedent', 'consequent'),
           remove = FALSE) %>%
  mutate(pc_phys_ratio = case_when(consequent == "1" ~ pc_phys_ratio,
                                   consequent == "0" ~ "0",
                                   is.na(pc_phys_ratio) ~ "0"),
         year = 2019
         )

```


## 4. Add data quality flags

First we need to get population estimates from the US Census Bureau Population Estimation Program. # Replace YOUR-KEY-HERE below with your unique Census API key. Request a key at https://api.census.gov/data/key_signup.html if you don't have one.
```{r}
# census_api_key("YOUR-KEY-HERE")

pop <- tidycensus::get_estimates("county", year = 2019, variables = "POP") %>%
  rename(fips = GEOID)

pop <- pop %>%
  # Drop PR observations
  filter(!str_detect(fips, '^72')) %>%
  select(fips,
         population = value)

```

Join the population data to the physicians data.
```{r}
joined_data <- left_join(physicians, pop, by = "fips")

joined_data <- joined_data %>%
  mutate(
    # Convert the antecedent of the primary care physicians:population ratio to numeric so we can use it for calculations
    antecedent = as.numeric(antecedent),
    # Divide the 2019 PEP population by the 2019 CHR Number of Physicians. Compare this value to the antecedent of the ratio.
    check_antecedent = population / number_pc_phys,
    diff_pop = abs(antecedent - check_antecedent)
  )
```

Any differences in the ratio are likely due to rounding. This means the population variable we pulled from PEP is a good proxy for the population variable used by County Health Rankings to construct this ratio.
```{r}
test <- joined_data %>%
  filter(number_pc_phys != 0)
  
quantile(test$diff_pop, probs = seq(0, 1, 0.25), na.rm = TRUE)

```

```{r}
joined_data <- joined_data %>%
  mutate(
    # Add 1 to the denominator of the ratio and reconstruct the ratio
    plus1 = population / (number_pc_phys + 1),
    # Calculate the differences between each of the new ratios and the original ratio
    plus1_diff = abs(plus1 - as.numeric(antecedent)),
    sensitivity = abs((plus1 - antecedent) / antecedent)
  )

```

```{r}
test <- joined_data %>%
  filter(number_pc_phys != 0)
  
quantile(test$sensitivity, probs = c(0, 0.2, 0.4, 0.6, 0.75, 0.8, 0.9, 1), na.rm = TRUE)

```


```{r}
joined_data %>%
  filter(!is.na(antecedent)) %>%
  ggplot(aes(number_pc_phys, sensitivity)) +
  geom_point(alpha = 0.2) +
  labs(title = "Counties with the lowest number of PC physicians have the highest sensitivity",
       subtitle = "") +
  scatter_grid()

```

To measure the data quality, we've added 1 to the number of primary care physicians in each county and re-created the ratio of the number of population per primary care physician. If the ratio changes less than 20 percentage points we assign a quality value of 1. If the ratio changes by more than 20 percentage points we assign a quality value of 2.

```{r}
# Assign a data quality flag
joined_data <- joined_data %>%
  mutate(pc_phys_quality = 
           case_when(
             sensitivity < 0.2 ~ 1,
             sensitivity < 0.4 ~ 2,
             sensitivity < 0.6 ~ 3
           )
  )

joined_data %>%
  count(pc_phys_quality)

```

## 5. Save the data

```{r}
all_counties <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv")) %>%
  tidylog::filter(year == 2019) %>%
  select(-year)

joined_data <- joined_data %>%
  select(-population) %>%
  left_join(all_counties, by = c("state", "county"))

test <- anti_join(all_counties, physicians, by = c("state", "county"))

joined_data <- joined_data %>%
  mutate(year = 2019) %>%
  select(year, state, county, pc_phys_ratio, pc_phys_quality) %>%
  arrange(year, state, county)

joined_data %>%
  write_csv(here::here("04_health",
                       "access-health-services",
                       "access-health-services-2019.csv"))

```


