---
title: "ELA LEARNING GROWTH: average annual learning growth between 3rd and 8th grade"
author: "Written originally by E Gutierrez and translated to R by JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
params:
  year: "2026"
---

*Program for the creation of the english language arts growth metric at the city level*

This file creates city/places level learning rate estimates for years (fall) 2013 - 2017 and subgroups by economic disadvantage, race, and gender. However, for the purposes of creating the community dashboards we focus on the 2015 race and economic disadvantage subgroups only. 

Uses MEPS: the share of students attending schools where 20% or more students are in Poverty  **
 - Uses 2014-2024 data
 - Produces city level data


## Housekeeping

Read in packages for use. Set the global year to the latest education data.

```{r, results='hide'}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)
library(srvyr)
library(scales)
library(educationdata)
library(haven)
library(broom)        # For regression output processing
library(future)       # For parallel processing
library(future.apply) # For parallel apply functions
library(glue)

options(scipen = 999)
options(timeout=60*10)

# Set up parallel processing for computational efficiency
plan(multisession, workers = parallel::detectCores() - 1)


# Configuration and file paths
config <- list(
  # Update these paths for your local environment
  git_folder = "C:/Users/jwalsh/Desktop/mobility-from-poverty",
  current_year = 2020,    # Spring of school year (2019-20)  
  latest_year = 2019,     # Fall of current SEDA year (2019-20)
  
  # Quality thresholds for learning rate estimates
  quality_thresholds = list(
    high_quality = list(min_grades = 5, min_sample = 30),      # Quality = 1
    medium_quality = list(min_grades = 4, min_sample = 30),    # Quality = 2
    low_quality = list(min_grades = 3, min_sample = 10)        # Quality = 3
  )
)


# Demographic subgroups to process
subgroups <- c("all", "wht", "blk", "hsp", "nec", "ecd", "mal", "fem")
```

Source the function `evaluate_final_data.R`. 

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```

## Import and save raw data

Read in the county crosswalk file 

```{r}
#| label: read-place-file
county <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv"))
```


Clean city names 

```{r}
#| label: clean-county-file
county_clean <- county %>% 
  mutate(county = str_pad(county, width = 3, side = "left", pad = "0"),
         state = str_pad(state, width = 2, side = "left", pad= "0"))
```

### Download SEDA

SEDA data are downloaded from Stanford's SEDA website: https://edopportunity.org/get-the-data/seda-archive-downloads/ 
	** SEDA data standardize EDFacts assessments data across states and years using NAEP data **


```{r}
#| label: download-seda

if(!file.exists(here::here("07_effective-public-education", "Data", "Raw", "seda_county.csv"))) {
  
download.file("https://stacks.stanford.edu/file/xh833nn4025/seda_county_long_gcs_6.0.csv", dest = here::here("07_effective-public-education", "Data", "Raw", "seda_county.csv"))
  
}

seda <- data.table::fread(here::here("07_effective-public-education", "Data", "Raw", "seda_county.csv"))
```


Filter for the relevant subject "rla" and define cohorts. Cohort are defined by the year a cohort reaches 8th grade. Eg, the 2016 cohort is the cohort that is in 8th grade in 2016, in 7th grade in 2015, in 6th grade in 2014, etc

```{r}
#| label: clean-seda
seda <- seda %>% 
  filter(subject == "rla") %>% 
  mutate(cohort = year - grade + 8) %>% 
  filter(cohort >= 2014, !is.na(cohort)) 

```


Combine Asian and native american students to "other"

```{r}
#| label: create-other-cat
seda <- seda %>% 
  mutate(tot_asmt_oth = tot_asmt_asn + tot_asmt_nam,
         temp_asn = gcs_mn_asn*tot_asmt_asn,
         temp_nam = gcs_mn_nam*tot_asmt_nam,
         gcs_mn_oth = temp_asn + temp_nam,
         gcs_mn_oth = gcs_mn_oth/tot_asmt_oth)
```

## Clean and Calculate Growth Estimates for each subgroup

*Outline the process for creating the data: SEDA data are manually downloaded and read in, 
*and a regression of mean assessment scores was run on grade (as a continuous variable) 
*interacted with each county in order to obtain county-specific grade slopes. 
*Regressions are weighted by the number of test-takers for each year, grade, and county. 
*95% confidence intervals are calculated as the slope estimate plus or minus 1.96 times the standard error of the estimate. 
*A flag indicates how many grades are included in each estimate.


```{r learning-rates}
#| label: learning-rates

calculate_learning_rates <- function(education_data, subgroups) {
  
  cat("Calculating learning rates by subgroup...\n")
  
  # Function to calculate learning rate for a single subgroup-cohort combination
  calc_subgroup_rates <- function(data, subgroup, cohort_year) {
    
    # Column names for this subgroup
    score_col <- glue("gcs_mn_{subgroup}")
    n_col <- glue("tot_asmt_{subgroup}")
    
    # Prepare data for regression
    reg_data <- data %>%
      filter(cohort == cohort_year) %>%
      select(sedacounty, grade, all_of(score_col), all_of(n_col)) %>%
      rename(score = !!score_col, n_students = !!n_col) %>%
      filter(!is.na(score), n_students > 0)
    
    if (nrow(reg_data) == 0) return(NULL)
    
    # Run weighted regression: score ~ grade × district 
    # This gives district-specific slopes (learning rates)
    model <- lm(score ~ grade:factor(sedacounty) + factor(sedacounty), 
                data = reg_data, weights = n_students)
    
    # Extract coefficients and standard errors
    coef_summary <- tidy(model) %>%
      filter(str_detect(term, "grade:factor\\(sedacounty\\)")) %>%
      mutate(
        sedacounty = as.numeric(str_extract(term, "(?<=sedacounty\\))\\d+"))
      ) %>%
      select(sedacounty, estimate, std.error) %>%
      rename(learning_rate = estimate, se = std.error)
    
    # Calculate quality metrics for each district
    quality_metrics <- reg_data %>%
      group_by(sedacounty) %>%
      summarise(
        num_grades_included = n(),
        min_sample_size = min(n_students, na.rm = TRUE),
        .groups = "drop"
      )
    
    # Combine results
    results <- coef_summary %>%
      left_join(quality_metrics, by = "sedacounty") %>%
      mutate(
        cohort = cohort_year,
        subgroup = subgroup,
        
        # Calculate confidence intervals
        learning_rate_lb = learning_rate - 1.96 * se,
        learning_rate_ub = learning_rate + 1.96 * se,
        
        # Assign quality flags based on sample size and grade coverage
        learning_rate_quality = case_when(
          (num_grades_included >= 5 & min_sample_size >= 30) ~ 1L,  # High quality
          (num_grades_included >= 4 & min_sample_size >= 30) ~ 2L,  # Medium quality  
          (!is.na(learning_rate)) ~ 3L,                             # Low quality
          TRUE ~ NA_integer_
        ),
        
        # Remove false zeros (artifact of regression procedure)
        learning_rate = ifelse(learning_rate == 0 & learning_rate_lb == 0 & learning_rate_ub == 0, 
                               NA_real_, learning_rate),
        learning_rate_lb = ifelse(is.na(learning_rate), NA_real_, learning_rate_lb),
        learning_rate_ub = ifelse(is.na(learning_rate), NA_real_, learning_rate_ub),
        learning_rate_quality = ifelse(is.na(learning_rate), NA_integer_, learning_rate_quality)
      ) %>%
      
      select(cohort, sedacounty, subgroup, learning_rate, learning_rate_lb, 
             learning_rate_ub, learning_rate_quality)
    
    return(results)
  }
  
  # Calculate learning rates for all subgroups and cohorts
  # Use parallel processing for efficiency
  all_combinations <- expand_grid(
    subgroup = subgroups,
    cohort = unique(education_data$cohort)
  )
  
  #Increase future max limit 
  
  oopts <- options(future.globals.maxSize = 1.0 * 1e9)  ## 1.0 GB
  on.exit(options(oopts))
  
  learning_rates <- furrr::future_map2_dfr(
    all_combinations$subgroup, 
    all_combinations$cohort,
    ~ calc_subgroup_rates(education_data, .x, .y),
    .progress = TRUE
  )
  
cat(glue("✓ Calculated learning rates for {format(nrow(learning_rates), big.mark = ",")} county-subgroup-cohort combinations\n"))

  return(learning_rates)
}

# Calculate learning rates
county_learning_rates <- calculate_learning_rates(seda, subgroups)

```

## City-Level Aggregation

Aggregate district-level learning rates to city level using enrollment weighting.

```{r city-aggregation}
#| label: city-aggregation

county_learning_rates_clean <- county_learning_rates %>% 
  rename(year = cohort) %>% 
  mutate(year = year - 1,
         county = str_pad(sedacounty, width = 5, side = "left", pad = "0"),
         state = str_sub(county, 1, 2),
         county = str_sub(county, 3, 5))
```

Join on crosswalk 

```{r}
county_learning_rates_merge <- county_clean %>% 
  left_join(county_learning_rates_clean, by = c("year", "state", "county")) %>% 
  filter(year >= 2014, year <= config$current_year-1)
```

## Data Quality Assessment

Evaluate the quality and coverage of the learning rate estimates.

```{r quality-assessment}
#| label: quality-assessment

assess_data_quality <- function(county_data) {
  
  cat("=== DATA QUALITY ASSESSMENT ===\n\n")
  
  # Overall summary statistics by quality level
  cat("Learning Rate Distributions by Quality Level:\n")
  quality_summary <- county_data %>%
    filter(subgroup == "all") %>%
    group_by(learning_rate_quality) %>%
    summarise(
      n_obs = n(),
      mean_rate = mean(learning_rate, na.rm = TRUE),
      median_rate = median(learning_rate, na.rm = TRUE),
      sd_rate = sd(learning_rate, na.rm = TRUE),
      .groups = "drop"
    )
  
  print(quality_summary)
  cat("\n")
  
  # Coverage by year and subgroup
  cat("Data Coverage by Year and Subgroup:\n")
  coverage <- county_data %>%
    group_by(year, subgroup) %>%
    summarise(
      total_counties = n(),
      counties_with_data = sum(!is.na(learning_rate)),
      pct_coverage = round(100 * counties_with_data / total_counties, 1),
      .groups = "drop"
    ) %>%
    pivot_wider(names_from = subgroup, values_from = c(counties_with_data, pct_coverage))
  
  print(coverage)
  cat("\n")
  
  # Check for logical consistency in confidence intervals  
  cat("Confidence Interval Consistency Checks:\n")
  ci_issues <- county_data %>%
    filter(!is.na(learning_rate)) %>%
    summarise(
      below_lower_bound = sum(learning_rate < learning_rate_lb, na.rm = TRUE),
      above_upper_bound = sum(learning_rate > learning_rate_ub, na.rm = TRUE)
    )
  
  cat(glue("• Estimates below lower bound: {ci_issues$below_lower_bound}\n"))
  cat(glue("• Estimates above upper bound: {ci_issues$above_upper_bound}\n\n"))
  
  # Quality flag completeness
  cat("Quality Flag Assessment:\n")
  quality_flags <- county_data %>%
    summarise(
      total_observations = n(),
      missing_quality_flags = sum(is.na(learning_rate_quality)),
      quality_flag_when_missing_rate = sum(is.na(learning_rate) & !is.na(learning_rate_quality))
    )
  
  cat(glue("• Missing quality flags: {quality_flags$missing_quality_flags}\n"))
  cat(glue("• Quality flags when rate is missing: {quality_flags$quality_flag_when_missing_rate}\n\n"))
  
  return(invisible(county_data))
}

# Assess data quality  
final_data <- assess_data_quality(county_learning_rates_merge)
```

## Data Visualization

Create diagnostic visualizations to understand the distribution and quality of learning rate estimates.

```{r visualizations}
#| label: visualizations

create_diagnostic_plots <- function(county_data) {
  
  # Distribution of learning rates by quality level for all students
  p1 <- county_data %>%
    filter(subgroup == "all", !is.na(learning_rate)) %>%
    ggplot(aes(x = learning_rate)) +
    geom_histogram(bins = 30, alpha = 0.7, fill = "steelblue") +
    facet_grid(learning_rate_quality ~ year, labeller = label_both) +
    labs(
      title = "Distribution of ELA Learning Rates by Quality Level and Year",
      subtitle = "All Students",
      x = "Learning Rate (Grade Equivalents per Year)",
      y = "Number of Cities"
    ) +
    theme_minimal()
  
  # Learning rates by demographic subgroup
  p2 <- county_data %>%
    filter(!is.na(learning_rate), year == 2019) %>%
    ggplot(aes(x = learning_rate, fill = subgroup)) +
    geom_density(alpha = 0.6) +
    facet_wrap(~ subgroup, scales = "free_y") +
    labs(
      title = "Learning Rate Distributions by Demographic Subgroup",
      subtitle = glue("Most Recent Year: {max(county_data$year, na.rm = TRUE)}"),
      x = "Learning Rate (Grade Equivalents per Year)",
      y = "Density"
    ) +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Quality distribution over time
  p3 <- county_data %>%
    filter(subgroup == "all", year != 2021) %>%
    count(year, learning_rate_quality) %>%
    ggplot(aes(x = factor(year), y = n, fill = factor(learning_rate_quality))) +
    geom_col(position = "stack") +
    labs(
      title = "Data Quality Distribution Over Time",
      subtitle = "All Students",
      x = "Year",
      y = "Number of Cities",
      fill = "Quality Level"
    ) +
    scale_fill_manual(
      values = c("1" = "#2E8B57", "2" = "#FFD700", "3" = "#CD5C5C"),
      labels = c("1" = "High", "2" = "Medium", "3" = "Low"),
      na.value = "grey80"
    ) +
    theme_minimal()
  
  # Return plots as a list
  return(list(
    quality_by_year = p1,
    subgroup_distributions = p2, 
    quality_over_time = p3
  ))
}

# Generate diagnostic plots
plots <- create_diagnostic_plots(final_data)

# Display plots (uncomment to show)
plots$quality_by_year
plots$subgroup_distributions  
plots$quality_over_time

```

## Data Export

Format the data long with subgroup titles

```{r data-export}
#| label: data-export

long_format <- final_data %>%
    
    # Create proper subgroup labels and categories
    mutate(
      subgroup_type = case_when(
        subgroup == "all" ~ "all",
        subgroup %in% c("wht", "blk", "hsp") ~ "race-ethnicity", 
        subgroup %in% c("mal", "fem") ~ "gender",
        subgroup %in% c("ecd", "nec") ~ "income",
        TRUE ~ "other"
      ),
      
      subgroup = case_when(
        subgroup == "all" ~ "All",
        subgroup == "wht" ~ "White, Non-Hispanic",
        subgroup == "blk" ~ "Black, Non-Hispanic", 
        subgroup == "hsp" ~ "Hispanic",
        subgroup == "mal" ~ "Male",
        subgroup == "fem" ~ "Female",
        subgroup == "ecd" ~ "Economically Disadvantaged",
        subgroup == "nec" ~ "Not Economically Disadvantaged",
        TRUE ~ subgroup
      )
    ) %>%
    
    # Final column ordering
    select(year, state, county, subgroup_type, subgroup, 
           learning_rate, learning_rate_lb, learning_rate_ub, learning_rate_quality) %>%
    
    arrange(desc(year), state, county, subgroup_type, subgroup)

  
```

Write out all file
```{r}

all <- long_format %>% 
  filter(subgroup_type %in% c("all")) %>% 
  select(-subgroup_type, -subgroup)

data.table::fwrite(all, here::here("07_effective-public-education", "Data", "Final", "seda_all_county.csv"))
                   
```


Write out subgroup files
```{r}

race_eth <- long_format %>% 
  filter(subgroup_type %in% c("all", "race-ethnicity"))

data.table::fwrite(race_eth, here::here("07_effective-public-education", "Data", "Final", "seda_race_ethnicity_all_county.csv"))

gender <- long_format %>% 
  filter(subgroup_type %in% c("all", "gender"))

data.table::fwrite(gender, here::here("07_effective-public-education", "Data", "Final", "seda_gender_all_county.csv"))

income <- long_format %>% 
  filter(subgroup_type %in% c("all", "income"))

data.table::fwrite(income, here::here("07_effective-public-education", "Data", "Final", "seda_income_all_county.csv"))
                   
```

