---
title: "English Language Arts Learning Growth Analysis"
subtitle: "City-Level ELA Growth Rates by Demographic Subgroups from SEDA Data"
author: "Translated from STATA to R with Enhanced Documentation"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    code-overflow: wrap
    embed-resources: true
execute:
  warning: false
  message: false
  eval: false
---

## Overview

This analysis calculates **English Language Arts (ELA) learning growth rates** for students between grades 3-8, aggregated by city/place and demographic subgroups. The analysis uses data from the Stanford Education Data Archive (SEDA) which standardizes state assessment data using NAEP benchmarks.

### Key Outputs

- **Learning Rate**: Average annual learning growth between 3rd and 8th grade
- **Confidence Intervals**: 95% confidence bounds around growth estimates  
- **Quality Flags**: Data quality indicators based on sample size and grade coverage
- **Demographic Breakdowns**: Results by race/ethnicity, gender, and economic status

### Methodology

Learning rates are calculated as **district-specific grade coefficients** from weighted regressions of standardized test scores on grade level. This approach captures how much students typically learn per grade within each school district, which are then aggregated to city level.

### Data Coverage

- **Years**: 2014-2019 (fall of academic year)
- **Geographic Level**: Cities/places with sufficient enrollment
- **Subgroups**: All students, Race/ethnicity, Gender, Economic disadvantage status

```{r setup}
#| label: setup
#| include: false

# Core packages for data manipulation and analysis
library(tidyverse)
library(data.table)
library(haven)        # For reading STATA files
library(broom)        # For regression output processing
library(future)       # For parallel processing
library(future.apply) # For parallel apply functions
library(glue)         # For string interpolation

# Set up parallel processing for computational efficiency
plan(multisession, workers = parallel::detectCores() - 1)

# Configuration and file paths
config <- list(
  # Update these paths for your local environment
  git_folder = "C:/Users/jwalsh/Desktop/mobility-from-poverty",
  current_year = 2019,    # Spring of school year (2018-19)  
  latest_year = 2018,     # Fall of current SEDA year (2018-19)
  
  # Quality thresholds for learning rate estimates
  quality_thresholds = list(
    high_quality = list(min_grades = 5, min_sample = 30),      # Quality = 1
    medium_quality = list(min_grades = 4, min_sample = 30),    # Quality = 2
    low_quality = list(min_grades = 3, min_sample = 10)        # Quality = 3
  )
)

# Construct file paths
paths <- list(
  city_file = file.path(config$git_folder, "geographic-crosswalks/data/place-populations.csv"),
  data_dir = file.path(config$git_folder, "08_education/data"),
  raw_dir = file.path(config$git_folder, "08_education/data/raw"),
  intermediate_dir = file.path(config$git_folder, "08_education/data/intermediate"), 
  built_dir = file.path(config$git_folder, "08_education/data/built")
)

# Create directories if they don't exist
walk(paths[str_detect(names(paths), "dir")], ~ dir.create(.x, recursive = TRUE))

# Set working directory
setwd(paths$data_dir)

# Demographic subgroups to process
subgroups <- c("all", "wht", "blk", "hsp", "nec", "ecd", "mal", "fem")
```

## Data Import and Preparation

### City Crosswalk File Processing

The city crosswalk file maps places to standardized geographic identifiers and handles city name variations that occur when matching to education data.

```{r city-crosswalk}
#| label: city-crosswalk

# Import and clean city crosswalk file
prepare_city_crosswalk <- function(city_file_path) {
  
  cat("Preparing city crosswalk data...\n")
  
  city_data <- fread(city_file_path, select = c("place", "state", "place_name")) %>%
    
    # Standardize place and state codes to proper length with leading zeros
    mutate(
      place = str_pad(as.character(place), width = 5, pad = "0"),
      state = str_pad(as.character(state), width = 2, pad = "0"),
      city_name = place_name
    ) %>%
    
    # Clean city names to match education data naming conventions 
    mutate(
      city_name = str_remove_all(city_name, " town| village| municipality| urban county| city")
    ) %>%
    
    # Apply hardcoded fixes for known naming mismatches
    mutate(
      city_name = case_when(
        city_name == "San Buenaventura (Ventura)" ~ "Ventura",
        city_name == "Athens-Clarke County unified government (balance)" ~ "Athens",
        city_name == "Augusta-Richmond County consolidated government (balance)" ~ "Augusta", 
        city_name == "Macon-Bibb County" ~ "Macon",
        city_name == "Boise City" ~ "Boise",
        city_name == "Lexington-Fayette" ~ "Lexington",
        city_name == "Louisville/Jefferson County metro government (balance)" ~ "Louisville",
        city_name == "Lee's Summit" ~ "Lees Summit",
        city_name == "Nashville-Davidson metropolitan government (balance)" ~ "Nashville",
        city_name == "O'Fallon" ~ "Ofallon",
        city_name == "McAllen" ~ "Mcallen",
        city_name == "McKinney" ~ "Mckinney",
        TRUE ~ city_name
      )
    ) %>%
    
    select(state, place, city_name)
  
  # Save intermediate file
  fwrite(city_data, file.path(paths$intermediate_dir, "cityfile.csv"))
  
  cat(glue("✓ Processed {nrow(city_data):,} cities/places\n"))
  return(city_data)
}

city_crosswalk <- prepare_city_crosswalk(paths$city_file)
```

### Education District Location Data

Download district-level data from the Urban Institute's Education Data Portal to map school districts to cities.

```{r district-data}
#| label: district-data
#| eval: false

# Note: This requires the educationdata package
# Install with: devtools::install_github("UrbanInstitute/education-data-package-r")

download_district_data <- function(start_year = 2008, end_year = config$latest_year) {
  
  cat("Downloading CCD district data...\n")
  
  # Download district directory data with city locations
  district_data <- educationdata::get_education_data(
    level = "district-directory",
    source = "ccd", 
    topic = "directory",
    filters = list(year = start_year:end_year),
    select = c("year", "leaid", "city_location", "fips")
  ) %>%
    
    # Create state FIPS code from LEAID  
    mutate(
      state = str_sub(as.character(leaid), 1, 2)
    )
  
  # Save raw data
  fwrite(district_data, file.path(paths$intermediate_dir, 
                                  glue("ccd_lea_city_{start_year}-{end_year}.csv")))
  
  cat(glue("✓ Downloaded {nrow(district_data):,} district-year observations\n"))
  return(district_data)
}

# Uncomment to download (requires educationdata package)
# district_data <- download_district_data()
```

### SEDA Assessment Data

Download standardized achievement data from the Stanford Education Data Archive.

```{r seda-data}
#| label: seda-data

download_and_process_seda <- function() {
  
  cat("Processing SEDA achievement data...\n")
  
  seda_file <- file.path(paths$raw_dir, "seda_geodist_long_gcs_5.0.dta")
  
  # Download SEDA data if not already available
  if (!file.exists(seda_file)) {
    cat("Downloading SEDA data (this may take several minutes)...\n")
    download.file(
      url = "https://stacks.stanford.edu/file/druid:cs829jn7849/seda_geodist_long_gcs_5.0_updated_20240319.dta",
      destfile = seda_file,
      mode = "wb"
    )
  }
  
  # Read and process SEDA data with optimized approach
  seda_data <- read_dta(seda_file) %>%
    
    # Filter to Reading/Language Arts only
    filter(subject == "rla") %>%
    
    # Create cohort variable (year cohort reaches 8th grade) 
    mutate(cohort = year - grade + 8) %>%
    
    # Filter to relevant cohorts and adjust year to fall semester
    filter(cohort >= 2014, !is.na(cohort)) %>%
    mutate(year = year - 1) %>%
    
    # Combine Asian and Native American into "Other" category for sufficient sample sizes
    mutate(
      tot_asmt_oth = tot_asmt_asn + tot_asmt_nam,
      gcs_mn_oth = case_when(
        tot_asmt_oth > 0 ~ (gcs_mn_asn * tot_asmt_asn + gcs_mn_nam * tot_asmt_nam) / tot_asmt_oth,
        TRUE ~ NA_real_
      )
    ) %>%
    
    # Clean LEAID for merging  
    mutate(
      leaid = str_pad(as.character(sedalea), width = 7, pad = "0"),
      state = str_sub(leaid, 1, 2)
    ) %>%
    
    select(year, cohort, leaid, state, grade, sedaleaname,
           starts_with("gcs_mn_"), starts_with("tot_asmt_"))
  
  cat(glue("✓ Processed {nrow(seda_data):,} district-grade-year observations for ELA\n"))
  return(seda_data)
}

# Process SEDA data
# seda_data <- download_and_process_seda()
```

## Data Merging and Geographic Matching

Combine SEDA achievement data with district locations and city crosswalks.

```{r data-merging}
#| label: data-merging

merge_education_data <- function(seda_data, district_data, city_crosswalk) {
  
  cat("Merging education datasets...\n")
  
  # Merge SEDA with district city locations
  merged_data <- seda_data %>%
    left_join(district_data, by = c("leaid", "state", "year")) %>%
    
    # Handle districts that appear in some years but not others
    group_by(leaid) %>%
    mutate(
      # Flag districts that ever match to CCD data
      has_location = any(!is.na(city_location))
    ) %>%
    # Forward/backward fill city location within district (only for districts with location data)
    # Using tidyr::fill() which is more robust than zoo::na.fill() for missing value patterns
    filter(has_location) %>%
    arrange(leaid, year) %>%
    group_by(leaid) %>%
    tidyr::fill(city_location, .direction = "downup") %>%
    ungroup() %>%
    
    # Clean city names to proper case for matching
    mutate(
      city_name = str_to_title(city_location)
    ) %>%
    
    # Merge with city crosswalk to get standardized place codes
    inner_join(city_crosswalk, by = c("city_name", "state")) %>%
    
    # Final data restrictions
    filter(
      cohort >= 2014,
      year <= config$latest_year
    ) %>%
    
    # Convert back to numeric LEAID
    mutate(leaid = as.numeric(leaid)) %>%
    
    select(-city_location, -has_location)
  
  cat(glue("✓ Merged data: {length(unique(merged_data$leaid)):,} districts in {length(unique(merged_data$city_name)):,} cities\n"))
  
  return(merged_data)
}

# Merge datasets
# merged_education_data <- merge_education_data(seda_data, district_data, city_crosswalk)
```

## Learning Rate Calculation

Calculate district-specific learning rates using weighted regression, then aggregate to city level.

```{r learning-rates}
#| label: learning-rates

calculate_learning_rates <- function(education_data, subgroups) {
  
  cat("Calculating learning rates by subgroup...\n")
  
  # Function to calculate learning rate for a single subgroup-cohort combination
  calc_subgroup_rates <- function(data, subgroup, cohort_year) {
    
    # Column names for this subgroup
    score_col <- glue("gcs_mn_{subgroup}")
    n_col <- glue("tot_asmt_{subgroup}")
    
    # Prepare data for regression
    reg_data <- data %>%
      filter(cohort == cohort_year) %>%
      select(leaid, grade, all_of(score_col), all_of(n_col)) %>%
      rename(score = !!score_col, n_students = !!n_col) %>%
      filter(!is.na(score), n_students > 0)
    
    if (nrow(reg_data) == 0) return(NULL)
    
    # Run weighted regression: score ~ grade × district 
    # This gives district-specific slopes (learning rates)
    model <- lm(score ~ grade:factor(leaid) + factor(leaid), 
                data = reg_data, weights = n_students)
    
    # Extract coefficients and standard errors
    coef_summary <- tidy(model) %>%
      filter(str_detect(term, "grade:factor\\(leaid\\)")) %>%
      mutate(
        leaid = as.numeric(str_extract(term, "(?<=leaid\\))\\d+"))
      ) %>%
      select(leaid, estimate, std.error) %>%
      rename(learning_rate = estimate, se = std.error)
    
    # Calculate quality metrics for each district
    quality_metrics <- reg_data %>%
      group_by(leaid) %>%
      summarise(
        num_grades_included = n(),
        min_sample_size = min(n_students, na.rm = TRUE),
        .groups = "drop"
      )
    
    # Combine results
    results <- coef_summary %>%
      left_join(quality_metrics, by = "leaid") %>%
      mutate(
        cohort = cohort_year,
        subgroup = subgroup,
        
        # Calculate confidence intervals
        learning_rate_lb = learning_rate - 1.96 * se,
        learning_rate_ub = learning_rate + 1.96 * se,
        
        # Assign quality flags based on sample size and grade coverage
        learning_rate_quality = case_when(
          (num_grades_included >= 5 & min_sample_size >= 30) ~ 1L,  # High quality
          (num_grades_included >= 4 & min_sample_size >= 30) ~ 2L,  # Medium quality  
          (!is.na(learning_rate)) ~ 3L,                             # Low quality
          TRUE ~ NA_integer_
        ),
        
        # Remove false zeros (artifact of regression procedure)
        learning_rate = ifelse(learning_rate == 0 & learning_rate_lb == 0 & learning_rate_ub == 0, 
                               NA_real_, learning_rate),
        learning_rate_lb = ifelse(is.na(learning_rate), NA_real_, learning_rate_lb),
        learning_rate_ub = ifelse(is.na(learning_rate), NA_real_, learning_rate_ub),
        learning_rate_quality = ifelse(is.na(learning_rate), NA_integer_, learning_rate_quality)
      ) %>%
      
      select(cohort, leaid, subgroup, learning_rate, learning_rate_lb, 
             learning_rate_ub, learning_rate_quality)
    
    return(results)
  }
  
  # Calculate learning rates for all subgroups and cohorts
  # Use parallel processing for efficiency
  all_combinations <- expand_grid(
    subgroup = subgroups,
    cohort = unique(education_data$cohort)
  )
  
  learning_rates <- future_map2_dfr(
    all_combinations$subgroup, 
    all_combinations$cohort,
    ~ calc_subgroup_rates(education_data, .x, .y),
    .progress = TRUE
  )
  
  cat(glue("✓ Calculated learning rates for {nrow(learning_rates):,} district-subgroup-cohort combinations\n"))
  
  return(learning_rates)
}

# Calculate learning rates
# district_learning_rates <- calculate_learning_rates(merged_education_data, subgroups)
```

## City-Level Aggregation

Aggregate district-level learning rates to city level using enrollment weighting.

```{r city-aggregation}
#| label: city-aggregation

aggregate_to_cities <- function(education_data, learning_rates, city_crosswalk) {
  
  cat("Aggregating learning rates to city level...\n")
  
  # Get enrollment weights for aggregation
  enrollment_weights <- education_data %>%
    group_by(leaid, cohort, city_name, state, place) %>%
    summarise(
      across(starts_with("tot_asmt_"), sum, na.rm = TRUE),
      .groups = "drop"
    )
  
  # Function to aggregate one subgroup at a time
  aggregate_subgroup <- function(subgroup_name) {
    
    n_col <- glue("tot_asmt_{subgroup_name}")
    
    # Merge learning rates with enrollment weights
    subgroup_data <- learning_rates %>%
      filter(subgroup == subgroup_name) %>%
      left_join(enrollment_weights, by = c("leaid", "cohort")) %>%
      filter(!is.na(.data[[n_col]]), .data[[n_col]] > 0)
    
    # Aggregate to city level using enrollment weighting
    city_aggregated <- subgroup_data %>%
      group_by(city_name, state, place, cohort) %>%
      summarise(
        learning_rate = weighted.mean(learning_rate, .data[[n_col]], na.rm = TRUE),
        learning_rate_lb = weighted.mean(learning_rate_lb, .data[[n_col]], na.rm = TRUE),
        learning_rate_ub = weighted.mean(learning_rate_ub, .data[[n_col]], na.rm = TRUE),
        learning_rate_quality = round(weighted.mean(learning_rate_quality, .data[[n_col]], na.rm = TRUE)),
        .groups = "drop"
      ) %>%
      mutate(subgroup = subgroup_name) %>%
      rename(year = cohort) %>%
      # Adjust year to reflect fall of academic year
      mutate(year = year - 1)
    
    return(city_aggregated)
  }
  
  # Aggregate all subgroups and combine
  city_results <- map_dfr(subgroups, aggregate_subgroup)
  
  cat(glue("✓ Aggregated to {length(unique(city_results$city_name)):,} cities\n"))
  
  return(city_results)
}

# Aggregate to city level
# city_learning_rates <- aggregate_to_cities(merged_education_data, district_learning_rates, city_crosswalk)
```

## Data Quality Assessment

Evaluate the quality and coverage of the learning rate estimates.

```{r quality-assessment}
#| label: quality-assessment

assess_data_quality <- function(city_data) {
  
  cat("=== DATA QUALITY ASSESSMENT ===\n\n")
  
  # Overall summary statistics by quality level
  cat("Learning Rate Distributions by Quality Level:\n")
  quality_summary <- city_data %>%
    filter(subgroup == "all") %>%
    group_by(learning_rate_quality) %>%
    summarise(
      n_obs = n(),
      mean_rate = mean(learning_rate, na.rm = TRUE),
      median_rate = median(learning_rate, na.rm = TRUE),
      sd_rate = sd(learning_rate, na.rm = TRUE),
      .groups = "drop"
    )
  
  print(quality_summary)
  cat("\n")
  
  # Coverage by year and subgroup
  cat("Data Coverage by Year and Subgroup:\n")
  coverage <- city_data %>%
    group_by(year, subgroup) %>%
    summarise(
      total_cities = n(),
      cities_with_data = sum(!is.na(learning_rate)),
      pct_coverage = round(100 * cities_with_data / total_cities, 1),
      .groups = "drop"
    ) %>%
    pivot_wider(names_from = subgroup, values_from = c(cities_with_data, pct_coverage))
  
  print(coverage)
  cat("\n")
  
  # Check for logical consistency in confidence intervals  
  cat("Confidence Interval Consistency Checks:\n")
  ci_issues <- city_data %>%
    filter(!is.na(learning_rate)) %>%
    summarise(
      below_lower_bound = sum(learning_rate < learning_rate_lb, na.rm = TRUE),
      above_upper_bound = sum(learning_rate > learning_rate_ub, na.rm = TRUE)
    )
  
  cat(glue("• Estimates below lower bound: {ci_issues$below_lower_bound}\n"))
  cat(glue("• Estimates above upper bound: {ci_issues$above_upper_bound}\n\n"))
  
  # Quality flag completeness
  cat("Quality Flag Assessment:\n")
  quality_flags <- city_data %>%
    summarise(
      total_observations = n(),
      missing_quality_flags = sum(is.na(learning_rate_quality)),
      quality_flag_when_missing_rate = sum(is.na(learning_rate) & !is.na(learning_rate_quality))
    )
  
  cat(glue("• Missing quality flags: {quality_flags$missing_quality_flags}\n"))
  cat(glue("• Quality flags when rate is missing: {quality_flags$quality_flag_when_missing_rate}\n\n"))
  
  return(invisible(city_data))
}

# Assess data quality  
# final_data <- assess_data_quality(city_learning_rates)
```

## Data Visualization

Create diagnostic visualizations to understand the distribution and quality of learning rate estimates.

```{r visualizations}
#| label: visualizations

create_diagnostic_plots <- function(city_data) {
  
  # Distribution of learning rates by quality level for all students
  p1 <- city_data %>%
    filter(subgroup == "all", !is.na(learning_rate)) %>%
    ggplot(aes(x = learning_rate)) +
    geom_histogram(bins = 30, alpha = 0.7, fill = "steelblue") +
    facet_grid(learning_rate_quality ~ year, labeller = label_both) +
    labs(
      title = "Distribution of ELA Learning Rates by Quality Level and Year",
      subtitle = "All Students",
      x = "Learning Rate (Grade Equivalents per Year)",
      y = "Number of Cities"
    ) +
    theme_minimal()
  
  # Learning rates by demographic subgroup
  p2 <- city_data %>%
    filter(!is.na(learning_rate), year == max(year, na.rm = TRUE)) %>%
    ggplot(aes(x = learning_rate, fill = subgroup)) +
    geom_density(alpha = 0.6) +
    facet_wrap(~ subgroup, scales = "free_y") +
    labs(
      title = "Learning Rate Distributions by Demographic Subgroup",
      subtitle = glue("Most Recent Year: {max(city_data$year, na.rm = TRUE)}"),
      x = "Learning Rate (Grade Equivalents per Year)",
      y = "Density"
    ) +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Quality distribution over time
  p3 <- city_data %>%
    filter(subgroup == "all") %>%
    count(year, learning_rate_quality) %>%
    ggplot(aes(x = factor(year), y = n, fill = factor(learning_rate_quality))) +
    geom_col(position = "stack") +
    labs(
      title = "Data Quality Distribution Over Time",
      subtitle = "All Students",
      x = "Year",
      y = "Number of Cities",
      fill = "Quality Level"
    ) +
    scale_fill_manual(
      values = c("1" = "#2E8B57", "2" = "#FFD700", "3" = "#CD5C5C"),
      labels = c("1" = "High", "2" = "Medium", "3" = "Low"),
      na.value = "grey80"
    ) +
    theme_minimal()
  
  # Return plots as a list
  return(list(
    quality_by_year = p1,
    subgroup_distributions = p2, 
    quality_over_time = p3
  ))
}

# Generate diagnostic plots
# plots <- create_diagnostic_plots(final_data)

# Display plots (uncomment to show)
# plots$quality_by_year
# plots$subgroup_distributions  
# plots$quality_over_time
```

## Data Export

Format and export the final datasets for use in community dashboards and analysis.

```{r data-export}
#| label: data-export

export_final_data <- function(city_data, export_path = paths$built_dir) {
  
  cat("Exporting final datasets...\n")
  
  # Prepare long format data with all subgroups
  long_format <- city_data %>%
    
    # Create proper subgroup labels and categories
    mutate(
      subgroup_type = case_when(
        subgroup == "all" ~ "all",
        subgroup %in% c("wht", "blk", "hsp") ~ "race-ethnicity", 
        subgroup %in% c("mal", "fem") ~ "gender",
        subgroup %in% c("ecd", "nec") ~ "income",
        TRUE ~ "other"
      ),
      
      subgroup = case_when(
        subgroup == "all" ~ "All",
        subgroup == "wht" ~ "White, Non-Hispanic",
        subgroup == "blk" ~ "Black, Non-Hispanic", 
        subgroup == "hsp" ~ "Hispanic",
        subgroup == "mal" ~ "Male",
        subgroup == "fem" ~ "Female",
        subgroup == "ecd" ~ "Economically Disadvantaged",
        subgroup == "nec" ~ "Not Economically Disadvantaged",
        TRUE ~ subgroup
      )
    ) %>%
    
    # Final column ordering
    select(year, state, place, subgroup_type, subgroup, 
           learning_rate, learning_rate_lb, learning_rate_ub, learning_rate_quality) %>%
    
    arrange(desc(year), state, place, subgroup_type, subgroup)
  
  # Export all subgroups data
  all_file <- file.path(export_path, 
                        glue("SEDA_all_subgroups_city_2014-{config$current_year-1}.csv"))
  fwrite(long_format, all_file)
  
  # Export "All students" only data  
  all_students <- long_format %>%
    filter(subgroup_type == "all") %>%
    select(-subgroup_type, -subgroup)
  
  all_students_file <- file.path(export_path,
                                glue("SEDA_all_city_2014-{config$current_year-1}.csv"))
  fwrite(all_students, all_students_file)
  
  cat(glue("✓ Exported all subgroups data: {all_file}\n"))
  cat(glue("✓ Exported all students data: {all_students_file}\n"))
  cat(glue("✓ Final dataset includes {nrow(long_format):,} city-subgroup-year observations\n"))
  
  return(long_format)
}

# Export final data
# exported_data <- export_final_data(final_data)
```

## Summary and Next Steps

This analysis provides city-level estimates of English Language Arts learning growth rates with the following key features:

### **Methodological Strengths**
- **Rigorous Statistical Approach**: Uses weighted regression to account for varying sample sizes across districts and grades
- **Quality Controls**: Implements systematic quality flags based on grade coverage and sample size
- **Demographic Equity**: Provides breakdowns by race/ethnicity, gender, and economic status
- **Geographic Aggregation**: Properly weights district estimates when aggregating to city level

### **Data Quality Considerations** 
- **Coverage**: Not all cities have sufficient data for all subgroups in all years
- **Quality Tiers**: Use quality flags to understand reliability of estimates
- **Sample Sizes**: Smaller demographic subgroups may have limited geographic coverage
- **Temporal Trends**: Consider year-to-year variation when interpreting results

### **Recommended Next Steps**
1. **Validation**: Compare results with state-level trends and known educational policies
2. **Robustness Checks**: Examine sensitivity to different quality thresholds and weighting schemes  
3. **Integration**: Link with other community indicators for comprehensive dashboard
4. **Updates**: Refresh annually as new SEDA data becomes available

### **File Organization**
```
08_education/data/
├── raw/               # Original SEDA and CCD data files
├── intermediate/      # Processed components for merging  
└── built/            # Final analytical datasets
    ├── SEDA_all_subgroups_city_2014-2018.csv
    └── SEDA_all_city_2014-2018.csv
```

---

*This analysis was translated from STATA to R with emphasis on reproducibility, documentation, and computational efficiency. The methodology follows the original SEDA research protocols while leveraging R's strengths in data processing and visualization.*