---
title: "Juvenile Arrests - Place"
author: "Vincent Pancini"
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: false
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    editor_options:
      chunk_output_type: console
---

<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup}
options(scipen = 999)

library(tidyverse)
library(tidylog)
library(urbnthemes)

set_urbn_defaults(style = "print")

```

This file creates the place-level metric for juvenile arrests per 100,000 juveniles (age 10-17).

This file relies on data created in `01_agency_geo_place.Rmd`. You must run that file before running this file.

The rest of this file is organized as follows:
1. Background
2. Manually download NIBRS data from openICPSR
3. Read in and clean NIBRS data
4. Aggregate agency-level arrests to the place level
5. Calculate reporting rate for all agencies
6. Calculate rates of juvenile justice arrests per 100,000 juveniles
7. Construct quality indicators
8. Save and write out data


## 1. Background
This section provides background for the data used to create this metric.

The data for this metric come from the FBI's [National Incident-Based Reporting System (NIBRS)](https://bjs.ojp.gov/national-incident-based-reporting-system-nibrs). This is the national standard for law enforcement crime data reporting in the United States. Processing the raw NIBRS data would take a lot of work.

Instead of downloading the raw data directly, we use criminologist Jacob Kaplan's Concatenated Files: National Incident-Based Reporting System (NIBRS) Data, 1991-2022. We use the latest version, Version 9. These data are available for download via [openICPSR](https://www.openicpsr.org/openicpsr/project/118281/version/V9/view).

The NIBRS data are separated into segments. For this metric, we use the following two segments:
* Arrestee segment
* Group B Arrest Report segment

To learn more about NIBRS data, consult Jacob Kaplan's book titled [National Incident-Based Reporting System (NIBRS) Data: A Practitioner's Guide](https://nibrsbook.com/). For this metric, the most important thing to know about these data is that most police agencies do not report NIBRS data prior to 2021. In 2019, only 8,500 police agencies (covering about 45% of the US population) reported NIBRS data. This represented less than half of the approximately 18,000 police agencies in the United States. Instead, agencies primarily reported [UCR data](https://ucrbook.com/). However, the FBI has moved entirely to NIBRS data starting in 2021 and no longer collects UCR data, which is why we use NIBRS data instead ([source](https://nibrsbook.com/overview-of-the-data.html)). We do not want to use NIBRS data for some years and UCR data for others. This means that **there will be a lot of missing values prior to 2021**.


## 2. Manually download NIBRS data from openICPSR
This section describes the steps for manually downloading the NIBRS data segments necessary for this metric.

Accessing these data requires an ICPSR account. Therefore, we do not directly download the data in the code. Manually downloading these data is a five-step process, but you will only need to do each of these steps once for the creation of this metric:
    1. Create an ICPSR account [here](https://www.icpsr.umich.edu/cgi-bin/newacct).
    2. Navigate to Jacob Kaplan's project page for his NIBRS Concatenated Files [here](https://www.openicpsr.org/openicpsr/project/118281/version/V9/view).
    3. Select `nibrs_1991_2022_arrestee_segment_rds.zip`
        + Select `Download this file`
        + Select `I Agree` on the `Terms of Use`
        + Move the zip file from your `Downloads` folder to the appropriate directory
    4. Navigate back to the [project page](https://www.openicpsr.org/openicpsr/project/118281/version/V9/view)
    5. Select `nibrs_group_b_arrest_report_segment_2022_rds.zip`
        + Select `Download this file`
        + Select `I Agree` on the `Terms of Use`
        + Move the zip file from your `Downloads` folder to the appropriate directory


## 3. Read in and clean NIBRS data
This section extracts the necessary files from each zip file downloaded in the previous step and reads the files into the R environment. Each zip file contains data files for 1991-2022. However, we only want 2016-2022.

### 3.1 Read in NIBRS data

#### 3.1a Extract files for 2016-2022 from the zip files for both segments
Extract each file for the years 2016-2022 from the Group A `arrestee` segment and from the Group B `arrest_report` segment.
```{r}
# Specify range of years
start_year <- 2016
end_year <- 2022

# Create a vector of years between start_year and end_year
my_years <- seq(start_year, end_year)

# Extract years 2016-2022 for the arrestee segment
for (year in my_years) {
  unzip(
  zipfile = here::here("07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_arrestee_segment_rds.zip"),
  files = paste0("nibrs_arrestee_segment_", year, ".rds"),
  exdir = here::here("07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_arrestee_segment_rds")
)
}

# Extract years 2016-2022 for the Group B arrest report segment
for (year in my_years) {
  unzip(
  zipfile = here::here("07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_group_b_arrest_report_segment_rds.zip"),
  files = paste0("nibrs_group_b_arrest_report_segment_", year, ".rds"),
  exdir = here::here("07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_group_b_arrest_report_segment_rds")
)
}

# Remove zip files
unlink(here::here("07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_arrestee_segment_rds.zip"))

unlink(here::here("07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_group_b_arrest_report_segment_rds.zip"))

# Clean up environment
rm(start_year, end_year)

```

#### 3.1b Read in and append 2016-2022 NIBRS data for both groups
For each segment, read in the files for years 2016-2022 and append them. We also limit the appended files to relevant variables.
    todo(): This section has a very long run time. Is there a quicker way to read in and append all files for each segment?
```{r}
# List all files in the Arrestee segment folder (Group A)
file_list <- list.files(here::here(
  "07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_arrestee_segment_rds"),
  pattern = "[.]rds$", full.names = TRUE)

# Read in Group A files and append them
arrests_a <- file_list %>%
  set_names(basename) %>%
  purrr::map(readRDS) %>%
  list_rbind(names_to = "year") %>%
  mutate(year = parse_number(year))

# Select only necessary variables from Group A
arrests_a <- arrests_a %>%
  select(year, ori, # ORI is a unique agency identifier
         incident_number, ucr_arrest_offense_code, age_of_arrestee, sex_of_arrestee, race_of_arrestee, ethnicity_of_arrestee) %>%
  mutate(group = "A",
         age = as.numeric(age_of_arrestee))
  

# List all files in the Arrest Report Segment folder (Group B)
file_list <- list.files(here::here(
  "07_safety", "juvenile-arrests", "data", "nibrs_1991_2022_group_b_arrest_report_segment_rds"),
  pattern = "[.]rds$", full.names = TRUE)

# Read in Group B files and append them
arrests_b <- file_list %>%
  set_names(basename) %>%
  purrr::map(readRDS) %>%
  list_rbind(names_to = "year") %>%
  mutate(year = parse_number(year))

# Select only necessary variables from Group B
arrests_b <- arrests_b %>%
  rename(incident_number = arrest_transaction_incident_num) %>%
  select(year, ori, # ORI is a unique agency identifier
         incident_number, ucr_arrest_offense_code, age_of_arrestee, sex_of_arrestee, race_of_arrestee, ethnicity_of_arrestee) %>%
  mutate(group = "B",
         age = as.numeric(age_of_arrestee))

```


#### 3.1c Append the 2016-2022 Group A file and the 2016-2022 Group B file.
Append Group A and Group B files for all arrests
```{r}
# Append files
arrests_all <- bind_rows(arrests_a, arrests_b)

```

Note that because the process of extracting and reading in these files takes so long, I write out the appended all arrests data here. Then I read this file back in when returning to this program so that I don't have to constantly re-run the previous steps.
```{r}
# Write out all arrests data
write_csv(arrests_all, file = here::here("07_safety", "juvenile-arrests", "data", "arrests_all_16_22.csv"))

# Read all arrests data back in for ease of future use
arrests_all <- read.csv(here::here("07_safety", "juvenile-arrests", "data", "arrests_all_16_22.csv"))

```

#### 3.1d Create subset of data for just juveniles
Subset the arrests data to just those under the age of 18
```{r}
# Create a subset of just juvenile arrests
arrests_juv <- arrests_all %>%
  filter(age < 18)

```

Note that because the process of extracting and reading in these files takes so long, I write out appended juvenile arrests data here. Then I read this file back in when returning to this program so that I don't have to constantly re-run the previous steps.
```{r}
# Write out juvenile arrests data
write_csv(arrests_juv, file = here::here("07_safety", "juvenile-arrests", "data", "arrests_juv_16_22.csv"))

# Read juvenile arrests data back in for ease of future use
arrests <- read.csv(here::here("07_safety", "juvenile-arrests", "data", "arrests_juv_16_22.csv"))

# Remove obsolete files
rm(arrests_a, arrests_b)
gc()

```


### 3.2 Identify violent crimes and property crimes
Group A offense codes are categorized into crimes against property, crimes against persons (violent crimes), or crimes against society. Group B offense codes do not receive any of these categorizations. We are interested in property crimes and violent crimes. Using Section 2.3 of the [2021.1 National Incident-Based Reporting System User Manual](https://bjs.ojp.gov/sites/g/files/xyckuh236/files/sarble/data_common/nibrs-user-manual-2021-1041521.pdf), we create a data frame of all offense, crime type, and group combinations. We then use this data to identify property crimes and violent crimes in our data. 
    
There are 65 distinct offense codes across all years.
```{r}
# Count number of observations in each offense code
arrests %>%
  count(ucr_arrest_offense_code)

```


Create a data frame that lists all offense, crime type, and group combinations using Section 2.3 of the [2021.1 National Incident-Based Reporting System User Manual](https://bjs.ojp.gov/sites/g/files/xyckuh236/files/sarble/data_common/nibrs-user-manual-2021-1041521.pdf). Note that the full frame includes crimes against society, but these are not used in our analysis.
```{r}
# Create data frame for all offense, type, and group combinations.
codes <- tribble(
  ~offense,	~crime_against,	~group,
  "animal cruelty",	                                                    "society",	"A",	
  "arson",	                                                            "property",	"A",	
  "assault offenses - aggravated assault",	                            "person",	"A",	
  "assault offenses - simple assault",	                                "person",	"A",	
  "assault offenses - intimidation",	                                  "person",	"A",	
  "bribery",	                                                          "property",	"A",	
  "burglary/breaking and entering",	                                    "property",	"A",	
  "commerce violations - import violations",	                          "society",	"A",	
  "commerce violations - export violations",	                          "society",	"A",	
  "commerce violations - federal liquor offenses",	                    "society",	"A",	
  "commerce violations - federal tobacco offenses",	                    "society",	"A",	
  "commerce violations - wildlife trafficking",	                        "society",	"A",	
  "counterfeiting/forgery",	                                            "property",	"A",	
  "destruction/damage/vandalism of property",	                          "property",	"A",	
  "drug/narcotic offenses - drug/narcotic violations",	                "society",	"A",	
  "drug/narcotic offenses - drug equipment violations",	                "society",	"A",	
  "embezzlement",	                                                      "property",	"A",	
  "espionage",	                                                        "society",	"A",	
  "extortion/blackmail",	                                              "property",	"A",	
  "fraud offenses - false pretenses/swindle/confidence game",	          "property",	"A",	
  "fraud offenses - credit card/atm fraud",	                            "property",	"A",	
  "fraud offenses - impersonation",	                                    "property",	"A",	
  "fraud offenses - welfare fraud",	                                    "property",	"A",	
  "fraud offenses - wire fraud",	                                      "property",	"A",	
  "fraud offenses - identity theft",	                                  "property",	"A",	
  "fraud offenses - hacking/computer invasion",	                        "property",	"A",	
  "fraud offenses - money laundering",	                                "society",	"A",	
  "fugitive offenses - harboring escapee/concealing from arrest",	      "society",	"A",	
  "fugitive offenses - flight to avoid prosecution",	                  "society",	"A",	
  "fugitive offenses - flight to avoid deportation",	                  "society",	"A",	
  "gambling offenses - betting/wagering",	                              "society",	"A",	
  "gambling offenses - operating/promoting/assisting gambling",	        "society",	"A",	
  "gambling offenses - gambling equipment violations",	                "society",	"A",	
  "gambling offenses - sports tampering",	                              "society",	"A",	
  "murder/nonnegligent manslaughter",	                                  "person",	"A",	
  "negligent manslaughter",	                                            "person",	"A",	
  "homicide offenses - justifiable homicide",	                          "not a crime",	"A",	
  "human trafficking - commercial sex acts",	                          "person",	"A",	
  "human trafficking - involuntary servitude",	                        "person",	"A",	
  "immigration violations - illegal entry into the united states",	    "society",	"A",	
  "immigration violations - false citizenship",	                        "society",	"A",	
  "immigration violations - smuggling aliens",	                        "society",	"A",	
  "immigration violations - reentry after deportation",               	"society",	"A",	
  "kidnapping/abduction",	                                              "person",	"A",	
  "larceny/theft offenses - pocket-picking",	                          "property",	"A",	
  "larceny/theft offenses - purse-snatching",	                          "property",	"A",	
  "larceny/theft offenses - shoplifting",	                              "property",	"A",	
  "larceny/theft offenses - theft from building",	                      "property",	"A",	
  "larceny/theft offenses - theft from coin-operated machine or device", "property",	"A",	
  "larceny/theft offenses - theft from motor vehicle",	                "property",	"A",	
  "larceny/theft offenses - theft of motor vehicle parts/accessories",	"property",	"A",	
  "larceny/theft offenses - all other larceny",	                        "property",	"A",	
  "motor vehicle theft",	                                              "property",	"A",	
  "pornography/obscene material",	                                      "society",	"A",	
  "prostitution offenses - prostitution",	                              "society",	"A",	
  "prostitution offenses - assisting or promoting prostitution",	      "society",	"A",	
  "prostitution offenses - purchasing prostitution",	                  "society",	"A",	
  "robbery",	                                                          "property",	"A",	
  "sex offenses - rape",	                                              "person",	"A",	
  "sex offenses - sodomy",	                                            "person",	"A",	
  "sex offenses - sexual assault with an object",	                      "person",	"A",	
  "sex offenses - fondling (incident liberties/child molest)",	        "person",	"A",	
  "sex offenses - incest",	                                            "person",	"A",	
  "sex offenses - statutory rape",	                                    "person",	"A",	
  "sex offenses - failure to register as a sex offender",	              "society",	"A",	
  "stolen property offenses (receiving, selling, etc.)",	              "property",	"A",	
  "treason",	                                                          "society",	"A",	
  "weapon law violations - weapon law violations",	                    "society",	"A",	
  "weapon law violations - violation of national firearm act of 1934",	"society",	"A",	
  "weapon law violations - weapons of mass destruction",	              "society",	"A",	
  "weapon law violations - explosives",	                                "society",	"A",	
  "failure to appear",	                                                 "",	"B",	
  "curfew/loitering/vagrancy violations",	                               "",	"B",	
  "disorderly conduct",	                                                 "",	"B",	
  "driving under the influence",	                                       "",	"B",	
  "family offenses, nonviolent",	                                       "",	"B",	
  "federal resource violations",	                                       "",	"B",	
  "liquor law violations",	                                             "",	"B",	
  "perjury",	                                                           "",	"B",	
  "trespass of real property",	                                         "",	"B",	
  "all other offenses",	                                                 "",	"B",	
  # The following three codes were removed in 2021
  "bad checks",	                                                         "",	"",
  "drunkenness",                                                      	 "",	"",	
  "peeping tom",	                                                       "",	"",	
  # The following code was removed in 2008
  "runaway",                                                             "",	"",	
)

```

Using the data frame above, group offenses into violent crimes and property crimes.
```{r}
# Extract offenses into vectors for property crimes and violent crimes
property <- subset(codes, crime_against=="property")$offense
violent <- subset(codes, crime_against=="person")$offense

# Create indicator variables for property and violent crimes
arrests <- arrests %>%
  mutate(property = ifelse(ucr_arrest_offense_code %in% property, 1, 0),
         violent = ifelse(ucr_arrest_offense_code %in% violent, 1, 0)
  )

# Remove obsolete objects
rm(codes, property, violent)

```

1,025,994 observations (51%) are not in either category of arrests. These are all crimes against society or Group B offenses.
```{r}
arrests %>%
  group_by(property, violent) %>%
  count()

arrests %>%
  filter(property==0 & violent==0) %>%
  count(ucr_arrest_offense_code)

```


### 3.3 Recode race and ethnicity
The values for `race_of_arrestee` and `ethnicity_of_arrestee` are different between the Group A and Group B files. The races and ethnicities for Group B observations are coded using single letters and are therefore not very clear. I can't find a codebook anywhere for these variables, so I make my best guess at recoding them to match the codes used for Group A. 
    Note that if there is a codebook available, we should confirm this is correct. It shouldn't matter for the violent crime or property crime arrests because Group B arrests are excluded from those categories, but it would matter for total arrests.
    
Recode race and ethnicity for Group B observations.
```{r}
# Group B codes for race and ethnicity do not match those used for Group A
arrests %>%
  group_by(group, race_of_arrestee, ethnicity_of_arrestee) %>%
  count()

# No arrests for Group B are classified as violent or property crimes
arrests %>%
  group_by(group, property, violent) %>%
  count()


# Making my best guess to match Group B race/ethnicity codes to Group A codes

## Recode race
arrests <- arrests %>%
  mutate(
    race_of_arrestee = if_else(race_of_arrestee=="w", "white", race_of_arrestee),
    race_of_arrestee = if_else(race_of_arrestee=="b", "black", race_of_arrestee),
    race_of_arrestee = if_else(race_of_arrestee=="i", "american indian/alaskan native", race_of_arrestee),
    race_of_arrestee = if_else(race_of_arrestee=="a", "asian", race_of_arrestee),
    race_of_arrestee = if_else(race_of_arrestee=="p", "native hawaiian or other pacific islander", race_of_arrestee),
    race_of_arrestee = if_else(race_of_arrestee=="u", "unknown", race_of_arrestee)
  )

## Recode ethnicity
arrests <- arrests %>%
  mutate(
    ethnicity_of_arrestee = if_else(ethnicity_of_arrestee=="h", "hispanic origin", ethnicity_of_arrestee),
    ethnicity_of_arrestee = if_else(ethnicity_of_arrestee=="n", "not of hispanic origin", ethnicity_of_arrestee),
    ethnicity_of_arrestee = if_else(ethnicity_of_arrestee=="u", "unknown", ethnicity_of_arrestee)
  )

```

207,231 observations have `unknown` ethnicity and 183,693 have missing ethnicity
```{r}
arrests %>% 
  group_by(race_of_arrestee, ethnicity_of_arrestee) %>% 
  count()

arrests %>% 
  group_by(ethnicity_of_arrestee) %>% 
  count()

```


### 3.4 Aggregate arrest counts to the reporting agency level
Aggregate arrests by agency and finish race construction
    todo(): For this metric, we want additional subgroups of sex (male, female) and age groups (age 10-14, age 15-17) for all years. Need to sum arrests for these groups as well.
```{r}
arrests_agency <- arrests %>%
  filter(age >= 10) %>%
  group_by(year, ori) %>%
  summarize(arr_total = n(),
            arr_prop = sum(property==1),
            arr_viol = sum(violent==1),
            age_10_14 = sum(age >= 10 & age <= 14),
            age_15_17 = sum(age >= 15 & age <= 17),
            # sex_male = sum(sex_of_arrestee == "male"),
            # sex_female = sum(sex_of_arrestee == "female"),
            race_aian = sum(race_of_arrestee == "american indian/alaskan native"),
            race_asian = sum(race_of_arrestee == "asian"),
            race_black = sum(race_of_arrestee == "black"),
            race_nhpi = sum(race_of_arrestee == "native hawaiian or other pacific islander"),
            race_unkno = sum(race_of_arrestee == "unknown"),
            race_white = sum(race_of_arrestee == "white"),
            ethn_hispanic = sum(ethnicity_of_arrestee == "hispanic origin"),
            ethn_nonhisp = sum(ethnicity_of_arrestee == "not of hispanic origin"),
            group_a = max(group=="A"),
            group_b = max(group=="B")
  ) %>%
  ungroup() %>%
  mutate(race_asian_other = race_aian + race_asian + race_nhpi)

```

3,842 agencies across all years only have Group B arrests of juveniles
```{r}
arrests_agency %>% 
  count(group_a, group_b)

```

Make narrow version with just needed variables
```{r}
arrests_agency <- arrests_agency %>%
  select(
    year,
    ori, 
    arr_total, 
    arr_viol, 
    arr_prop, 
    # starts_with("age"),
    # starts_with("sex")
    starts_with("race"), 
    ethn_hispanic
  )

```


## 4. Aggregate agency-level arrests to the place-level
This section aggregates the agency-level arrest counts from the agency-level to the place-level.

Our data are currently long by year/agency. However, we want to know the arrest counts at the place level, not the agency level. In the `01_agency_geo_place.Rmd` program, we constructed a file that linked reporting agencies to places. We can use that file to aggregate counts from agencies to places.

### 4.1 Load universe of place/agency pairs
First load the universe of agencies linked to places. This file was created in `01_agency_geo_place.Rmd` and is long by unique agency/place pairs.
```{r}
bhf_place_urb <- read_csv(here::here("07_safety", "modified data", "all_agency_place.csv"))

```


### 4.2 Join agency-level arrests for all years onto the place-agency file
Join arrests onto the full universe of place-agencies.
    Note that the place-agency file does not contain `year`.
```{r}
arrests_agency_place <- left_join(
  x = bhf_place_urb,
  y = arrests_agency,
  by = c("ori")
)
  
```

### 4.3 Summarize agency-level arrests to the place level
Now that we have joined our agency-level juvenile arrests data onto the full universe of place-agencies, we can aggregate the counts from the agency to the place level.
```{r}
# Count number of arrests from reporting agencies in each place
arrests_place <- arrests_agency_place %>%
  group_by(state, place) %>%
  # AR: there are some agencies with no data and this summarize statement
  # makes it so that the value is zero instead of missing (ori - CA0300400)
  summarize(across(arr_total:ethn_hispanic, ~sum(.x, na.rm=TRUE))) %>%
  mutate(GEOID = str_c(state, place)) %>%
  ungroup()

```



## 5. Calculate reporting rate for all agencies
Agencies may have reported zero juvenile arrests. We use the full arrests file (i.e., all arrests and not just juvenile arrests) to calculate agency reporting rates.
```{r}
arrests_all_agency <- arrests_all %>%
  group_by(year, ori) %>%
  summarize(arr_total_any = n()) %>%
  ungroup()

```

Join total number of arrests for each agency onto the place/agency file.
    todo(): This should be for all years, not just 2022
```{r}
arrests_all_agency_place <- left_join(
   x = bhf_place_urb,
   y = arrests_all_agency %>%
     filter(year==2022),
   by = c("ori")
 ) %>%
  mutate(
    # Indicate reporting by place-agency
    reporting = ifelse(is.na(arr_total_any), 0, 1)
    )

```

Aggregate variables to the place level.
```{r}
# Summarize by place
arrests_all_place <- arrests_all_agency_place %>%
  group_by(state, place) %>%
  summarize(
    n = n(), 
    n_reporting = sum(reporting),
    n_core_city = sum(core_city),
    n_core_city_rpt = sum(core_city==1 & reporting==1),
    across(arr_total_any, ~sum(.x, na.rm=TRUE))
  ) %>%
  mutate(
    agencies_reporting = n_reporting / n,
    core_reporting = n_core_city_rpt / n_core_city,
    core_reporting = ifelse(is.nan(core_reporting), NA, core_reporting),
    GEOID = str_c(state, place)
  ) %>%
  ungroup()

```

Join total arrests back to juvenile arrest file
```{r}
# Join all arrests and juvenile arrests
arrests_place <- left_join(
  x = arrests_place,
  y = arrests_all_place,
  by = c("GEOID", "state", "place")
)

```



## 6. Calculate rates of juvenile arrests per 100,000 juveniles
Each instance of an arrest, citation, or summons for an offense is counted as an arrest. Note that an individual can be arrested more than once and therefore the rates constructed are of arrests, not of persons arrested. Arrest rates are calculated for all offense types available in NIBRS data, including offenses against property, persons, and society.

First load the demographic information for each place and for each year 2016-2022. This file was created using 5 year ACS data in `01_agency_geo_place.Rmd`.
```{r}
place_demo <- read_csv(here::here("07_safety", "modified data", "all_place_demo.csv"))

```

Join the place-level juvenile arrests data onto place-level population and demographic information from ACS.
```{r}
arrests_place_demo <- left_join(
  x = place_demo,
  y = arrests_place,
  by = c("GEOID")
)

```

Calculate arrest rates per 100,000 juveniles 
```{r}
arrest_rates_place <- arrests_place_demo %>%
  mutate(
    rate_juv_arrest = arr_total / age_1017 * 100000,
    rate_juv_arrest_violent = arr_viol / age_1017 * 100000,
    rate_juv_arrest_property = arr_prop / age_1017 * 100000,
    rate_juv_arrest_white = race_white / age_1017_white * 100000,
    rate_juv_arrest_black = race_black / age_1017_black * 100000,
    # rate_juv_arrest_aian  = race_aian / age_1017_aian * 100000,
    rate_juv_arrest_asian_other = race_asian_other / age_1017_asian_other * 100000,
    # rate_juv_arrest_nhpi  = race_nhpi / age_1017_nhpi * 100000,
    rate_juv_arrest_hispanic  = ethn_hispanic / age_1017_hispanic * 100000
  )

```

Replace the rates of observations with a population less than 30 with `NA`.
```{r}
# Suppress data using populations < 30 people
arrest_rates_place <- arrest_rates_place %>%
  mutate(
    rate_juv_arrest = ifelse(age_1017 < 30, NA, rate_juv_arrest),
    rate_juv_arrest_violent = ifelse(age_1017 < 30, NA, rate_juv_arrest_violent),
    rate_juv_arrest_property = ifelse(age_1017 < 30, NA, rate_juv_arrest_property),
    rate_juv_arrest_white = ifelse(age_1017_white < 30, NA, rate_juv_arrest_white),
    rate_juv_arrest_black = ifelse(age_1017_black < 30, NA, rate_juv_arrest_black),
    # rate_juv_arrest_aian = ifelse(age_1017_aian < 30, NA, rate_juv_arrest_aian),
    rate_juv_arrest_asian_other = ifelse(age_1017_asian_other < 30, NA, rate_juv_arrest_asian_other),
    # rate_juv_arrest_nhpi = ifelse(age_1017_nhpi < 30, NA, rate_juv_arrest_nhpi),
    rate_juv_arrest_hispanic = ifelse(age_1017_hispanic < 30, NA, rate_juv_arrest_hispanic)
  )

```

Check the distribution of rates
```{r}
# Check rates
arrest_rates_place %>%
  select(starts_with("rate_juv_arrest")) %>%
  skim()

```


## 7. Construct quality indicators
This section constructs an indicator used to assess the quality of data for each place. Places with 100% of agencies reporting are coded as 1; places with 80% or more of agencies reporting OR 100% of core agencies reporting are coded as 2; places with less than 80% of agencies reporting are coded as 3; places with 0% of agencies reporting (or no agencies in the place) are coded as `NA`.

Construct the quality indicator.
```{r}
# Compare all reporting plus core reporting
arrest_rates_place <- arrest_rates_place %>%
  mutate(
    rate_juv_arrest_quality = case_when(
      agencies_reporting == 1 ~ 1,
      agencies_reporting >= 0.8 ~ 2,
      agencies_reporting > 0 ~ 3,
      agencies_reporting == 0 ~ NA_real_
    ),
    # Quality for core reporting agencies
    rate_juv_arrest_core_quality = case_when(
      agencies_reporting == 1 ~ 1,
      agencies_reporting >= 0.8 | core_reporting==1 ~ 2,
      agencies_reporting > 0 ~ 3,
      agencies_reporting == 0 ~ NA_real_
    ),
    rate_juv_arrest_quality = ifelse(age_1017 < 30, NA, rate_juv_arrest_quality),
    rate_juv_arrest_core_quality = ifelse(age_1017 < 30, NA, rate_juv_arrest_core_quality)
  )

```


With core reporting, XX counties move from a quality indicator of 3 to a quality indicator of 2.
    Note the 2021 file used the `rate_juv_arrest_core_quality` version of the quality metric. I rename it below to adhere to our naming conventions
```{r}
# Check distribution of quality for all agencies and for core reporting agencies
arrest_rates_place %>%
  count(rate_juv_arrest_quality, rate_juv_arrest_core_quality)

# Remove all quality metric and rename core reporting quality metric
arrest_rates_place <- arrest_rates_place %>%
  select(-rate_juv_arrest_quality) %>%
  rename(rate_juv_arrest_quality = rate_juv_arrest_core_quality)

```


## 8. Save and write out data
```{r}
arrest_rates_place_all <- arrest_rates_place %>%
  mutate(across(starts_with("rate"), 
                ~ifelse(is.na(rate_juv_arrest_quality), NA, .x)),
         ) %>%
  select(c(year, state, place, starts_with("rate"), starts_with("age_1017")))
```

We expect a lot of missingness because most agencies did not use NIBRS to 2021.
```{r}
skim(arrest_rates_place_all)

```

Create a subset without subgroup rates and write out this data set
```{r}
# File with only total juvenile arrest rates
juv_arrest_place <- arrest_rates_place_all %>%
  select(year,
         state,
         place,
         rate_juv_arrest,
         rate_juv_arrest_violent,
         rate_juv_arrest_property,
         rate_juv_arrest_quality)

skim(juv_arrest_place)

# Write out this data set
juv_arrest_place %>%
  write_csv(here::here("07_safety",
                       "modified data",
                       "ja_metric_all_place.csv"))

```

Create a file that is long by subgroup and write it out
```{r}
# long form file with juvenile arrest rates by subgroup
# pivot longer subgroup populations
# create all row with overall juvenile population and arrest rate
juv_pop_all_subgroup <- arrest_rates_place_all %>%
  mutate(age_1017_all = age_1017) %>%
  select(
    year, 
    state,
    place, 
    age_1017_all, 
    age_1017_white, 
    age_1017_black, 
    age_1017_asian_other,
    age_1017_hispanic,
    rate_juv_arrest_quality
  ) %>%
  pivot_longer(cols = contains("age_1017"),
               names_to = "subgroup",
               values_to = "age_1017",
               values_drop_na = TRUE,
               names_pattern = "age_1017_(.*)")

# pivot longer subgroup arrest rates
juv_arrest_all_subgroup <- arrest_rates_place_all %>%
  mutate(rate_juv_arrest_all = rate_juv_arrest) %>%
  select(
    year, 
    state, 
    place, 
    rate_juv_arrest_all, 
    rate_juv_arrest_white, 
    rate_juv_arrest_black,
    rate_juv_arrest_asian_other,
    rate_juv_arrest_hispanic
  ) %>%
  pivot_longer(cols = contains("rate_juv_arrest_"), 
               names_to = "subgroup",
               values_to = "rate_juv_arrest",
               values_drop_na = TRUE,
               names_pattern = "rate_juv_arrest_(.*)")

# combine into one file
# replace quality indicators with missing if subgroup population < 30
juv_arrest_place_all_subgroup <- left_join(
  x = juv_pop_all_subgroup,
  y = juv_arrest_all_subgroup,
  by = c("year", "state", "place", "subgroup")
) %>%
  mutate(subgroup_type = "race-ethnicity",
         rate_juv_arrest_quality = ifelse(age_1017 < 30, NA, rate_juv_arrest_quality)) %>%
  select(-age_1017)

skim(juv_arrest_place_all_subgroup)

juv_arrest_place_all_subgroup %>% 
  count(rate_juv_arrest_quality, is.na(rate_juv_arrest))

# Write out this data set
juv_arrest_place_all_subgroup %>%
  write_csv(here::here("07_safety",
                       "modified data",
                       "ja_metric_all_place_subgroup.csv"))
```




