---
title: "Share of households with broadband access"
author: "Ridhi Purohit"
date: now
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
    code-fold: show
    code-tools: true
editor_options: 
  chunk_output_type: console
execute:
  warning: false
---

```{=html}
<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>
```

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

This document programs the **Digital Access** metric which encapsulates a ratio
of households with the presence of a computer (laptop, desktop, notebook computer, 
tablet, or portable wireless computer) and broadband (high speed) internet service
in the household. An overall ratio is calculated, as well as by racial subgroup 
- measured according to the race of each head of the household.

## Process

-   [Housekeeping](#housekeeping)
-   [Get IPUMS Data](#get-ipums-data)
-   [Format Data](#format-data)
-   [Merge Crosswalk](#merge-crosswalk)
-   [Create Metric](#create-metric)
-   [Explore Metric](#explore-metric)
-   [Metric Quality](#metric-quality)
-   [Evaluate Metric](#evaluate-metric)
-   [Final Data](#final-data)

## Housekeeping {#sec-housekeeping}

```{r}
#| label: setup

options(scipen = 999)
librarian::shelf(
  tidyverse,
  ipumsr,
  here,
  UrbanInstitute / urbnthemes
  )
set_urbn_defaults(style = "print")

source(here::here("functions", "testing", "evaluate_final_data.R"))

```


## Get IPUMS Data {#sec-get-ipums-data}

This code book uses the 2017-2021 5-year ACS extract from the IPUMS API using
R package `ipumsr`. To access this data, an IPUMS API key is required. This can
be obtained by registering on the [IPUMS
USA](https://uma.pop.umn.edu/usa/user/new?return_url=https%3A%2F%2Fusa.ipums.org%2Fusa-action%2Fmenu)
website and creating an API key. For more information see the `ipumsr` [getting started
vignette](https://tech.popdata.org/ipumsr/articles/ipums.html).


```{r}
#| label: query-ipums-api
#| cache: true
#| eval: false

# Run the code in this block the first time this notebook is executed to get the 
# data and avoid repeated downloads 

# define variables to be queried from IPUMS API

# NOTE: running var_spec("CIHISPEED", data_quality_flags = TRUE) produced following error: Error in `ipums_api_request()`: 
# ! API request failed with status 400.
# ✖ CIHISPEED has data quality flags, but none are available in the selected samples

acs_vars <- list("YEAR", "MULTYEAR", "SAMPLE", "SERIAL", "CBSERIAL", "HHWT", 
              "CLUSTER", "STRATA", "GQ", "REGION", "STATEFIP", "COUNTYFIP", "PUMA",
              "HHINCOME", var_spec("CILAPTOP", data_quality_flags = TRUE), 
              var_spec("CITABLET", data_quality_flags = TRUE),
              "CIHISPEED", "CINETHH",
              "PERNUM", "PERWT", "RACE", "HISPAN")


# define the IPUMS extract
# acs_2021_extract <- ipumsr::define_extract_micro(
#  collection = "usa",
#  description = "2017-2021 ACS 5-yr Sample",
#  samples = "us2021c",
#  variables = acs_vars)

# submit and wait for IPUMS extract
# acs_ext_submit <- ipumsr::submit_extract(acs_2021_extract)
# acs_ext_complete <- ipumsr::wait_for_extract(acs_ext_submit)

# check if data download directory is present
fs::dir_create(here("08_education/data/raw"))

#download IPUMS extract
# data_path <- ipumsr::download_extract(
#  acs_ext_complete, 
#  download_dir = here("08_education/data/raw")
#  )

```

## Format Data {#sec-format-data}

Now, the downloaded data needs to be cleaned and formatted so that it can be used
further to generate the metric. The 2017-2021 5-year ACS microdata extract for computer 
and internet use is in the form of household level data set with person level variables 
added for this analysis. 

To analyze the data at the household level, only the first person sampled in the household
is retained in the data.

```{r}
#| label: load-acs
#| cache: true

# Manually set the file name downloaded in the previous code chunk which is 
# 'usa_00007.xml' for me but will have some different value for whoever re-runs this

micro_filepath <- "usa_00008.xml"

# load acs data and filter for household level analysis
acs5_2021_raw <- ipumsr::read_ipums_micro(
  here(paste0("08_education/data/raw/", micro_filepath))
  ) |>
  filter(PERNUM == 1)

```

The IPUMS data consists of labelled vectors ([ipumsr](https://tech.popdata.org/ipumsr/articles/value-labels.html#remove-unused-value-labels)). 
We find the missing labels using the `ipums_var_info(acs5_2021_raw)` and inspecting the 
`val_labels` column.

For the "Total household income" i.e. `HHINCOME` column, "N/A" values are denoted
by `9999999` which needs to be updated before proceeding.

```{r}
#| label: handle-missingness

# re-label `HHINCOME` when value is 9999999 denoting "N/A "

acs5_2021_raw <- acs5_2021_raw |>
  mutate(
    HHINCOME = lbl_na_if(HHINCOME, ~ .lbl %in% c("N/A "))
  )

```

Since the digital access metric is at the household level, individuals who live in
group quarters need to be excluded from the data. For more information, see
the [variable description](https://usa.ipums.org/usa-action/variables/GQ#codes_section) on
IPUMS USA.

Running `ipums_var_info(acs5_2021_raw)` shows the value labels for the group 
quarters variable `GQ`:

val lbl                                        
0   Vacant unit                                
1   Households under 1970 definition           
2   Additional households under 1990 definition
3   Group quarters--Institutions               
4   Other group quarters                       
5   Additional households under 2000 definition
6   Fragment

```{r}
#| label: remove-group-quarters

# Only retain people who live in housing units in the data
acs5_2021_raw <- acs5_2021_raw |>
  filter((GQ %in% c(1,2,5)))

```


FIPS codes require re-formatting by padding with leading zeros so that state codes 
are two digits, county codes are three digits, and place codes are five digits.

```{r}
#| label: format-fips-codes

acs5_2021_raw <-
  acs5_2021_raw |>
  mutate(
    across(c(STATEFIP, COUNTYFIP, PUMA), as.character),
    STATEFIP = str_pad(STATEFIP, width = 2, pad = "0", side = "left"),
    COUNTYFIP = str_pad(COUNTYFIP, width = 3, pad = "0", side = "left"),
    PUMA = str_pad(PUMA, width = 5, pad = "0", side = "left")
  ) |>
  rename(
    STATE = STATEFIP,
    COUNTY = COUNTYFIP
  ) |>
  # Reformat the YEAR variable
  mutate(YEAR = as.integer(YEAR))
```

The race-ethnicity subgroup is created by collapsing the `RACE` and 
`HISPAN` variables into following categories:
- Black, Non-Hispanic
- White, Non-Hispanic
- Other Races and Ethnicities
- Hispanic

Running `ipums_val_labels(acs5_2021_raw$RACE)` shows the value labels for the 
`RACE` variable:

val   lbl                                        
                           
1     White                           
2     Black/African American          
3     American Indian or Alaska Native
4     Chinese                         
5     Japanese                        
6     Other Asian or Pacific Islander 
7     Other race, nec                 
8     Two major races                 
9     Three or more major races

```{r}
#| label: create-subgroup-race-ethnicity

# Create new column `RACE_mod` identifying White, Black , and Other races
acs5_2021_raw <- acs5_2021_raw |>
  mutate( 
    RACE_mod = lbl_relabel(
      RACE, lbl(10, "Other") ~ .val %in% c(3, 4, 5, 6, 7, 8, 9)),
    RACE_mod = as_factor(RACE_mod),
    HISPAN = as_factor(HISPAN)
    ) 

# Create `RACE_ETHNICITY` subgroup by collapsing `RACE_mod` and `HISPAN`
acs5_2021_raw <- acs5_2021_raw |>
  mutate(
    RACE_ETHNICITY = case_when(
      (RACE_mod == "Black/African American" & HISPAN == "Not Hispanic") ~ "Black, Non-Hispanic",
      (RACE_mod == "White" & HISPAN == "Not Hispanic") ~ "White, Non-Hispanic",
      (RACE_mod == "Other" & HISPAN == "Not Hispanic") ~ "Other, Non-Hispanic",
      (HISPAN %in% c("Mexican", "Puerto Rican", "Cuban", "Other")) ~ "Hispanic",
      .default = NA_character_
    )
  )

```


Income subgroup is created at two levels:
- Less than $50,000
- $50,000 or More

```{r}
#| label: create-subgroup-income

# Create `INCOME` subgroup using `HHINCOME`
acs5_2021_raw <- acs5_2021_raw |>
  mutate(
    INCOME = case_when(
      HHINCOME < 50000 ~ "Less than $50,000",
      HHINCOME >= 50000 ~ "$50,000 or More",
      .default = NA_character_)
    )

```


For measuring digital access, a composite variable `DIGITAL_ACCESS` is created 
using `CIHISPEED`, `CILAPTOP`, and `CITABLET` IPUMS variables. 

We consider the `CINETHH` variable to find the universe of observations for households 
with digital access. This variable indicates whether any member of a household 
has access to the internet (by either paying or without paying a cellphone 
company) or nobody in the household has access to the internet.

```{r}
#| label: data-exploration

# Find unique values and corresponding label values in the data

# In this data, we only observe values "1" and "2" corresponding to labels
# "Yes" and "No" respectively for `CITABLET`
unique(acs5_2021_raw$CITABLET)


# In this data, we only observe values "1" and "2" corresponding to labels
# "Yes" and "No" respectively for `CILAPTOP`
unique(acs5_2021_raw$CILAPTOP)


# In this data, we only observe values "0", "10" , and "20" corresponding to labels
# "N/A (GQ)", "Yes (Cable modem, fiber optic or DSL service)", and "No" 
# respectively for `CIHISPEED`

#### NOTE: It seems like "N/A (GQ)" value exists in the data even after group quarters 
# are removed from the data. Casual inspection of data frame suggests that even 
# if a household is classified as a housing unit, it might still have value "N/A (GQ)"
# in the `CIHISPEED` variable.
unique(acs5_2021_raw$CIHISPEED)


# In this data, we observe values "1", "2", and "3" corresponding to labels 
# 'Yes, with a subscription to an Internet Service', 'Yes, without a subscription 
# to an Internet Service' and 'No Internet access at this house, apartment, or 
# mobile home' for `CINETHH`
unique(acs5_2021_raw$CINETHH)

```

Since, group quarters are filtered out from the data, we need to investigate why
`CIHISPEED` has N/A(GQ) values present. Investigating based on [this](https://forum.ipums.org/t/cihispeed-group-quarters/3453) discussion on
IPUMS forum.

```{r}
#| label: investigate-data-issue-1

# Subset dataframe so that only `CIHISPEED` = 0 values are retained and check
# the values that exist for `CINETHH`
test_cihispeed_gq <- acs5_2021_raw |>
  filter(CIHISPEED==0) |>
  distinct(CIHISPEED, CINETHH)

print(paste0("When `CIHISPEED` has label 'N/A (GQ)' then `CINETHH` takes up ", 
              "following values:"))
print(test_cihispeed_gq)

```


```{r}
#| label: investigate-data-issue-2

# Subset dataframe so that `CIHISPEED` = 0 values are removed and check
# the values that exist for `CINETHH`
test_cihispeed <- acs5_2021_raw |>
  filter(CIHISPEED!=0) |>
  distinct(CIHISPEED, CINETHH)

print(paste0("When `CIHISPEED` does not have label 'N/A (GQ)' then `CINETHH` ", 
              "takes up following values:"))
print(test_cihispeed)

```

The results imply that `CIHISPEED` takes on 'N/A(GQ)' values when `CINETHH` 
indicates either 'No Internet access at this house, apartment, or mobile home' or
'Yes, without a subscription to an Internet Service'. 

`CINETHH` is derived from Question 9 in the ACS listed below:

    9. At this house, apartment, or mobile home - do you or any member of this 
    household have access to the internet?
    [ ] Yes, by paying a cell phone company or Internet service provider
    [ ] Yes, without paying a cell phone company of Internet service provider -> 
        SKIP to question 11
    [ ] No access to the Internet at this house, apartment, or mobile home -> 
        SKIP to question 11

`CIHISPEED` is derived from Question 10, part (b) in the ACS listed below:

    10. Do you or any member of this household have access to the Internet using a -

    b) broadband (high speed) Internet service such as cable, fiber optic, or DSL 
        service installed in this household?
        [ ]Yes
        [ ]No

It is reasonable to assume based on the data values in the IPUMS extract that 
whenever a household skipped to question 11 (i.e. answered access to internet 
without subscription or no subscription for question 9), the corresponding value 
assigned to `CIHISPEED` was 'N/A(GQ)' as for these households the question of
having access to paid high speed broadband was inapplicable. 

This finding tracks with the [IPUMS forum discussion](https://forum.ipums.org/t/cihispeed-group-quarters/3453) indicating that a household
could have 'N/A(GQ)' values for reasons other than being categorized as a group
quarter.

In light of this, our assumption stands that the universe of households to answer
questions regarding digital access comes from the data encapsulated by the 
`CINETHH` variable. This will feed into the denominator to compute share of digital
access in a geography. 

Create `DIGITAL_ACCESS`as a composite binary variable that takes on the value `1` 
when a household has access to high speed internet and either a tablet or laptop 
computer or both.

```{r}
#| label: create-composite-digital-access

# Create `DIGITAL_ACCESS` composite binary variable using `CIHISPEED`, `CILAPTOP`, 
# and `CITABLET`

acs5_2021_raw <- acs5_2021_raw |>
  mutate(
    DIGITAL_ACCESS = case_when(
      ((CILAPTOP == 1 | CITABLET == 1) & CIHISPEED == 10) ~ 1,
      CIHISPEED == 20 ~ 0,
      .default = NA_integer_)
    )

```


```{r}
#| label: universe-for-digital-access

# Run this code only if `N/A (GQ)` label corresponding to '0' value is present in the 
# data for `CINETHH`
# Remove values labelled `N/A (GQ)` as they do not represent households of interest

#acs5_2021_raw |> acs5_2021_raw |>
#  filter(CINETHH!=0)

```


## Merge Crosswalk {#sec-merge-crosswalk}

The 5-yr 2017-21 ACS data from IPUMS needs to be merged with PUMA to county and 
PUMA to place geocodes to get county and place boundaries.

Since, the data IPUMS data is pre-2022, a crosswalk from 2012 PUMA to 2014 county # QUESTION: should this be 2014 or 2016 in the Wiki?)
is used. The weight of each household will be adjusted by the proportion of the 
PUMA area that falls into a given County/Place

::: panel-tabset
### PUMA to County

```{r}
#| label: crosswalk-puma-county-1

# Read the relevant crosswalk file
crosswalk_puma_county <- read_csv((here::here("geographic-crosswalks", 
                                              "data", 
                                              "crosswalk_puma_to_county.csv"))) |>
  # filter the crosswalk for the data period 2017-2021
  filter(crosswalk_period == "pre-2022") |>
  rename(STATE = statefip,
         COUNTY = county,
         PUMA = puma) |>
  # drop Puerto Rico from the crosswalk
  filter(STATE != "72")
```


```{r}
#| label: crosswalk-puma-county-2

# Merge crosswalk PUMA to county with the IPUMS data
acs5_2021_county <- acs5_2021_raw |>
  right_join(crosswalk_puma_county,
             by = c("STATE", "PUMA"),
             relationship = "many-to-many") |>
  # Use county codes from crosswalk if they're not present in the IPUMS data
  mutate(
    COUNTY = if_else(COUNTY.x != "000", COUNTY.x, COUNTY.y),
    HHWT = if_else(COUNTY.x != "000", HHWT, HHWT * afact),
    sum_products = if_else(COUNTY.x != "000", 1, sum_products) #Question - is this right??
  ) |>
  select(-c(COUNTY.x, COUNTY.y)) |>
  mutate(COUNTY = str_pad(COUNTY, width = 3, pad = "0", side = "left")) 

```

### PUMA to Place

```{r}
#| label: crosswalk-puma-place-1

# Read the relevant crosswalk file
crosswalk_puma_place <- read_csv((here::here("geographic-crosswalks", 
                                              "data", 
                                              "crosswalk_puma_to_place.csv"))) |>
  # filter the crosswalk for the data period 2017-2021
  filter(crosswalk_period == "pre-2022") |>
  rename(STATE = statefip,
         PUMA = puma) 
```

```{r}
#| label: crosswalk-puma-place-2

# Merge crosswalk PUMA to county with the IPUMS data
acs5_2021_place <- acs5_2021_raw |>
  right_join(crosswalk_puma_place,
             by = c("STATE", "PUMA"),
             relationship = "many-to-many") |>
  # Use county codes from crosswalk if they're not present in the IPUMS data
  mutate(
    HHWT = HHWT * afact,
  ) 

```

:::

Filter out observations where HHWT is N/A or zero as these will not be counted
in the metric calculation.

```{r}
#| label: clean-up-crosswalked-county

# Remove observations where HHWT is N/A
count(acs5_2021_county, is.na(HHWT))

acs5_2021_county <- acs5_2021_county %>% 
  drop_na(HHWT)

# Remove observations where HHWT is zero
count(acs5_2021_county, HHWT == 0)

acs5_2021_county <- acs5_2021_county %>% 
  filter(HHWT > 0)

```


```{r}
#| label: clean-up-crosswalked-place

# Remove observations where HHWT is N/A
count(acs5_2021_place, is.na(HHWT))

acs5_2021_place <- acs5_2021_place %>% 
  drop_na(HHWT)

# Remove observations where HHWT is zero
count(acs5_2021_place, HHWT == 0)

acs5_2021_place <- acs5_2021_place %>% 
  filter(HHWT > 0)
```

## Create Metric {#sec-create-metric}

Now digital access metric can be created as share of households having access to 
either a tablet or laptop and a high speed internet connection for a given county/place.
However, in an IPUMS sample a household may be representative of many households 
which is gauged through the `HHWT` variable. So, to find the share of households 
with digital access, a weighted measure is needed.

```{r}
#| label: weighted-digital-access

# Create weighted digital access variable

acs5_2021_place <- acs5_2021_place |>
mutate(HHWTD_DIGITAL_ACCESS = if_else(DIGITAL_ACCESS == 1, HHWT, 0))

acs5_2021_county <- acs5_2021_county |>
mutate(HHWTD_DIGITAL_ACCESS = if_else(DIGITAL_ACCESS == 1, HHWT, 0))

```


For quality measures, compute effective sample size count for each county/place
which is the unweighted sum of households present in the data for each geography. 
The effective sample will indicate the actual number of survey records that will
factor into metric calculation as we do not want to include geographies with less
than 30 records for quality purposes. 

```{r}
#| label: create-effective-sample-county

# Create effective sample count variable for county
acs5_2021_county <- acs5_2021_county |>
    group_by(YEAR, STATE, COUNTY) |>
    mutate(
      effective_sample_count = n()
    ) |>
    ungroup()

# Summarize effective sample count information for county
acs5_2021_county |>
  group_by(YEAR, STATE, COUNTY) |>
  summarise(
    effective_sample_count = first(effective_sample_count)
  ) |>
  ungroup() |>
  summarise(
    total_counties = n(),
    effective_30_sample_counties = sum(effective_sample_count > 30)
  )
```


```{r}
#| label: create-effective-sample-place

# Create effective sample count variable for place
acs5_2021_place <- acs5_2021_place |>
    group_by(YEAR, STATE, place) |>
    mutate(
      effective_sample_count = n()
    ) |>
    ungroup()

# Summarize effective sample count information for place
acs5_2021_place |>
  group_by(YEAR, STATE, place) |>
  summarise(
    effective_sample_count = first(effective_sample_count)
  ) |>
  ungroup() |>
  summarise(
    total_places = n(),
    effective_30_sample_places = sum(effective_sample_count > 30)
  )

```


Using the weighted digital access measure, we will compute the county and place 
level metric. The digital access metric will represent the number of households
that have digital access divided by the total number of households for a given 
geography at the county/place level. The numerator represents the households which 
have access to high speed internet and a tablet/laptop computer while the denominator
represents all the households that do (either with subscription or without) and 
do not have access to the internet.

### County

```{r}
#| label: digital-access-county

# Compute share of households with digital access in a given county

share_dig_acc_county <- 
  acs5_2021_county |>
  # Filter counties with less than 30 samples
  filter(effective_sample_count >= 30) |>
  select("YEAR", "STATE", "COUNTY", "HHWTD_DIGITAL_ACCESS", "HHWT",
          "effective_sample_count") |>
  group_by(YEAR, STATE, COUNTY) |>
  summarise(
    # Sum households with digital access, remove households assigned `NA`
    county_dig_acc = sum(HHWTD_DIGITAL_ACCESS, na.rm=TRUE),
    # Total households include all households that do/do not have access
    total = sum(HHWT),
    effective_sample_count = first(effective_sample_count),
    .groups = 'drop'
  ) |>
  # Calculate share of households with digital access
  mutate(
    share_digital_access = county_dig_acc/total
  ) |>
  select(YEAR, STATE, COUNTY, share_digital_access, effective_sample_count) |>
  janitor::clean_names()

```


### County Subgroups: Race-Ethnicity & Income


```{r}
#| label: function-share-digital-access

# Compute share of households with digital access based on geography and subgroup
# Geography can take values "COUNTY" or "place"
# Subgroup can take values "RACE_ETHNICITY" or "INCOME"

compute_share_dig_acc_subgroup <- function(data, geo, subgroup_vars) {
  
  data <- data |>
  select(c("YEAR", "STATE", all_of(geo), "HHWTD_DIGITAL_ACCESS", "HHWT", 
            all_of(subgroup_vars))) |>
  # Group data by geography and subgroup
  group_by(across(c("YEAR", "STATE", all_of(geo), all_of(subgroup_vars)))) |>
  summarise(
    # Sum households with digital access, remove households assigned `NA`
    subgroup_dig_acc = sum(HHWTD_DIGITAL_ACCESS, na.rm=TRUE),
    # Total households include all households that do/do not have access
    total_subgroup = sum(HHWT),
    # Compute subgroup effective sample size,
    effective_sample_count = n(),
    .groups = 'drop'
  ) |>
  # Calculate share of households with digital access
  mutate(
    share_digital_access = subgroup_dig_acc/total_subgroup
  ) |> 
  # Mark digital access as N/A for subgroups with sample size < 30 
  mutate(
   share_digital_access = if_else(effective_sample_count < 30, NA_real_, 
                                                            share_digital_access)
  ) |>
  select(c("YEAR", "STATE", geo, subgroup_vars,"effective_sample_count", 
    "share_digital_access")) |>
  janitor::clean_names()

  return (data)
}

```


```{r}
#| label: digital-access-county-subgroups

# Compute share of households with digital access in a given county by race-ethnicity
share_dig_acc_county_race_ethnicity <- 
    compute_share_dig_acc_subgroup(acs5_2021_county, "COUNTY", "RACE_ETHNICITY")


# Compute share of households with digital access in a given county by income
share_dig_acc_county_income <- 
  compute_share_dig_acc_subgroup(acs5_2021_county, "COUNTY", "INCOME")

```


### Place

```{r}
#| label: digital-access-place

# Compute share of households with digital access in a given place
share_dig_acc_place <- 
  acs5_2021_place |>
  # Filter places with less than 30 samples
  filter(effective_sample_count >= 30) |>
  select("YEAR", "STATE", "place", "HHWTD_DIGITAL_ACCESS", "HHWT",
         "effective_sample_count") |>
  group_by(YEAR, STATE, place) |> 
  summarise(
    # Sum households with digital access, remove households assigned `NA`
    place_dig_acc = sum(HHWTD_DIGITAL_ACCESS, na.rm=TRUE),
    # Total households include all households that do/do not have access
    total = sum(HHWT),
    effective_sample_count = first(effective_sample_count),
    .groups = 'drop'
  ) |>
  # Calculate share of households with digital access
  mutate(
    share_digital_access = place_dig_acc/total
  ) |>
  select(YEAR, STATE, place, share_digital_access, effective_sample_count) |>
  janitor::clean_names()

```


### Place Subgroups: Race-Ethnicity & Income

```{r}
#| label: digital-access-place-subgroups

# Compute share of households with digital access in a given place by race-ethnicity
share_dig_acc_place_race_ethnicity <- 
    compute_share_dig_acc_subgroup(acs5_2021_place, "place", "RACE_ETHNICITY")

  
# Compute share of households with digital access in a given place by income
share_dig_acc_place_income <- 
    compute_share_dig_acc_subgroup(acs5_2021_place, "place", "INCOME")

```


## Explore Metric {#sec-evaluate-metric}

```{r} 

#| label: metric-range-checks

# Check the range of digital access metric is between 0 and 1

all_df <- list(share_dig_acc_county, share_dig_acc_county_income, 
                share_dig_acc_county_race_ethnicity, share_dig_acc_place,
                share_dig_acc_place_income, share_dig_acc_place_race_ethnicity)

# Find min and max values for digital access metric
max_vals <- sapply(all_df, function(x) max(x$share_digital_access, na.rm=TRUE))
min_vals <- sapply(all_df, function(x) min(x$share_digital_access, na.rm=TRUE))

# All minimum metric values should be less than or equal to 1 and greater than or
# equal to 0
min_max_check <- data.frame(
    df_name = c('share_dig_acc_county', 'share_dig_acc_county_income', 
                'share_dig_acc_county_race_ethnicity', 'share_dig_acc_place',
                'share_dig_acc_place_income', 'share_dig_acc_place_race_ethnicity'),
    min_greater_than_equal_zero = sapply(min_vals, function(x) x >= 0),
    max_less_than_equal_one = sapply(max_vals, function(x) x <= 1)

)

print(min_max_check)

rm(all_df)
```


To get a better sense of the distribution of the digital access metric, we visualize
it for both county and place.


```{r} 

#| label: metric-visualization-function

metric_visualization_subgroup <- function(data, subgroup, subtitle) {
  
  data <- data |> drop_na()

  data |>
    ggplot(aes(x = share_digital_access)) +
    geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 0.02) +
    facet_wrap(.~ data[[subgroup]]) +
    labs(
      title = paste0("Share of Digital Access by ", 
                      str_to_title(gsub("_", " and ", subgroup))),
      subtitle = subtitle,
      y = "Count",
      x = "Share of Digital Access"
    )
}

```


::: panel-tabset

### Visualizing Digital Access for County by Race and Ethnicity

```{r}

metric_visualization_subgroup(
  share_dig_acc_county_race_ethnicity, 
  "race_ethnicity",
  "County")

quantile(share_dig_acc_county_race_ethnicity$share_digital_access, na.rm=TRUE)
skimr::skim(share_dig_acc_county_race_ethnicity)

```


### Visualizing Digital Access for County by Income

```{r}

metric_visualization_subgroup(
  share_dig_acc_county_income, 
  "income",
  "County")

quantile(share_dig_acc_county_income$share_digital_access, na.rm=TRUE)
skimr::skim(share_dig_acc_county_income)

```

### Visualizing Digital Access for Place by Race and Ethnicity

```{r}

metric_visualization_subgroup(
  share_dig_acc_place_race_ethnicity, 
  "race_ethnicity",
  "Place")

quantile(share_dig_acc_place_race_ethnicity$share_digital_access, na.rm=TRUE)
skimr::skim(share_dig_acc_place_race_ethnicity)

```

### Visualizing Digital Access for Place by Income

```{r}

metric_visualization_subgroup(
  share_dig_acc_place_income, 
  "income",
  "Place")

quantile(share_dig_acc_place_income$share_digital_access, na.rm=TRUE)
skimr::skim(share_dig_acc_place_income)

```

:::

## Metric Quality {#sec-metric-quality}

Determining the quality of the digital access metric is a two step process that
considers an effective sample test and geographic allocation test. 

- Effective sample test entails that for the IPUMS micro data each geography must 
have a sample of at least 30 survey responses otherwise we report the metric as
NA for that geography. This is useful because counties/places with an extremely 
small number of households sampled would have a poorer data quality. 

- Geographic allocation test leverages the project crosswalks (either from PUMA 
to county or PUMA to place for IPUMS data) which include a measure for quality 
based on the area of the destination geography that falls within the IPUMS geography.
For geographies that have a large number of households, this helps figure out how
much of the data associated with a county/place actually came from that county/place.
The measure for this comes from the `sum_products` variable in the crosswalks
(for more information see ["Generate Puma to County Crosswalk"](geographic-crosswalks/generate_puma_county_crosswalks.qmd) notebook).


Combining the above two guidelines for quality testing, the following measure of
digital access quality is determined:

 - Data quality is Strong (`1`): When number of households in a geography is 
    greater than or equal to 30 and 75% or more of the observations are from 
    the county/place.
 
 - Data quality is Adequate (`2`): When number of households in a geography is 
   greater than or equal to 30 but only 35% to 74% of the observations are from 
   the county/place.
  
 - Data quality is Weak (`3`): When number of households in a geography is 
   greater than or equal to 30 but less than 35% of the observations are from 
   the county/place.
  
 - Data quality is NA: When number of households in a geography is less than 30.
 
 USE CINETHH for DENOMINATOR being the universe and then 9 and 11 correspond to numerator. get rid of NA based on CINETHH
 Put some language together to add to the Git issue on this decision
 
 Check whether each unique county/place has a singular `sum_products` value associated.
 
```{r}
#| label: check-sum-products-county

# Check whether each unique combination of state and county has the same `sum_products`
# value associated

# For county

# In this data, each county is observed to have one `sum_products` value which 
# checks out with the logic of crosswalks
quality_county <- acs5_2021_county |>
      select("YEAR", "STATE", "COUNTY", "sum_products") |>
      group_by(YEAR, STATE, COUNTY) |>
      summarise(
          distinct_sum_products = n_distinct(sum_products),
          sum_products = list(unique((sum_products)))
        ) |>
      ungroup() |>
      janitor::clean_names()

print("Are there any counties with multiple `sum_products` values?")
any(quality_county$distinct_sum_products > 1)
```


```{r}
#| label: check-sum-products-place

# For place

# In this data, each place is observed to have one `sum_products` value which 
# checks out with the logic of crosswalks
quality_place <- acs5_2021_place |>
      select("YEAR", "STATE", "place", "sum_products") |>
      group_by(YEAR, STATE, place) |>
      summarise(
          distinct_sum_products = n_distinct(sum_products),
          sum_products = list(unique((sum_products)))
        ) |>
      ungroup() |>
      janitor::clean_names()

print("Are there any places with multiple `sum_products` values?")
any(quality_place$distinct_sum_products > 1)

```


Putting together the `sum_products` and effective sample size values together
will help create the digital access quality variable.

```{r}
#| label: function-data-quality-metric

create_quality_metric <- function(metric_df, quality_df, geo) {

    # Create data quality metric
    data <- metric_df |>
              right_join(quality_df,
                          by = c("year", "state", geo)) |>
              select(-c("distinct_sum_products")) |>
              mutate(quality_digital_access = case_when(
              (effective_sample_count < 30) ~ NA_integer_,
              (effective_sample_count >=30 & sum_products >= 0.75) ~ 1,
              (effective_sample_count >= 30 & sum_products >= 0.35) ~ 2,
              (effective_sample_count >= 30 & sum_products < 0.35) ~ 3,
              .default = NA_integer_
              ))
    
    return (data)
}

```


```{r}
#| label: create-data-quality-variable

# Add data quality variable to digital access metric 

# For counties
share_dig_acc_county <- create_quality_metric(share_dig_acc_county, 
                                              quality_county, "county")

share_dig_acc_county_income <- create_quality_metric(share_dig_acc_county_income, 
                                                      quality_county, "county")

share_dig_acc_county_race_ethnicity <- create_quality_metric(share_dig_acc_county_race_ethnicity, 
                                                      quality_county, "county")

# For places
share_dig_acc_place <- create_quality_metric(share_dig_acc_place, 
                                              quality_place, "place")

share_dig_acc_place_income <- create_quality_metric(share_dig_acc_place_income, 
                                              quality_place, "place")

share_dig_acc_place_race_ethnicity <- create_quality_metric(share_dig_acc_place_race_ethnicity, 
                                              quality_place, "place")

```


## Evaluate Metric {#sec-evaluate-metric}

Check number of unique counties and places in the crosswalks.

```{r}

# Unique counties in PUMA to county crosswalk
n_distinct(crosswalk_puma_county[c("STATE", "COUNTY")])

# We observe 3143 unique counties in the crosswalk
counties_pre_2022 <- crosswalk_puma_county |>
                janitor::clean_names() |>
                distinct(state, county) |>
                arrange(state, county)

# Unique counties in PUMA to place crosswalk
n_distinct(crosswalk_puma_place[c("STATE", "place")])

# We observe 485 unique places in the crosswalk
places_pre_2022 <- crosswalk_puma_place |>
                janitor::clean_names() |>
                distinct(state, place) |>
                arrange(state, place)
```

Check whether counties present during digital access metric calculation for county 
and place are the same ones present in the crosswalks.

```{r} 
#| label: function-geography-equivalence

check_geography_equivalence <- function(crosswalk, metric, geo) {
  return (assertthat::assert_that(
      all.equal(
            crosswalk,
            metric |>
                distinct(across(c("state", geo))) |>
                arrange(across(c("state", geo)))
    )
 ))
}

```


```{r} 
#| label: check-missing-counties

print("Are all 3143 counties present for overall county metric?")
check_geography_equivalence(counties_pre_2022, share_dig_acc_county, "county")

print("Are all 3143 counties present for race-ethnicity subgroup county metric?")
check_geography_equivalence(counties_pre_2022, share_dig_acc_county_race_ethnicity
                            , "county")

print("Are all 3143 counties present for income subgroup county metric?")
check_geography_equivalence(counties_pre_2022, share_dig_acc_county_income
                            , "county")

print("Are all 485 places present for overall place metric?")
check_geography_equivalence(places_pre_2022, share_dig_acc_place, "place")

print("Are all 485 places present for race-ethnicity subgroup place metric?")
check_geography_equivalence(places_pre_2022, share_dig_acc_place_race_ethnicity
                            , "place")

print("Are all 485 places present for income subgroup place metric?")
check_geography_equivalence(places_pre_2022, share_dig_acc_place_income
                            , "place")

```

The above results suggest that all of the unique counties/places from the crosswalks 
are present in the digital access metric calculation.


Check whether required number of observations exist for each subgroup associated
with county and place digital access metric.

```{r}

# For county race-ethnicity subgroup there should be 3143*4 observations

# We observe missingness for some counties for the `Black, Non-Hispanic` subgroup type
assertthat::are_equal((3143 * 4), nrow(share_dig_acc_county_race_ethnicity))
share_dig_acc_county_race_ethnicity |> 
      group_by(race_ethnicity) |> 
      summarise(n = n())

# For county income subgroup there should be 3143*2 observations
assertthat::are_equal((3143 * 2), nrow(share_dig_acc_county_income))
share_dig_acc_county_income |> 
      group_by(income) |> 
      summarise(n = n())

# For place race-ethnicity subgroup there should be 485*4 observations
assertthat::are_equal((485 * 4), nrow(share_dig_acc_place_race_ethnicity))
share_dig_acc_place_race_ethnicity |> 
      group_by(race_ethnicity) |> 
      summarise(n = n())

# For place income subgroup there should be 485*2 observations
assertthat::are_equal((485 * 2), nrow(share_dig_acc_place_income))
share_dig_acc_place_income |> 
      group_by(income) |> 
      summarise(n = n())
```

The above results suggest that there is missingness in digital access metric observation
for `Black, Non-Hispanic` race-ethnicity subgroup data for some counties. Expected 
number of observations are present in the data for county income subgroup, place 
race-ethnicity subgroup, and place income subgroup.

```{r}

# Find counties missing in the race-ethnicity subgroup data set

missing_counties <- anti_join(counties_pre_2022, 
                              share_dig_acc_county_race_ethnicity |>
                                filter(race_ethnicity == "Black, Non-Hispanic"), 
                              by = c("state", "county"))

```


Add missing counties back to the metric data.

```{r}

# Add rows corresponding to missing counties for the county `Black, Non-Hispanic` 
# race-ethnicity subgroup

# Create data with missing county data for the subgroup with N/A values for metric
missing_data <- data.frame(
                  year = 2021,
                  state = missing_counties$state,
                  county = missing_counties$county,
                  race_ethnicity = "Black, Non-Hispanic",
                  effective_sample_count = NA,
                  share_digital_access = NA_real_,
                  sum_products = NA,
                  quality_digital_access = NA_integer_
                )

# Add the missing data to applicable subgroup data set 
share_dig_acc_county_race_ethnicity <-
    rbind(share_dig_acc_county_race_ethnicity, missing_data)

# Check data is not missing anymore
assertthat::are_equal((3143 * 4), nrow(share_dig_acc_county_race_ethnicity))
share_dig_acc_county_race_ethnicity |> 
      group_by(race_ethnicity) |> 
      summarise(n = n())

check_geography_equivalence(counties_pre_2022, 
                            share_dig_acc_county_race_ethnicity, 
                            "county")
```


## Final Data {#sec-final-data}

The final output files will include digital access metric calculated for county 
and place at the overall and subgroup level.

The subgroups level data files should include a row indicating the metric value
for the overall population.

```{r} 
#| label: function-overall-subgroup-row

# Add row indicating `all` metric value corresponding to entire population for
# subgroup datasets
# In this function `subgroup` can take values "race_ethnicity" and "income"
# `overall_df` corresponds to data frame with the digital access metric for all 
# the counties and `subgroup_df` corresponds to the data frame  with the digital 
# access metric for county subgroups

create_all_subgroup <- function(overall_df, subgroup_df, subgroup) {
  
  # Add "All" as subgroup in the overall population metric data
  overall_df <- overall_df |>
                mutate(!!subgroup := "All")
  
  # Re-arrange metric columns based on subgroup data
  overall_df <- overall_df[, names(subgroup_df)]
  
  # Combine overall and subgroup data
  subgroup_df <- rbind(subgroup_df, overall_df)
  
  return (subgroup_df)
}

```


```{r} 
#| label: add-overall-subgroup-row-county

# For county

# Add `All` metric value to subgroup data sets
share_dig_acc_county_race_ethnicity <- create_all_subgroup(
                                              share_dig_acc_county,
                                              share_dig_acc_county_race_ethnicity,
                                              "race_ethnicity")

share_dig_acc_county_income <- create_all_subgroup(
                                              share_dig_acc_county,
                                              share_dig_acc_county_income,
                                              "income")

# Check all counties are present in updated subgroups data
check_geography_equivalence(counties_pre_2022, 
                            share_dig_acc_county_race_ethnicity |> 
                                filter(race_ethnicity == "All"),
                            "county")

check_geography_equivalence(counties_pre_2022, 
                            share_dig_acc_county_income |> 
                                filter(income == "All"),
                            "county")
```


```{r} 
#| label: add-overall-subgroup-row-place
# For place

# Add `All` metric value to subgroup data sets
share_dig_acc_place_race_ethnicity <- create_all_subgroup(
                                              share_dig_acc_place,
                                              share_dig_acc_place_race_ethnicity,
                                              "race_ethnicity")

share_dig_acc_place_income <- create_all_subgroup(
                                              share_dig_acc_place,
                                              share_dig_acc_place_income,
                                              "income")

# Check all places are present in updated subgroups data
check_geography_equivalence(places_pre_2022, 
                            share_dig_acc_place_race_ethnicity |> 
                                filter(race_ethnicity == "All"),
                            "place")

check_geography_equivalence(places_pre_2022, 
                            share_dig_acc_place_income |> 
                                filter(income == "All"),
                            "place")
```

The final data files need to be created according to UMF guidelines mentioned
in the GitHub Wiki:

[2_Data Structure and Expectations](https://github.com/UI-Research/mobility-from-poverty/wiki/2_Data-Structure-and-Expectations%E2%80%AF)
[9_Final File Output](https://github.com/UI-Research/mobility-from-poverty/wiki/9_Final-File-Output)


```{r}
#| label: function-final-data-export

# Function to create the final data csv file
# `subgroups` is a indicator variable for the presence of subgroup in the final data
# `subgroup_var` is a character variable for the name of the subgroup in the data 
# and can take values "race_ethnicity" and "income", defaults to NA
# `geo` indicates whether data is for "county" or "place"

create_final_data_file <- function(data, subgroups, subgroup_var = NA, geo) {

    if (!subgroups) {
        
      data <- data |> 
                # Order the data
                arrange(year, state, geo) |>
                # Rename variables
                rename(share_digital_access_quality = quality_digital_access) |>
                select(year, state, geo, share_digital_access, share_digital_access_quality)
      
      file_name <- paste0("digital_access_", geo, "_all_longitudinal_", 
                            data$year[1], ".csv")
      }
      
    else {

      data <- data |> 
                # Order the data
                arrange(year, state, geo, subgroup_var) |>
                # Rename variables
                rename(
                        share_digital_access_quality = quality_digital_access,
                        subgroup = {{subgroup_var}}
                ) |>
                # Create subgroup_type variable indicating overall & subrgoup metric
                mutate( subgroup_type = 
                          case_when(
                                    (subgroup == "All") ~ "all",
                                    TRUE ~ as.character({{subgroup_var}}) )
                ) |>
                  select(year, state, geo, subgroup, subgroup_type, 
                          share_digital_access, share_digital_access_quality)

      file_name <- paste0("digital_access_", geo, "_", subgroup_var, 
                            "_longitudinal_", data$year[1], ".csv")
    }
  
  # Define file path to save the final data file
  file_path <- here::here("08_education", "data", "final", file_name)
  
  # Save the csv file
  write.csv(data, file_path, row.names = FALSE)
 }

```


```{r}
#| label: generate-final-data

# For county
create_final_data_file(share_dig_acc_county, FALSE, geo = "county")
create_final_data_file(share_dig_acc_county_race_ethnicity, TRUE, 
                        "race_ethnicity", "county")
create_final_data_file(share_dig_acc_county_income, TRUE, "income", "county")

# For place
create_final_data_file(share_dig_acc_place, FALSE, geo = "place")
create_final_data_file(share_dig_acc_place_race_ethnicity, TRUE, 
                        "race_ethnicity", "place")
create_final_data_file(share_dig_acc_place_income, TRUE, "income", "place")

```

After generating the final data files in folder `08_Education/data/final`, 
we need to evaluate the outputs using the `evaluate_final_data` function per 
UMF [guidelines](https://github.com/UI-Research/mobility-from-poverty/wiki/8_Final-Data-Expectations-Form%E2%80%AF).

```{r, eval=FALSE}
#| label: final-data-evaluation-county

# For county files

overall_county <- read_csv(
        here::here("08_Education/data/final/digital_access_county_all_longitudinal_2021.csv")
        ,show_col_types = FALSE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_digital_access_overall_county.csv",
    data = overall_county, geography = "county",
    subgroups = FALSE, confidence_intervals = FALSE)




```

