---
title: "Exploring IPUMS Data to Construct a Measure for Digital Access"
author: "Ridhi Purohit"
date: now
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
    code-fold: show
    code-tools: true
editor_options: 
  chunk_output_type: console
execute:
  warning: false
---

```{=html}
<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>
```
<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

This document investigates and programs the **Digital Access** variable which encapsulates households with the presence of a computer (laptop, desktop, notebook computer, tablet, or portable wireless computer) and access to broadband (high speed) internet service. In this program, data from 2017-21 ACS 5 year sample available through IPUMS is used to investigate the potential of creating a digital access variable by combining ACS computer and internet use variables from IPUMS.

## Process

-   [Housekeeping](#housekeeping)
-   [Get IPUMS Data](#get-ipums-data)
-   [Format Data](#format-data)

## Housekeeping {#housekeeping}

```{r}
#| label: setup

options(scipen = 999)
librarian::shelf(
  tidyverse,
  ipumsr,
  here,
  UrbanInstitute / urbnthemes,
  tidylog,
  warn.conflicts = FALSE
  )
set_urbn_defaults(style = "print")

```

**Instructions to run this code book**

-   This Quarto markdown document uses the `here` R library to set and access file paths. Please check your root directory by executing `here::here()` to verify that it points to the root folder of the `mobility-from-poverty` GitHub repository.
-   To avoid manually setting up file paths, open the GitHub repository as an R project by leveraging the `mobility-from-poverty.Rproj` file present in the root folder of the `mobility-from-poverty` GitHub repository when you clone it.
-   Un-comment all of the code in @lst-query-ipums-api code block (lines 70 through 105) when you run the code in this file for the first time to download the IPUMS data.
-   In @lst-load-ipums code block, you'd be asked to update the microdata file name (line number 126) that you download through the @lst-query-ipums-api block, so please make a note of it.

## Get IPUMS Data {#get-ipums-data}

This code book uses the 2017-2021 5-year ACS extract from the IPUMS API using R package `ipumsr`. To access this data, an IPUMS API key is required. This can be obtained by registering on the [IPUMS USA](https://uma.pop.umn.edu/usa/user/new?return_url=https%3A%2F%2Fusa.ipums.org%2Fusa-action%2Fmenu) website and creating an API key. For more information see the `ipumsr` [getting started vignette](https://tech.popdata.org/ipumsr/articles/ipums.html).

```{r}
#| label: query-ipums-api
#| lst-label: lst-query-ipums-api
#| lst-cap: Query IPUMS API

# Run the code in this block the first time this notebook is executed to get the 
# data and avoid repeated downloads 

# define variables to be queried from IPUMS API

# NOTE: running var_spec("CIHISPEED", data_quality_flags = TRUE) produced following error: Error in `ipums_api_request()`: 
# ! API request failed with status 400.
# ✖ CIHISPEED has data quality flags, but none are available in the selected samples

# acs_vars <- list("YEAR", "MULTYEAR", "SAMPLE", "SERIAL", "CBSERIAL", "HHWT", 
#              "CLUSTER", "STRATA", "GQ", "REGION", "STATEFIP", "COUNTYFIP", "PUMA",
#              "HHINCOME", var_spec("CILAPTOP", data_quality_flags = TRUE), 
#              var_spec("CITABLET", data_quality_flags = TRUE),
#              "CIHISPEED", "CINETHH",
#              "PERNUM", "PERWT", "RACE", "HISPAN")


# define the IPUMS extract
# acs_2021_extract <- ipumsr::define_extract_micro(
#  collection = "usa",
#  description = "2017-2021 ACS 5-yr Sample",
#  samples = "us2021c",
#  variables = acs_vars)

# submit and wait for IPUMS extract
# acs_ext_submit <- ipumsr::submit_extract(acs_2021_extract)
# acs_ext_complete <- ipumsr::wait_for_extract(acs_ext_submit)

# check if data download directory is present
fs::dir_create(here("08_education/data/raw"))

#download IPUMS extract
# data_path <- ipumsr::download_extract(
#  acs_ext_complete, 
#  download_dir = here("08_education/data/raw")
#  )

```

## Format Data {#format-data}

Now, the downloaded data needs to be cleaned and formatted so that it can be used further to generate the metric. The 2017-2021 5-year ACS microdata extract for computer and internet use is in the form of household level data set with person level variables added for this analysis.

### Clean Household-Level Data

To analyze the data at the household level, only the first person sampled in the household is retained in the data.

```{r}
#| label: load-ipums
#| lst-label: lst-load-ipums
#| lst-cap: Load IPUMS Data

# Manually set the file name downloaded in the previous code chunk which is 
# 'usa_00018.xml' for me but will have some different value for whoever re-runs this

micro_filepath <- "usa_00018.xml"

# load acs data and filter for household level analysis
acs5_2021_raw <- ipumsr::read_ipums_micro(
  here(paste0("08_education/data/raw/", micro_filepath))
  ) |>
  filter(PERNUM == 1)

```

The IPUMS data consists of labelled vectors ([ipumsr](https://tech.popdata.org/ipumsr/articles/value-labels.html#remove-unused-value-labels)). We find the missing labels using the `ipums_var_info(acs5_2021_raw)` and inspecting the `val_labels` column.

For the "Total household income" i.e. `HHINCOME` column, "N/A" values are denoted by `9999999` which needs to be updated before proceeding.

```{r}
#| label: handle-missingness

# re-label `HHINCOME` when value is 9999999 denoting "N/A "

acs5_2021_raw <- acs5_2021_raw |>
  mutate(
    HHINCOME = lbl_na_if(HHINCOME, ~ .lbl %in% c("N/A "))
  )

```

Since the digital access metric is at the household level, individuals who live in group quarters need to be excluded from the data. For more information, see the [variable description](https://usa.ipums.org/usa-action/variables/GQ#codes_section) on IPUMS USA.

Running `ipums_var_info(acs5_2021_raw)` shows the value labels for the group quarters variable `GQ`:

val lbl\
0 Vacant unit\
1 Households under 1970 definition\
2 Additional households under 1990 definition 3 Group quarters--Institutions\
4 Other group quarters\
5 Additional households under 2000 definition 6 Fragment

```{r}
#| label: remove-group-quarters

# Only retain people who live in housing units in the data
acs5_2021_raw <- acs5_2021_raw |>
  filter((GQ %in% c(1,2,5)))

# Check if only required GQ values remain in data
assertthat::assert_that(all(acs5_2021_raw$GQ %in% c(1, 2, 5)))
```

FIPS codes require re-formatting by padding with leading zeros so that state codes are two digits, county codes are three digits, and place codes are five digits.

```{r}
#| label: format-fips-codes

acs5_2021_raw <-
  acs5_2021_raw |>
  mutate(
    across(c(STATEFIP, COUNTYFIP, PUMA), as.character),
    STATEFIP = str_pad(STATEFIP, width = 2, pad = "0", side = "left"),
    COUNTYFIP = str_pad(COUNTYFIP, width = 3, pad = "0", side = "left"),
    PUMA = str_pad(PUMA, width = 5, pad = "0", side = "left")
  ) |>
  rename(
    STATE = STATEFIP,
    COUNTY = COUNTYFIP
  ) |>
  # Reformat the YEAR variable
  mutate(YEAR = as.integer(YEAR))
```

### Computer and Internet Use Variables

-   For measuring digital access, a composite variable `DIGITAL_ACCESS` is created using `CIHISPEED`, `CILAPTOP`, and `CITABLET` IPUMS variables.

- `CILAPTOP` indicates whether the respondent or any member of their household owned or used a desktop, laptop, netbook, or notebook computer. 

- `CITABLET` indicates whether the respondent or any member of their household owned or used a tablet or other portable wireless computer.

-   `CINETHH` is derived from Question 9 in the ACS listed below:

```         
  9. At this house, apartment, or mobile home - do you or any member of this 
      household have access to the internet?
    [ ] Yes, by paying a cell phone company or Internet service provider
    [ ] Yes, without paying a cell phone company of Internet service provider -> 
    SKIP to question 11
    [ ] No access to the Internet at this house, apartment, or mobile home -> 
    SKIP to question 11
```

-   `CIHISPEED` is derived from Question 10, part (b) in the ACS listed below:

```         
  10. Do you or any member of this household have access to the Internet using a -

  b) broadband (high speed) Internet service such as cable, fiber optic, or DSL 
      service installed in this household?
      [ ]Yes
      [ ]No
```

We first investigate the unique values and labels that exist for the computer and internet use variables in the IPUMS data.

```{r}
#| label: data-exploration

# Find unique values and corresponding label values in the data

# In this data, we only observe values "1" and "2" corresponding to labels
# "Yes" and "No" respectively for `CITABLET`
unique(acs5_2021_raw$CITABLET)


# In this data, we only observe values "1" and "2" corresponding to labels
# "Yes" and "No" respectively for `CILAPTOP`
unique(acs5_2021_raw$CILAPTOP)


# In this data, we only observe values "0", "10" , and "20" corresponding to labels
# "N/A (GQ)", "Yes (Cable modem, fiber optic or DSL service)", and "No" 
# respectively for `CIHISPEED`

#### NOTE: It seems like "N/A (GQ)" value exists in the data even after group quarters 
# are removed from the data. Casual inspection of data frame suggests that even 
# if a household is classified as a housing unit, it might still have value "N/A (GQ)"
# in the `CIHISPEED` variable.
unique(acs5_2021_raw$CIHISPEED)


# In this data, we observe values "1", "2", and "3" corresponding to labels 
# 'Yes, with a subscription to an Internet Service', 'Yes, without a subscription 
# to an Internet Service' and 'No Internet access at this house, apartment, or 
# mobile home' for `CINETHH`
unique(acs5_2021_raw$CINETHH)
```

In the data, it seems that `CIHISPEED` has N/A(GQ) values present. This needs to be investigated because group quarters were filtered out from the data so it is not clear why the `CIHISPEED` variable would have `N/A(GQ)` values. Investigating based on [this](https://forum.ipums.org/t/cihispeed-group-quarters/3453) discussion on IPUMS forum.

> We consider the `CINETHH` variable to find the universe of observations for households with digital access. This variable indicates whether any member of a household has access to the internet (by either paying or without paying a cellphone company) or nobody in the household has access to the internet.

Since, `CINETHH` is the universe of observations, first, let's observe the relationship between `CINETHH` and `CIHISPEED` in relation to the `CIHISPEED` variable taking the value `0` corresponding to label `N/A(GQ)`.

```{r}
#| label: investigate-data-issue-1

# Subset dataframe so that only `CIHISPEED` = 0 values are retained and check
# the values that exist for `CINETHH`

print(paste0("When `CIHISPEED` equals `0` i.e. 'N/A (GQ)' then `CINETHH` takes up ", 
              "following values:"))

acs5_2021_raw |>
  filter(CIHISPEED==0) |>
  distinct(CIHISPEED, CINETHH)
```


```{r}
#| label: investigate-data-issue-2

# Subset dataframe so that `CIHISPEED` = 0 values are removed and check
# the values that exist for `CINETHH`

print(paste0("When `CIHISPEED` does not equal `0` i.e. 'N/A (GQ)' then `CINETHH` ", 
              "takes up following values:"))
acs5_2021_raw |>
  filter(CIHISPEED!=0) |>
  distinct(CIHISPEED, CINETHH)
```

```{r}
# Assert that when `CIHISPEED` == 0 then there aren't any instances where `CINETHH` is not equal to
# either 3 or 2
assertthat::assert_that(!any(
                             acs5_2021_raw$CIHISPEED==0 & 
                             !(acs5_2021_raw$CINETHH %in% c(3, 2))  
                             ))

```

```{r}
# Assert that when `CIHISPEED` != 0 then there aren't any instances where `CINETHH` is not equal to 1
assertthat::assert_that(!any(
                             acs5_2021_raw$CIHISPEED!=0 & 
                             !(acs5_2021_raw$CINETHH %in% 1)  
                             ))

```

The above results imply that `CIHISPEED` takes on 'N/A(GQ)' value when `CINETHH` indicates either 'No Internet access at this house, apartment, or mobile home' or 'Yes, without a subscription to an Internet Service'.

It is reasonable to assume based on the data values in the IPUMS extract that whenever a household skipped to question 11 while answering question 9 referenced earlier (i.e. answered access to internet without subscription or no subscription for question 9), the corresponding value assigned to `CIHISPEED` was 'N/A(GQ)' as for these households the question of having access to paid high speed broadband was inapplicable.

This finding tracks with the [IPUMS forum discussion](https://forum.ipums.org/t/cihispeed-group-quarters/3453) indicating that a household could have 'N/A(GQ)' values for reasons other than being categorized as a group quarter.

In light of this, our assumption stands that the universe of households to answer questions regarding digital access comes from the data encapsulated by the `CINETHH` variable. This will feed into the denominator to compute share of digital access in a geography. The target households are those which answered Question 9 in the ACS. We will remove any households that have `CINETHH` equals `0` i.e. `N/A(GQ)` as they will not be a part of our target universe of households. 

```{r}
#| label: universe-for-digital-access

# Remove `0` values labelled `N/A (GQ)` for `CINETHH` as they do not represent households of interest

# Check if any `0` values exist in `CINETHH`
if ( any(acs5_2021_raw$CINETHH == 0) ) {
  
  acs5_2021_raw <- acs5_2021_raw |> 
                                 filter(CINETHH!=0)
} else {
  print("Variable `CINETHH` represents households of interest.")
}

```

### Digital Access

Create `DIGITAL_ACCESS`as a composite binary variable that takes on the value `1` when a household has access to high speed internet and either a tablet or laptop computer or both.

For computing `DIGITAL_ACCESS`, we consider that access relies on whether `CIHISPEED` indicates household has access to high speed internet or not. If `CIHISPEED` has value `N/A(GQ)` after filtering for group quarters, we assume that the corresponding `DIGITAL_ACCESS` for those households would be `N/A` as they did not answer question 10 and instead skipped to Question 11.

```{r}
#| label: create-composite-digital-access

# Create `DIGITAL_ACCESS` composite binary variable using `CIHISPEED`, `CILAPTOP`, 
# and `CITABLET`

acs5_2021_raw <- acs5_2021_raw |>
  mutate(
    DIGITAL_ACCESS = case_when(
      ((CILAPTOP == 1 | CITABLET == 1) & CIHISPEED == 10) ~ 1,
      CIHISPEED == 20 ~ 0,
      .default = NA_integer_)
    )

skimr::skim(acs5_2021_raw$DIGITAL_ACCESS)

```

