---
title: "Debt in Collections City-Level Shares"
author: "Manuel Alcalá Kovalski"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

## Overview

This document processes city-level debt in collections data from the 
Urban Institute's [Financial Health & Wealth dashboard](https://apps.urban.org/features/financial-health-wealth-dashboard/?location=4803000&year=2021&metric=delinquent_debt&subgroup=Residents+of+color)  for 2019, 2021, 2022, and 2023. 

The processed data is used for the Upward Mobility Metrics Dashboard.

```{r setup}
library(tidyverse)
library(readxl)
library(here)
library(tigris)
# Set output path
output_path <- here("04_financial-security/final")
```

Get state abbreviations & FIPS codes using `tigris` and place FIPS codes from IPUMS
```{r}
state_fips <- fips_codes |>
    distinct(state, state_code, state_name) |>
    rename(stusab = state, state = state_code)


temp_file_ipums <- tempfile(fileext = ".xlsx")
download.file(
    url = "https://usa.ipums.org/usa/resources/volii/large_place_PUMA2010_match_summary.xlsx",
    destfile = temp_file_ipums,
    mode = "wb"
)


place_fips <- read_excel(temp_file_ipums) |>
    rename(
        place_fips = `Place FIPS Code`,
        city = `IPUMS CITY Label`
    ) |>
    # Fix specific city labels to match dashboard data
    mutate(
        city = case_when(
            `State Name` == "Arizona" & `Census Place Name` == "Gilbert town" ~ "Gilbert, AZ",
            `State Name` == "Hawaii" & `Census Place Name` == "Urban Honolulu CDP" ~ "Honolulu, HI",
            `State Name` == "Kansas" & `Census Place Name` == "Overland Park city" ~ "Overland Park, KS",
            `State Name` == "Nevada" & `Census Place Name` == "Paradise CDP" ~ "Paradise, NV",
            `State Name` == "Arizona" & `Census Place Name` == "Scottsdale city" ~ "Scottsdale, AZ",
            `State Name` == "Nevada" & `Census Place Name` == "Sunrise Manor CDP" ~ "Sunrise Manor, NV",
            TRUE ~ city
        )
    ) |>
    filter(city != "[Not identified in any sample]") |>
    select(place_fips, city)

```

## Import and process Urban Data Catalog file 

Download and read Excel file from Financial Health & Wealth Dashboard

```{r}
overall_debt_city <-
  read_csv("https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/related-documents/city_2019_23.csv") |> 
  filter(!is.na(state)) |> 
  select(year,
         city = city_name,
         stusab = state,
         subgroup,
         share_debt_coll = delinquent_debt) |> 
  filter(subgroup %in% c("Overall", "Residents of color", "White")) |> 
     mutate(
        subgroup = case_match(
            subgroup,
            "Overall" ~ "All",
            "Residents of color" ~ "Majority Non-White",
            "White" ~ "Majority White"
        ),
        subgroup_type = "race-ethnicity",
        # Handle missing/suppressed values
        share_debt_coll = if_else(
            share_debt_coll == "NA",
            NA_real_,
            share_debt_coll
        ),
        share_debt_coll = as.numeric(share_debt_coll),
        # Create data quality field: 1 = valid, 3 = suppressed/missing
        share_debt_coll_quality = if_else(
            is.na(share_debt_coll),
            3L,
            1L
        )
    ) |> 
     # Merge state FIPS codes
    left_join(state_fips, by = "stusab") |>
    select(-stusab) |>
    # Merge place FIPS codes
    left_join(place_fips, by = "city") |>
    select(
        year, state, place = place_fips, subgroup_type, subgroup,
        share_debt_coll,
        share_debt_coll_quality, 
    ) |>
    arrange(year, state, place, subgroup) |> 
    # There are 63 cities in the dashboard data that are not present
    # in the IPUMS place FIP codes files
    filter(!is.na(place))
  
```


## Export data

```{r}
write_csv(
    overall_debt_city,
    here(output_path, "metrics_overall_debt_coll_race_ethnicity_city.csv")
)

write_csv(overall_debt_city |> filter(subgroup == "All") |> 
            select(-c(subgroup_type, subgroup)),
          here(output_path, "metrics_overall_debt_coll_all_city.csv"))

```


# Final Data Test Function

The data team for the Mobility Metrics project has created a function that tests baseline requirements for the structure and content of data produced for each metric. This function relies in part on information provided in a metric-specific final data expectations form, which have been filled out prior to final approval by their technical reviewer. 

These evaluation forms are saved in `10a_final-evaluation`


Detail:

 - *Metric: Debt in Collections* 
 - *Tester: Manuel Alcalá Kovalski*
 - *Test date: November 2025*

# Run Test

Put the data through the final evaluation test.

## Read Function

The first step is to source the function `evaluate_final_data.R`. 

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```


## Read Final Data

This file tests the final data for the debt in collections metric. There are four final files to be tested:

 - `04_financial-security/final/metrics_overall_debt_coll_race_ethnicity_city.csv`
 - `04_financial-security/final/metrics_overall_debt_coll_all_city.csv`

Read these final files in.

```{r}
final_data_overall_city <- read_csv(here("04_financial-security/final/metrics_overall_debt_coll_all_city.csv"), show_col_types = FALSE)

final_data_overall_city_race <- read_csv(here("04_financial-security/final/metrics_overall_debt_coll_race_ethnicity_city.csv"), 
                                   show_col_types = FALSE)

```

## Run Evaluation Test

Enter the information into the `evaluate_final_data` function. 

In order to run the function, we fill out the relevant arguments. The debt in collections data does not have subgroups or confidence intervals.

County overall data.
```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_overall_debt_coll_all_city.csv",
                data = final_data_overall_city,  geography = "place",
                subgroups = FALSE, confidence_intervals = FALSE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_overall_debt_coll_race_ethnicity_city.csv",
                data = final_data_overall_city_race,  geography = "place",
                subgroups = TRUE, confidence_intervals = FALSE)
```
