---
title: "Debt in Collections City-Level Shares"
author: "Manuel Alcalá Kovalski"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

## Overview

This document processes city-level debt in collections data from the 
Urban Institute's **Debt in America** project for 2021. Subsequent years of data are only available at the county level.

The processed data is used for the Upward Mobility Metrics Dashboard.

```{r setup}
library(tidyverse)
library(readxl)
library(here)
library(tigris)
# Set output path
output_path <- here("04_financial-security/final")
```

Get state abbreviations & FIPS codes using `tigris` and place FIPS codes from IPUMS
```{r}
state_fips <- fips_codes |>
    distinct(state, state_code, state_name) |>
    rename(stusab = state, state = state_code)


temp_file_ipums <- tempfile(fileext = ".xlsx")
download.file(
    url = "https://usa.ipums.org/usa/resources/volii/large_place_PUMA2010_match_summary.xlsx",
    destfile = temp_file_ipums,
    mode = "wb"
)


place_fips <- read_excel(temp_file_ipums) |>
    rename(
        place_fips = `Place FIPS Code`,
        city = `IPUMS CITY Label`
    ) |>
    # Fix specific city labels to match dashboard data
    mutate(
        city = case_when(
            `State Name` == "Arizona" & `Census Place Name` == "Gilbert town" ~ "Gilbert, AZ",
            `State Name` == "Hawaii" & `Census Place Name` == "Urban Honolulu CDP" ~ "Honolulu, HI",
            `State Name` == "Kansas" & `Census Place Name` == "Overland Park city" ~ "Overland Park, KS",
            `State Name` == "Nevada" & `Census Place Name` == "Paradise CDP" ~ "Paradise, NV",
            `State Name` == "Arizona" & `Census Place Name` == "Scottsdale city" ~ "Scottsdale, AZ",
            `State Name` == "Nevada" & `Census Place Name` == "Sunrise Manor CDP" ~ "Sunrise Manor, NV",
            TRUE ~ city
        )
    ) |>
    filter(city != "[Not identified in any sample]") |>
    select(place_fips, city)

```

## Import and process Urban Data Catalog file 

Download and read Excel file from Financial Health & Wealth Dashboard

```{r}

temp_file <- tempfile(fileext = ".xlsx")
download.file(
    url = "https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2022/12/05/3_city_financial_health_metrics_b.xlsx",
    destfile = temp_file,
    mode = "wb"
)

df <- read_excel(temp_file) |>
    # Keep only city-level debt in collections metrics
    filter(!is.na(`state code (fips)`)) |>
    select(
        city = `city name`,
        stusab = `state name`,
        share_debt_colloverall = `Residents with debt in collections*; Overall`,
        share_debt_collresidentsofcolor = `Residents with debt in collections*; Residents of color`,
        share_debt_collwhite = `Residents with debt in collections*;  White`
    ) |>
    # Remove header row
    filter(city != "city name") |>
    mutate(across(starts_with("share_debt_coll"), ~as.numeric(.x))) |> 
    # Reshape to long format by subgroup
    pivot_longer(
        cols = starts_with("share_debt_coll"),
        names_to = "subgroup",
        names_prefix = "share_debt_coll",
        values_to = "share_debt_coll"
    ) |>
    # Standardize subgroup labels
    mutate(
        subgroup = case_match(
            subgroup,
            "overall" ~ "All",
            "residentsofcolor" ~ "Majority Non-White",
            "white" ~ "Majority White"
        ),
        subgroup_type = "race-ethnicity",
        # Handle missing/suppressed values
        share_debt_coll = if_else(
            share_debt_coll == "NA",
            NA_real_,
            share_debt_coll
        ),
        share_debt_coll = as.numeric(share_debt_coll),
        # Create data quality field: 1 = valid, 3 = suppressed/missing
        share_debt_coll_quality = if_else(
            is.na(share_debt_coll),
            3L,
            1L
        )
    ) |>
    # Merge state FIPS codes
    left_join(state_fips, by = "stusab") |>
    select(-stusab) |>
    # Merge place FIPS codes
    left_join(place_fips, by = "city") |>
    # Add year and reorder
    mutate(year = 2021) |>
    select(
        year, place_fips, state, city, share_debt_coll,
        share_debt_coll_quality, subgroup_type, subgroup
    ) |>
    arrange(year, place_fips, state, city, subgroup)

# Clean up temporary file
unlink(temp_file)

```


## Export data

```{r}
write_csv(
    df,
    file.path(output_path, "city-debt-coll-shares-2021.csv")
)

```

