---
title: "Debt in Collections County-Level Shares"
author: "Manuel Alcalá Kovalski"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

## Overview

This document processes county-level debt in collections data from the 
Urban Institute's **Debt in America** project for years 2022, 2023, and 2025. 
The processed data is used for the Upward Mobility Metrics Dashboard.

**Data Source**: [Debt in America Data Catalog](https://datacatalog.urban.org/dataset/debt-america-2025)

**Key Methodology Changes Across Years:**

- **2022**: 60% threshold for racial communities, 2% sample, overall debt only
- **2023**: 50% threshold (not comparable to 2022 for racial subgroups), 4% sample,
overall + medical debt
- **2025**: Same as 2023 methodology


**Important Notes:**

- **Connecticut**: Debt in America does not have data for CT planning regions
in 2023/2025 data, only original counties
- **Medical Debt Reporting**: Starting in 2022, the three nationwide 
credit-reporting companies made significant changes to medical debt reporting.
These changes reduced the share of people with medical debt in collections
*reported on their credit records* but not necessarily the share of people 
with *actual* medical debt in collections.

For more details, see the [Technical Appendix](https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2025/11/19/Debt%20in%20America%20Technical%20Appendix%202025.pdf)


```{r setup}
library(tidyverse)
library(readxl)
library(glue)
library(here)

# Set output path
output_path <- here("04_financial-security/final")
```

## Helper Function

This function handles the shared transformation logic used across all years.

```{r helper-function}
#' Process debt collections data
#' 
#' Transforms raw debt data into standardized long format with subgroups
#' 
#' @param df Raw dataframe from Excel import
#' @param year Numeric year for the data
#' @param category Debt category ("overall" or "medical")
process_debt_data <- function(df, year, category) {
  df |>
    # Extract state and county FIPS codes
    mutate(
      county_fips = as.character(
        if ("County FIPS" %in% names(df)) `County FIPS`
        else if ("CountyFIPS" %in% names(df)) CountyFIPS
        else if ("GEOID" %in% names(df)) GEOID
        else stop("No FIPS column found")
      ),
      state = str_sub(county_fips, 1, 2),
      county = str_sub(county_fips, 3, 5),
      # Preserve leading zeros
      state = str_pad(state, 2, pad = "0"),
      county = str_pad(county, 3, pad = "0")
    ) |>
    # Rename debt collection share columns (using positions since names vary)
    rename(
      share_debt_coll_all = 4,           
      share_debt_coll_majnonwhite = 5,
      share_debt_coll_majwhite = 6
    ) |>
    select(state, county, starts_with("share_debt_coll")) |>
    filter(county != "OID") |>
    # Reshape to long format by subgroup
    pivot_longer(
      cols = starts_with("share_debt_coll"),
      names_to = "subgroup",
      names_prefix = "share_debt_coll",
      values_to = "share_debt_coll"
    ) |>
    # Standardize subgroup labels
    mutate(
      subgroup = case_match(
        subgroup,
        "_all" ~ "All",
        "_majnonwhite" ~ "Majority Non-White",
        "_majwhite" ~ "Majority White"
      ),
      subgroup_type = "race-ethnicity",
      # Handle missing/suppressed values
      share_debt_coll = if_else(
        share_debt_coll %in% c("n/a*", "n/a**", "NA"),
        NA_character_,
        share_debt_coll
      ),
      share_debt_coll = as.numeric(share_debt_coll),
      # Create data quality field
      share_debt_coll_quality = if_else(
        is.na(share_debt_coll), 
        NA_character_,
        "1"
      ),
      year = year,
      category = category
    ) |>
    # Reorder columns and sort
    select(year, state, county, subgroup_type, subgroup, 
           share_debt_coll, share_debt_coll_quality, category) |>
    arrange(year, state, county, subgroup)
}
```

## 2022 Data (Overall Only)


```{r data-2022}
# Download and read 2022 data
temp_file_2022 <- tempfile(fileext = ".xlsx")
download.file(
  url = "https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2022/06/16/county_dia_delinquency_%207%20Jun%202022.xlsx",
  destfile = temp_file_2022,
  mode = "wb"
)

data_2022 <- read_excel(temp_file_2022) |>
  process_debt_data(year = 2022, category = "overall")

unlink(temp_file_2022)

message("✓ Successfully processed 2022 debt collections data (overall only)")
```

## 2023 Data (Overall + Medical)

**Source**: Direct download from Debt in America using 2023 data updated in July 2024

**Output Files**: 4 files (race-ethnicity breakdown + "All" only for both 
overall and medical)

```{r data-2023}
# Process both overall and medical debt categories
debt_categories_2023 <- c("overall", "medical")

data_2023 <- map(debt_categories_2023, \(category) {
  temp_file <- tempfile(fileext = ".xlsx")
  download.file(
    url = glue("https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2024/09/dia_lbls_all_{category}_county_2023_1Jul2024.xlsx"),
    destfile = temp_file,
    mode = "wb"
  )
  
  df <- read_excel(temp_file) |>
    process_debt_data(year = 2023, category = category)
  
  unlink(temp_file)
  return(df)
}) |>
  bind_rows()

message("✓ Successfully processed 2023 debt collections data (overall + medical)")
```


## 2025 Data (Overall + Medical)

**Source**: Direct download from Debt in America (data from August 2025, 
released November 2024)

**Output Files**: 4 files (race-ethnicity breakdown + "All" only for both overall and medical)

```{r data-2025}
# Process both overall and medical debt categories
debt_categories_2025 <- c("Overall", "Medical")  # Note: Titlecase for 2025

data_2025 <- map(debt_categories_2025, \(category) {
  temp_file <- tempfile(fileext = ".xlsx")
  download.file(
    url = glue("https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2025/11/19/Debt%20in%20America%20County-Level%20{category}%20Debt.xlsx"),
    destfile = temp_file,
    mode = "wb"
  )
  
  df <- read_excel(temp_file) |>
    process_debt_data(year = 2025, category = tolower(category))
  
  unlink(temp_file)
  return(df)
}) |>
  bind_rows()

message("✓ Successfully processed 2025 debt collections data (overall + medical)")
```

## Create Multi-Year Files

```{r combine-years}
# Combine all overall debt data across years
debt_overall <- bind_rows(
  data_2022,
  data_2023 |> filter(category == "overall"),
  data_2025 |> filter(category == "overall")
) |>
  select(-category)

# Combine all medical debt data (2023 and 2025 only)
debt_medical <- bind_rows(
  data_2023 |> filter(category == "medical"),
  data_2025 |> filter(category == "medical")
) |>
  select(-category) |>
  rename(
    share_medical_debt_coll = share_debt_coll,
    share_medical_debt_coll_quality = share_debt_coll_quality
  )

# Export combined race-ethnicity files
debt_overall |>
  write_csv(file.path(output_path, "metrics_overall_debt_coll_race_ethnicity_county.csv"))

debt_medical |>
  write_csv(file.path(output_path, "metrics_medical_debt_coll_race_ethnicity_county.csv"))

# Export combined "Overall" files
debt_overall |>
  filter(subgroup == "All") |>
  select(year, state, county, share_debt_coll, share_debt_coll_quality) |>
  write_csv(file.path(output_path, "metrics_overall_debt_coll_all_county.csv"))

debt_medical |>
  filter(subgroup == "All") |>
  select(year, state, county, share_medical_debt_coll, share_medical_debt_coll_quality) |>
  write_csv(file.path(output_path, "metrics_medical_debt_coll_all_county.csv"))

message("✓ Successfully created combined multi-year datasets")
```

# Final Data Test Function

The data team for the Mobility Metrics project has created a function that tests baseline requirements for the structure and content of data produced for each metric. This function relies in part on information provided in a metric-specific final data expectations form, which have been filled out prior to final approval by their technical reviewer. 

These evaluation forms are saved in `10a_final-evaluation`


Detail:

 - *Metric: Debt in Collections* 
 - *Tester: Manuel Alcalá Kovalski*
 - *Test date: November 2025*

# Run Test

Put the data through the final evaluation test.

## Read Function

The first step is to source the function `evaluate_final_data.R`. 

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```


## Read Final Data

This file tests the final data for the debt in collections metric. There are four final files to be tested:

 - `04_financial-security/final/metrics_medical_debt_coll_all_county.csv`
 - `04_financial-security/final/metrics_medical_debt_coll_race_ethnicity_county.csv`
 - `04_financial-security/final/metrics_overall_debt_coll_all_county.csv`
 - `04_financial-security/final/metrics_overall_debt_coll_race_ethnicity_county.csv`

Read these final files in.

```{r}
final_data_med_county <- read_csv(here("04_financial-security/final/metrics_medical_debt_coll_all_county.csv"), show_col_types = FALSE)

final_data_med_county_race <- read_csv(here("04_financial-security/final/metrics_medical_debt_coll_race_ethnicity_county.csv"), 
                                   show_col_types = FALSE)

final_data_overall_county <- read_csv(here("04_financial-security/final/metrics_overall_debt_coll_all_county.csv"), 
                                   show_col_types = FALSE)

final_data_county_overall_race <- read_csv(here("04_financial-security/final/metrics_overall_debt_coll_race_ethnicity_county.csv"), 
                                   show_col_types = FALSE)
```

## Run Evaluation Test

Enter the information into the `evaluate_final_data` function. 

In order to run the function, we fill out the relevant arguments. The debt in collections data does not have subgroups or confidence intervals.

County overall data.
```{r,error=TRUE}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_overall_debt_coll_all_county_MZ.csv",
                data = final_data_overall_county,  geography = "county",
                subgroups = FALSE, confidence_intervals = FALSE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_medical_debt_coll_all_county_MZ.csv",
                data = final_data_med_county,  geography = "county",
                subgroups = FALSE, confidence_intervals = FALSE)
```

County race data.
```{r, error=TRUE}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_overall_debt_coll_race_ethnicity_county_MZ.csv",
                data = final_data_county_overall_race,  geography = "county",
                subgroups = TRUE, confidence_intervals = FALSE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_medical_debt_coll_race_ethnicity_county_MZ.csv",
                data = final_data_med_county_race,  geography = "county",
                subgroups = TRUE, confidence_intervals = FALSE)

```

Confirmation of passage.
