---
title: "Voter Turnout 2024"
author: ""
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html:
    embed-resources: true
    number_sections: false
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: !expr here::here("21_political-participation", "www", "web_report.css")
editor_options:
  chunk_output_type: console
---

<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

# Voter Turnout

This metric is a county-level estimate of voter turnout. We use Presidential Election turnout as a measure of "Highest Office" for the numerator. We use the Citizen Voting Age Population (CVAP) for the denominator. 

See the [United States Elections Project](https://election.lab.ufl.edu/voter-turnout/) for more information.

## Authors

Originally written in 2024, updated by JP Walsh in 2026.

**Process:**

1. Calculate votes in the 2024 Presidential election
2. Calculate the Citizen Voting Age Population
3. Divide 1. by 2. to calculate voter turnout
4. Add data quality flags
5. Save the data


```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r setup}
options(scipen = 999)

library(tidyverse)
library(urbnthemes)
library(sf)
library(rvest)

set_urbn_defaults(style = "print")

```

## 1. Numerator

### 1.1 MIT Election Data and Science Lab

We use the MIT Election Data and Science Lab for county-level votes in the 2024 Presidential election ([data here](https://dataverse.harvard.edu/file.xhtml?fileId=6104822&version=10.0)). First, we download the data. 

To download the data manually, go to [this link](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ). On the right side of the page, click "Access Dataset" and under "Download Options" select "Original Format ZIP." Complete the License/Data Use Agreement by providing your name, email, institution, and position, and then click "Accept." Once you've downloaded the data, unzip the file and move it into the `21_political-participation/voter-turnout/data/mit` directory.

Alternatively, from the [landing page](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ), scroll down to the file list and right click on the tabular data file (e.g., `countypres_2000-2024.tab` for 2024 - the only other file is an `md` file). Select "Copy Link Address." This link should include a 7 digit string after `fileID=`. Replace the 7 digit string in the `download.file()` command below with the new string (between `datafile/` and `?format=`). This will download the data directly. 

```{r}
# create a data folder for intermediate data files
intermediate_files <- here::here("21_political-participation",
                                 "data")

if (!dir.exists(intermediate_files)) {
  
  dir.create(intermediate_files)
  
}

# download the data from the Harvard dataverse
mit_directory <- here::here("21_political-participation", 
                            "data",
                            "mit")

mit_data <- here::here("21_political-participation", 
                       "data",
                       "mit",
                       "countypres_2000-2024.csv")

if (!dir.exists(mit_directory)) {
  
  dir.create(mit_directory)
  
}

if (!file.exists(mit_data)) {
  
  download.file(
    "https://dataverse.harvard.edu/api/access/datafile/13573089?format=original&gbrecs=true",
    destfile = mit_data
  )
  
  
  
}


```

Next we load and clean the data. 

```{r}
mit <- read_csv(here::here("21_political-participation", 
                           "data",
                           "mit", 
                           "countypres_2000-2024.csv")) %>%
  # rename the FIPS code variable to match 2016
  rename(FIPS = county_fips) %>% 
  mutate(FIPS = as.character(FIPS),
         FIPS = str_pad(FIPS, width = 5, side = "left", pad = "0"))

```

Filter for year 2024 and keep only relevant variables.

```{r}
mit <- mit %>%
  tidylog::filter(year == 2024) %>% 
  select(state, county_name, year, state_po, FIPS, mode, candidatevotes, totalvotes) 
```

[Shannon County, South Dakota (FIPS 46113) changed to Oglala Lakota County, South Dakota (FIPS 46102) in 2015.](https://www.census.gov/programs-surveys/geography/technical-documentation/county-changes.html) The MIT data have the correct county name but they do not have the correct FIPS code.

In 2019 there were 3,220 counties in the US including Washington DC and Puerto Rico. In 2020, the Valdez-Cordova Census Area in Alaska (FIPS 02261) has been split into two new Census Areas/FIPS: Copper River Census Area (FIPS 02066) and Chugach Census Area (FIPS 02063), so there are now [3,221 counties.](https://www.esri.com/arcgis-blog/products/arcgis-living-atlas/mapping/acs-2016-2020-updated-boundaries/) However, as discussed later, Alaska is reported as Districts instead of counties, so we standardize the FIPS codes for all Alaska districts to "02000" - thus, we don't need to worry about the county change in Alaska.


```{r}
mit %>%
  filter(FIPS %in% c("46113", "46102"))

mit <- mit %>%
  mutate(FIPS = if_else(FIPS == "46113", "46102", FIPS))

```

In prior years there were nine observations without FIPS codes originating from the District of Columbia and a county in Rhode Island formerly titled "FEDERAL PRECINCT." The census corrected this in the 2024 data and there are no longer any cases with missing fips.

There are some missing observations for `candidatevotes` but none for `totalvotes` and no observations for which `totalvotes` is 0.

```{r}
mit %>%
  filter(is.na(candidatevotes))

mit %>% 
  filter(is.na(totalvotes))

mit %>%
  filter(totalvotes == 0)

```

Alaska is reported as Districts instead of counties. We replace the FIPS codes for all Alaska observations with `02000`. 
```{r}

mit <- mit %>%
  mutate(FIPS = if_else(state == "ALASKA", "02000", FIPS))

mit %>%
  filter(state == "ALASKA") %>%
  select(state, FIPS)

```

There are several states where the summarized candidate votes consistently don't sum to the county votes.

```{r}
mit %>%
  group_by(state, county_name, FIPS) %>%
  summarize(candidatevotes = sum(candidatevotes),
            totalvotes = max(totalvotes),
            .groups = 'drop') %>%
  filter(totalvotes != candidatevotes)

```

This is because in 2024 MIT introduced the vote mode variable. The voting `mode` categories differ by state in the 2024 data. For some states this includes an additional "total vote" row which results in duplicates when summed.

The table below shows the differences in state mode categories. Some states have "total votes" alone while others have a combination of modes. Some have NA.
```{r}
mit %>% 
  select(state, mode, candidatevotes) %>% 
  unique() %>% 
  group_by(state, mode) %>% 
  summarise(category_apperance = n(),
            candidate_votes = sum(candidatevotes)) %>% 
  reactable::reactable(
  filterable = TRUE,
  defaultPageSize = 6
)
```

The 'totalvotes' column provides a clean version of the county total votes. To clean the data we drop the candidate votes from the dataset and keep unique rows. This results in one row per county.

We also create `state` and `county`.

```{r}
mit_counties <- mit %>% 
  select(year, state, county_name, FIPS, totalvotes) %>% 
  unique() %>% 
  rename(votes = totalvotes)

# create state and county
mit_counties <- mit_counties %>%
  mutate(state = str_sub(FIPS, 1, 2),
         county = str_sub(FIPS, 3, 5)) %>%
  select(year, 
         state, 
         county,
         votes)
```

## Turnout comparisons

We compare the aggregated totals to the state total posted on [Wikipedia](https://en.wikipedia.org/wiki/2024_United_States_presidential_election) and with totals published by the [Census](https://www.census.gov/data/tables/time-series/demo/voting-and-registration/p20-587.html). 

### Wikipedia 

This code scrapes the vote count from Wikipedia: 


```{r}
# scrape data from Wikipedia with rvest
wiki <- read_html("https://en.wikipedia.org/wiki/2024_United_States_presidential_election")

wiki_votes <- wiki %>%
  html_elements("td:nth-child(23)") %>%
  html_text() %>%
  # clean numbers as characters to numerics
  str_replace_all(",", "") %>%
  str_replace_all("\n", "") %>%
  as.numeric()

# the name column in the data doesn't scrape well so we manually state it
state <- c(state.name[1:8],
           'District of Columbia',
           state.name[9:19],
           rep('subset', 2),
           state.name[20:27],
           rep('subset', 3),
           state.name[28:50]
           )    

state_votes <- tibble(state, wiki_votes) %>%
  tidylog::filter(state != "subset")

rm(wiki, wiki_votes, state)

```

The totals are very close for most states. Vermont and Missouri have the biggest differences. 

```{r}
# compare the scraped data to the mit data
mit_counties %>%
  group_by(state) %>%
  summarize(mit_votes = sum(votes, na.rm = TRUE)) %>%
  bind_cols(state_votes) %>%
  mutate(mit_votes / wiki_votes) %>%
  select(state = state...3, mit_votes, wiki_votes, `mit_votes/wiki_votes`) %>%
  arrange(`mit_votes/wiki_votes`) %>%
  knitr::kable(digits = 3)

```

The aggregated nationwide total from MIT of 155,209,552 is also very close to the 155,240,955 votes reported on Wikipedia. We have just over 34k less votes than Wikipedia.

```{r}
state_votes %>%
  summarize(sum(wiki_votes))

mit_counties %>%
  summarize(sum(votes, na.rm = TRUE))

```

There are no missing values.

```{r}
mit_counties %>%
  map_dbl(~sum(is.na(.)))

```

### Census 

Read in Census data on 2024 election turnout. The census releases the "P20" table every two years following national level elections. 

"Since 1964, the U.S. Census Bureau has fielded the Voting and Registration supplement to the CPS every 2 years in November. While the Census Bureau has collected voting and registration data since 1964, the CPS has gathered  citizenship data for congressional voting only since 1978. The CPS is the only source of voting and registration data produced by the U.S. Census Bureau."

```{r}
#Specify the URL of the Excel file from the Census website 
  excel_url <- "https://www2.census.gov/programs-surveys/cps/tables/p20/587/vote04a_2024.xlsx" # Replace with latest year if updating

#Download the file to a temporary location
httr::GET(excel_url, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))

#Read the data from the temporary file into a data frame
census_voting_pop <- readxl::read_xlsx(tf, skip = 3) %>% 
  select(state = Characteristic, voting_age_pop = `Total population`, census_votes = Voted) %>% 
  filter(!state %in% c("UNITED STATES", NA_character_),
         !is.na(census_votes)) %>% 
  mutate(state = str_to_title(state),
         state = ifelse(str_detect(state, "District"), "District of Columbia", state),
         census_votes = as.numeric(census_votes),
         census_votes = census_votes * 1000,
         voting_age_pop = as.numeric(voting_age_pop),
         voting_age_pop = voting_age_pop * 1000)
```

The census vote counts show more variability and often do not directionally agree with the Wikipedia votes.
```{r}
# compare the data sources

mit_counties %>%
  group_by(state) %>%
  summarize(mit_votes = sum(votes, na.rm = TRUE)) %>%
  bind_cols(state_votes, census_voting_pop) %>%
  mutate(mit_votes / wiki_votes,
         mit_votes / census_votes) %>%
  select(state = state...3, mit_votes, wiki_votes, `mit_votes/wiki_votes`, census_votes, `mit_votes/census_votes`) %>%
  arrange(`mit_votes/wiki_votes`) %>%
  knitr::kable(digits = 3)


```

Overall the census shows closer to 1 million less votes than both the MIT and Wikipedia data.
```{r}
census_voting_pop %>%
  summarize(sum(census_votes))

mit_counties %>%
  summarize(sum(votes, na.rm = TRUE))

```

## 2. Denominator

The denominator for the analysis should be the total age-eligible citizen population ([data here](https://www.census.gov/programs-surveys/decennial-census/about/voting-rights/cvap.2020.html))

The US Census Bureau creates a special tabulation of the 2020-2024 ACS that includes county-level estimates of the total number of United States citizens 18 years of age or older. They also report estimates for subgroups within counties and for other geographic areas. 

If the data are not downloaded, then we download and unzip the data.

```{r}
cvap_zip <- here::here("21_political-participation", 
                       "data",
                       "CVAP_2020-2024_ACS_csv_files.zip")

cvap_data <- here::here("21_political-participation", 
                        "data",
                        "CVAP_2020-2024_ACS_csv_files")

if (!file.exists(cvap_data)) {
  
  download.file(url = "https://www2.census.gov/programs-surveys/decennial/rdo/datasets/2024/2024-cvap/CVAP_2020-2024_ACS_csv_files.zip",
                destfile = cvap_zip)
  
  unzip(zipfile = cvap_zip,
        exdir = cvap_data)
  
  file.remove(cvap_zip)  
  
}

```

Next, we load and clean the data. The variable of interest is `cvap_est`.

> cvap_est: The rounded estimate of the total number of United States citizens 18 years of age or older for that geographic area and group.

```{r}
cvap <- read_csv(here::here("21_political-participation", 
                            "data",
                            "CVAP_2020-2024_ACS_csv_files", 
                            "County.csv"))

cvap <- cvap %>%
  tidylog::filter(lntitle == "Total") %>%
  select(-lntitle, -lnnumber)

cvap <- cvap %>%
  mutate(state = str_sub(string = geoid, start = 10, end = 11),
         county = str_sub(string = geoid, start = 12, end = 14),
         FIPS = str_c(state, county))

cvap <- cvap %>%
  tidylog::filter(state != "72") %>% # Drop Puerto Rico
  select(state, 
         county, 
         FIPS,
         cvap = cvap_est,
         cvap_moe)

```

We need to crosswalk the Connecticut counties in the CVAP data to Planning Regions in order to merge with the MIT data

```{r}
cw_connecticut <- read_csv(here::here("geographic-crosswalks", "data", "connecticut_planning_regions_to_pre_2023_counties.csv")) %>% 
  select(FIPS = county, FIPS_old = CTcounty, afact) %>% 
   mutate(FIPS = as.character(FIPS),
         FIPS = str_pad(FIPS, width = 5, side = "left", pad = "0"),
         FIPS_old = as.character(FIPS_old),
         FIPS_old = str_pad(FIPS_old, width = 5, side = "left", pad = "0"))
```

Join onto CVAP data and sum the eligible voting population weighted by the AFACT

```{r}
cvap_join <- cvap %>% 
  left_join(cw_connecticut, by = c("FIPS")) %>% 
  mutate(FIPS = ifelse(state == "09", FIPS_old, FIPS),
         afact = ifelse(is.na(afact), 1, afact),
         county = str_sub(FIPS, -3)) %>% 
  group_by(state, county, FIPS) %>% 
  summarise(cvap = sum(cvap * afact),
            cvap_moe = sum(cvap_moe * afact))
```

## 3. Combine and calculate turnout

We combine the data. The join works for all counties except for counties in Alaska and [Kalawao County, Hawaii (FIPS 15005 which is a judicial district of Maui County (FIPS 15009))](https://en.wikipedia.org/wiki/Kalawao_County,_Hawaii). Alaska is dropped entirely because it doesn't have a county that exists in `cvap`.

```{r}
joined_data <- left_join(cvap_join, mit_counties, by = c("state", "county"))

anti_join(cvap_join, mit_counties, by = c("state", "county"))

```

We calculate turnout and the standard error for the CVAP estimate. 

```{r}
joined_data <- joined_data %>%
  mutate(
    election_turnout = votes / cvap,
    cvap_se = cvap_moe / 1.645
  ) %>%
  mutate(cv = cvap_se / cvap)

```

6 observations have voter turnout above `1`. This is likely because of sampling error in the denominator. We set these values to one and all of these cases will be flagged. 

```{r}
joined_data %>%
  filter(votes > cvap)

joined_data <- joined_data %>%
  mutate(election_turnout = if_else(condition = election_turnout > 1, true = 1, false = election_turnout))

```

**Check:** Is voter turnout bounded by 0 and 1 inclusive

```{r}
stopifnot(
  max(joined_data$election_turnout, na.rm = TRUE) <= 1
)

stopifnot(
  min(joined_data$election_turnout, na.rm = TRUE) >= 0
)

```

```{r}
joined_data %>%
  ggplot(aes(votes, election_turnout)) +
  geom_point(alpha = 0.1) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0, 0.1))) +
  scatter_grid() +
  labs(title = "There Isn't Much Relationship Between Turnout and Votes")

joined_data %>%
  ggplot(aes(cvap, election_turnout)) +
  geom_point(alpha = 0.1) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0, 0.1))) +
  scatter_grid() +
  labs(title = "There Isn't Much Relationship Between CVAP and Votes")

```

## 4. Quality flags

Except for the cases excluded above, the quality of the numerator does not seem to be a concern. Sampling error in the denominator is definitely a concern for small counties. 

We flag cases with high and very high coefficients of variation in the denominator. 

* `1` No issue
* `2` CV >= 0.05
* `3` CV >= 0.15

There isn't much consensus on critical values for coefficients of variation. We use `0.15` because it is mentioned [A Compass for Understanding and Using American Community Survey Data](https://www.census.gov/content/dam/Census/library/publications/2009/acs/ACSstateLocal.pdf).

>  While there is no hard-and-fast rule, for the purposes of this handbook, estimates with CVs of more than 15 percent are considered cause for caution when interpreting patterns in the data.

If anything, a stricter threshold is necessary because the estimates are used in denominators. Thus, we use `0.05` for a `2`. 

```{r}
joined_data <- joined_data %>%
  mutate(
    election_turnout_quality = case_when(
      is.na(election_turnout) ~ NA_real_,
      cv >= 0.15 ~ 3,
      cv >= 0.05 ~ 2,
      TRUE ~ 1
    )
  )

count(joined_data, election_turnout_quality)

```

```{r}
joined_data %>%
  ggplot(aes(factor(election_turnout_quality), election_turnout)) +
  geom_point(alpha = 0.1) +
  scale_y_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0, 0.1))) +
  scatter_grid() +
  labs(title = "Very High Turnout is Associated with Poor Data Quality")
```

## 5. Save the data

```{r}
# load the 2020 county file
all_counties <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv")) %>%
  tidylog::filter(year == 2024) %>%
  select(-year)

all_counties <- left_join(all_counties, joined_data, by = c("state", "county"))

all_counties <- all_counties %>%
  mutate(year = 2024) %>%
  select(year, state, county, election_turnout, election_turnout_quality, population) 

all_counties %>%
  select(-population) %>%
  write_csv(here::here("21_political-participation",
                       "data",
                       "final",
                       "voter-turnout-2024.csv"))
  
```

Create combined file with all years (2016, 2020, 2024)

```{r}
all_counties %>%
  select(-population) %>%
  bind_rows(
    read_csv(here::here("21_political-participation", "data", "final", "voter-turnout-2016.csv")),
    read_csv(here::here("21_political-participation", "data", "final", "voter-turnout-2020.csv"))
  ) %>% 
  arrange(year) %>% 
  write_csv(here::here("21_political-participation",
                       "data",
                       "final",
                       "voter-turnout-2016-2024.csv"))
```

## Benchmark against 2020

### Compare turnout

```{r}
turnout2020 <- read_csv(here::here("21_political-participation",
                                    "data",
                                    "final",
                                   "voter-turnout-2020.csv"))


turnout_combined <- left_join(
  all_counties,
  turnout2020,
  by = c("state", "county"),
  suffix = c("_2024", "_2020")
)

turnout_combined <- turnout_combined %>%
  mutate(change = election_turnout_2024 - election_turnout_2020)

turnout_combined %>%
  ggplot(aes(election_turnout_2024, election_turnout_2020, color = population < 10000)) +
  geom_abline() +
  geom_point(alpha = 0.1) +
  labs(
    title = "Most voter turnouts decreased modestly",
    subtitle = "Counties with fewer than 10,000 people in yellow"
  ) +
  coord_equal() +
  scatter_grid()

turnout_combined %>%
  filter(population < 1000000) %>%
  ggplot(aes(population, change)) +
  geom_point(alpha = 0.1) +
  labs(
    title = "Large changes in turnout are in low-population areas",
    subtitle = "Exclusing counties with 1+ million people"
    ) +
  scatter_grid()

turnout_combined %>%
  arrange(desc(abs(change))) %>%
  select(-year_2024, -year_2020)

```

### Compare quality

```{r}
turnout_combined %>%
  count(election_turnout_quality_2020, election_turnout_quality_2024)

```

