---
title: "SHARE OF STUDENTS IN HIGH POVERTY SCHOOLS BY RACE/ETHNICITY "
author: "Written originally by E Gutierrez and translated to R by JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
params:
  year: "2025"
---

*Program for the creation of the students in high poverty schools metric at the city level*

11/28/2022 
		Uses MEPS: the share of students attending schools where 20% or more students are in Poverty  **
		Uses 2014-2018 data
	  Produces city level data
10/15/2025 
		Produces data for 2014-2022 (School Years 2014-15 through 2022-23)

## Housekeeping

Read in packages for use. Set the global year to the latest education data.

```{r, results='hide'}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)
library(srvyr)
library(scales)
library(educationdata)
library(haven)

options(scipen = 999)

year_ed=2022
year_meps=2022

```

Source the function `evaluate_final_data.R`. 

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```

## Import and save raw data

Read in the City crosswalk file 

```{r}
place <- read_csv(here::here("geographic-crosswalks", "data", "place-populations.csv"))
```

Clean city names 

```{r}
place_clean <- place %>% 
  mutate(place_name = str_replace_all(place_name, " city| town| municipality| village| urban county", ""),
         city_name = case_when(place_name == "San Buenaventura (Ventura)" ~ "Ventura",
                                      place_name == "Athens-Clarke County unified government (balance)" ~ "Athens",
                                      place_name == "Augusta-Richmond County consolidated government (balance)" ~ "Augusta",
                                      place_name == "Macon-Bibb County" ~ "Macon",
                                      place_name == "Boise City" ~ "Boise",
                                      place_name == "Lexington-Fayette" ~ "Lexington",
                                      place_name == "Louisville/Jefferson County metro government (balance)" ~ "Louisville",
                                      place_name == "Nashville-Davidson metropolitan government (balance)" ~ "Nashville",
                                      place_name == "McAllen" ~ "Mcallen",
                                      place_name == "McKinney" ~ "Mckinney",
         TRUE ~ place_name),
         city_name = str_to_lower(city_name)) %>% 
  select(-place_name, -population)
```

Now read in the necessary CCD district data

### School level enrollment

Download the CCD school level enrollment by race/ethnicity from Urban's Education Data Portal
Filter grade for 99 - 99 is the total enrollment in the school, regardless of grade.

Note this download is computationally large, will take time to finish.

```{r}
input_files <- here::here("08_school-economic-diversity/data/raw/", paste0("ccd_enr_2014-", year_ed, ".rds"))

if (!file.exists(input_files)) {
  education_data <- get_education_data(level = 'schools', 
                                       source = 'ccd', 
                                       topic = 'enrollment', 
                                       subtopic = 'race',
                                       filters = list(year = 2014:year_ed,
                                                      grade = 99),
                                       add_labels = TRUE) %>% 
    select(sex, enrollment, year, ncessch, ncessch_num, leaid, fips, race) 
  
  write_rds(education_data, input_files)
}  else {
  
  education_data <- read_rds(input_files) %>% 
    select(sex, enrollment, year, ncessch, ncessch_num, leaid, fips, race) 

}
```

Add the number of enrolled male and female students within race

```{r}
education_data <- education_data %>% 
  filter(sex != 99) %>% 
  mutate(enrollment = ifelse(enrollment < 0, NA_integer_, enrollment)) %>% 
  group_by(year, ncessch, ncessch_num, leaid, fips, race) %>% 
  summarise(enrollment = sum(enrollment, na.rm = TRUE))  %>% 
  ungroup()
```

Reshape the the data wide by race

```{r}

ed_data_wide <- education_data %>% 
  pivot_wider(id_cols = c(year, ncessch), names_from = race, values_from = enrollment) 

```

### CCD district data

Download district data from CCD, includes total enrollment, city location and county codes

```{r}
input_files <- here::here("08_school-economic-diversity/data/raw/", paste0("ccd_dir_2014-", year_ed, ".rds"))


if (!file.exists(input_files)) {
  ccd_directory <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'directory',
                         filters = list(year = 2014:year_ed),
                         add_labels = FALSE)  %>% 
    select(enrollment, year, ncessch, city_location, county_code)
  write_rds(ccd_directory, input_files)
}  else {
    ccd_directory <- read_rds(input_files) %>% 
    select(enrollment, year, ncessch, city_location, county_code)
}
```

### Meps data

Read in the latest MEPS data (using temporary internal Urban data while the Ed Portal is being updated)

```{r}
meps <- read_dta(here::here("08_school-economic-diversity/data/raw/", "Abrv Set of Portal Variables.dta")) %>% 
  select(year, ncessch, meps_poverty_pct)
```

For future users - download when latest MEPs is available on the ed portal (remove comments from the line of code below)

```{r}
#input_files <- here::here("08_school-economic-diversity/data/raw/", paste0("ccd_meps_2014-", year_meps, ".rds"))

# if(!file.exists(input_files)) {
#   meps <- get_education_data(level = 'schools', 
#                          source = 'meps', 
#                          filters = list(year = 2014:year_meps)) %>% 
#  select(year, ncessch, meps_poverty_pct)
#   write_rds(meps, input_files)
# }  else {
#     meps <- read_rds(input_files) 
# }

```

## Data merge

Join the MEPs and wide education data to get the full ccd directory data to get a dataframe race enrollment and poverty share for every school.

First check if any schools from the ed data wide do not have a match in MEPs. There are missing cases in every year.

```{r}
anti_join(ed_data_wide, meps, by = c("year", "ncessch")) %>% 
  group_by(year) %>% 
  count()
```

Now check missing with ccd directory dat. There are also mismatches for year and school codes between these data sets.

```{r}
anti_join(ccd_directory, ed_data_wide, by = c("year", "ncessch")) %>% 
  group_by(year) %>% 
  count()
```

Eliminate negative enrollment and join the ed race data set and MEPS data on to the directory. 

```{r}
ccd_directory <- ccd_directory %>% 
  mutate(enrollment = if_else(enrollment < 0, 0, enrollment))

ccd_merged <- ccd_directory %>% 
  left_join(ed_data_wide, by = c("year", "ncessch")) %>% 
  left_join(meps, by = c("year", "ncessch")) 

```

Check merge value 

Directory and ed data merge is high: between 93 and 97 percent 

```{r}
ccd_merged %>% 
  mutate(merge_indicator = ifelse(is.na(Total), 0, 1)) %>% 
  group_by(year) %>% 
  summarise(merge_rate = mean(merge_indicator))
```

Directory and meps data merge is good outside of years 2021 and 2022 (waiting on updated data)

```{r}
ccd_merged %>% 
  mutate(merge_indicator = ifelse(is.na(meps_poverty_pct), 0, 1)) %>% 
  group_by(year) %>% 
  summarise(merge_rate = mean(merge_indicator))
```

Check extent of missing and negative enrollment values
```{r}
ccd_merged %>% 
  mutate(enrollment_value = case_when(is.na(enrollment) ~ "Missing",
                                      enrollment == 0 ~ "Zero",
                                      TRUE ~ "Populated")) %>% 
  group_by(year, enrollment_value) %>% 
  count() %>% 
  pivot_wider(names_from = enrollment_value, values_from = n) %>% 
  mutate(share_missing_zero = (Missing + Zero) / (Missing + Zero + Populated)) %>% 
  select(year, share_missing_zero)
```

Drop values where enrollment is missing or equal to zero

```{r}
ccd_merged <- ccd_merged %>% 
  filter(!is.na(enrollment), enrollment != 0)
```

## Clean merged data

*Using MEPS - edit MEPS share and binary if the share is greater than or equal to 20%

```{r}
ccd_merged_clean <- ccd_merged %>% 
  mutate(meps_share = meps_poverty_pct / 100,
         meps_20 = ifelse(meps_share >= 0.2, 1, 0),
         numerator = ifelse(meps_20 == 0 | is.na(meps_20), meps_20, enrollment),
         across(White:Hispanic, ~ifelse(meps_20 == 0 | is.na(meps_20), meps_20, .), .names = "numerator_{.col}"))
```

Look at distribution of MEPS share

```{r}
ccd_merged_clean %>% 
  select(meps_share) %>% 
  ggplot(aes(x = meps_share)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 0.01) +
  theme_minimal() +
  ggtitle("MEPS mod poverty pct share") +
  ylab("Count")
```
 
 Count at the 20 percentage threshold 
 
```{r}
ccd_merged_clean %>% 
  group_by(meps_20) %>% 
  count()
```
 

## Clean geo

Pad MEPS data couty fips to match the overall county file


```{r}

ccd_merged_clean <- ccd_merged_clean %>% 
  mutate(county_code = ifelse(county_code == -2, NA_character_, county_code),
         county_code = str_pad(county_code, width = 5, side = "left", pad = "0"),
         state = str_sub(county_code, 1, 2))
```

Drop rows with missing county information 

```{r}
ccd_merged_clean <- ccd_merged_clean %>% 
  filter(!is.na(county_code))
```

Create lower_case version of city name 

```{r}
ccd_merged_clean <- ccd_merged_clean %>% 
  mutate(city_name = tolower(city_location))
```

## Create city level data

Create enrollment summaries for each year by unique state and city combinations

```{r}
ccd_city <- ccd_merged_clean %>% 
  group_by(year, state, city_name) %>% 
  summarise(enrollment = sum(enrollment),
            enrollment_black = sum(Black),
            enrollment_white = sum(White),
            enrollment_hisp = sum(Hispanic),
           numerator = sum(numerator, na.rm = TRUE),
           numerator_black = sum(numerator_Black, na.rm = TRUE),
           numerator_white = sum(numerator_White, na.rm = TRUE),
           numerator_hisp = sum(numerator_Hispanic, na.rm = TRUE)) 
```

Create MEPS20 shares by enrollment 
When the denominator is zero replace with NA integer

```{r}
meps_city <- ccd_city %>% 
   mutate(meps20_total = numerator / enrollment,
         share_meps20_white = numerator_white/enrollment_white,
        share_meps20_black = numerator_black/enrollment_black,
         share_meps20_hispanic = numerator_hisp/enrollment_hisp,
        share_meps20_white = ifelse(enrollment_white == 0, NA_integer_, share_meps20_white),
         share_meps20_black = ifelse(enrollment_black == 0, NA_integer_, share_meps20_black),
         share_meps20_hispanic = ifelse(enrollment_hisp == 0, NA_integer_, share_meps20_hispanic))
```

## Merge onto place file

Join MEPs data onto the overall palce file
```{r}
meps_city_geo <- place_clean %>% 
  left_join(meps_city, by = c("year", "state", "city_name"))

```

## Add quality flags 

Add quality flag
```{r}
meps_city_clean <- meps_city_geo %>% 
  mutate(share_meps20_white_quality = case_when(enrollment_white >= 30 & !is.na(share_meps20_white) ~ 1,
                                          enrollment_white >= 15 & !is.na(share_meps20_white) ~ 2,
                                          !is.na(share_meps20_white) ~ 3,
                                          TRUE ~ NA_integer_),
         share_meps20_black_quality = case_when(enrollment_black >= 30 & !is.na(share_meps20_black) ~ 1,
                                          enrollment_black >= 15 & !is.na(share_meps20_black) ~ 2,
                                          !is.na(share_meps20_black) ~ 3,
                                          TRUE ~ NA_integer_),
         share_meps20_hispanic_quality = case_when(enrollment_hisp >= 30 & !is.na(share_meps20_hispanic) ~ 1,
                                          enrollment_hisp >= 15 & !is.na(share_meps20_hispanic) ~ 2,
                                          !is.na(share_meps20_hispanic) ~ 3,
                                          TRUE ~ NA_integer_)) %>% 
  select(year, state, place, share_meps20_black, share_meps20_black_quality, share_meps20_hispanic, share_meps20_hispanic_quality,
         share_meps20_white, share_meps20_white_quality) %>% 
  filter(!year %in% c(2023))
```

Sort by year, state and place

```{r}
meps_city_clean <- meps_city_clean %>% 
  arrange(year, state, place)
```


## Run quality checks and write out 

Check missingness for each variable  
Missingness is generally reflective of counties missing data from the merge to the crosswalk rather that MEPS missing data (see line 164)

```{r}
colSums(is.na(meps_city_clean))
```

Run city-level quality check

```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_meps_overall_place.csv",
                data = meps_city_clean,  geography = "place",
                subgroups = FALSE, confidence_intervals = FALSE)
```

write out final csv

```{r}
write_csv(meps_city_clean, paste0(here::here(), "/08_school-economic-diversity/data/final/meps_place_2014-", year_meps, ".csv"))
```


