---
title: "Final Data Evaluation MEPs Data"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

# Final Data Test Function

The data team for the Mobility Metrics project has created a function that tests baseline requirements for the structure and content of data produced for each metric. This function relies in part on information provided in a metric-specific final data expectations form, which have been filled out prior to final approval by their technical reviewer. 

These evaluation forms are saved in `10a_final-evaluation`

For programs written in STATA the evaluation process is run externally in this QMD file. 

Detail:

 - *Metric: Students in high poverty schools* 
 - *Tester: JP Walsh*
 - *Test date: March 2025*

# Run Test

Put the data through the final evaluation test.

## Read Function

The first step is to source the function `evaluate_final_data.R`. 

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```


## Read Final Data

This file tests the final data for the MEPs metric on students in high poverty schools. There are two final files to be tested:

 - `08_education/data/final/meps_city_2014-2021.csv`
 - `08_education/data/final/meps_county_2014-2021.csv`

Read these final files in.

```{r}
final_data_city <- read_csv(here("08_education/data/final/meps_city_2014-2021.csv"), show_col_types = FALSE)

final_data_county <- read_csv(here("08_education/data/final/meps_county_2014-2021.csv"), show_col_types = FALSE)
```

## Run Evaluation Test

Enter the information into the `evaluate_final_data` function. 

In order to run the function, we fill out the relevant arguments. MEPs data does not have subgroups or confidence intervals.

City level data.
```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_meps_overall_place.csv",
                data = final_data_city,  geography = "place",
                subgroups = FALSE, confidence_intervals = FALSE)
```

County level data.
```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_meps_overall_county.csv",
                data = final_data_county,  geography = "county",
                subgroups = FALSE, confidence_intervals = FALSE)

```

Confirmation of passage.
