---
title: "SHARE OF STUDENTS IN HIGH POVERTY SCHOOLS BY RACE/ETHNICITY "
author: "Written originally by E Gutierrez and translated to R by JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
params:
  year: "2025"
---

*Program for the creation of the students in high poverty schools metric at the city level*

11/28/2022 
		Uses MEPS: the share of students attending schools where 20% or more students are in Poverty  **
		Uses 2014-2018 data
	  Produces city level data
10/15/2025 
		Produces data for 2014-2022 (School Years 2014-15 through 2022-23)

## Housekeeping

Read in packages for use. Set the global year to the latest education data.

```{r, results='hide'}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)
library(srvyr)
library(scales)
library(educationdata)

options(scipen = 999)

year_ed=2022

```

## Import and save raw data

Read in the City crosswalk file 

```{r}
place <- read_csv(here::here("geographic-crosswalks", "data", "place-populations.csv"))
```

Read in the CCD district data

Download the CCD school level enrollment by race/ethnicity from Urban's Education Data Portal

```{r}
df <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'enrollment', 
                         subtopic = 'race',
                         filters = list(year = 2014:year_ed,
                                        grade = 99,
                                        sex = 99),
                         add_labels = TRUE)
```

