---
title: "SHARE OF STUDENTS IN HIGH POVERTY SCHOOLS BY RACE/ETHNICITY "
author: "Written originally by E Gutierrez and translated to R by JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
params:
  year: "2025"
---

*Program for the creation of the students in high poverty schools metric at the city level*

11/28/2022 
		Uses MEPS: the share of students attending schools where 20% or more students are in Poverty  **
		Uses 2014-2018 data
	  Produces city level data
10/15/2025 
		Produces data for 2014-2022 (School Years 2014-15 through 2022-23)

## Housekeeping

Read in packages for use. Set the global year to the latest education data.

```{r, results='hide'}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)
library(srvyr)
library(scales)
library(educationdata)

options(scipen = 999)

year_ed=2022

```

## Import and save raw data

Read in the City crosswalk file 

```{r}
place <- read_csv(here::here("geographic-crosswalks", "data", "place-populations.csv"))
```

Now read in the necessary CCD district data

### School level enrollment

Download the CCD school level enrollment by race/ethnicity from Urban's Education Data Portal
Filter grade for 99 - 99 is the total enrollment in the school, regardless of grade.

Note this download is computationally large, will take time to finish.

```{r}
if(!file.exists(here::here("08_education/data/raw/", paste0("ccd_enr_2014-", year_ed, ".rds"))))
{
  education_data <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'enrollment', 
                         subtopic = 'race',
                         filters = list(year = 2014:year_ed,
                                        grade = 99),
                         add_labels = TRUE)
  write_rds(education_data, 
            here::here("08_education/data/raw/", paste0("ccd_enr_2014-", year_ed, ".rds")))
}  else {
    education_data <- read_rds(here::here("08_education/data/raw/", paste0("ccd_enr_2014-", year_ed, ".rds")))
}

```

Add the number of enrolled male and female students within race

```{r}
education_data <- education_data %>% 
  select(-grade) %>% 
  filter(sex != 99) %>% 
  mutate(enrollment = ifelse(enrollment < 0, NA_integer_, enrollment)) %>% 
  group_by(year, ncessch, ncessch_num, leaid, fips, race) %>% 
  summarise(enrollment = sum(enrollment, na.rm = TRUE))
```

Reshape the the data wide by race
1-white, 2-black, 3-hispanic

```{r}

ed_data_wide <- education_data %>% 
  pivot_wider(id_cols = c(year, ncessch), names_from = race, values_from = enrollment) 

```

### CCD district data

Download district data from CCD, includes total enrollment, city location and county codes

```{r}

if(!file.exists(here::here("08_education/data/raw/", paste0("ccd_dir_2014-", year_ed, ".rds"))))
{
  ccd_directory <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'directory',
                         filters = list(year = 2014:year_ed),
                         add_labels = FALSE)
  write_rds(ccd_directory, 
            here::here("08_education/data/raw/", paste0("ccd_dir_2014-", year_ed, ".rds")))
}  else {
    ccd_directory <- read_rds(here::here("08_education/data/raw/", paste0("ccd_dir_2014-", year_ed, ".rds")))
}

```

### Meps data

MEPS data

```{r}
year_meps <- 2020
if(!file.exists(here::here("08_education/data/raw/", paste0("ccd_meps_2014-", year_meps, ".rds"))))
{
  meps <- get_education_data(level = 'schools', 
                         source = 'meps', 
                         filters = list(year = 2014:year_meps))
  write_rds(meps, 
            here::here("08_education/data/raw/", paste0("ccd_meps_2014-", year_meps, ".rds")))
}  else {
    meps <- read_rds(here::here("08_education/data/raw/", paste0("ccd_meps_2014-", year_meps, ".rds")))
}

```

## Data merge

```{r}

ccd_directory <- ccd_directory %>% 
  mutate(enrollment = if_else(enrollment < 0, 0, enrollment))


ccd_merged <- ccd_directory %>% 
  left_join(ed_data_wide, by = c("year", "ncessch")) %>% 
  left_join(meps, by = c("year", "ncessch"))


```

Check merge value 

Directory and ed data merge is high: between 93 and 97 percent 
```{r}
ccd_merged %>% 
  mutate(merge_indicator = ifelse(is.na(Total), 0, 1)) %>% 
  group_by(year) %>% 
  summarise(merge_rate = mean(merge_indicator))
```

Directory and meps data merge is good outside of years 2021 and 2022 (waiting on updated data)

```{r}
ccd_merged %>% 
  mutate(merge_indicator = ifelse(is.na(meps_mod_poverty_pct), 0, 1)) %>% 
  group_by(year) %>% 
  summarise(merge_rate = mean(merge_indicator))
```

Check extent of missing and negative enrollment values
```{r}
ccd_merged %>% 
  mutate(enrollment_value = case_when(is.na(enrollment) ~ "Missing",
                                      enrollment == 0 ~ "Zero",
                                      TRUE ~ "Populated")) %>% 
  group_by(year, enrollment_value) %>% 
  count() %>% 
  pivot_wider(names_from = enrollment_value, values_from = n) %>% 
  mutate(share_missing_zero = (Missing + Zero) / (Missing + Zero + Populated)) %>% 
  select(year, share_missing_zero)
```

Drop values where enrollment is missing or equal to zero

```{r}
ccd_merged <- ccd_merged %>% 
  filter(!is.na(enrollment), enrollment != 0)
```

## Clean merged data


*Using MEPS - edit MEPS share and binary if the share is greater than or equal to 20%

```{r}


ccd_merged_clean <- ccd_merged %>% 
  mutate(meps_share = meps_mod_poverty_pct / 100,
         meps_20 = ifelse(meps_share >= 0.2, 1, 0),
         numerator = ifelse(meps_20 == 0 | is.na(meps_20), meps_20, enrollment),
         across(White:Hispanic, ~ifelse(meps_20 == 0 | is.na(meps_20), meps_20, .), .names = "numerator_{.col}"))
```

Look at distribution of MEPS share

```{r}
ccd_merged_clean %>% 
  select(meps_share) %>% 
  ggplot(aes(x = meps_share)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 0.01) +
  theme_minimal() +
  ggtitle("MEPS mod poverty pct share") +
  ylab("Count")
```
 
 Count at the 20 percentage threshold 
 
```{r}
ccd_merged_clean %>% 
  group_by(meps_20) %>% 
  count()
```
 

## Clean geo

```{r}

ccd_merged_clean <- ccd_merged_clean %>% 
  mutate(county_code = ifelse(county_code == -2, NA_character_, county_code),
         county_code = str_pad(county_code, width = 5, side = "left", pad = "0"),
         state = str_sub(county_code, 1, 2))
```

Drop rows with missing county information 

```{r}
ccd_merged_clean <- ccd_merged_clean %>% 
  filter(!is.na(county_code))
```

Create lower_case version of city name 

```{r}
ccd_merged_clean <- ccd_merged_clean %>% 
  mutate(city_name = tolower(city_location))
```

## Create city level data

```{r}
ccd_city <- ccd_merged_clean %>% 
  group_by(year, state, city_name) %>% 
  summarise(enrollment = sum(enrollment),
            enrollment_black = sum(Black),
            enrollment_white = sum(White),
            enrollment_hisp = sum(Hispanic),
           numerator = sum(numerator, na.rm = TRUE),
           numerator_black = sum(numerator_Black, na.rm = TRUE),
           numerator_white = sum(numerator_White, na.rm = TRUE),
           numerator_hisp = sum(numerator_Hispanic, na.rm = TRUE)) 
```


Create shares by enrollment 

```{r}
meps_city <- ccd_city %>% 
  mutate(meps20_total = numerator / enrollment,
         meps20_white = numerator_white/enrollment_white,
         meps20_black = numerator_black/enrollment_black,
         meps20_hisp = numerator_hisp/enrollment_hisp) %>% 
  filter(str_detect(city_name, "atlanta"))
```

