---
title: "Final Data Evaluation SEDA Data"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

# Final Data Test Function

The data team for the Mobility Metrics project has created a function that tests baseline requirements for the structure and content of data produced for each metric. This function relies in part on information provided in a metric-specific final data expectations form, which have been filled out prior to final approval by their technical reviewer. 

These evaluation forms are saved in `10a_final-evaluation`

For programs written in STATA the evaluation process is run externally in this QMD file. 

Detail:

 - *Metric: English Language Arts (ELA)* 
 - *Tester: JP Walsh*
 - *Test date: March 2025*

# Run Test

Put the data through the final evaluation test.

## Read Function

The first step is to source the function `evaluate_final_data.R`. 

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```


## Read Final Data

This file tests the final data for the SEDA metric on students English Language Arts scores (ELA). There are four final files to be tested:

 - `08_education/data/built/SEDA_all_city_2014-2018.csv`
 - `08_education/data/built/SEDA_all_county_2014-2018.csv`
 - `08_education/data/built/SEDA_all_subgroups_city_2014-2018.csv`
 - `08_education/data/built/SEDA_all_subgroups_county_2014-2018.csv`

Read these final files in.

```{r}
final_data_city <- read_csv(here("08_education/data/built/SEDA_all_city_2014-2018.csv"), show_col_types = FALSE)

final_data_county <- read_csv(here("08_education/data/built/SEDA_all_county_2014-2018.csv"), show_col_types = FALSE)

final_data_city_sub <- read_csv(here("08_education/data/built/SEDA_all_subgroups_city_2014-2018.csv"), show_col_types = FALSE)

final_data_county_sub <- read_csv(here("08_education/data/built/SEDA_all_subgroups_county_2014-2018.csv"), show_col_types = FALSE)
```

Separate out the individual subgroups for the city and county files.

City subgroups.
```{r}


final_data_city_income <- final_data_city_sub %>% 
  filter(subgroup_type %in% c("all", "income"))

final_data_city_gender <- final_data_city_sub %>% 
  filter(subgroup_type %in% c("all", "gender"))

final_data_city_race <- final_data_city_sub %>% 
  filter(subgroup_type  %in% c("all", "race-ethnicity"))
```

County subgroups.
```{r}
final_data_county_income <- final_data_county_sub %>% 
  filter(subgroup_type %in% c("all", "income"))

final_data_county_gender <- final_data_county_sub %>% 
  filter(subgroup_type %in% c("all", "gender"))

final_data_county_race <- final_data_county_sub %>% 
  filter(subgroup_type %in% c("all", "race-ethnicity"))
```


## Run Evaluation Test

Enter the information into the `evaluate_final_data` function. 

In order to run the function, we fill out the relevant arguments. The SEDA data have confidence intervals and there are subgroup files.

### All

City level data.
```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_overall_place.csv",
                data = final_data_city,  geography = "place",
                subgroups = FALSE, confidence_intervals = TRUE)
```

County level data.
```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_overall_county.csv",
                data = final_data_county,  geography = "county",
                subgroups = FALSE, confidence_intervals = TRUE)

```

### Subgroups

City level subgroup data.
```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_race_ethnicity_place.csv",
                data = final_data_city_race,  geography = "place",
                subgroups = TRUE, confidence_intervals = TRUE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_income_place.csv",
                data = final_data_city_income,  geography = "place",
                subgroups = TRUE, confidence_intervals = TRUE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_gender_place.csv",
                data = final_data_city_gender,  geography = "place",
                subgroups = TRUE, confidence_intervals = TRUE)
```

County level data.
```{r}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_race_ethnicity_county.csv",
                data = final_data_county_race,  geography = "county",
                subgroups = TRUE, confidence_intervals = TRUE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_income_county.csv",
                data = final_data_county_income,  geography = "county",
                subgroups = TRUE, confidence_intervals = TRUE)

evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_ela_gender_county.csv",
                data = final_data_county_gender,  geography = "county",
                subgroups = TRUE, confidence_intervals = TRUE)

```

Confirmation of passage.
