---
title: "Number and share of public-school children who are ever homeless during the school year — County"
author: "Manuel Alcalá Kovalski"
date: now
format:
  html:
    self-contained: true
    toc: true
    css: ../12_housing-stability/www/web_report.css
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---


This program creates the number and share of homeless students by county for 2019-2022 (school years 2019-20 through 2022-23), including breakdowns by race/ethnicity.

**Data notes:**

- Data suppression: data are suppressed when values are between 0-2, but if only one value is suppressed the next smallest number is also suppressed
- Original code replaced all suppressed data with the midpoint (1) but this does not yield numbers that align perfectly with [federal reports](https://nche.ed.gov/wp-content/uploads/2020/01/Federal-Data-Summary-SY-15.16-to-17.18-Published-1.30.2020.pdf)
- Data are unduplicated by LEA but may not be unduplicated by county if students switch between LEAs
- Raw data is now posted on EdDataExpress instead of EdFacts
- Starting in 2019-20, subgrant_status is no longer used to determine suppression information

## Housekeeping

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(here)
library(janitor)
library(educationdata)
library(glue)

options(scipen = 999)


# Define years for analysis
years <- 2019:2023

# Define paths
county_file <- here("geographic-crosswalks", "data", "county-populations.csv")
data_dir <- here("12_housing-stability", "data")
raw_dir <- here(data_dir, "raw")
intermediate_dir <- here(data_dir, "intermediate")
built_dir <- here(data_dir, "built")
final_dir <- here("12_housing-stability", "data", "final")

# Create directories if they don't exist
dir.create(data_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(raw_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(intermediate_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(built_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(final_dir, showWarnings = FALSE, recursive = TRUE)

```

## Import and Prepare County Crosswalk

```{r}
#| label: county-crosswalk

# Import county crosswalk file
county_crosswalk <- read_csv(
  here("geographic-crosswalks", "data", "county-populations.csv"),
  show_col_types = FALSE
) |>
  select(-population, -state_name, -county_name) |>
  mutate(
    # Convert county to string with leading zeros (3 characters)
    county = str_pad(county, width = 3, pad = "0"),
    # Convert state to string with leading zeros (2 characters)
    state = str_pad(state, width = 2, pad = "0")
  )



```

## Download CCD District Data

The Common Core of Data (CCD) provides district-level enrollment data.

```{r}
#| label: download-ccd-data

# Download CCD district directory data with county codes and enrollment
ccd_data <- map_dfr(years, function(year) {
  message(glue("Downloading CCD data for {year}..."))
  
  get_education_data(
    level = "school-districts",
    source = "ccd",
    topic = "directory",
    filters = list(year = year),
    add_labels = FALSE
  ) |>
    select(year, leaid, county_code, enrollment) |>
    filter(!is.na(county_code))
})

# Process county codes
ccd_data_processed <- ccd_data |>
  mutate(
    county_code = str_pad(county_code, width = 5, pad = "0"),
    state = str_sub(county_code, 1, 2),
    county = str_sub(county_code, 3, 5)
  ) |>
  filter(str_length(county) == 3) |>
  select(-county_code)



```

## Download CCD Enrollment by Race

```{r}
#| label: download-ccd-race

# Download CCD district enrollment data by race
ccd_race_data <- map_dfr(years, function(year) {
  message(glue("Downloading CCD enrollment by race for {year}..."))
  
  get_education_data(
    level = "school-districts",
    source = "ccd",
    topic = "enrollment",
    subtopic = "race",
    filters = list(
      year = year,
      grade = 99,  # All grades
      sex = 99     # All sexes
    ),
    add_labels = FALSE
  ) |>
    select(year, leaid, race, enrollment) |>
    filter(race != 99)  # Remove "All races"
})

# Reshape to wide format
ccd_race_wide <- ccd_race_data |>
  mutate(
    race_cat = case_when(
      race == 1 ~ "enroll_white",
      race == 2 ~ "enroll_black",
      race == 3 ~ "enroll_hispanic",
      race == 4 ~ "enroll_asian",
      race == 5 ~ "enroll_amin_an",
      race == 6 ~ "enroll_nh_pi",
      race == 7 ~ "enroll_twomore",
      TRUE ~ NA_character_
    )
  ) |>
  select(-race) |>
  pivot_wider(
    names_from = race_cat,
    values_from = enrollment,
    values_fill = 0
  ) |>
  mutate(
    # Combine Asian, American Indian/Alaska Native, Native Hawaiian/Pacific Islander, 
    # and Two or More into "Other"
    enroll_other = enroll_asian + enroll_amin_an + enroll_nh_pi + enroll_twomore
  ) |>
  select(year, leaid, enroll_white, enroll_black, enroll_hispanic, enroll_other)

# Save intermediate file
write_rds(ccd_race_wide, here(intermediate_dir, "ccd_lea_recent_county_race.rds"))

```

## Merge CCD Datasets

```{r}
#| label: merge-ccd

ccd_merged <- ccd_data_processed |>
  left_join(ccd_race_wide, by = c("year", "leaid"))

# Save merged file
write_rds(ccd_merged, here(intermediate_dir, "ccd_lea_recent_county_merged.rds"))

```

## Download EdDataExpress Homelessness Data

EdDataExpress publishes homeless student enrollment data. We download data for 2019-20 through 2022-23.

**Note:** 2022-23 data must be manually downloaded due to changes in EdDataExpress website. 

Instructions for manual download:
1. Go to: https://eddataexpress.ed.gov/download/data-builder/data-download-tool
2. Set filters: Homeless Students, 2022-2023, Local Education Agency
3. Ensure data group = 655 ("Homeless students enrolled")
4. Download and save to raw folder as "2022-23 LEA homeless student data.csv"

```{r}
#| label: download-eddataexpress

# Download URLs for 2019-2021
download_urls <- tribble(
  ~year, ~url,
  2019, "https://eddataexpress.ed.gov/sites/default/files/data_download/EID_6526/SY1920_FS118_DG655_LEA_data_files.zip",
  2020, "https://eddataexpress.ed.gov/sites/default/files/data_download/EID_8321/SY2021_FS118_DG655_LEA_data_files.zip",
  2021, "https://eddataexpress.ed.gov/sites/default/files/data_download/EID_11718/SY2122_FS118_DG655_LEA_data_files.zip"
)

# Download and extract files
walk2(download_urls$year, download_urls$url, function(year, url) {
  zip_file <- here(raw_dir, glue("EdDataEx Homelessness {year}.zip"))
  
  if (!file.exists(zip_file)) {
    message(glue("Downloading EdDataExpress data for {year}..."))
    download.file(url, zip_file)
    unzip(zip_file, exdir = raw_dir)
  }
})

```

## Import and Clean EdDataExpress Data

```{r}
#| label: import-eddataexpress

# Function to read and process EdDataExpress files
read_eddataexpress <- function(year) {
  message(glue("Processing EdDataExpress data for {year}..."))
  
  # File names vary by year
  file_name <- case_when(
    year == 2019 ~ "SY1920_FS118_DG655_LEA.csv",
    year == 2020 ~ "SY2021_FS118_DG655_LEA.csv",
    year == 2021 ~ "SY2122_FS118_DG655_LEA.csv",
    year == 2022 ~ "2022-23 LEA homeless student data.csv"
  )
  
  read_csv(here(raw_dir, file_name), show_col_types = FALSE) |>
    clean_names() |>
    mutate(year = year)
}

# Read all years
eddataexpress_raw <- map_dfr(years, read_eddataexpress)

# Filter and reshape
eddataexpress_clean <- eddataexpress_raw |>
  filter(
    datagroup == 655,  # Homeless students enrolled
    !subgroup %in% c(
      "", "Children with disabilities", "English Learner", 
      "Migratory students", "Unaccompanied Youth",
      "Children with one or more disabilities (IDEA)"
    )
  ) |>
  mutate(
    subgroup = case_when(
      subgroup == "American Indian or Alaska Native" ~ "amin_an",
      subgroup == "Black or African American" ~ "black",
      subgroup == "Hispanic/Latino" ~ "hispanic",
      subgroup == "White" ~ "white",
      subgroup == "Two or more races" ~ "twomore",
      subgroup == "Native Hawaiian or Other Pacific Islander" ~ "nh_pi",
      subgroup == "Asian" ~ "asian",
      subgroup == "All Students in LEA" ~ "homeless",
      TRUE ~ subgroup
    )
  ) |>
  select(state, ncesleaid, lea, year, subgroup, value) |>
  pivot_wider(
    names_from = subgroup,
    values_from = value
  ) |>
  rename(leaid = ncesleaid)

# Save intermediate file
write_rds(eddataexpress_clean, here(intermediate_dir, "eddataexpress_all_years.rds"))

```

## Process Suppression and Create Bounds

Data suppression follows specific rules. Suppressed values (marked as "S") represent 1-2 students and are replaced with 1 for aggregation purposes.

```{r}
#| label: process-suppression

# Process suppressions and create confidence bounds
homeless_data <- eddataexpress_clean |>
  mutate(
    # Create FIPS state code
    fipst = str_pad(str_sub(as.character(leaid), 1, 2), width = 2, pad = "0")
  )

# Function to process suppression for each variable
process_suppression <- function(data, var_name) {
  var_sym <- sym(var_name)
  
  data |>
    mutate(
      supp_{var_name} := if_else(!!var_sym == "S", 1, NA_real_),
      !!var_sym := if_else(!!var_sym == "S", "1", !!var_sym),
      !!var_sym := as.numeric(!!var_sym)
    ) |>
    group_by(year, fipst) |>
    mutate(
      min_{var_name} := min(!!var_sym, na.rm = TRUE),
      count_supp_{var_name} := sum(supp_{var_name}, na.rm = TRUE),
      {var_name}_lower_ci := if_else(!is.na(supp_{var_name}), 0, !!var_sym),
      {var_name}_upper_ci := case_when(
        !is.na(supp_{var_name}) & count_supp_{var_name} <= 2 ~ min_{var_name},
        !is.na(supp_{var_name}) ~ 2,
        TRUE ~ !!var_sym
      )
    ) |>
    ungroup()
}

# Apply suppression processing to all race/ethnicity variables
race_vars <- c("homeless", "black", "hispanic", "white", "amin_an", 
               "twomore", "nh_pi", "asian")

homeless_processed <- reduce(race_vars, process_suppression, .init = homeless_data)

# Create "other" category combining multiple races
homeless_processed <- homeless_processed |>
  mutate(
    other = rowSums(select(., twomore, nh_pi, asian, amin_an), na.rm = TRUE),
    other = if_else(is.na(other), 1, other)
  ) |>
  process_suppression("other")

# Keep only needed variables
homeless_final <- homeless_processed |>
  select(
    year, leaid, 
    matches("homeless|black|hispanic|white|other"),
    -matches("amin_an|twomore|nh_pi|asian")
  ) |>
  mutate(
    leaid = str_pad(as.character(leaid), width = 7, pad = "0")
  )

```

## Merge with CCD and Aggregate to County Level

```{r}
#| label: merge-and-aggregate

# Read CCD merged data
ccd_merged <- read_rds(here(intermediate_dir, "ccd_lea_recent_county_merged.rds"))

# Merge homeless data with CCD
homeless_merged <- homeless_final |>
  left_join(
    ccd_merged,
    by = c("year", "leaid")
  ) |>
  filter(!is.na(county))

# Replace missing enrollments with zeros
homeless_merged <- homeless_merged |>
  mutate(
    enrollment = if_else(enrollment < 0 | is.na(enrollment), 0, enrollment),
    across(
      c(homeless, black, white, hispanic, other,
        matches("_upper_ci$|_lower_ci$")),
      ~if_else(enrollment == 0, 0, .x)
    ),
    across(
      matches("^supp_"),
      ~if_else(enrollment == 0, 0, .x)
    )
  )

# Create enrollment variables based on suppression
homeless_merged <- homeless_merged |>
  mutate(
    across(
      c(homeless, black, white, hispanic, other),
      list(
        enroll_nonsupp = ~if_else(get(paste0("supp_", cur_column())) != 1, enrollment, NA_real_),
        enroll_supp = ~if_else(get(paste0("supp_", cur_column())) == 1, enrollment, NA_real_)
      ),
      .names = "{.fn}_{.col}"
    )
  )

# Aggregate to county level
county_homeless <- homeless_merged |>
  group_by(year, state, county) |>
  summarize(
    across(
      c(homeless, black, hispanic, white, other,
        matches("_lower_ci$|_upper_ci$|^supp_|^enroll_")),
      ~sum(.x, na.rm = TRUE)
    ),
    enrollment = sum(enrollment, na.rm = TRUE),
    .groups = "drop"
  ) |>
  rename_with(
    ~str_replace(.x, "^(homeless|black|hispanic|white|other)$", "\\1_count"),
    matches("^(homeless|black|hispanic|white|other)$")
  ) |>
  rename_with(
    ~str_replace(.x, "_lower_ci$", "_count_lb"),
    matches("_lower_ci$")
  ) |>
  rename_with(
    ~str_replace(.x, "_upper_ci$", "_count_ub"),
    matches("_upper_ci$")
  ) |>
  rename_with(
    ~str_replace(.x, "^supp_", "\\1_districts_suppress"),
    matches("^supp_")
  )

```

## Calculate Shares and Quality Metrics

```{r}
#| label: calculate-metrics

# Read CCD enrollment by race for denominators
ccd_race <- read_rds(here(intermediate_dir, "ccd_lea_recent_county_race.rds"))

# Aggregate CCD race enrollment to county
county_enroll_race <- homeless_merged |>
  group_by(year, state, county) |>
  summarize(
    across(
      c(enroll_white, enroll_black, enroll_hispanic, enroll_other),
      ~sum(.x, na.rm = TRUE)
    ),
    .groups = "drop"
  )

# Join race enrollment data
county_homeless <- county_homeless |>
  left_join(county_enroll_race, by = c("year", "state", "county"))

# Calculate shares and coverage
county_homeless <- county_homeless |>
  mutate(
    # Share of all students who are homeless
    homeless_share = homeless_count / enrollment,
    coverage_homeless = enroll_nonsupp_homeless / enrollment,
    
    # Share of each race/ethnicity that is homeless (using race-specific enrollment)
    black_share = black_count / enroll_black,
    white_share = white_count / enroll_white,
    hispanic_share = hispanic_count / enroll_hispanic,
    other_share = other_count / enroll_other,
    
    # Coverage by race
    coverage_black = enroll_nonsupp_black / enroll_black,
    coverage_white = enroll_nonsupp_white / enroll_white,
    coverage_hispanic = enroll_nonsupp_hispanic / enroll_hispanic,
    coverage_other = enroll_nonsupp_other / enroll_other
  )

# Create quality variables
county_homeless <- county_homeless |>
  mutate(
    across(
      c(homeless, black, white, hispanic, other),
      list(
        quality_count = ~case_when(
          get(paste0(cur_column(), "_count_ub")) / get(paste0(cur_column(), "_count_lb")) <= 1.05 ~ 1,
          get(paste0(cur_column(), "_count_ub")) / get(paste0(cur_column(), "_count_lb")) <= 1.1 ~ 2,
          !is.na(get(paste0(cur_column(), "_count"))) ~ 3,
          TRUE ~ NA_real_
        )
      ),
      .names = "{.col}_{.fn}"
    )
  )

# Adjust quality for small enrollment
county_homeless <- county_homeless |>
  mutate(
    homeless_quality_count = if_else(enrollment < 30, 3, homeless_quality_count)
  )

```

## Apply Data Quality Rules

```{r}
#| label: quality-rules

# Replace with -1 (NA) when homeless count < 10
county_homeless <- county_homeless |>
  mutate(
    across(
      c(homeless_count, black_count, white_count, hispanic_count, other_count),
      ~if_else(homeless_count < 10, -1, .x)
    ),
    across(
      matches("_count_lb$|_count_ub$"),
      ~if_else(homeless_count < 10, -1, .x)
    ),
    across(
      c(homeless_share, black_share, white_share, hispanic_share, other_share),
      ~if_else(homeless_count < 10, -1, .x)
    ),
    across(
      matches("_quality_count$"),
      ~if_else(homeless_count < 10, -1, .x)
    ),
    across(
      matches("_quality_count$"),
      ~if_else(homeless_count >= 10 & homeless_count < 30, 3, .x)
    )
  )

# Set all to -1 if share > 1
county_homeless <- county_homeless |>
  mutate(
    across(
      c(homeless, black, white, hispanic, other),
      list(
        count = ~if_else(get(paste0(cur_column(), "_share")) > 1 & 
                          !is.na(get(paste0(cur_column(), "_share"))), -1, .x),
        count_lb = ~if_else(get(paste0(cur_column(), "_share")) > 1 & 
                             !is.na(get(paste0(cur_column(), "_share"))), -1, .x),
        count_ub = ~if_else(get(paste0(cur_column(), "_share")) > 1 & 
                             !is.na(get(paste0(cur_column(), "_share"))), -1, .x),
        share = ~if_else(get(paste0(cur_column(), "_share")) > 1 & 
                          !is.na(get(paste0(cur_column(), "_share"))), -1, .x),
        quality_count = ~if_else(get(paste0(cur_column(), "_share")) > 1 & 
                                  !is.na(get(paste0(cur_column(), "_share"))), -1, .x)
      ),
      .names = "{.col}_{.fn}"
    )
  )

# Check subgroup totals vs total homeless
county_homeless <- county_homeless |>
  mutate(
    # Create non-missing versions for summing
    across(
      c(black_count, white_count, hispanic_count, other_count),
      ~if_else(.x == -1, NA_real_, .x),
      .names = "{.col}_missing"
    ),
    # Calculate total of race subgroups
    total_race = rowSums(select(., ends_with("_count_missing")), na.rm = TRUE),
    share_race_total = (total_race / homeless_count) * 100,
    share_race_total = if_else(is.na(black_count_missing), NA_real_, share_race_total)
  )

# Replace with -1 when subgroup totals are < 70% or > 110% of total homeless
county_homeless <- county_homeless |>
  mutate(
    across(
      c(black, white, hispanic, other),
      list(
        count = ~if_else((share_race_total < 70 | share_race_total > 110) & 
                          !is.na(share_race_total), -1, .x),
        count_lb = ~if_else((share_race_total < 70 | share_race_total > 110) & 
                             !is.na(share_race_total), -1, .x),
        count_ub = ~if_else((share_race_total < 70 | share_race_total > 110) & 
                             !is.na(share_race_total), -1, .x),
        share = ~if_else((share_race_total < 70 | share_race_total > 110) & 
                          !is.na(share_race_total), -1, .x),
        quality_count = ~if_else((share_race_total < 70 | share_race_total > 110) & 
                                  !is.na(share_race_total), -1, .x)
      ),
      .names = "{.col}_{.fn}"
    )
  )

# Apply suppression rules: set to -1 when count <= 2
county_homeless <- county_homeless |>
  mutate(
    across(
      c(homeless, black, white, hispanic, other),
      list(
        count_lb = ~if_else(get(paste0(cur_column(), "_count")) <= 2, -1, .x),
        count_ub = ~if_else(get(paste0(cur_column(), "_count")) <= 2, -1, .x),
        share = ~if_else(get(paste0(cur_column(), "_count")) <= 2, -1, .x),
        quality_count = ~if_else(get(paste0(cur_column(), "_count")) <= 2, -1, .x),
        count = ~if_else(get(paste0(cur_column(), "_count")) <= 2, -1, .x)
      ),
      .names = "{.col}_{.fn}"
    )
  )

```

## Merge with County Crosswalk

```{r}
#| label: merge-crosswalk

# Read county crosswalk
county_crosswalk <- read_rds(here(intermediate_dir, "countyfile.rds"))

# Merge and filter to relevant years
county_final <- county_crosswalk |>
  filter(year %in% 2019:2022) |>
  left_join(
    county_homeless,
    by = c("year", "state", "county")
  )

# Clean up columns
county_final <- county_final |>
  select(
    -enrollment, -starts_with("coverage"), -starts_with("enroll_"),
    -ends_with("_districts_suppress"), -starts_with("min_"), 
    -starts_with("count_supp_"), -ends_with("_missing"),
    -total_race
  ) |>
  arrange(desc(year), state, county)

```

## Quality Checks

```{r}
#| label: quality-checks

# Summary statistics by year
county_final |>
  group_by(year) |>
  summarize(
    n_counties = n(),
    homeless_missing = sum(is.na(homeless_count)),
    mean_homeless_share = mean(homeless_share[homeless_share > 0], na.rm = TRUE),
    median_homeless_share = median(homeless_share[homeless_share > 0], na.rm = TRUE)
  )

# Check coverage by quality level
county_final |>
  group_by(homeless_quality_count) |>
  summarize(
    n = n(),
    mean_share = mean(homeless_share[homeless_share > 0], na.rm = TRUE)
  )

# Verify data quality: count should always be between lower and upper bounds
county_final |>
  filter(
    homeless_count > 0,
    (homeless_count < homeless_count_lb | homeless_count > homeless_count_ub)
  ) |>
  nrow() # Should be 0

# Check race/ethnicity totals
county_final |>
  summarize(
    mean_share_race_total = mean(share_race_total[!is.na(share_race_total)]),
    median_share_race_total = median(share_race_total[!is.na(share_race_total)])
  )

```

## Visual Checks

```{r}
#| label: visual-checks

# Distribution of homeless share by year
county_final |>
  filter(homeless_share > 0, homeless_share < 1) |>
  ggplot(aes(x = homeless_share)) +
  geom_histogram(bins = 50) +
  facet_wrap(~year) +
  labs(
    title = "Distribution of Homeless Share by Year",
    x = "Share of Students Homeless",
    y = "Count of Counties"
  )

# Distribution by quality level
county_final |>
  filter(homeless_share > 0, homeless_share < 1, homeless_quality_count > 0) |>
  ggplot(aes(x = homeless_share)) +
  geom_histogram(bins = 50) +
  facet_grid(year ~ homeless_quality_count, 
             labeller = label_both) +
  labs(
    title = "Distribution of Homeless Share by Year and Quality",
    x = "Share of Students Homeless",
    y = "Count of Counties"
  )

```

## Export Final Data

```{r}
#| label: export

# Convert -1 to "NA" and missing to "NA" for export
county_export <- county_final |>
  mutate(
    across(
      where(is.numeric),
      ~case_when(
        .x == -1 ~ "NA",
        is.na(.x) ~ "NA",
        TRUE ~ as.character(.x)
      )
    )
  )

# Create quality_share variables (same as quality_count)
county_export <- county_export |>
  mutate(
    homeless_quality_share = homeless_quality_count,
    black_quality_share = black_quality_count,
    white_quality_share = white_quality_count,
    hispanic_quality_share = hispanic_quality_count,
    other_quality_share = other_quality_count
  )

# Export "all" (total homeless) data
county_export_all <- county_export |>
  select(
    year, state, county,
    count_homeless = homeless_count,
    count_homeless_lb = homeless_count_lb,
    count_homeless_ub = homeless_count_ub,
    share_homeless = homeless_share,
    count_homeless_quality = homeless_quality_count,
    share_homeless_quality = homeless_quality_share
  )

write_csv(
  county_export_all,
  here(final_dir, "homelessness_2019-2022_county.csv")
)

# Prepare subgroup data for export
county_export_subgroup <- county_export |>
  select(
    year, state, county,
    homeless_count, homeless_count_lb, homeless_count_ub, 
    homeless_share, homeless_quality_count, homeless_quality_share,
    black_count, black_count_lb, black_count_ub, 
    black_share, black_quality_count, black_quality_share,
    white_count, white_count_lb, white_count_ub, 
    white_share, white_quality_count, white_quality_share,
    hispanic_count, hispanic_count_lb, hispanic_count_ub, 
    hispanic_share, hispanic_quality_count, hispanic_quality_share,
    other_count, other_count_lb, other_count_ub, 
    other_share, other_quality_count, other_quality_share
  ) |>
  rename(code_county = county) |>
  pivot_longer(
    cols = -c(year, state, code_county),
    names_to = c(".value", "subgroup"),
    names_pattern = "(.+)_(homeless|black|white|hispanic|other)$"
  ) |>
  mutate(
    subgroup_type = if_else(subgroup == "homeless", "all", "race-ethnicity"),
    subgroup = case_when(
      subgroup == "homeless" ~ "All",
      subgroup == "black" ~ "Black, Non-Hispanic",
      subgroup == "white" ~ "White, Non-Hispanic",
      subgroup == "hispanic" ~ "Hispanic",
      subgroup == "other" ~ "Other Races and Ethnicities",
      TRUE ~ subgroup
    )
  ) |>
  rename(
    county = code_county,
    count_homeless = count,
    count_homeless_lb = count_lb,
    count_homeless_ub = count_ub,
    share_homeless = share,
    count_homeless_quality = quality_count,
    share_homeless_quality = quality_share
  ) |>
  select(
    year, state, county, subgroup_type, subgroup,
    count_homeless, count_homeless_lb, count_homeless_ub,
    share_homeless, count_homeless_quality, share_homeless_quality
  ) |>
  arrange(desc(year), state, county, subgroup_type, subgroup)

write_csv(
  county_export_subgroup,
  here(final_dir, "homelessness_2019-2022_subgroups_county.csv")
)

message("Export complete!")
message("Files saved to: ", final_dir)

```

## Session Info

```{r}
#| label: session-info

sessionInfo()

```

