---
title: "Compare Local Data with Data Catalog"
author: "Aaron R. Williams"
date: now
abstract: This document compares the UMI data on the Urban Institute catalog with updated versions of the data that are local. This is meant to help with quality control when updating UMI data. 
date-format: "MMMM D, YYYY h:mm a z"
format: 
  html:
    toc: true
embed-resources: true
editor: source
execute: 
  code-fold: true
editor_options: 
  chunk_output_type: console
---

## Setup

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(here)

source(here("functions", "construction", "compare_data.R"))

```

## Load Data

This section loads the remote and local versions of the data and combines them into a list. The URLs for the remote data need to be updated before each run.

```{r}
#| label: load-data

# Increase timeout for large file downloads
options(timeout = 300)

read_data <- function(file, date, local_path, root_url) {

  message("Reading: ", file)

  data <- 
    list(
      `local` = read_csv(here(local_path, str_remove(file, "_0")), show_col_types = FALSE),
      `catalog` = read_csv(paste0(root_url, date, file), show_col_types = FALSE)
    )
  
  return(data)
  
}

root_url <- "https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/"
local_path <- here("data")

data_urls <- tribble(
  ~date, ~file,
  "2026/04/07/", "11_mobility-metrics_county_longitudinal.csv",
  "2026/04/07/", "12_mobility-metrics_county_race-ethnicity_longitudinal.csv",
  "2026/04/07/", "13_mobility-metrics_county_race-share_longitudinal.csv",
  "2026/04/07/", "14_mobility-metrics_county_income_longitudinal.csv",
  "2026/04/07/", "15_mobility-metrics_county_age_longitudinal.csv",
  "2026/04/07/", "18_mobility-metrics_county_disability_longitudinal.csv",
  "2026/04/07/", "19_mobility-metrics_county_industry_longitudinal.csv",
  "2026/04/07/", "20_mobility-metrics_county_mothers_education_longitudinal.csv",
  "2026/04/07/", "61_mobility-metrics_place_longitudinal.csv",
  "2026/04/07/", "62_mobility-metrics_place_race-ethnicity_longitudinal.csv",
  "2026/04/07/", "63_mobility-metrics_place_race-share_longitudinal_0.csv",
  "2026/04/07/", "64_mobility-metrics_place_income_longitudinal_0.csv",
  "2026/04/07/", "65_mobility-metrics_place_age_longitudinal_0.csv",
  "2026/04/07/", "66_mobility-metrics_place_gender_longitudinal_0.csv",
  "2026/04/07/", "67_mobility-metrics_place_tenure_longitudinal_0.csv",
  "2026/04/07/", "68_mobility-metrics_place_disability_longitudinal_0.csv"
)

data <- map2(
  .x = data_urls$file |>
    purrr::set_names(),
  .y = data_urls$date,
  .f = ~read_data(
    file = .x, 
    date = .y,
    local_path = local_path, 
    root_url = root_url
  )
)

```

## Data Quality

This iterates a function over each pair of data sets and calculates several comparisons:

- Identifies observations (rows) unique to local or catalog datasets
- Counts differing observations by year
- Lists variables (columns) present in only one dataset
- Identifies shared variables that have different values

```{r}
#| label: compare-data
#| results: asis

walk2(
  .x = data,
  .y = names(data),
  .f = ~compare_data(.x, .y)
)

```

## Medians

This sections repeats the comparison for the medians file.

```{r}
#| label: compare-medians

medians <- 
  list(
    `local` = read_csv(here("99_construct-database", "county_medians_year_suppressed.csv"), show_col_types = FALSE),
    `catalog` = read_csv("https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2026/04/07/county_medians_year_suppressed.csv", show_col_types = FALSE)
  )


compare_data(data_list = medians, "medians")



```


