---
title: "Housing Value"
author: " Manu Alcalá Kovalski"
date: now
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
    code-fold: show
    code-tools: true
editor_options:
  chunk_output_type: console
execute:
  warning: false
---

```{=html}
<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>
```
<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

This program creates the **Housing Value** metric defined as the ratio between
the share of a community’s housing wealth held by a racial or ethnic group to
the share of households of the same group metric

Initially written by Jung Hyun Choi and updated by Manu Alcalá Kovalski
in 2023-2024.

## Process

-   [Housekeeping](#housekeeping)
-   [Read Data](#read-data)
-   [Clean Data](#clean-data)
-   [Merge Crosswalk](#merge-crosswalk)
-   [Create Housing Value Metric](#create-metric)
-   [Evaluate Results](#evaluate-results)
-   [Export](#export)

## Housekeeping {#sec-housekeeping}

```{r}
#| label: setup

options(scipen = 999)
librarian::shelf(
  tidyverse,
  ipumsr,
  here,
  glue,
  labelled,
  skimr,
  UrbanInstitute / urbnthemes,
  tidylog,
  assertr, 
  fs
)
set_urbn_defaults(style = "print")

source(here::here("functions", "testing", "evaluate_final_data.R"))
```

## Methodology Overview

### Metric Definition

The housing value metric captures racial/ethnic disparities in home value accumulation:

**Formula:** 
$$\text{Ratio} = \frac{\text{Share of Housing Value}}{\text{Share of Households}}$$

Where:
- **Share of Housing Value** = Group's total home values / Total home values in geography
- **Share of Households** = Group's households / Total households in geography

### Interpretation

- **Ratio = 1.0**: Group's housing wealth share matches their population share (proportional)
- **Ratio > 1.0**: Group holds more housing wealth than expected by population (overrepresentation)
- **Ratio < 1.0**: Group holds less housing wealth than expected by population (underrepresentation)

**Example:** If Black households are 20% of county households but hold only 10% of housing wealth, the ratio is 0.5:1.0, indicating significant wealth underrepresentation.

### Population Scope

- **Homeowners only** for housing value calculations (OWNERSHP = owned or being bought)
- **All households** included in household share denominator
- **Group quarters excluded** (correctional facilities, nursing homes, dorms, barracks)
- **Age subgroups:** <45 years vs. ≥45 years (captures wealth accumulation over life course)

### Geographic Aggregation

**Challenge:** ACS microdata provides PUMA-level geography (100,000+ population), but we need county/place estimates.

**Solution:** PUMA-to-county/place crosswalks with allocation factors (`afact`):
- Household weights multiplied by `afact` to distribute across overlapping geographies
- `afact` represents the proportion of a PUMA's population in each geography
- **Pre-2022:** Uses 2012 PUMA to 2014 county boundaries
- **2022+:** Uses 2020 PUMA to 2020 county boundaries (reflects 2020 Census updates)

### Race/Ethnicity Categories

IPUMS detailed categories collapsed into four groups:
- **white_nh:** White, non-Hispanic
- **black_nh:** Black/African American, non-Hispanic
- **hispanic:** Hispanic (any race)
- **other_nh:** Other races, non-Hispanic (includes Asian, AIAN, Pacific Islander, multiracial)

### Read Data {#sec-read-data}

### Configuration Parameters

**ACS Years:** 2014-2024 (excluding 2020 due to data quality issues)

**IPUMS Microdata File:** Update the `microdata` variable (around line 148) to match your downloaded XML filename after initial extract.

**Extract Settings:** The data extract request is cached (`eval: false` in chunk below). Set `eval: true` only when requesting new data or updating variables.

### IPUMS API Setup

We extract the 1-year ACS files from the IPUMS API using `ipumsr`. You will need an API key to access IPUMS data.

**First-time setup:**
1. Register at [IPUMS USA](https://uma.pop.umn.edu/usa/user/new?return_url=https%3A%2F%2Fusa.ipums.org%2Fusa-action%2Fmenu)
2. Generate an API key from your account settings
3. Run once: `set_ipums_api_key(api_key = "paste-your-key-here", save = TRUE)`
4. Restart R session for changes to take effect

**How it works:** By default, `ipumsr` API functions assume your key is stored in the **IPUMS_API_KEY** environment variable. Storing it in `.Renviron` saves typing and prevents accidentally sharing your key (e.g., on GitHub).

For more information, see the `ipumsr` [getting started vignette](https://tech.popdata.org/ipumsr/articles/ipums.html).

```{r}
#| label: query-ipums-api
#| cache: true
#| eval: false
## 2.1. ACS: keep households living not living in gq

# Run this the first time you run this code to avoid requesting the same data
# multiple times

acs_vars <- c(
  "YEAR", "SAMPLE", "SERIAL", "CBSERIAL", "HHWT", "CLUSTER",
  "STATEFIP", "COUNTYFIP", "MET2013", "CITY", "PUMA", "STRATA",
  "GQ", "OWNERSHP", "OWNERSHPD", "VALUEH", "PERNUM", "PERWT",
  "RACE", "RACED", "HISPAN", "HISPAND", "AGE"
)

years <- 2014:2024

if(length(fs::dir_ls(here::here("01_financial-well-being", "data"), glob = "*.xml")) == 0){
  
  usa_ext_def <- define_extract_usa(
  description = "ACS 1 Year 2014-2024",
  samples = paste0("us", years, "a"),
  variables = acs_vars
)

usa_ext_submitted <- submit_extract(usa_ext_def)
usa_ext_complete <- wait_for_extract(usa_ext_submitted)

# download_extract() will return a character vector including the file paths
# to all downloaded files.
# Create a data folder before running this if it doesn't exist

fs::dir_create(here("01_financial-well-being/data"))

filepath <- download_extract(usa_ext_complete,
  download_dir = here("01_financial-well-being/data")
)
}

```

### Clean data {#sec-clean-data}

We load the 2014-2024 1 Year ACS microdata we downloaded from IPUMS above. IPUMS
loads the data as a person-level data set with some household
variables. Therefore, we can analyze the data at the household level by keeping
the first person sampled in the household.

**Important:** 1-year ACS data for 2020 is excluded because Census Bureau warns these data [should be considered experimental](https://www.census.gov/programs-surveys/acs/technical-documentation/table-and-geography-changes/2020/1-year.html) due to COVID-19 disruptions affecting data collection quality and reliability.

```{r}
#| label: load-acs
#| cache: true
#| dependson: "import-acs-one-year"
#| include: false

# Manually set the microdata object equal to whatever the file you downloaded in
# the previous chunk is called. In my case it is, "usa_00018.xml" but that will
# be different for you.
# **ACTION REQUIRED:** Update this filename to match your downloaded .xml file
microdata <- "usa_00337.xml"

acs1yr_raw <-
  read_ipums_micro(here(glue("01_financial-well-being/data/{microdata}"))) |>
  janitor::clean_names() |>
  rename(
    ownership = ownershp,
    state = statefip,
    county = countyfip,
  ) |>
  filter(
    year != 2020,
    pernum == 1
  )
```

The
[ipumsr](https://tech.popdata.org/ipumsr/articles/value-labels.html#remove-unused-value-labels)
documentation says that:

> "labelled vectors are not intended for use throughout the entire analysis
> process. Instead, they should be used during the initial data preparation
> process to convert raw data into values that are more meaningful. These can
> then be converted to other variable types (often factors) for analysis.
> Unfortunately, this isn’t a process that can typically be automated, as it
> depends primarily on the research questions the data will be used to address.
> However, ipumsr provides several functions to manipulate value labels to make
> this process easier.

A useful way of finding missing value labels is to call `ipums_var_info(acs1yr)`
and inspect the `val_labels`.


```{r}
#| label: handle-missingness


acs1yr <-
  acs1yr_raw |>
  mutate(
    valueh = lbl_na_if(valueh, ~ .lbl %in% c("Missing", "N/A")),
  )

```

Next, we exclude individuals living in group quarters because we only want
households for this metric. In particular, we exclude people living in places
other than housing units (e.g., correctional facilities, nursing homes, mental
hospitals, dorms, barracks, group homes, missions). For more information, see
the [variable
description](https://usa.ipums.org/usa-action/variables/GQ#codes_section) on
IPUMS USA.

```{r}
#| label: keep-households-only
acs1yr <-
  acs1yr |>
  mutate(gq = as_factor(gq)) |>
  filter(!(gq %in% c("Group quarters--Institutions", "Other group quarters")))
```

FIPS codes should always contain leading zeros so that state codes are two
digits, county codes are three digits, and place codes are five digits.

```{r}
#| label: clean-fips-codes

acs1yr <-
  acs1yr |>
  mutate(
    across(c(state, county, city, puma), as.character),
    state = str_pad(state, width = 2, pad = "0", side = "left"),
    county = str_pad(county, width = 3, pad = "0", side = "left"),
    city = str_pad(city, width = 5, pad = "0", side = "left"),
    puma = str_pad(puma, width = 5, pad = "0", side = "left")
  )
```

We create two age subgroups: households over 45 and households under 45

```{r}
#| label: add-age-subgroup

acs1yr <-
  acs1yr |>
  mutate(subgroup = if_else(age >= 45, "Age 45 and Over", "Under Age 45"))
```

We collapse IPUMS race and ethnicity categories into white non-Hispanic, Black
non-Hispanic, other non-Hispanic, and Hispanic.

```{r}
#| label: add-race-ethnicity-subgroup


acs1yr <-
  acs1yr |>
  mutate(
    # Collapse detailed race categories: AIAN, Asian subgroups, multiracial → "Other"
    race = lbl_relabel(race, lbl(10, "Other") ~ .lbl %in% c("American Indian or Alaska Native", "Chinese", "Japanese", "Other Asian or Pacific Islander", "Other race, nec", "Two major races", "Three or more major races")),
    across(c(race, hispan), as_factor)
  ) |>
  mutate(
    race_ethnicity = case_when(
      race == "White" & hispan == "Not Hispanic" ~ "white_nh",
      race == "Black/African American" & hispan == "Not Hispanic" ~ "black_nh",
      race == "Other" & hispan == "Not Hispanic" ~ "other_nh",
      hispan != "Not Hispanic" ~ "hispanic",
      .default = NA_character_
    ),
    race_ethnicity = factor(race_ethnicity,
      levels = c(
        "white_nh",
        "black_nh",
        "hispanic",
        "other_nh"
      )
    )
  )
```

## Merge Crosswalk {#sec-merge-crosswalk}

We merge our 1-year ACS data with PUMA-county and PUMA-place geocodes to get
county and place boundaries.

**Crosswalk Period Assignment**

For 2022 ACS and onward we use a crosswalk going from 2020 PUMA to 2020 county.
For pre-2022 ACS data, we use a crosswalk from 2012 PUMA to 2016 county.

The weight of each household is adjusted by the allocation factor (`afact`), representing the proportion of the PUMA's population that falls into a given county/place.

### Special Geographic Considerations: Connecticut

**Connecticut Planning Regions (2022+)**

Starting in 2022, Connecticut transitioned from 8 counties to 9 planning regions:
- **Pre-2022:** Traditional county boundaries (8 counties, FIPS codes starting with "0")
- **2022+:** Nine planning regions replace counties in data (includes county "000" plus 8 regions)
- **Impact:** County counts increase from 3,143 to 3,144 starting in 2022

**Handling in code:**
- Connecticut counties should be present in the data before 2022 and planning regions *only* after 2022
- Old county codes filtered out for 2022+ CT data
- Multiple assertions throughout code verify correct CT geography handling 

```{r}
#| label: define-crosswalk-years

acs1yr <-
  acs1yr |>
  mutate(crosswalk_period = if_else(year < 2022, "pre-2022", "2022"))

# Check that only new planning regions are present after 2022 for CT
assertthat::assert_that(
  setequal(
    acs1yr |>
      select(year, state, county) |>
      filter(state == "09", year %in% c(2022, 2023, 2024)) |>
      distinct(county) |>
      pull(),
    c("170", "000", "190", "140", "180", "130", "120", "160")
  )
)

county_file <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv"))

county_file |>
  filter(year == 2021, state == "09")


```

::: panel-tabset
#### County
To assess the data quality of counties/places with
larger N, we need to consider what percent of data for the county/place actually
came from the county/place itself. We have this information in the `sum_products` column obtained from the crosswalk file. For more information on `sum_products`
see the script that generates the [PUMA to County crosswalk](geographic-crosswalks/generate_puma_county_crosswalks.qmd)
```{r}
#| label: puma-to-county-crosswalk

puma_county_crosswalk <-
  read_csv(here::here(
    "geographic-crosswalks",
    "data",
    "crosswalk_puma_to_county.csv"
  )) |>
  rename(
    state = statefip,
  ) |>
  # Filter out Puerto Rico since this will be NA for the ACS data
  filter(state != "72")


households_county <-
  acs1yr |>
  right_join(puma_county_crosswalk,
    by = c("state", "puma", "crosswalk_period")
  ) |>

  mutate(
    # if county isn't present in the ACS use county codes from PUMA-County crosswalk
    county = if_else(county.x != "000", county.x, county.y),
    # the weight of each household is adjusted by the area of the PUMA that falls into a given County
    hhwt = if_else(county.x != "000", hhwt, hhwt * afact), 
    sum_products = if_else(county.x != "000", 1, sum_products)
  ) |>
  select(-c(county.x, county.y)) |>
  # Remove old counties from 2022 Connecticut data, but keep county "000"
  filter(!(year == 2022 & state == "09" & str_starts(county, "0") & county != "000"))
  puma_county_crosswalk |>
  filter(state == "09", crosswalk_period == "2022") |>
  distinct(county)
```

#### Place

```{r}
#| label: puma-to-place-crosswalk

puma_place_crosswalk <-
  read_csv(here::here(
    "geographic-crosswalks",
    "data",
    "crosswalk_puma_to_place.csv"
  )) |>
  rename(state = statefip)

households_place <-
  acs1yr |>
  right_join(puma_place_crosswalk,
    by = c("state", "puma", "crosswalk_period")
  ) |>
  # there aren't any places missing from the ACS so we just adjust household
  # weight to account for PUMA-to-place mapping (those where the PUMA is not entirely inside the place)
  mutate(hhwt = hhwt * afact)
```

:::

## Create Housing Value Metric {#sec-create-metric}

### Housing Share

We will now compute the share of households by race and ethnicity (i.e., total
number of households for each race and ethnic group divided by the total number
of households). In other words, for each race and ethnicity, we compute the 
share of total households in a county/place by dividing the total number of households
of that race and ethnicity by the total number of households in the county/place.

Further, for each race-ethnicity and subgroup (e.g., age), we compute the share
of total households in that subgroup by dividing the total number of households 
of that race-ethnicity and subgroup by the total number of households in that 
subgroup in the county/place.

::: panel-tabset
### County

```{r}
#| label: share-of-households-by-race-ethnicity-county

share_households_county <-
  households_county |>
  select(year, state, county, hhwt, race_ethnicity) |>
  summarise(
    household = sum(hhwt),
    .by = c(year, state, county, race_ethnicity)
  ) |>
  mutate(
    total = sum(household),
    .by = c(year, state, county)
  ) |>
  mutate(share_households = household / total) |>
  select(year, state, county, race_ethnicity, share_households)

share_households_county_subgroup <-
  households_county |>
  select(year, state, county, hhwt, race_ethnicity, subgroup) |>
  summarise(
    household = sum(hhwt),
    .by = c(year, state, county, race_ethnicity, subgroup)
  ) |>
  mutate(
    total = sum(household),
    .by = c(year, state, county, subgroup)
  ) |>
  mutate(share_households = household / total) |>
  select(year, state, county, race_ethnicity, subgroup, share_households)
```

**Check:** Is `share_households` between 0 and 1?

```{r}
#| label: check-share-hh-county-check
stopifnot(min(share_households_county$share_households) >= 0)
stopifnot(max(share_households_county$share_households) <= 1)
```

We visualize the distribution of the share of households by race and ethnicity

```{r}
#| label: check-share-households-county

share_households_county |>
  drop_na() |>
  ggplot(aes(x = share_households)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 0.02) +
  facet_wrap(. ~ race_ethnicity) +
  labs(
    title = "Share of Households by Race and Ethnicity",
    subtitle = "County",
    y = "Count",
    x = "Share of Households"
  )

summary(share_households_county)
skimr::skim(share_households_county)
```


```{r}
#| label: check-share-hh-county-subgroup-check
stopifnot(min(share_households_county_subgroup$share_households) >= 0)
stopifnot(max(share_households_county_subgroup$share_households) <= 1)
```

```{r}
#| label: check-share-households-county-age

share_households_county_subgroup |>
  ggplot(aes(x = share_households)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 0.02) +
  facet_wrap(race_ethnicity ~ subgroup, ncol = 2) +
  labs(
    title = "Share of Households by Race and Ethnicity",
    subtitle = "County",
    y = "Count",
    x = "Share of Households"
  )

summary(share_households_county_subgroup)
skimr::skim(share_households_county_subgroup)
```

### Place

```{r}
#| label:  share-of-households-by-race-ethnicity-place

share_households_place <-
  households_place |>
  select(year, state, place, hhwt, race_ethnicity) |>
  summarise(
    households = sum(hhwt),
    .by = c(year, state, place, race_ethnicity)
  ) |>
  mutate(
    total_households = sum(households),
    .by = c(year, state, place)
  ) |>
  mutate(share_households = households / total_households) |>
  select(year, state, place, race_ethnicity, share_households)

share_households_place_subgroup <-
  households_place |>
  select(year, state, place, hhwt, race_ethnicity, subgroup) |>
  summarise(
    households = sum(hhwt),
    .by = c(year, state, place, race_ethnicity, subgroup)
  ) |>
  mutate(
    total_households = sum(households),
    .by = c(year, state, place, subgroup)
  ) |>
  mutate(share_households = households / total_households) |>
  select(year, state, place, race_ethnicity, subgroup, share_households)
```

**Check:** Is `share_households` between 0 and 1?

```{r}
#| label: check-share-households-race-ethnicity-place

stopifnot(min(share_households_place$share_households) >= 0)
stopifnot(max(share_households_place$share_households) <= 1)

share_households_place |>
  ggplot(aes(x = share_households)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 0.02) +
  facet_wrap(. ~ race_ethnicity) +
  labs(
    title = "Share of Households by Race and Ethnicity",
    subtitle = "Place",
    y = "Count",
    x = "Share of Households"
  )

skimr::skim(share_households_place)
summary(share_households_place)
```

```{r}
#| label: check-share-households-place-age

stopifnot(min(share_households_place_subgroup$share_households) >= 0)
stopifnot(max(share_households_place_subgroup$share_households) <= 1)

share_households_place_subgroup |>
  ggplot(aes(x = share_households)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 0.02) +
  facet_wrap(subgroup ~ race_ethnicity, ncol = 2) +
  labs(
    title = "Share of Households by Race and Ethnicity",
    subtitle = "Place",
    y = "Count",
    x = "Share of Households"
  )

skimr::skim(share_households_place_subgroup)
summary(share_households_place_subgroup)
```

:::

### Housing Value Share

We keep observations for homeowners only in order to calculate the housing value share. I.e., the total aggregated
housing wealth for homeowners for each race and ethnic group divided by the
total aggregated housing value.

```{r}
#| label: total-house-values-by-race

house_value_by_race <- function(data, ...) {
  data |>
    mutate(ownership = as_factor(ownership)) |>
    filter(ownership == "Owned or being bought (loan)") |>
    summarise(
      house_value = sum(valueh * hhwt),
      home_owners = n(),
      .by = c(year, state, race_ethnicity, ...)
    )
}

house_value_county_subgroup <- house_value_by_race(households_county, county, subgroup)
house_value_county <- house_value_by_race(households_county, county)


house_value_place_subgroup <- house_value_by_race(households_place, place, subgroup)
house_value_place <- house_value_by_race(households_place, place)
```

### Data Quality Flags

The metrics for the ACS indices are based on:

  - (1) sample size of each race and ethnic household at the county/place and
  - (2) the number of observations that comes from the specific county/place
    calculated from the PUMA-county/place crosswalk.

Therefore, the data quality is poor for counties/places with a extremely small
number of households (N). To assess the data quality of counties/places with
larger N, we need to consider what percent of data for the county/place actually
came from the county/place itself. We have this information in the `sum_products` column obtained from the crosswalk file. To read more about
how it was constructed, see the "Create quality variable" section of the ["Generate Puma to County Crosswalk"](geographic-crosswalks/generate_puma_county_crosswalks.qmd) notebook.

In particular,

  - Data quality is good (`1`): N is greater than 30 and more than 75% of
    observations are from the county/place.
  - Data quality is marginal (`2`): N is greater than 30 but more than 35% but less than 75% or equal of
    observations are from the county/place.
  - Data quality is poor (`3`): If less
    than or equal 35% of observations are from the county/place
  - Data quality is NA if N is less than or equal to 30.


```{r}
#| label: quality-check-flag-functions

construct_quality_flag <- function(x, y) {
  case_when(
    x > 30 & y > 0.75 ~ 1,
    x > 30 & (y > 0.35 & y <= 0.75) ~ 2,
    x > 30 & y <= 0.35 ~ 3,
    x <= 30 ~ NA_integer_,
    .default = NA_integer_
  )
}
```

Now, we implement the data quality flags explained above 

::: {.panel-tabset}

### County


```{r}
#| label: add-quality-check-flags-county

county_quality <-
  households_county |>
  distinct(year, state, county, .keep_all = TRUE) |>
  select(year, state, county, population, sum_products, geographic_allocation_quality)


house_value_county <-
  house_value_county |>
  left_join(county_quality, by = c("year", "state", "county")) |>
  mutate(
    ratio_house_value_households_quality = construct_quality_flag(population, sum_products)
  )

house_value_county |>
  ggplot(aes(x = ratio_house_value_households_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  labs(
    title = "Quality Flag",
    subtitle = "County",
    x = "Quality Flag",
    y = "Count"
  )

janitor::tabyl(house_value_county, ratio_house_value_households_quality)
```


```{r}
#| label: add-quality-check-flags-county-subgroup

house_value_county_subgroup <-
  house_value_county_subgroup |>
  left_join(county_quality, by = c("year", "state", "county")) |>
  mutate(
    ratio_house_value_households_quality = construct_quality_flag(population, sum_products)
  )

house_value_county_subgroup |>
  ggplot(aes(x = ratio_house_value_households_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  labs(
    title = "Quality Flag",
    subtitle = "County",
    x = "Quality Flag",
    y = "Count"
  ) +
  facet_wrap(. ~ subgroup, ncol = 2)
```

### Place

```{r}
#| label: add-quality-check-flags-place

place_quality <-
  households_place %>%
  distinct(state, place, .keep_all = TRUE) |>
  select(state, place, sum_products, population)

house_value_place <-
  house_value_place |>
  left_join(place_quality,
    by = c("state", "place")
  ) |>
  mutate(
    ratio_house_value_households_quality = construct_quality_flag(population, sum_products)
  )


house_value_place |>
  ggplot(aes(x = ratio_house_value_households_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  labs(
    title = "Quality Flag",
    subtitle = "Place",
    x = "Quality Flag",
    y = "Count"
  )
```


```{r}
#| label: add-quality-check-flags-place-subgroup

house_value_place_subgroup <-
  house_value_place_subgroup |>
  left_join(place_quality,
    by = c("state", "place")
  ) |>
  mutate(
    ratio_house_value_households_quality = construct_quality_flag(population, sum_products)
  )


house_value_place_subgroup |>
  ggplot(aes(x = ratio_house_value_households_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  theme_minimal() +
  labs(
    title = "Quality Flag",
    subtitle = "Place, Age Over 45 Subgroup",
    x = "Quality Flag",
    y = "Count"
  ) +
  facet_wrap(race_ethnicity ~ subgroup, ncol = 2)
```
:::

Now, we calculate the housing value share by computing house value divided by
the aggregate house value for a county/place.

```{r}
#| label: house-value-by-race-ethnicity

house_value_by_race <- function(data, ...) {
  data |>
    mutate(total_house_value = sum(house_value), .by = c(year, state, ...)) |>
    mutate(share_house_value = house_value / total_house_value) |>
    select(year, state, ..., race_ethnicity, house_value, total_house_value, share_house_value, ratio_house_value_households_quality)
}



house_value_county <- house_value_by_race(house_value_county, county)
house_value_county_subgroup <- house_value_by_race(house_value_county_subgroup, county, subgroup)
house_value_place <- house_value_by_race(house_value_place, place)
house_value_place_subgroup <- house_value_by_race(house_value_place_subgroup, place, subgroup)
```


```{r}
house_value_county |>
  ggplot() +
  geom_density(aes(x = share_house_value),
    color = "blue", fill = alpha("blue", 0.3)
  ) +
  facet_wrap(. ~ race_ethnicity) +
  theme_minimal() +
  labs(
    title = "Share House Value by Race/Ethnicity",
    subtitle = "County",
    x = "Share House Value",
    y = "Density"
  )

house_value_county_subgroup |>
  ggplot() +
  geom_density(aes(x = share_house_value),
    color = "blue", fill = alpha("blue", 0.3)
  ) +
  facet_wrap(race_ethnicity ~ subgroup, ncol = 2) +
  theme_minimal() +
  labs(
    title = "Share House Value by Race/Ethnicity and Age",
    subtitle = "County, Age over 45 Subgroup",
    x = "Share House Value",
    y = "Density"
  )


house_value_place |>
  ggplot() +
  geom_density(aes(x = share_house_value),
    color = "blue", fill = alpha("blue", 0.3)
  ) +
  facet_wrap(. ~ race_ethnicity) +
  theme_minimal() +
  labs(
    title = "Share House Value by Race/Ethnicity",
    subtitle = "Place",
    x = "Share House Value",
    y = "Density"
  )


house_value_place_subgroup |>
  ggplot() +
  geom_density(aes(x = share_house_value),
    color = "blue", fill = alpha("blue", 0.3)
  ) +
  facet_wrap(race_ethnicity ~ subgroup, ncol = 2) +
  theme_minimal() +
  labs(
    title = "Share House Value by Race/Ethnicity and Age",
    subtitle = "Place, Age over 45 Subgroup",
    x = "Share House Value",
    y = "Density"
  )
```

## Create Final Output Data: Merge Household & Housing Value Share

```{r}
#| label: compute-ratio-house-value-households

construct_ratio <- function(share_house_value, share_households) {
  paste0(
    format(round(share_house_value, 3), nsmall = 1),
    ":",
    format(round(share_households, 3), nsmall = 1)
  ) |>
    str_remove_all(pattern = " ")
}

ratio_house_value_households <- function(households_data, house_value_data, ...) {
  geo_var <- if_else("county" %in% colnames(households_data), "county", "place")
  households_data |>
    left_join(house_value_data,
      by = c("year", "state", geo_var, "race_ethnicity", ...)
    ) |>
    mutate(
      ratio_house_value_households = construct_ratio(share_house_value, share_households),
      .by = c(race_ethnicity, ...)
    ) |>
    arrange(year, state, {{ geo_var }}, ...) |> 
    mutate(ratio_house_value_households = if_else(is.na(ratio_house_value_households_quality), NA_character_, ratio_house_value_households))
}


reshape_to_wide <- function(data, ...) {
  data |>
    pivot_wider(
      id_cols = c(year, state, ...),
      names_from = race_ethnicity,
      values_from = c(ratio_house_value_households, ratio_house_value_households_quality),
      names_glue = "{race_ethnicity}_{.value}"
    ) %>%
    rename_with(~ str_remove(.x, "ratio_") %>% paste0("ratio_", .), contains("ratio")) %>%
    # order variables according to request
    select(
      year, state, ...,
      ratio_black_nh_house_value_households, ratio_black_nh_house_value_households_quality,
      ratio_hispanic_house_value_households, ratio_hispanic_house_value_households_quality,
      ratio_other_nh_house_value_households, ratio_other_nh_house_value_households_quality,
      ratio_white_nh_house_value_households, ratio_white_nh_house_value_households_quality
    ) %>%
    arrange(year, state, ...)
}
```

Before computing the ratio of house value to households, we need to merge the household and house value data. We do not expect perfect matches between the two datasets because the house value data is based on households that own their homes, while the household data includes all households.

```{r}
#| label: calculate-metric

ratio_house_value_households_county <- ratio_house_value_households(
  share_households_county,
  house_value_county
) |>
  reshape_to_wide(county)


ratio_house_value_households_county_subgroup <- ratio_house_value_households(
  share_households_county_subgroup,
  house_value_county_subgroup,
  "subgroup"
) |>
  reshape_to_wide(county, subgroup)

ratio_house_value_households_place <-
  ratio_house_value_households(
    share_households_place,
    house_value_place
  ) |>
  reshape_to_wide(place)

ratio_house_value_households_place_subgroup <-
  ratio_house_value_households(
    share_households_place_subgroup,
    house_value_place_subgroup,
    "subgroup"
  ) |>
  reshape_to_wide(place, subgroup)
```

**Check:** Are there 3,143 counties and 486 places in the final output?

Our metric county data has 3,143 counties as expected for all years prior to
2022. The metric places data also has the expected 486 places.

```{r}
#| label: check-number-of-counties-places

stopifnot(
  ratio_house_value_households_county |>
    distinct(year, state, county) |>
    count(year) |>
    filter(!(year %in% c(2022, 2023, 2024))) |>
    pull(n) == 3143
)


stopifnot(
  ratio_house_value_households_place |>
    distinct(year, state, place) |>
    count(year) |>
    filter(year %in% c(2022, 2023, 2024)) |>
    pull(n) == 486
)
```


There are 3 counties in the county populations file that are
not present in our metric data. These are the planning regions of Connecticut. We will merge these in by performing as `right_join` using the county population file from the geographic crosswalks directory before exporting. Metric and quality values for these counties will be left as NA.


```{r}
#| label: find-missing-counties


all_counties <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv")) %>%
  tidylog::filter(year %in% c(2022, 2023, 2024)) %>%
  select(-year)

(
  missing_counties <- ratio_house_value_households_county %>%
    filter(year %in% c(2022, 2023, 2024)) |>
    distinct(state, county) |>
    anti_join(x = all_counties, y = _, by = c("state", "county"))
)

all_counties |>
  filter(state == "09")
```

At this stage, the number of counties in 2022 is 3,141.

```{r}
#| label: check-number-of-counties-2022
#| error: true

stopifnot(
  ratio_house_value_households_county |>
    distinct(year, state, county) |>
    count(year) |>
    filter(year >= 2022) |>
    pull(n) |> unique() == 3144
)
```

Check number of missings and variable summaries before exporting:

```{r}
skimr::skim(ratio_house_value_households_county)
skimr::skim(ratio_house_value_households_county_subgroup)
skimr::skim(ratio_house_value_households_place)
skimr::skim(ratio_house_value_households_place_subgroup)
```


## Data Dictionary

### Key Variables in Final Output

| Variable | Type | Description | Range/Values |
|----------|------|-------------|-------------|
| `year` | Integer | ACS survey year | 2014-2024 (excl 2020) |
| `state` | Character | State FIPS code (2-digit) | 01-56 |
| `county` / `place` | Character | County (3-digit) or Place (5-digit) FIPS | Varies |
| `ratio_[group]_house_value_households` | Character | Ratio formatted as "X.XXX:X.XXX" | e.g., "0.5:1.0" |
| `ratio_[group]_house_value_households_quality` | Integer | Data quality flag (1-3) | 1=best, 3=poor, NA=suppressed |
| `subgroup_type` | Character | Type of subgroup (only in subgroup files) | "age" |
| `subgroup` | Character | Subgroup category | "All", "Age 45 and Over", "Under Age 45" |

### Racial/Ethnic Group Codes

- `black_nh` = Black/African American, non-Hispanic
- `hispanic` = Hispanic (any race)
- `other_nh` = Other races, non-Hispanic (includes Asian, AIAN, Pacific Islander, multiracial)
- `white_nh` = White, non-Hispanic

### Intermediate Variables

| Variable | Description | Source |
|----------|-------------|--------|
| `afact` | Allocation factor for PUMA overlap | Crosswalk files |
| `sum_products` | Proportion of data from target geography | Crosswalk calculation |
| `population` | Population count in geography | Crosswalk files |
| `valueh` | Owner-occupied home value | IPUMS ACS (VALUEH) |
| `hhwt` | Household weight (adjusted by afact) | IPUMS ACS (HHWT, modified) |
| `ownership` | Homeownership status | IPUMS ACS (OWNERSHP) |
| `share_households` | Group's proportion of all households | Calculated |
| `share_house_value` | Group's proportion of total housing value | Calculated |

## Export {#sec-export}


```{r}
#| label: export-data

export_metric <- function(data, file_type) {
  
  geography <- ifelse("place" %in% colnames(data), "city", "county")

  geo_var <- ifelse(geography == "city", "place", "county")

  pop_file <-
    read_csv(here::here("geographic-crosswalks", "data", paste0(geo_var, "-populations.csv"))) |>
    filter(year != 2020)
  data <- data |> right_join(pop_file, by = c("year", "state", geo_var))



  if (file_type == "subgroup") {
    if (geo_var == "county") {

      data <-
        data |>
        mutate(subgroup_type = "age", .before = "subgroup") |>
        bind_rows(ratio_house_value_households_county |>
          mutate(
            subgroup_type = "age",
            subgroup = "All"
          )) |>
        mutate(subgroup = factor(subgroup,
          levels = c(
            "All",
            "Age 45 and Over",
            "Under Age 45"
          )
        )) |>
        arrange(year, state, county, subgroup_type, subgroup) |>
        relocate("subgroup_type", .before = subgroup) 
      
    } else if (geo_var == "place") {

      data <-
        data |>
        mutate(subgroup_type = "age", .before = "subgroup") |>
        bind_rows(ratio_house_value_households_place |>
          mutate(
            subgroup_type = "age",
            subgroup = "All"
          )) |>
        mutate(subgroup = factor(subgroup,
          levels = c(
            "All",
            "Age 45 and Over",
            "Under Age 45"
          )
        )) |>
        arrange(year, state, place, subgroup_type, subgroup) |>
        relocate("subgroup_type", .before = subgroup) |>
        group_by(year, state, place) |>
        fill(c(state_name, place_name), .direction = "downup") |>
        ungroup()
    }
  }

  output_path <- here(glue("01_financial-well-being/final/households_house_value_race_ethnicity_{file_type}_{geography}.csv"))

  data |>
    select(-c(population, state_name, rlang::sym(paste0(geo_var, "_name")))) |>
    write_csv(file = output_path)
}


export_metric(ratio_house_value_households_county, file_type = "all")
export_metric(ratio_house_value_households_county_subgroup, file_type = "subgroup")

export_metric(ratio_house_value_households_place, file_type = "all")
export_metric(ratio_house_value_households_place_subgroup, file_type = "subgroup")
```

Check that final files have the expected number of observations per year

```{r}
#| label: check-ct-counties-all
county_all <- read_csv(here::here("01_financial-well-being", "final", "households_house_value_race_ethnicity_all_county.csv"))

county_file <- read_csv(here::here("geographic-crosswalks", "data", "county-populations.csv"))
county_all |> count(year)
county_file |> count(year)

# Check that only new planning regions are present in 2022 for CT and all 9 are prese t
assertthat::assert_that(
  all.equal(
    # CT counties in final county metric file
    county_all |>
      select(year, state, county) |>
      filter(state == "09", year >= 2022) |>
      distinct(county) |>
      arrange(county),
    # CT county in population file
    county_file |>
      select(year, state, county) |>
      filter(state == "09", year >= 2022) |>
      distinct(county) |>
      arrange(county)
  )
)

assertthat::assert_that(
  county_all |>
    select(year, state, county) |>
    filter(state == "09", year >= 2022) |>
    distinct(county) |>
    nrow() == 9
)
```




```{r}
#| label: check-ct-counties-subgroup
county_subgroup <- read_csv(here::here("01_financial-well-being", "final", "households_house_value_race_ethnicity_subgroup_county.csv"))


county_all |> count(year)
county_file |> count(year)

# Check that only new planning regions are present in 2022 for CT and all 9 are prese t
assertthat::assert_that(
  all.equal(
    # CT counties in final county metric file
    county_subgroup |>
      select(year, state, county) |>
      filter(state == "09", year >= 2022) |>
      distinct(county) |>
      arrange(county),
    # CT county in population file
    county_file |>
      select(year, state, county) |>
      filter(state == "09", year >= 2022) |>
      distinct(county) |>
      arrange(county)
  )
)

assertthat::assert_that(
  county_subgroup |>
    select(year, state, county) |>
    filter(state == "09", year >= 2022) |>
    distinct(county) |>
    nrow() == 9
)
```

```{r}
place_all <- read_csv(here::here("01_financial-well-being", "final", "households_house_value_race_ethnicity_all_city.csv"))

place_file <- read_csv(here::here("geographic-crosswalks", "data", "place-populations.csv")) |> filter(year != 2020)
place_all |> count(year)
place_file |> count(year)

# Check that place files have the same number of observations per year
assertthat::assert_that(
  all.equal(
place_all |> as.data.frame() |>  count(year),
place_file |> as.data.frame() |>  count(year)
  )
)

```

## Expected Outputs

### Files Created

All files saved to `01_financial-well-being/final/`:

1. **`households_house_value_race_ethnicity_all_county.csv`**
   - Rows: ~3,143-3,144 counties × 11 years = ~34,575 rows
   - All racial/ethnic groups, no age subgroups
   - 8 metric columns (4 groups × 2 columns: ratio + quality)

2. **`households_house_value_race_ethnicity_subgroup_county.csv`**
   - Rows: ~3,143 counties × 11 years × 3 age categories = ~103,725 rows
   - Includes "All", "Age 45 and Over", "Under Age 45"
   - Additional columns: `subgroup_type`, `subgroup`

3. **`households_house_value_race_ethnicity_all_city.csv`**
   - Rows: 486 places × 11 years = ~5,346 rows
   - Largest 486 U.S. places

4. **`households_house_value_race_ethnicity_subgroup_city.csv`**
   - Rows: 486 places × 11 years × 3 age categories = ~16,038 rows

### Column Structure

Each file contains ratio and quality columns for 4 racial/ethnic groups:
- Format: `ratio_[group]_house_value_households` and `ratio_[group]_house_value_households_quality`
- Order: black_nh, hispanic, other_nh, white_nh
- 8 metric columns total + geography identifiers

## Testing

```{r}
final_county_data <- read_csv(here::here("01_financial-well-being/final/households_house_value_race_ethnicity_all_county.csv"), show_col_types = FALSE)

final_place_data <- read_csv(here::here("01_financial-well-being/final/households_house_value_race_ethnicity_all_city.csv"), show_col_types = FALSE)
evaluate_final_data(
  exp_form_path = "99a_final-evaluation/evaluation_form_house_value_overall_county.csv",
  data = final_county_data, geography = "county",
  subgroups = FALSE, confidence_intervals = FALSE
)

evaluate_final_data(
  exp_form_path = "99a_final-evaluation/evaluation_form_house_value_overall_place.csv",
  data = final_place_data, geography = "place",
  subgroups = FALSE, confidence_intervals = FALSE
)

final_county_data_subgroups <- read_csv(here::here("01_financial-well-being/final/households_house_value_race_ethnicity_subgroup_county.csv"), show_col_types = FALSE)

final_place_data_subgroups <-
  read_csv(here::here("01_financial-well-being/final/households_house_value_race_ethnicity_subgroup_city.csv"), show_col_types = FALSE)

evaluate_final_data(
  exp_form_path = "99a_final-evaluation/evaluation_form_house_value_age_county.csv",
  data = final_county_data_subgroups, geography = "county",
  subgroups = TRUE, confidence_intervals = FALSE
)

evaluate_final_data(
  exp_form_path = "99a_final-evaluation/evaluation_form_house_value_age_county.csv",
  data = final_place_data_subgroups, geography = "place",
  subgroups = TRUE, confidence_intervals = FALSE
)
```

## Processing Workflow Summary

Complete workflow for reference:

1. **Extract ACS Data** → Pull 1-year microdata (2014-2024) from IPUMS API
2. **Clean Microdata** → Filter to households, handle missing values, format FIPS codes
3. **Recode Demographics** → Collapse race/ethnicity into 4 groups, create age subgroups
4. **Geographic Crosswalk** → Merge PUMA-to-county and PUMA-to-place mappings with allocation factors
5. **Adjust Weights** → Multiply household weights by `afact` to distribute across geographies
6. **Calculate Household Shares** → Sum weighted households by race/ethnicity and geography
7. **Calculate Housing Value Shares** → Sum housing values for homeowners only by group
8. **Apply Quality Flags** → Assess data quality based on population and geographic precision
9. **Compute Ratios** → Divide housing value share by household share for each group
10. **Format Output** → Reshape to wide format with separate columns per racial/ethnic group
11. **Add Missing Geographies** → Right join with population files to include all counties/places (with NAs)
12. **Validate and Export** → Run quality checks, export final CSVs to `final/` directory

## References and Documentation

### Primary Data Source
- **IPUMS USA:** Steven Ruggles, Sarah Flood, Matthew Sobek, et al. IPUMS USA: Version 13.0 [dataset]. Minneapolis, MN: IPUMS, 2023. https://doi.org/10.18128/D010.V13.0

### Methodology Sources
- **Original code:** Jung Hyun Choi (initial development)
- **Updates:** Manu Alcalá Kovalski (2023-2024)

### Related Documentation
- **PUMA Crosswalks:** `geographic-crosswalks/generate_puma_county_crosswalks.qmd` and `generate_puma_place_crosswalks.qmd`
- **Quality flag methodology:** See crosswalk scripts for `sum_products` calculation details
- **Project documentation:** `mobility-from-poverty-documentation/` folder
- **Evaluation forms:** `99a_final-evaluation/` folder

### External Resources
- **IPUMS USA documentation:** https://usa.ipums.org/usa/
- **ACS 2020 data quality warning:** https://www.census.gov/programs-surveys/acs/technical-documentation/table-and-geography-changes/2020/1-year.html
- **ipumsr package:** https://tech.popdata.org/ipumsr/
- **Value labels guide:** https://tech.popdata.org/ipumsr/articles/value-labels.html

### Support
For questions about this metric, contact the Upward Mobility Framework team at the Urban Institute.
