---
title: "Opportunity for Income Predictor, county-level"
author: "JP Walsh and Aaron R. Williams"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

*Program for the creation of the Opportunity for Income Predictor at the county level*

ACS Code: Initially written by Tina Chelidze and updated by JP Walsh 2022-2023.

Primary data derived from the IPUMS API.

Based on processes developed by Paul Johnson and Kevin Werner in SAS.

## Housekeeping

Read in packages for use. If you do not have an IPUMS API key you will need to make one via the [IPUMS website](https://www.ipums.org/).

```{r}
library(tidyverse)
library(Hmisc)
library(ipumsr)
library(reactable)
library(srvyr)
library(scales)
library(tictoc)

options(scipen = 999)

theme_set(theme_minimal())

# DO NOT PUSH YOUR API KEY. You only have to run this once and then comment it out as below.
#set_ipums_api_key("INSERT KEY HERE", save = TRUE, overwrite = TRUE)

source(here::here("functions", "API", "extract_ipums.R"))
source(here::here("functions", "API", "ipums_repwt_household.R"))
source(here::here("01_financial-well-being", "R", "calc_income_quantiles.R"))
source(here::here("01_financial-well-being", "R", "calc_income_quantiles_subgroup.R"))

```

## Load Data

### Household Data

Read in the ACS extracts using the `extract_ipums()` function. This function only loads the data if it isn't already saved on the computer with the appropriate name. The name is determined by `extract_name`. 

We currently include three years of 1-year data (2018a, 2021a and 2022a) and two years of 5-year data (2018c and 2021c). 5-year 2022 data will be added pending its release.

```{r}
acs <- extract_ipums(
  extract_name = "umf_data_18_22",
  extract_description = "Microdata pull for Mobility Metric Predictors. 
  American Community Survey, years 2018, 2019 and 2022 (1 and 5-year).",
  survey = list("us2018a", "us2021a", "us2022a")#, "us2018c", "us2021c")
) 

```

todo(aaron): split into two programs. One with 5-year data and one with 1-year data

Look at the distribution of samples in the data. The 5-year data sets should comprise a larger share of the total sample.

```{r}
count(acs, sample)

```

Isolate data to include each household only once.

```{r}
acs <- acs %>%
  filter(pernum == 1)

```

### Household Replicate Weights

Read in the ACS household replicate weights using the `ipums_repwt_household()` function. This function only loads the data if it isn't already saved on the computer with the appropriate name. The name is determined by `extract_name`. The replicate weights are needed to construct confidence intervals. 

```{r}
repwts_household <- ipums_repwt_household(
  extract_name = "household_replicate_weights",
  extract_description = "Household level replicate weights for income quantiles 
  metric. American Community Survey, years 2018, 2019 and 2022 (1 and 5-year).",
  survey = list("us2018a", "us2021a", "us2022a")#, "us2018c", "us2021c")
)  %>% 
  select(-cbserial, -cbpernum)

```

todo(aaron): split into two programs. One with 5-year data and one with 1-year data

Look at the distribution of ACS survey samples in the data. The number of different samples should align with the number of surveys selected in the extract ipums function.

```{r}
count(repwts_household, sample)

```

Isolate data to include each household only once.

```{r}
repwts_household <- repwts_household %>%
  filter(pernum == 1)

```

Remove the sample variable so it isn't duplicated in the join. 
```{r}
repwts_household <- repwts_household %>% 
  select(-sample)

```

Join the replicate weights to the ACS household-level data. This join should be one-to-one based on the unique person identifier. 

```{r}
length(unique(pull(acs, unique_person_id)))

length(unique(pull(repwts_household, unique_person_id)))

acs_combined <-  
  left_join(
    acs,
    repwts_household, 
    by = "unique_person_id"
  )

length(unique(pull(acs_combined, unique_person_id)))

rm(acs)
rm(repwts_household)

```

## Clean Data

Missing data is reported in a variety of ways by IPUMS - listed below:

* `hhincome` is missing it and the missing value is reported as `9999999`. All missing incomes are for individuals in group quarters.
* `vacancy` is reported as "0". 

There is a large share of records missing ``hhincome`. However, note there are no vacancy results by default of the structure of census data read in.

```{r}
acs_combined %>% 
  ggplot(aes(x = hhincome)) +
  geom_density(color = "blue", fill = alpha("blue", 0.3)) +
  labs(
    title = "Household Income Variable Results",
    y = "Density"
  )

acs_combined %>% 
  ggplot(aes(x = vacancy)) +
  geom_density(color = "blue", fill = alpha("blue", 0.3)) +
  labs(
    title = "Vacancy Variable Results",
    y = "Density"
  )

```

Turn the missing/"not in universe" values for the `hhincome` variable into `NA`s.

```{r}
acs_clean <- acs_combined %>%
  mutate(
    hhincome = if_else(
      hhincome == 9999999, 
      NA_integer_,
      hhincome
    )
  )

```

Look at distribution of `hhincome` after adjustment.

```{r}

acs_clean %>% 
  ggplot(aes(x = hhincome)) +
  geom_density(color = "blue", fill = alpha("blue", 0.3)) +
  labs(
    title = "Household Income Variable Results",
    y = "Density"
  )

```

Select our group quarters, we want to keep only households including additional households under updated definitions.

```{r}
count(acs_clean, gq)

```

```{r}
acs_clean <- acs_clean %>% 
  filter(gq %in% c("Households under 1970 definition", 
                   "Additional households under 1990 definition",
                   "Additional households under 2000 definition"))

```

Check that group quarters fall only into household categories.

```{r}
acs_clean %>% 
  count(gq) %>% 
  ggplot(mapping = aes(x = gq, y = n)) +
  geom_col() +
  geom_text(mapping = aes(label = n), vjust = -1) +   
  scale_x_discrete(labels = label_wrap(10)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "GQ, All Records",
    y = NULL
  )
  
```

### Create race variable

Evaluate the values in the RACE and HISPAN variables.

IPUMS documented values for RACE: 

* 1 White
* 2 Black/African American/Negro
* 3 American Indian or Alaska Native 
* 4 Chinese 
* 5 Japanese
* 6 Other Asian or Pacific Islander 
* 7 Other race 
* 8 Two major races 
* 9 Three or more major races

IPUMS documented values for HISPAN: 

* 0 Not Hispanic 
* 1 Mexican 
* 2 Puerto Rican 
* 3 Cuban 
* 4 Other 
* 9 Not Reported 

Check that group quarters fall only into household categories.

```{r}
acs_clean %>% 
  count(race) %>% 
  ggplot(mapping = aes(x = n, y = race)) +
  geom_col() +
  geom_text(mapping = aes(label = n), hjust = -0.25) +  
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_y_discrete(labels = label_wrap(15)) +
  labs(
    title = "Race, All Records"
  ) 

acs_clean %>% 
  count(hispan) %>% 
  ggplot(mapping = aes(x = n, y = hispan)) +
  geom_col() +
  geom_text(mapping = aes(label = n), vjust = -0.1) +  
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_y_discrete(labels = label_wrap(15)) +
  labs(
    title = "Hispan, All Records",
  )

```

Create the race/ethnicity variable. For other race/ethnicity categories not coded as Hispanic, Hispan must be equal to "Not Hispanic". 

Rename the values from number category to race label: 

* Hispanic heads of households are Hispanic
* White, Non-Hispanic heads of household are White, Non-Hispanic
* Black/African American, Non-Hispanic heads of household are Black, Non-Hispanic
* All other heads of households are Other Races and Ethnicities
  
```{r}
acs_clean <- acs_clean %>%
  mutate(
    subgroup_race_ethnicity = case_when(
      hispan != "Not Hispanic" ~ "Hispanic",
      race == "White" ~ "White, Non-Hispanic",
      race == "Black/African American" ~ "Black, Non-Hispanic",
      !race %in% c("Black/African American", "White") ~ "Other Races and Ethnicities"
    )
  )

```

Look at subgroup distribution.

```{r}
acs_clean %>% 
  count(subgroup_race_ethnicity) %>% 
  ggplot(mapping = aes(x = subgroup_race_ethnicity, y = n)) +
  geom_col() +
  geom_text(mapping = aes(label = n), vjust = -1) + 
  scale_x_discrete(labels = label_wrap(15)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Subgroup, All Records",
    x = "Subgroup",
    y = NULL
  )

```

Create a binary variable to distinguish between 1 and 5 year ACS datasets. 5-year samples include the term "5-year".

```{r}
acs_clean <- acs_clean %>%
  mutate(sample_structure = if_else(str_detect(sample, "5-year"), "5year", "1year")) 

```

Confirm there are no missing values remaining in the data set for the household income variable. 

```{r}
acs_clean %>%
  filter(is.na(hhincome)) %>% 
  count()

```

Confirm there are no vacant properties included in the data.

```{r}
acs_clean %>%
  filter(is.na(vacancy)) %>% 
  count()

```

```{r}
rm(acs_combined)

```

## Crosswalk

Read in the PUMA to county crosswalk file. This file is created by the program generate_puma_county_crosswalks.rmd in the geographic-crosswalks folder. 

```{r}
puma_county_crosswalk <- read_csv(
  here::here("geographic-crosswalks", "data", "crosswalk_puma_to_county.csv")
)

```

Crate a version of the crosswalk with just the county-level values of data quality variables for each crosswalk period.

```{r}
county_puma <- puma_county_crosswalk %>%
  group_by(crosswalk_period, statefip, county) %>% 
  summarise(geographic_allocation_quality = mean(geographic_allocation_quality)) %>% 
  ungroup()

```

## Merge Crosswalk 

Create a variable in the ACS data indicating whether the survey is pre or post-2022. This will be used to join on the correct crosswalk information.

```{r}
acs_clean <- acs_clean %>% 
  mutate(crosswalk_period = if_else(year < 2022, "pre-2022", "2022"))

```

Check that the assignment worked. All years prior to 2022 should be designated "pre-2022".

```{r}
count(acs_clean, crosswalk_period, year)

```

Join the cleaned ACS data onto the crosswalk. This is a many-to-many join because we are aggregating from PUMA to county.
```{r}
acs_crosswalked <- left_join(
  acs_clean, 
  puma_county_crosswalk, 
  by = c("crosswalk_period", "statefip","puma"), 
  relationship = "many-to-many"
)

rm(acs_clean)

```

Drop any observations with `NA` for `afact` (i.e. there is no county of interest overlapping this PUMA).

```{r}
acs_crosswalked <- acs_crosswalked %>% 
  drop_na(afact)

```

Also filter out cases where `afact` is equal to zero.

```{r}
acs_crosswalked <- acs_crosswalked %>% 
  filter(afact > 0)

```

Adjust weight to account for PUMA-to-county mapping (those where unique_types do not equal 1).Drop PUMA flag variable from PUMA_county.

`adjust` converts the dollar amounts to the amount that they would have been had they been earned entirely during the calendar year. This is to deal with households being surveyed at different times during the year.

```{r}
select(acs_crosswalked, hhincome, hhwt, repwt1, repwt80)

acs_crosswalked <- acs_crosswalked %>%
  mutate(
    hhwt = hhwt * afact, 
    across(matches("repwt[0-9]+"), ~.x * afact),
    hhincome = hhincome * adjust
  )

select(acs_crosswalked, hhincome, hhwt, repwt1, repwt80)

```

## Calculate Metric: Overall

### Pre-drop suppressed cells

To save time when processing the metric, calculate the effective sample size in each county, defined as the sum of the `afact` variable. This is used to create a cutoff for data quality based on the actual number of survey records being used in the calculation. We will not include counties with less than 30 effective samples.

```{r}
#| label: all-effective-samples

acs_all <- acs_crosswalked %>%
  group_by(year, sample, crosswalk_period, statefip, county) %>%
  mutate(effective_sample = sum(afact)) %>% 
  ungroup()

# calculate the number of statistics after suppressing values
acs_all %>%
  group_by(year, sample, crosswalk_period, statefip, county) %>%
  summarize(
    effective_sample = max(effective_sample)
  ) %>%
  ungroup() %>%
  summarize(
    original_stats = n(),
    unsuppressed_stats = sum(effective_sample >= 30)
  )

```

Remove records from counties below the size quality cutoff. 

```{r}
#| label: subset-all

acs_all <- acs_all %>% 
  filter(effective_sample >= 30)

```

### Calculate metric

Calculate the county-level metrics. 

```{r}
#| label: calc-all

tic()
results_all <- acs_all %>%
  calc_income_quantiles()
toc()

```

We tabulated effective sample counts for each county, defined as the sum of the `afact` for all records in that county. Look at the distribution of counties by potential cutoffs (30, 100, 300). This is to measure data quality for each county based on the number of records reported. 

In the 1-year data for all there are no cases of a county with less than 100 observations for 2018, 2021 and 2022.

Generate the quality variable. Mark cases where the effective sample is less than 100, these will be assigned a data quality of 3.

```{r}
results_all <- left_join(
  results_all, 
  county_puma, 
  by = c("crosswalk_period", "statefip","county")
)

results_all <- results_all %>% 
  mutate(
    pctl_20_quality = case_when(
      effective_sample < 100 ~ 3,
      geographic_allocation_quality == 1 ~ 1,
      geographic_allocation_quality == 2 ~ 2,
      geographic_allocation_quality == 3 ~ 3
    ),
    pctl_50_quality = case_when(
      effective_sample < 100 ~ 3,
      geographic_allocation_quality == 1 ~ 1,
      geographic_allocation_quality == 2 ~ 2,
      geographic_allocation_quality == 3 ~ 3
    ),
    pctl_80_quality = case_when(
      effective_sample < 100 ~ 3,
      geographic_allocation_quality == 1 ~ 1,
      geographic_allocation_quality == 2 ~ 2,
      geographic_allocation_quality == 3 ~ 3
    )
  )
```

Keep only relevant variables.

```{r}
clean_names <- function(name) {
  
  name %>%
    str_replace_all(pattern = "_q2", "_2") %>%
    str_replace_all(pattern = "_q5", "_5") %>%
    str_replace_all(pattern = "_q8", "_8") %>%
    str_replace_all(pattern = "_low", "_lb") %>%
    str_replace_all(pattern = "_upp", "_ub")
    
}


results_all %>% 
  rename_with(clean_names)
  
  
  
  # select(year, statefip, county, pctl_income_20, pctl_income_20_lb = pctl_income_20_low, 
  #        pctl_income_20_ub = pctl_income_20_upp, pctl_20_quality, 
  #        pctl_income_50, pctl_income_50_lb = pctl_income_50_low, 
  #        pctl_income_50_ub = pctl_income_50_upp, pctl_50_quality, 
  #        pctl_income_80, pctl_income_80_lb = pctl_income_80_low, 
  #        pctl_income_80_ub = pctl_income_80_upp,
  #        pctl_80_quality)

```









## Calculate Metric: Race/Ethnicity

### Pre-drop suppressed cells

```{r}
#| label: race-ethnicity-effective-samples

acs_race_ethnicity <- acs_crosswalked %>%
  group_by(year, sample, crosswalk_period, statefip, county, subgroup_race_ethnicity) %>%
  mutate(effective_sample = sum(afact)) %>% 
  ungroup()

# calculate the number of statistics after suppressing values
acs_race_ethnicity %>%
  group_by(year, sample, crosswalk_period, statefip, county, subgroup_race_ethnicity) %>%
  summarize(
    effective_sample = max(effective_sample)
  ) %>%
  ungroup() %>%
  summarize(
    original_stats = n(),
    unsuppressed_stats = sum(effective_sample >= 30)
  )

```

```{r}
#| label: subset-race-ethnicity

acs_race_ethnicity <- acs_race_ethnicity %>% 
  filter(effective_sample >= 30)

```

### Calculate metric

Calculate the county-level metrics. 

```{r}
#| label: calc-race-ethnicity

tic()
results_race_ethnicity <- acs_race_ethnicity %>%
  calc_income_quantiles_subgroup(.subgroup = subgroup_race_ethnicity)
toc()

```

## Add Supressed Rows

## Data Quality Flags

## Validation

Summarize the predictors for the 1-year and 5-year data.
```{r}
metrics_income_all %>% 
  select(pctl_income_20:pctl_80_quality) %>% 
  summary()

metrics_income_sub %>% 
  select(pctl_income_20:pctl_80_quality) %>% 
  summary()
```


Look at distributions of income percentiles by county for all (2022 and 2021).
```{r}
metrics_income_all %>% 
  filter(year == 2022) %>% 
  select(pctl_income_20, pctl_income_50, pctl_income_80) %>% 
  pivot_longer(cols = c(pctl_income_20, pctl_income_50, pctl_income_80), names_to = "income", values_to = "percentile") %>% 
  ggplot(aes(x = percentile, color = income, fill = income)) +
  geom_density(alpha = 0.15) +
  ggtitle("Income percentiles (1-year data)") +
  ylab("Density")

metrics_income_sub %>% 
  filter(year == 2021, subgroup == "All") %>% 
  select(pctl_income_20, pctl_income_50, pctl_income_80) %>% 
  pivot_longer(cols = c(pctl_income_20, pctl_income_50, pctl_income_80), names_to = "income", values_to = "percentile") %>% 
  ggplot(aes(x = percentile, color = income, fill = income)) +
  geom_density(alpha = 0.15) +
  ggtitle("Income percentiles (5-year data)") +
  ylab("Density")
  
```

Break out distributions by subgroup (2021).
```{r}
metrics_income_sub %>% 
  filter(year == 2021, subgroup == "Black, Non-Hispanic") %>% 
  select(county, pctl_income_20, pctl_income_50, pctl_income_80) %>% 
  pivot_longer(cols = c(pctl_income_20, pctl_income_50, pctl_income_80), names_to = "income", values_to = "percentile") %>% 
  ggplot(aes(x = percentile, color = income, fill = income)) +
  geom_density(alpha = 0.15) +
  ggtitle("Income percentiles, Black") +
  ylab("Density")
  
metrics_income_sub %>% 
  filter(year == 2021, subgroup == "Hispanic") %>% 
  select(county, pctl_income_20, pctl_income_50, pctl_income_80) %>% 
  pivot_longer(cols = c(pctl_income_20, pctl_income_50, pctl_income_80), names_to = "income", values_to = "percentile") %>% 
  ggplot(aes(x = percentile, color = income, fill = income)) +
  geom_density(alpha = 0.15) +
  ggtitle("Income percentiles, Hispanic") +
  ylab("Density")
  
metrics_income_sub %>% 
  filter(year == 2021, subgroup == "Other Races and Ethnicities")  %>% 
  select(county, pctl_income_20, pctl_income_50, pctl_income_80) %>% 
  pivot_longer(cols = c(pctl_income_20, pctl_income_50, pctl_income_80), names_to = "income", values_to = "percentile") %>% 
  ggplot(aes(x = percentile, color = income, fill = income)) +
  geom_density(alpha = 0.15) +
  ggtitle("Income percentiles, Other") +
  ylab("Density")
  
metrics_income_sub %>% 
  filter(year == 2021, subgroup == "White, Non-Hispanic")  %>% 
  select(county, pctl_income_20, pctl_income_50, pctl_income_80) %>% 
  pivot_longer(cols = c(pctl_income_20, pctl_income_50, pctl_income_80), names_to = "income", values_to = "percentile") %>% 
  ggplot(aes(x = percentile, color = income, fill = income)) +
  geom_density(alpha = 0.15) +
  ggtitle("Income percentiles, white") +
  ylab("Density")
```

Look at the counts of the quality flag (all years).
```{r}
metrics_income_all %>% 
  select(pctl_20_quality) %>% 
  ggplot(aes(x = pctl_20_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  ggtitle("Quality Flag (1-year data") +
  ylab("Count")

metrics_income_sub %>% 
  select(pctl_20_quality) %>% 
  ggplot(aes(x = pctl_20_quality)) +
  geom_histogram(color = "blue", fill = alpha("blue", 0.3), binwidth = 1.0) +
  ggtitle("Quality Flag (5-year data") +
  ylab("Count")
```

Tabulate share of data suppressed by year.
```{r}
metrics_income_all %>% 
  group_by(year) %>% 
  filter(is.na(pctl_income_20)) %>% 
  count()

metrics_income_sub %>% 
  group_by(year) %>% 
  filter(is.na(pctl_income_20)) %>% 
  count()
```


## Export

Sort by year, state and county.

```{r}
metrics_income_sub <- metrics_income_sub %>%
  arrange(year, statefip, county)
```

Rename state.

```{r}
metrics_income_all <- metrics_income_all %>% 
  rename("state" = "statefip")

metrics_income_sub <- metrics_income_sub %>% 
  rename("state" = "statefip")
```

Order the variables how we want.

```{r}
metrics_income_all <- metrics_income_all %>% 
  select(year, state, county, pctl_income_20, pctl_income_20_lb, 
         pctl_income_20_ub, pctl_20_quality, 
         pctl_income_50, pctl_income_50_lb, 
         pctl_income_50_ub, pctl_50_quality, 
         pctl_income_80, pctl_income_80_lb, 
         pctl_income_80_ub,
         pctl_80_quality)

metrics_income_sub <- metrics_income_sub %>% 
  select(year, state, county, subgroup_type, subgroup, pctl_income_20, pctl_income_20_lb, 
         pctl_income_20_ub, pctl_20_quality, 
         pctl_income_50, pctl_income_50_lb, 
         pctl_income_50_ub, pctl_50_quality, 
         pctl_income_80, pctl_income_80_lb, 
         pctl_income_80_ub,
         pctl_80_quality)
```

Export as CSV

```{r}
write_csv(metrics_income_all, here::here("01_financial-well-being", "data", "final", "metrics_income_county_all_longitudinal.csv"))
write_csv(metrics_income_sub, here::here("01_financial-well-being", "data", "final", "metrics_income_county_race-ethnicity_longitudinal.csv"))
```













