---
title: ""
author: ""
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output:
  html_document:
    number_sections: TRUE
    self_contained: TRUE
    code_folding: show
    toc: TRUE
    toc_float: TRUE
    css: !expr here::here("05_local-governance", "www", "web_report.css")
    editor_options:
      chunk_output_type: console
---

<style>
@import url('https://fonts.googleapis.com/css?family=Lato&display=swap');
</style>

# Descriptive Representation

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

This script calculates four measures of descriptive representation for the Urban Institute Mobility Metrics project. The metrics are constructed using the [ICMA Municipal Form of Government 2018 Survey](https://icma.org/2018-municipal-fog-survey) and 5-year ACS data. 

```{r setup}
options(scipen = 999)

library(tidyverse)
library(readxl)
library(censusapi)

source(here::here("05_local-governance", "R", "census_api_key.R"))
source(here::here("05_local-governance", "R", "get_vars.R"))

theme_set(theme_minimal())

```

The Urban Institute purchased the [ICMA Municipal Form of Government 2018 Survey](https://icma.org/2018-municipal-fog-survey). It is available on Box and needs to be added to this directory in a way that matches the filepath below. 

```{r read-icma}
if (!dir.exists("data")) {
  
  dir.create("data")  
  
}

# read in the ICMA survey
icma2018 <- read_excel(here::here("05_local-governance", 
                                  "descriptive-representation", 
                                  "data", 
                                  "Municipal Form of Government 2018 Survey", 
                                  "Municipal Form of Government 2018 Data.xlsx"))

# only keep the relevant ICMA variables
analysis <- icma2018 %>%
  select(ID, 
         LocalGovernment, 
         state_name = State, 
         county_name = County, 
         U18POPest, 
         FIPS_StateCounty, 
         FIPS_Place, 
         starts_with("Q35"))

```

### Recreate Table 35.

We begin by recreating Table 35 from the 2018 Municipal Form of Government Survey Summary of Survey Results. The results match exactly. 

```{r recreate-table35}
# move from wide to long and recode the variable names to race/eth groups
table35 <- analysis %>%
  pivot_longer(cols = starts_with("Q35"),
               names_to = "race_ethnicity",
               values_to = "count") %>%
  mutate(race_ethnicity = case_when(
    race_ethnicity == "Q35_a" ~ "American Indian or Alaska Native",
    race_ethnicity == "Q35_b" ~ "Hispanic or Latino",
    race_ethnicity == "Q35_c" ~ "Asian or Pacific Islander",
    race_ethnicity == "Q35_d" ~ "White, not of Hispanic Origin",
    race_ethnicity == "Q35_e" ~ "Black or African American",
    race_ethnicity == "Q35_f" ~ "Native Hawaiin or Other Pacific Islander",
    race_ethnicity == "Q35_g" ~ "Middle Eastern or North African",
    race_ethnicity == "Q35_h" ~ "Some other race, ethnicity, or origin",
    )
  )

# confirm a = 21,466
table35 %>%
  summarize(sum(count, na.rm = TRUE))

# confirm frequencies from table 35
table35 %>%
  group_by(race_ethnicity) %>%
  summarize(count = sum(count, na.rm = TRUE)) %>%
  mutate(prop = count / sum(count))

```

## Descriptive Representation Metrics

The metric is the ratio of the share of local elected officials of a racial/ethnic group to the share of residents of the same racial/ethnic group. The four racial/ethnic groups of interest are Asian/other, Black non-Hispanic, Latinx, and white non-Hispanic. 

We will use eight steps to create the four metrics:

1. Get shares of 4 race/ethnicity groups at the council level from the ICMA survey
2. Pull demographics for Census places and Census County Subdivisions 
3. Join the Census data to the ICMA data
4. Calculate the metrics
5. Aggregate to the county level
6. Identify issues
7. Create quality flags
8. Save the data

### 1. Get shares of 4 race/ethnicity groups at the council level from the ICMA survey

The counts of councilmembers of different racial and ethnic groups are expressed as a series of variables beginning with `Q35_`. We rename the variables. The total number of councilmembers is not expressed as an integer in the survey, so we sum the groups. 

```{r}
# replace missing values with zeros
analysis <- analysis %>%
  mutate(across(starts_with("Q35"), .fns = ~replace_na(data = .x, replace = 0))) %>%
  rename(
    american_indian_alaskan_native = Q35_a,
    hispanic_or_latino = Q35_b,
    asian_or_pacific_islander = Q35_c,
    white_nonhispanic = Q35_d,
    black =  Q35_e,
    native_hawaiin_or_pacific_islander = Q35_f,
    middle_eastern_or_north_african = Q35_g,
    other = Q35_h
  )

# sum up the councilmembers
analysis <- analysis %>%
  mutate(council_members = 
           american_indian_alaskan_native +
           hispanic_or_latino +
           asian_or_pacific_islander +
           white_nonhispanic +
           black +
           native_hawaiin_or_pacific_islander +
           middle_eastern_or_north_african +
           other) 

```

We collapse the eight racial/ethnic groups from ICMA into the four groups of interest. 

```{r}
# collapse into 4 groups
analysis <- analysis %>%
  mutate(
    council_asian_other = american_indian_alaskan_native +
      asian_or_pacific_islander +
      native_hawaiin_or_pacific_islander +
      middle_eastern_or_north_african +
      other,
    council_black_nonhispanic = black,
    council_hispanic = hispanic_or_latino, 
    council_white_nonhispanic = white_nonhispanic
  ) 

# drop old variables
analysis <- analysis %>%
  select(-american_indian_alaskan_native,
         -hispanic_or_latino,
         -asian_or_pacific_islander,
         -black,
         -native_hawaiin_or_pacific_islander,
         -middle_eastern_or_north_african,
         -other,
         -white_nonhispanic)

```

The ICMA survey does not distinguish between `0` and `NA`. We calculate the share of councilmembers by race and turn council shares into `NA` for municipalities that have zero councilmembers. 

```{r}
calc_council_proportion <- function(council_race, council_members) {
  
  if_else(condition = council_members == 0, 
              true = as.numeric(NA), 
              false = {{council_race}} / council_members)
  
}

analysis <- analysis %>%
  mutate(
    council_asian_other = calc_council_proportion(council_asian_other, 
                                                  council_members),
    council_black_nonhispanic = calc_council_proportion(council_black_nonhispanic, 
                                                  council_members),
    council_hispanic = calc_council_proportion(council_hispanic, 
                                                  council_members),
    council_white_nonhispanic = calc_council_proportion(council_white_nonhispanic, 
                                                  council_members)
  )

```

We test to ensure that the ratios are bounded by 0 and 1 inclusive. 

```{r}
# bounded by 0 and 1
stopifnot(
  analysis %>%
    filter(
      council_asian_other < 0 | council_asian_other > 1 |
        council_black_nonhispanic < 0 | council_black_nonhispanic > 1|
        council_hispanic < 0 | council_hispanic > 1 |
        council_white_nonhispanic < 0 | council_white_nonhispanic > 1
    ) %>% 
    nrow() == 0
)

```

* todo(aaron): Test number of councilmembers
* todo(aaron): Compare with other questions 

### 2. Pull demographics for Census places and Census County Subdivisions

We pull the share of the population that belongs to different racial/ethnic in Census places and Census county subdivisions. 

We pull all of the race/ethnicity counts for 2018 using `library(censusapi)`. **Note:** This will require a [Census API key](https://api.census.gov/data/key_signup.html). Add the key to `census_api_key-template.R` and then delete then delete "template". It is sourced above. 

```{r}
# list variable of interest
vars <- c(
  # Hispanic or Latino
  "DP05_0071E", # Estimate!!HISPANIC OR LATINO AND RACE!!Total population!!Hispanic or Latino (of any race)
  "DP05_0071M", 
  # Not Hispanic or Latino
  "DP05_0077E", # White alone
  "DP05_0077M",
  "DP05_0078E", # Black or African American alone
  "DP05_0078M",
  "DP05_0079E", # American Indian and Alaska Native alone
  "DP05_0079M",
  "DP05_0080E", # Asian alone
  "DP05_0080M",
  "DP05_0081E", # Native Hawaiian and Other Pacific Islander alone
  "DP05_0081M",
  "DP05_0082E", # Some other race alone
  "DP05_0082M",
  "DP05_0083E", # Two or more races
  "DP05_0083M"
)

# pull ACS data at the Census Place and Census County Subdivision levels
places <- get_vars(year = 2018, vars = vars, geography = "place")
county_subdivisions <- get_vars(year = 2018, vars = vars, geography = "county subdivision")

# combine the two geographies
geographies <- bind_rows(places, county_subdivisions)

rm(places, county_subdivisions)

# give meaningful names
geographies <- geographies %>%
  rename(
    people = B01003_001E,
    hispanic = DP05_0071E, 
    hispanic_moe = DP05_0071M,
    white_nonhispanic = DP05_0077E, 
    white_nonhispanic_moe = DP05_0077M,
    black_nonhispanic = DP05_0078E, 
    black_nonhispanic_moe = DP05_0078M, 
    aian_nh = DP05_0079E, 
    aian_nh_moe = DP05_0079M,
    asian_nh = DP05_0080E, 
    asian_nh_moe = DP05_0080M,
    nhpi_nh = DP05_0081E, 
    nhpi_nh_moe = DP05_0081M,
    other_nh = DP05_0082E, 
    other_nh_moe = DP05_0082M, 
    two_or_more_nh = DP05_0083E,
    two_or_more_nh_moe = DP05_0083M
  )

```

The Census reports race and ethnicity as many different groups. Table 2 in [this publication](https://www.census.gov/prod/cen2010/briefs/c2010br-02.pdf) is helpful. Here, we collapse the detailed groups into the same four groups of interest from the above section. 

```{r}
# construct asian_other 
geographies <- geographies %>%
  mutate(asian_other = aian_nh + asian_nh + nhpi_nh + other_nh + two_or_more_nh) %>%
  select(state, 
         county, 
         place, 
         county_subdivision, 
         people, asian_other, 
         black_nonhispanic, 
         hispanic, 
         white_nonhispanic, 
         geography)

# test that the new groups sum to the original people total
stopifnot(
  geographies %>%
    mutate(people2 = asian_other + black_nonhispanic + hispanic + white_nonhispanic) %>%
    filter(people != people2) %>%
    nrow() == 0
)

```

We calculate the share of municipalities by race. 

```{r}
geographies <- geographies %>%
  mutate(
    asian_other = asian_other / people, 
    black_nonhispanic = black_nonhispanic / people, 
    hispanic = hispanic / people,
    white_nonhispanic = white_nonhispanic / people
  )

# test to see if bounded by 0 and 1
stopifnot(
  geographies %>%
    filter(
      asian_other < 0 | asian_other > 1 |
        black_nonhispanic < 0 | black_nonhispanic > 1|
        hispanic < 0 | hispanic > 1 |
        white_nonhispanic < 0 | white_nonhispanic > 1
    ) %>% 
    nrow() == 0
)

# test to see if the 4 categories sum to 1
stopifnot(
  geographies %>%
    mutate(total = round(asian_other + black_nonhispanic + hispanic + white_nonhispanic, 10)) %>%
    filter(total != 1) %>%
    nrow() == 0
)

```

### 3. Join the Census data to the ICMA data

`FIPS_Place` in the ICMA data is actually combination of Census Places and Census County Subdivisions. Here's a note from Will Fricke at ICMA:

> I meant to add this in the READ ME txt file: The FIPS_place column includes both place and county subdivision codes (I suppose the column should have a different name) – if you go back through ICMA datasets you’ll find it’s consistently done that way regardless of the wisdom of it.
 
> The method I use to merge census data into our datasets is to use a vlookup in excel from the FIPS_place column to the Census data place codes column, and then the same thing from the N/A rows to the COUSUB codes column. (Check to make sure you have enough RAM first…)

We pull a list of all Census geographies from [here](https://www.census.gov/geographies/reference-files.html). 

```{r}
# load a list of all FIPS
fips <- read_excel(here::here("05_local-governance", 
                              "descriptive-representation", 
                              "fips", "all-geocodes-v2018.xlsx"), 
                   skip = 4) %>%
  mutate(FIPS_Place = paste0(`State Code (FIPS)`, `Place Code (FIPS)`))

# create 7- and 10- digit codes in fips
fips <- fips %>%
  mutate(place_code = paste0(`State Code (FIPS)`, `Place Code (FIPS)`),
         county_subdivision_code = paste0(`State Code (FIPS)`, 
                                          `County Code (FIPS)`, 
                                          `County Subdivision Code (FIPS)`))

# create 7- and 10- digit codes in geographies
geographies <- geographies %>%
  mutate(place_code = paste0(state, place),
         county_subdivision_code = str_c(state, county, county_subdivision))
```

If a locality from the ICMA data is a Census place then assign Census place. If a locality isn't a Census place then assign Census County Subdivision. 

```{r}
# if a place, use place
analysis <- analysis %>%
  # create a potential code
  mutate(potential_place = str_c(str_sub(FIPS_StateCounty, start = 1, end = 2), FIPS_Place)) %>%
  mutate(place_code = if_else(potential_place %in% pull(fips, place_code),
                              true = potential_place,
                              false = as.character(NA)))

# if not a place, then a county subdivision
analysis <- analysis %>%
  # create a potential code
  mutate(potential_county_subdivision_code = str_c(FIPS_StateCounty, FIPS_Place)) %>%
  mutate(county_subdivision_code = if_else(potential_county_subdivision_code %in% pull(fips, county_subdivision_code),
                                          true = potential_county_subdivision_code,
                                          false = as.character(NA)))

# create geography variable
analysis <- analysis %>%
  select(-potential_place, -potential_county_subdivision_code) %>%
  mutate(geography = case_when(
    !is.na(place_code) ~ "place",
    !is.na(county_subdivision_code) ~ "county subdivision"
    ))

# cases without a place or a county subdivision
trouble <- tidylog::filter(analysis, is.na(geography)) %>%
  mutate(geography = "trouble")

# join the Census places
places <- geographies %>%
  tidylog::filter(geography == "place") %>%
  select(-county, -county_subdivision, -county_subdivision_code)

places <- analysis %>%
  select(-county_subdivision_code) %>%
  tidylog::filter(geography == "place") %>%
  left_join(places, by = c("place_code", "geography"))
```

Combine the three data (places, county subdivisions, trouble) into one file. 

```{r}
# join the Census county subdivisions
county_subdivisions <- geographies %>% 
  tidylog::filter(geography == "county subdivision") %>%
  select(-place, -place_code)

county_subdivisions <- analysis %>%
  tidylog::filter(geography == "county subdivision") %>%
  select(-place_code) %>%
  left_join(county_subdivisions, by = c("county_subdivision_code", "geography"))

# combine three joined data sets into one data set
joined_data <- bind_rows(
  places,
  county_subdivisions,
  trouble
) %>%
  mutate(state = str_sub(FIPS_StateCounty, start = 1, end = 2),
         county = str_sub(FIPS_StateCounty, start = 3, end = 5))

```

Quality checks:

```{r}
ggplot(joined_data, aes(U18POPest, people)) +
  geom_point(alpha = 0.2) +
  labs(title = "The ICMA and Census data have similar estimated populations")

```

### 4. Calculate the metrics

Find the ratio of the proportions. Conditional logic is needed to avoid division by `0`.

```{r}
joined_data <- joined_data %>%
  mutate(
    asian_other = 
      if_else(condition = asian_other == 0, 
              true = 0, 
              false = council_asian_other / asian_other), 
    black_nonhispanic = 
      if_else(condition = black_nonhispanic == 0, 
              true = 0, 
              false = council_black_nonhispanic / black_nonhispanic), 
    hispanic = 
      if_else(condition = hispanic == 0, 
              true = 0, 
              false = council_hispanic / hispanic),
    white_nonhispanic = 
      if_else(condition = white_nonhispanic == 0, 
              true = 0, 
              false = council_white_nonhispanic / white_nonhispanic)
  ) %>%
  select(-council_asian_other, 
         -council_black_nonhispanic, 
         -council_hispanic, 
         -council_white_nonhispanic)

```

Ensure that there are no undefined numbers.

```{r}
stopifnot(
  filter(joined_data, is.nan(black_nonhispanic)) %>%
    nrow() == 0
)

```

### 5. Aggregate to the county level

We now have the four metrics of interest for municipalities. We need to roll these up to counties. 

```{r}
# drop municipalities that don't report ICMA data and drop municipalities without matches
joined_data <- joined_data %>%
  tidylog::filter(council_members > 0) %>%
  tidylog::filter(geography != "trouble")

# find the person-weighted average of the four metrics at the county level
county_data <- joined_data %>%
  group_by(state, county) %>%
  summarize(municipalities = n(),
            asian_other = sum(asian_other * people) / sum(people), 
            black_nonhispanic = sum(black_nonhispanic * people) / sum(people),
            hispanic = sum(hispanic * people) / sum(people),
            white_nonhispanic = sum(white_nonhispanic * people) / sum(people),
            people = sum(people)) %>%
  ungroup()
```

We only have a fraction of the county population for many counties. We calculate the proportion represented.

```{r}
# pull the county population using tidycensus
county_pop <- tidycensus::get_acs(geography = "county",
                                  variable = c(people = "B01003_001E"),
                                  year = 2018) %>%
  mutate(state = str_sub(GEOID, start = 1, end = 2),
         county = str_sub(GEOID, start = 3, end = 5)) %>%
  select(county_state = NAME,
         state,
         county,
         people_county = estimate)

# calculate the share of the county population represented
county_data <- left_join(county_data, county_pop, by = c("state", "county"))

county_data <- county_data %>%
  mutate(population_prop = people / people_county)

```

Quality checks:

```{r}
ggplot(county_data, aes(people_county, people)) +
  geom_point(alpha = 0.2) +
  geom_abline(slope = 1) +
  coord_equal() +
  labs(title = "The Propotion of Population Represented in Counties is Underwhelming",
       subtitle = "Comparison of County Population to Represented Population")

county_data %>%
  ggplot(aes(population_prop)) +
  geom_histogram() +
  labs(title = "The Propotion of Population Represented in Counties is Underwhelming",
       subtitle = "Distribution of Population Representation")
```

### 6. Identify issues

### 7. Create quality flags

The municipality-level data are accurate and useful. The representativeness of the county-level data is questionable because many municipalities are missing from the summarized data. 

* `1` if > 80% of population is captured. 
* `2` if > 50% of the population is captured. 
* `3` otherwise.

```{r}
# create a helper function for data quality
set_quality <- function(proportion) {
  
  quality <- case_when(
    proportion >= 0.8 ~ 1,
    proportion >= 0.5 ~ 2,
    proportion >= 0 ~ 3
  )
  
  return(quality)
  
}

# quality flags
county_data <- county_data %>%
  mutate(
    asian_other_quality = set_quality(proportion = population_prop),
    black_nonhispanic_quality = set_quality(proportion = population_prop),
    hispanic_quality = set_quality(proportion = population_prop),
    white_nonhispanic_quality = set_quality(proportion = population_prop)
  )

count(county_data, asian_other_quality)
count(county_data, black_nonhispanic_quality)
count(county_data, hispanic_quality)
count(county_data, white_nonhispanic_quality)

```

### 8. Save the data

We need to include all counties in the published data even if we don't have a metric for the county. We load the county file and join our metrics to the county file. 

```{r}
# load the 2018 county file
all_counties <- read_csv(here::here("geographic-crosswalks", "data", "county-file.csv")) %>%
  tidylog::filter(year == 2018)

# join the calculated metrics to the full file
all_counties <- left_join(all_counties, county_data, by = c("state", "county"))

# final data set
all_counties <- all_counties %>%
  mutate(year = 2018) %>%
  select(year, 
         state, 
         county, 
         asian_other, 
         asian_other_quality, 
         black_nonhispanic, 
         black_nonhispanic_quality, 
         hispanic, 
         hispanic_quality, 
         white_nonhispanic, 
         white_nonhispanic_quality)

write_csv(all_counties, here::here("05_local-governance", 
                                   "descriptive-representation", 
                                   "descriptive-representation.csv"))

```
