---
title: "Final Data Evaluation Homelessness Data"
author: "JP Walsh"
date: today
format:
  html:
    toc: true
    toc_float: true
    embed-resources: true
    code-fold: show
execute: 
  warning: false
editor_options: 
  chunk_output_type: console
---

# Final Data Test Function

The data team for the Mobility Metrics project has created a function that tests baseline requirements for the structure and content of data produced for each metric. This function relies in part on information provided in a metric-specific final data expectations form, which have been filled out prior to final approval by their technical reviewer. 

These evaluation forms are saved in `10a_final-evaluation`

For programs written in STATA the evaluation process is run externally in this QMD file. 

Detail:

 - *Metric: Homeless Students* 
 - *Tester: JP Walsh*
 - *Test date: March 2025*

# Run Test

Put the data through the final evaluation test.

## Read Function

The first step is to source the function `evaluate_final_data.R`. 

```{r}
source(here::here("functions", "testing", "evaluate_final_data.R"))
```


## Read Final Data

This file tests the final data for the students experiencing homelessness metric. There are four final files to be tested:

 - `02_housing/data/final/homelessness_2019-2022_city.csv`
 - `02_housing/data/final/homelessness_2019-2022_county.csv`
 - `02_housing/data/final/homelessness_2019-2022_subgroups_city.csv`
 - `02_housing/data/final/homelessness_2019-2022_subgroups_county.csv`

Read these final files in.

```{r}
final_data_city <- read_csv(here("02_housing/data/final/homelessness_2019-2022_city.csv"), show_col_types = FALSE)

final_data_county <- read_csv(here("02_housing/data/final/homelessness_2019-2022_county.csv"), show_col_types = FALSE)

final_data_city_race <- read_csv(here("02_housing/data/final/homelessness_2019-2022_subgroups_city.csv"), show_col_types = FALSE)

final_data_county_race <- read_csv(here("02_housing/data/final/homelessness_2019-2022_subgroups_county.csv"), show_col_types = FALSE)

```


## Run Evaluation Test

Enter the information into the `evaluate_final_data` function. 

In order to run the function, we fill out the relevant arguments. Homeless students data do have confidence intervals and there are subgroup files.

### All

City level data.
```{r, error = TRUE}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_homeless_student_place.csv",
                data = final_data_city,  geography = "place",
                subgroups = FALSE, confidence_intervals = TRUE)
```

County level data.
```{r, error = TRUE}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_homeless_student_county.csv",
                data = final_data_county,  geography = "county",
                subgroups = FALSE, confidence_intervals = TRUE)

```

### Subgroups

City level race data.
```{r, error = TRUE}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_homeless_student_race_eth_place.csv",
                data = final_data_city_race,  geography = "place",
                subgroups = TRUE, confidence_intervals = TRUE)
```

County level race data.
```{r, error = TRUE}
evaluate_final_data(exp_form_path = "10a_final-evaluation/evaluation_form_homeless_student_race_eth_county.csv",
                data = final_data_county_race,  geography = "county",
                subgroups = TRUE, confidence_intervals = TRUE)

```


Failed for a passable reason. The test was built to check the sum of missing confidence interval values against the sum of missing quality values. These data have two quality variables (share and count) but only one set of confidence intervals. Tweak this in the future. All other parts pass.
