---
title: "Comparison of Ratios of Housing Units Affordable and Available Outputs Across Updates"
format: 
  html:
    toc: true
    code-line-numbers: true
    embed-resources: true
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This file includes quick scatterplots to compare the changes between the 2023-2024 version (led by Amy Rogin) of the housing affordability/availability metrics and the updated 2024-2025 version (led by Will Curran-Groome). The goal is to summarize the effect of updates, specifically adjusting for the number of bedrooms and household size in generating the AMI indicators.

These scatterplots only look at the availability metric (i.e., the ratio between housing units affordable and available at various income levels and the number of households at or below those income levels) for 2022. The plots are broken out by AMI level (30%, 50%, 80%), subgroup (owner, renter), and for both counties and places/cities. 

## Configuration and setup

```{r}
library(tidyverse)
library(here)
library(urbnthemes)
options(scipen = 99999) ## no scientific notation
```

```{r}
create_comparison_scatterplot <- function(geo, version) {
  # Last year, the "place" file name used "city" 
  if (geo == "place") {geo_previous <- "city"} else {geo_previous <- geo}
  
  # Specify output files/paths and 
  if (version == "bed_adjust") {
    years <- years <- c(2014:2019, 2021:2023) |> paste0(collapse = "_")
    label <- "Bedroom adjustment"
    path <- "final"
  } else if (version == "no_adjust") {
    years <- 2022 # Only outputted 2022 values without bedroom adjustment for now 
    label <- "No bedroom adjustment"
    path <- "no_bedadj"
  } else if (version == "bed_hhsize_adjust") {
    years <- 2022 # Only outputted 2022 values without bedroom adjustment for now 
    label <- "Bedroom and household size adjustment"
    path <- "bedadj_hhsizeadj"
  }

  # Read previous year's data 
  previous <- read_csv(
  str_c(
    "https://raw.githubusercontent.com/UI-Research/mobility-from-poverty/refs/heads/main/02_housing/data/", 
    "available_2022_subgroups_", geo_previous, ".csv"
  )) |> 
  # Update variable names to match current structure 
  rename_with(~ str_replace(.x, "share_affordable_available_(\\d+)_ami", "share_available_\\1ami")) |> 
    pivot_longer(cols = starts_with("share_")) |> 
    rename("value_previous" = "value")

  # Read updated data 
  updated <- read_csv(
    here("02_housing", "data", path, str_c("housing_", geo, "_", years, "_subgroups.csv"))) |> 
    pivot_longer(cols = starts_with("share_")) |> 
    rename("value_updated" = "value")
  
  # Combine and plot 
  previous |> 
    left_join(updated, by = c("year", "state", geo, "subgroup_type", "subgroup", "name")) |> 
    filter(!str_detect(name, "quality")) |> 
    filter(year == 2022, subgroup != "All") |> 
    ggplot() +
    geom_point(aes(x = value_previous, y = value_updated, color = subgroup), alpha = 0.5, size = 0.5) + 
    facet_wrap(~ name) + 
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
    labs(
      title = str_c("Comparison of ratios (", label, ", ", geo, ", 2022)"), 
      y = "Updated", 
      x = "Previous") +
    theme_urbn_print() 
}
```

## No bedroom adjustment or household size adjustment 

```{r, fig.width=10}
create_comparison_scatterplot(geo = "county", version = "no_adjust") 
create_comparison_scatterplot(geo = "place", version = "no_adjust")
```


## With bedroom adjustment but without household size adjustment 

```{r, fig.width=10}
create_comparison_scatterplot(geo = "county", version = "bed_adjust")
create_comparison_scatterplot(geo = "place", version = "bed_adjust") 
```

## With bedroom adjustment and household size adjustment 

```{r, fig.width=10}
create_comparison_scatterplot(geo = "county", version = "bed_hhsize_adjust")
create_comparison_scatterplot(geo = "place", version = "bed_hhsize_adjust") 
```

